### 2024-08-30 code/manuscript.Rmd
Restructured code, added rmd file and paper folder, updated readfile and output folder path

---
title: "Reliably estimating the impact of a new control strategy in a building"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA] ^[Correspondence to aoyuzou@berkeley.edu], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Durate ^1^
abstract: Conventional measurement and verification (M&V) methods for whole-building energy savings estimation are both time-consuming and unreliable, especially when non-routine events occur during the M&V process. Those events are unrelated to the proposed intervention strategy but have substaintial impacts on the building energy consumption. In this study, we argue that for switchable interventions (e.g. most of the control retrofits) can benefit from random sampling where the analyst randomly decide which strategy (i.e. baseline or intervention) to implement each day. We tested the novel randomized M&V method on a large public dataset which covers multiple climate zones and types of commercial buildings. We applied a virtual chilled water supply temperature reset based on outdoor weather as a control retrofit intervention. Our study shows that the new M&V method can estimate the savings accurately much quicker than the conventional method and most importantly, the estimation results are much more robust compared to the conventional method when non-routine events are present.

output:
  bookdown::word_document2:
    reference_docx: "../paper/template.docx"
editor_options: 
  markdown: 
    wrap: 72
# bibliography: references.bib
# biblio-style: apalike

knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../paper"
    )
  })
---

```{r setup, include = FALSE, cache = FALSE}

# knitr setup
knitr::opts_chunk$set(echo = F, 
                      message = F,
                      warning = F,
                      dev = "jpeg",
                      cache = T,
                      dpi = 300,
                      fig.path = "figs",
                      fig.show = "hold",
                      fig.pos = "b", 
                      fig.path = "../figs/results/manuscript")

# str(knitr::opts_chunk$get()) # see for all options

```

```{r libraries, include = FALSE, cache = FALSE}

### 2024-08-30 code/0_messydata.R

@@ -35,14 +35,19 @@ perform_t_test <- function(df_pre, df_post, building) {
   t.test(df_pre[[building]], df_post[[building]], alternative = "two.sided")
 }
 
+perform_mean_diff <- function(df_pre, df_post, building) {
+  abs(mean(df_pre[[building]], na.rm = T) - mean(df_post[[building]], na.rm = T)) / mean(df_pre[[building]], na.rm = T) * 100
+}
 
-
+perform_mean_est <- function(df_pre, df_post, building) {
+  0.5 * (sum(df_pre[[building]], na.rm = T) + sum(df_post[[building]], na.rm = T))
+}
 
 #### READ DATA ####
-data_path <- "../buds-lab-building-data-genome-project-2/data/meters/cleaned/"
-meta_path <- "../buds-lab-building-data-genome-project-2/data/metadata/"
-weather_path <- "../buds-lab-building-data-genome-project-2/data/weather/"
-output_path <- "./readfiles/tidy/"
+data_path <- "../../buds-lab-building-data-genome-project-2/data/meters/cleaned/"
+meta_path <- "../../buds-lab-building-data-genome-project-2/data/metadata/"
+weather_path <- "../../buds-lab-building-data-genome-project-2/data/weather/"
+output_path <- "../readfiles/messy/"
 
 df_elec <- read_csv(paste0(data_path, "electricity_cleaned.csv"))
 df_meta <- read_csv(paste0(meta_path, "metadata.csv"))
@@ -53,8 +58,8 @@ df_weather <- read_csv(paste0(weather_path, "weather.csv"))
 
 #### FILTER ####
 # Focus
-remove_type <- c("parking", "warehouse", "utility")
-remove_site <- c("Eagle", "Bobcat", "Swan", "Hog")
+remove_type <- c("parking", "warehouse", "other")
+remove_site <- c("Eagle", "Bobcat", "Swan", "Hog", "Gator")
 
 # NAs
 na_counts <- sapply(df_elec[-1], function(x) sum(is.na(x)))
@@ -86,11 +91,28 @@ p_values <- sapply(buildings, function(building) {
   t_test_result$p.value
 })
 
-results <- data.frame(Building = buildings, P_Value = p_values) %>% 
-  filter(P_Value >= 0.05)
+mean_diff <- sapply(buildings, function(building) {
+  perform_mean_diff(df_pre, df_post, building)
+})
+
[…]

### 2024-08-30 code/0_tidydata.R

@@ -35,19 +35,14 @@ perform_t_test <- function(df_pre, df_post, building) {
   t.test(df_pre[[building]], df_post[[building]], alternative = "two.sided")
 }
 
-perform_mean_diff <- function(df_pre, df_post, building) {
-  abs(mean(df_pre[[building]], na.rm = T) - mean(df_post[[building]], na.rm = T)) / mean(df_pre[[building]], na.rm = T) * 100
-}
 
-perform_mean_est <- function(df_pre, df_post, building) {
-  0.5 * (sum(df_pre[[building]], na.rm = T) + sum(df_post[[building]], na.rm = T))
-}
+
 
 #### READ DATA ####
-data_path <- "../buds-lab-building-data-genome-project-2/data/meters/cleaned/"
-meta_path <- "../buds-lab-building-data-genome-project-2/data/metadata/"
-weather_path <- "../buds-lab-building-data-genome-project-2/data/weather/"
-output_path <- "./readfiles/messy/"
+data_path <- "../../buds-lab-building-data-genome-project-2/data/meters/cleaned/"
+meta_path <- "../../buds-lab-building-data-genome-project-2/data/metadata/"
+weather_path <- "../../buds-lab-building-data-genome-project-2/data/weather/"
+output_path <- "../readfiles/tidy/"
 
 df_elec <- read_csv(paste0(data_path, "electricity_cleaned.csv"))
 df_meta <- read_csv(paste0(meta_path, "metadata.csv"))
@@ -58,8 +53,8 @@ df_weather <- read_csv(paste0(weather_path, "weather.csv"))
 
 #### FILTER ####
 # Focus
-remove_type <- c("parking", "warehouse", "other")
-remove_site <- c("Eagle", "Bobcat", "Swan", "Hog", "Gator")
+remove_type <- c("parking", "warehouse", "utility")
+remove_site <- c("Eagle", "Bobcat", "Swan", "Hog")
 
 # NAs
 na_counts <- sapply(df_elec[-1], function(x) sum(is.na(x)))
@@ -91,28 +86,11 @@ p_values <- sapply(buildings, function(building) {
   t_test_result$p.value
 })
 
-mean_diff <- sapply(buildings, function(building) {
-  perform_mean_diff(df_pre, df_post, building)
-})
-
-mean_est <- sapply(buildings, function(building) {
-  perform_mean_est(df_pre, df_post, building)
-})
-
[…]

### 2024-08-30 code/1_analysis.R

@@ -40,7 +40,7 @@ ls_colors <- c("Baseline" = "#99d8c9",
                "True normalized savings" = "black")
 
 # load functions
-function_path <- "./functions/"
+function_path <- "../functions/"
 source(paste0(function_path, "blocking.R"))
 source(paste0(function_path, "model_fit.R"))
 source(paste0(function_path, "model_pred.R"))
@@ -289,9 +289,9 @@ get_frsaving <- function(timeline, dataframe){
 
 
 #### READ DATA ####
-readfile_path <- str_glue("./readfiles/{run_params$type}/")
-summaryfigs_path <- str_glue("./figs/{run_params$type}/site_summary/")
-combifigs_path <- str_glue("./figs/{run_params$type}/comb_analysis/")
+readfile_path <- str_glue("../readfiles/{run_params$type}/")
+summaryfigs_path <- str_glue("../figs/{run_params$type}/site_summary/")
+combifigs_path <- str_glue("../figs/{run_params$type}/comb_analysis/")
 
 df_energy <- read_rds(paste0(readfile_path, "df_energy.rds"))
 df_meta <- read_rds(paste0(readfile_path, "df_meta.rds"))
@@ -345,8 +345,8 @@ for (n in 1:(nrow(all_names))){
   
   site <- site_info$site
   
-  ifelse(!dir.exists(file.path(str_glue("./figs/{run_params$type}/site_analysis/{site}/{name}"))), dir.create(file.path(str_glue("./figs/{run_params$type}/site_analysis/{site}/{name}"))), FALSE)
-  sitefigs_path <- str_glue("./figs/{run_params$type}/site_analysis/{site}/{name}")
+  ifelse(!dir.exists(file.path(str_glue("../figs/{run_params$type}/site_analysis/{site}/{name}"))), dir.create(file.path(str_glue("../figs/{run_params$type}/site_analysis/{site}/{name}"))), FALSE)
+  sitefigs_path <- str_glue("../figs/{run_params$type}/site_analysis/{site}/{name}")
   
   site_weather <- df_weather %>%
     filter(site == site_info$site) %>%
@@ -887,8 +887,8 @@ if (run_params$nsprt) {
     
     site <- site_info$site
     
-    ifelse(!dir.exists(file.path(str_glue("./figs/{run_params$type}/site_analysis/{site}/{name}"))), dir.create(file.path(str_glue("./figs/{run_params$type}/site_analysis/{site}/{name}"))), FALSE)
[…]

### 2024-08-30 code/2_messyvisual.R

@@ -39,7 +39,7 @@ ls_colors <- c("Baseline" = "#99d8c9",
                "True normalized savings" = "black")
 
 # define parameters
-run_params <- list(type = "tidy", 
+run_params <- list(type = "messy", 
                    site = T)
 
 sprt_param <- list(baseline = "Baseline",
@@ -52,7 +52,6 @@ block_params <- list(start_date = "2016-01-01",
                      n_weeks = 108,
                      n_seasons = 9, 
                      block_unit = 12)
-
 
 
 
@@ -78,31 +77,24 @@ get_tmy <- function(all_sites){
   return(df_tmy)
 }
 
-
-
-
 #### READ DATA ####
-readfile_path <- str_glue("./readfiles/{run_params$type}/")
-summaryfigs_path <- str_glue("./figs/{run_params$type}/site_summary/")
-combifigs_path <- str_glue("./figs/{run_params$type}/comb_analysis/")
+readfile_path <- str_glue("../readfiles/{run_params$type}/")
+summaryfigs_path <- str_glue("../figs/{run_params$type}/site_summary/")
+combifigs_path <- str_glue("../figs/{run_params$type}/comb_analysis/")
 
 df_energy <- read_rds(paste0(readfile_path, "df_energy.rds"))
 df_meta <- read_rds(paste0(readfile_path, "df_meta.rds"))
 df_weather <- read_rds(paste0(readfile_path, "df_weather.rds"))
 df_sprt_all <- read_rds(paste0(readfile_path, "df_sprt_all.rds"))
 df_seq_FS <- read_rds(paste0(readfile_path, "df_seq_FS.rds"))
-
-if (run_params$type == "tidy"){
-  
-  df_NRE_occ <- read_rds(paste0(readfile_path, "df_NRE_occ.rds"))
-  
-}
-
 df_MD <- read_rds(paste0(readfile_path, "df_MD.rds"))
 df_FS <- read_rds(paste0(readfile_path, "df_FS.rds"))
 df_eui <- read_rds(paste0(readfile_path, "df_eui.rds"))
 df_cont_MD <- read_rds(paste0(readfile_path, "df_cont_MD.rds"))
 df_cont_FS <- read_rds(paste0(readfile_path, "df_cont_FS.rds"))
+df_FS_nsprt <- read_rds(paste0(readfile_path, "df_FS_nsprt.rds"))
[…]

### 2024-08-30 code/2_tidyvisual.R

@@ -39,7 +39,7 @@ ls_colors <- c("Baseline" = "#99d8c9",
                "True normalized savings" = "black")
 
 # define parameters
-run_params <- list(type = "messy", 
+run_params <- list(type = "tidy", 
                    site = T)
 
 sprt_param <- list(baseline = "Baseline",
@@ -52,6 +52,7 @@ block_params <- list(start_date = "2016-01-01",
                      n_weeks = 108,
                      n_seasons = 9, 
                      block_unit = 12)
+
 
 
 
@@ -77,10 +78,13 @@ get_tmy <- function(all_sites){
   return(df_tmy)
 }
 
+
+
+
 #### READ DATA ####
-readfile_path <- str_glue("./readfiles/{run_params$type}/")
-summaryfigs_path <- str_glue("./figs/{run_params$type}/site_summary/")
-combifigs_path <- str_glue("./figs/{run_params$type}/comb_analysis/")
+readfile_path <- str_glue("../readfiles/{run_params$type}/")
+summaryfigs_path <- str_glue("../figs/{run_params$type}/site_summary/")
+combifigs_path <- str_glue("../figs/{run_params$type}/comb_analysis/")
 
 df_energy <- read_rds(paste0(readfile_path, "df_energy.rds"))
 df_meta <- read_rds(paste0(readfile_path, "df_meta.rds"))
@@ -133,7 +137,7 @@ if (run_params$site){
     mutate(type = fct_reorder(type, n, .desc = F)) %>%
     ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
     geom_col() +
-    geom_text(aes(label = ifelse(n > 10, as.character(n), "")), color = "black", position = position_stack(vjust = 0.5)) +
+    geom_text(aes(label = n), color = "black", position = position_stack(vjust = 0.5)) +
     scale_y_continuous(expand = c(0, 0),
                        breaks = breaks_pretty(n = 4)) +
     scale_x_discrete(expand = c(0, 0.1)) +
@@ -165,7 +169,7 @@ if (run_params$site){
     geom_rect() +
     coord_polar(theta = "y") +
     xlim(c(2, 4)) +
-    facet_wrap(~ site, nrow = 3) +
+    facet_wrap(~ site, nrow = 2) +
     labs(x = NULL,
          y = NULL,
          subtitle = "For each site",
@@ -184,7 +188,7 @@ if (run_params$site){
             legend = "bottom") +
     plot_annotation(title = "Case study building type summary")
   
[…]

### 2024-08-30 functions/model_fit.R

@@ -2,8 +2,8 @@ library(tidyverse)
 library(lubridate)
 library(nmecr)
 library(dplyr)
-source("./functions/resample.R")
-source("./functions/find_occ_unocc.R")
+source("../functions/resample.R")
+source("../functions/find_occ_unocc.R")
 
 model_fit <- function(train_data){
   

### 2024-08-30 functions/model_pred.R

@@ -2,7 +2,7 @@ library(tidyverse)
 library(lubridate)
 library(stats)
 library(dplyr)
-source("./functions/create_temp_matrix.R")
+source("../functions/create_temp_matrix.R")
 
 model_pred <- function(dataframe, model){
   

### 2024-08-30 functions/saving_norm.R

@@ -1,9 +1,9 @@
 library(tidyverse)
 library(lubridate)
 library(dplyr)
-source("./functions/model_fit.R")
-source("./functions/model_pred.R")
-source("./functions/resample.R")
+source("../functions/model_fit.R")
+source("../functions/model_pred.R")
+source("../functions/resample.R")
 
 saving_norm <- function(dataframe, tmy, method){
   

### 2024-08-30 functions/seq_run.R

@@ -2,7 +2,7 @@ library(tidyverse)
 library(lubridate)
 library(nmecr)
 library(dplyr)
-source("./functions/saving_norm.R")
+source("../functions/saving_norm.R")
 
 seq_run <- function(param, dataframe, tmy){
   

### 2024-09-08 code/manuscript.Rmd
Updated manuscript structure

@@
   bookdown::word_document2:
-    reference_docx: "../paper/template.docx"
+    reference_docx: "../paper/TF_Template_Word_Mac_2011.dotx"
 editor_options: 
@@
                       dev = "jpeg",
-                      cache = T,
+                      # cache = T,
                       dpi = 300,
@@
                       fig.pos = "b", 
-                      fig.path = "../figs/results/manuscript")
+                      fig.path = "../figs/manuscript")
 
@@
 
-```{r libraries, include = FALSE, cache = FALSE}
+```{r prep, include = FALSE, cache = FALSE}
+#### LIBRARIES ####
+require(pacman)
+
+# load packages using pacman
+pacman::p_load(tidyverse, lubridate, here) #effect size
+
+# turn off scientific notation
+options(scipen = 999, digits = 15)
+
+# set directory
+here::i_am("manuscript.rmd")
+
+# set default theme for ggplot
+theme_set(theme_minimal())
+
+# define base ggplot theme
+theme_update(plot.title = element_text(size = 14, colour = "grey20", face = "bold", hjust = 0.5),
+             plot.subtitle = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
+             plot.caption = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5),
+             plot.background = element_rect(fill = "white", colour = NA),
+             panel.grid.minor = element_blank(),
+             panel.grid.major = element_blank(),
+             axis.text = element_text(size = 10),
+             strip.text = element_text(size = 10, color = "grey20", face = "bold"),
+             strip.background = element_blank())
+
+ls_colors <- c("Baseline" = "#99d8c9",
+               "Measured baseline" = "#99d8c9",
+               "Adjusted baseline" = "#99d8c9",
+               "Projected baseline\n(no change)" = "#1b9e77",
+               "Intervention" = "#fdbb84",
+               "Measured interv" = "#fdbb84",
+               "True savings" = "black",
+               "True normalized savings" = "black")
+
+#### FUNCTIONS ####
+# Function defined to read downloaded tmy files
+get_tmy <- function(all_sites, readfile_path){
+
+  df_tmy <- data.frame()
+
+  for (site in all_sites){
+    df <- read_csv(paste0(readfile_path, "tmy/", str_glue("{site}.epw")),
+                   skip = 8, col_types = "ddddd-d---------------------------------",
+                   col_names = c("year", "month", "day", "hour", "min", "tmy")) %>%
+      mutate(year = 2017,
+             time = ymd_h(paste(paste(year, month, day, sep = "-"), hour, sep = " ")),
+             temp = tmy) %>%
+      dplyr::select(time, temp) %>% 
+      mutate(site = site)
+
+    df_tmy <- rbind(df_tmy, df)
+  }
+
+  return(df_tmy)
+}
+
+# define parameters
+sprt_param <- list(baseline = "Baseline",
+                   strategy = "Intervention",
+                   parameter = "power_ave",
+                   label = "power",
+                   n_weeks = 48)
+
+block_params <- list(start_date = "2016-01-01",
+                     n_weeks = 108,
+                     n_seasons = 9, 
+                     block_unit = 12)
+```
+
+```{r readdata, include=FALSE}
+
+#### READ DATA ####
+# Tidy dataset
+readfile_tidy <- str_glue("../readfiles/tidy/")
+
+df_energy_tidy <- read_rds(paste0(readfile_tidy, "df_energy.rds"))
+df_meta_tidy <- read_rds(paste0(readfile_tidy, "df_meta.rds"))
+df_weather_tidy <- read_rds(paste0(readfile_tidy, "df_weather.rds"))
+df_sprt_all_tidy <- read_rds(paste0(readfile_tidy, "df_sprt_all.rds"))
+df_seq_FS_tidy <- read_rds(paste0(readfile_tidy, "df_seq_FS.rds"))
+df_NRE_occ <- read_rds(paste0(readfile_tidy, "df_NRE_occ.rds"))
+df_MD_tidy <- read_rds(paste0(readfile_tidy, "df_MD.rds"))
+df_FS_tidy <- read_rds(paste0(readfile_tidy, "df_FS.rds"))
+df_eui_tidy <- read_rds(paste0(readfile_tidy, "df_eui.rds"))
+df_cont_MD_tidy <- read_rds(paste0(readfile_tidy, "df_cont_MD.rds"))
+df_cont_FS_tidy <- read_rds(paste0(readfile_tidy, "df_cont_FS.rds"))
+
+all_sites_tidy <- df_energy_tidy %>%
+  select(site) %>%
+  distinct() %>%
+  arrange(site)
+
+all_types_tidy <- df_energy_tidy %>%
+  select(type) %>%
+  mutate(type = as.factor(type)) %>%
+  distinct()
+
+all_names_tidy <- df_energy_tidy %>%
+  select(name) %>%
+  distinct(name)
+
+# Messy dataset
+readfile_messy <- str_glue("../readfiles/messy/")
+
+df_energy_messy <- read_rds(paste0(readfile_messy, "df_energy.rds"))
+df_meta_messy <- read_rds(paste0(readfile_messy, "df_meta.rds"))
+df_weather_messy <- read_rds(paste0(readfile_messy, "df_weather.rds"))
+df_sprt_all_messy <- read_rds(paste0(readfile_messy, "df_sprt_all.rds"))
+df_seq_FS_messy <- read_rds(paste0(readfile_messy, "df_seq_FS.rds"))
+df_MD_messy <- read_rds(paste0(readfile_messy, "df_MD.rds"))
+df_FS_messy <- read_rds(paste0(readfile_messy, "df_FS.rds"))
+df_eui_messy <- read_rds(paste0(readfile_messy, "df_eui.rds"))
+df_cont_MD_messy <- read_rds(paste0(readfile_messy, "df_cont_MD.rds"))
+df_cont_FS_messy <- read_rds(paste0(readfile_messy, "df_cont_FS.rds"))
+df_FS_nsprt <- read_rds(paste0(readfile_messy, "df_FS_nsprt.rds"))
+df_MD_nsprt <- read_rds(paste0(readfile_messy, "df_MD_nsprt.rds"))
+df_seq_FS_nsprt <- read_rds(paste0(readfile_messy, "df_seq_FS_nsprt.rds"))
+
+all_sites_messy <- df_energy_messy %>%
+  select(site) %>%
+  distinct() %>%
+  arrange(site)
+
+all_types <- df_energy_messy %>%
+  select(type) %>%
+  mutate(type = as.factor(type)) %>%
+  distinct()
+
+all_names <- df_energy_messy %>%
+  select(name) %>%
+  distinct(name)
+```
+
+# Introduction
+
+## Background
+
+### Conventional M&V
+
+### Randomized (rapid) M&V
+
+### Non-routine events
+
+## Literature review
+
+## Objectives
+
+The objectives of this paper include:
+
+1.  Demonstrate the implementation of the proposed randomized rapid M&V
+    method using a public available dataset. We ensured the
+    reproducibility of the M&V method by making the analysis code open
+    source including randomized schedule generation, sequential
+    statistical analysis, energy modeling and normalized saving
+    calculation. Using the available open resources, building analysts
+    should be able to seamlessly integrate and apply them in their own
+    M&V projects.
+
+2.  Compare the energy saving estimation accuracy between the
+    conventional and the randomized method. In particular, this study
+    extends the comparison to large samples of buildings of various
+    types and across multiple climate zones.
+
+3.  Verify the superior robustness of the randomized method over the
+    conventional method. By using the realistic measurements from
+    real-world buildings, which contains various sources of noises,
+    could largely reflect the challenges that a building analyst would
+    be facing in any real project. Particularly, as we will demonstrate
+    in the following sections, non-routine events (i.e. 'noises' in the
+    measurements) have less impact in the energy saving estimation when
+    the analyst uses the randomized approach.
+
+# Method
+
+As mentioned, this study leverages a large public dataset to demonstrate
+the energy saving estimation results of a novel M&V method inspired by
+other scientific research fields. We outlined the methodology of the
+study in figure and outlined several key steps in this section.
+
+```{r flowchart}
+
+```
+
+## Data filtering process
+
+### 'Tidy' dataset - buildings with the same two-year electricity usage
+
+### 'Messy' dataset - buildings with reasonable change
+
+## Control intervention - chilled water supply temperature reset
+
+## Building Energy modeling
+
+# Results
+
+## Randomized M&V results
+
+### Saving estimation accuracy comparison ('tidy' dataset results)
+
+### Required estimation time
+
+### Continue sampling results
+
+## Influence of non-routine events
+
+### Example non-routine event: occupancy change
+
+### Saving estimation accuracy comparison ('messy' dataset results)
+
+# Discussion
+
+# Conclusion

### 2024-09-21 code/manuscript.Rmd
Added manuscript content

@@
 ---
-title: "Reliably estimating the impact of a new control strategy in a building"
+title: "Reliability Demonstration of A Novel Randomized Measurement and Verification Method for Switchable Control Retrofit Using Large-scale Public Dataset"
 author:
@@
                       dpi = 300,
-                      fig.path = "figs",
                       fig.show = "hold",
                       fig.pos = "b", 
-                      fig.path = "../figs/manuscript")
+                      fig.path = "../figs/manuscript/figs/")
 
@@
 # load packages using pacman
-pacman::p_load(tidyverse, lubridate, here) #effect size
+pacman::p_load(tidyverse, lubridate, here, stats, zoo, scales, ggpubr, patchwork, RColorBrewer)
 
@@
 
-ls_colors <- c("Baseline" = "#99d8c9",
-               "Measured baseline" = "#99d8c9",
-               "Adjusted baseline" = "#99d8c9",
-               "Projected baseline\n(no change)" = "#1b9e77",
+# colors
+ls_colors <- c("Baseline" = "#a6bddb",
+               "Measured baseline" = "#a6bddb",
+               "Adjusted baseline" = "#a6bddb",
+               "Projected baseline\n(no change)" = "#2b8cbe",
                "Intervention" = "#fdbb84",
@@
                "True normalized savings" = "black")
+
+# parameters
+ctr_params <- list(peak_hours = 10:16,
+                   chwl_perc = 0.25,
+                   step_perc = 0.08,
+                   conv_swt = 6,
+                   weather_knots = c(15, 25),
+                   swt_knots = c(12, 7),
+                   coe_peak = 0.8,
+                   coe_off = 1.2,
+                   enable_temp = 8)
+
+occ_params <- list(change_start = 5,
+                   change_end = 8,
+                   change = 20)
 
@@
 
-# define parameters
-sprt_param <- list(baseline = "Baseline",
-                   strategy = "Intervention",
-                   parameter = "power_ave",
-                   label = "power",
-                   n_weeks = 48)
+# Function defined to add chwst reset intervention
+run_reset <- function(df_baseline){
 
-block_params <- list(start_date = "2016-01-01",
-                     n_weeks = 108,
-                     n_seasons = 9, 
-                     block_unit = 12)
+  mean <- mean(df_baseline$base_eload, na.rm = T) * ctr_params$chwl_perc
+
+  grad <- (ctr_params$swt_knots[2] - ctr_params$swt_knots[1]) / 
+    (ctr_params$weather_knots[2] - ctr_params$weather_knots[1])
+
+  interc <- ctr_params$swt_knots[2] - (ctr_params$weather_knots[2] * grad)
+
+  df_interv <- df_baseline %>% 
+    mutate(swt = t_out * grad + interc, 
+           chwl = mean,
+           hour = hour(datetime)) %>% 
+    mutate(swt = ifelse(swt > ctr_params$swt_knots[1], ctr_params$swt_knots[1], ifelse(swt < ctr_params$swt_knots[2], ctr_params$swt_knots[2], swt)), 
+           temp_savings = ifelse(t_out >= ctr_params$enable_temp, (swt - ctr_params$conv_swt) * ctr_params$step_perc, 0), 
+           time_adj = ifelse(hour %in% ctr_params$peak_hours, ctr_params$coe_peak, ctr_params$coe_off), 
+           perc_savings = temp_savings * time_adj, 
+           savings = chwl * perc_savings, 
+           interv_eload = base_eload - savings) %>% 
+    select(datetime, base_eload, interv_eload, t_out)
+
+  return(df_interv)
+}
+
+# Function defined to interpolate NAs
+run_interpo <- function(df_all){
+
+  na_counts <- df_all %>%
+    mutate(date = date(timestamp)) %>%
+    group_by(date) %>%
+    summarize(na_hours = sum(is.na(eload)))
+
+  # Filter out days with more than half of the hours having NAs
+  valid_days <- na_counts %>%
+    filter(na_hours <= 12) %>%
+    pull(date)
+
+  df_filtered <- df_all %>%
+    filter(date(timestamp) %in% valid_days)
+
+  df_filtered <- df_filtered %>%
+    mutate(across(c(eload, t_out), ~ zoo::na.approx(., na.rm = FALSE))) %>% 
+    rename(datetime = timestamp, 
+           base_eload = eload)
+
+  return(df_filtered)
+}
+
+# Function defined to adjust the plot scale
+get_scale <- function(eload, range = 2){
+
+  min_y <- mean(eload, na.rm = T) - range * sd(eload, na.rm = T)
+  max_y <- mean(eload, na.rm = T) + range * sd(eload, na.rm = T)
+
+  return(c(min_y, max_y))
+}
 ```
@@
 readfile_tidy <- str_glue("../readfiles/tidy/")
+fig_path = "../figs/manuscript/"
 
@@
 
-all_types <- df_energy_messy %>%
+all_types_messy <- df_energy_messy %>%
   select(type) %>%
@@
 
-all_names <- df_energy_messy %>%
+all_names_messy <- df_energy_messy %>%
   select(name) %>%
   distinct(name)
+
+# read functions
+function_path <- "../functions/"
+source(paste0(function_path, "model_fit.R"))
+source(paste0(function_path, "model_pred.R"))
+source(paste0(function_path, "prepost_plot.R"))
 ```
@@
 
-### Randomized (rapid) M&V
+The conventional M&V process can be found in ASHRAE Guideline 14
+<!--# TODO:Add reference -->, which are summarized in Figure
+\@ref(fig:convmethod). The process normally starting with baseline
+measurement for a whole year before retrofit. After deploying the
+intervention, building performance measurement continues for another
+year. At the end of the two-year period,
+
+### Randomized M&V
 
 ### Non-routine events
+
+A common non-routine event in an energy-saving M&V project is a change
+in occupancy or a substantial change in occupant behavior. Those changes
+have a significant impact on measured building energy consumption and
+are typically not caused by the intervention strategy. If the M&V
+analyst are not aware of the change and has no reasonable approach to
+make adjustment on the energy measurements, the saving estimation result
+is largely biased. For example, most commercial buildings are unoccupied
+during the pandemic in 2020 and thus building managers have observed a
+drastically decrease in monthly energy bills despite no energy-efficient
+measures were implemented. In addition, common adjustments by an
+building analyst only consider the variations of outdoor weather
+conditions such as by fitting a regression model (e.g. Time-Of-Week
+Temperature model).
+
+### BDG2 dataset
 
@@
 
-The objectives of this paper include:
+As mentioned, the goal of a M&V project is to determine the effect
+(normally savings) of an energy-efficient intervention. And in this
+study, we further limit the study scope to switchable interventions,
+which mostly encapsulates control retrofits. An example intervention of
+such type can be a control retrofit developed by a software-as-a-service
+company that enables the chilled water plant to reset its supply water
+temperature based on outdoor weather condition. Therefore, we defined
+the M&V scenario as follows:
 
-1.  Demonstrate the implementation of the proposed randomized rapid M&V
-    method using a public available dataset. We ensured the
-    reproducibility of the M&V method by making the analysis code open
-    source including randomized schedule generation, sequential
-    statistical analysis, energy modeling and normalized saving
-    calculation. Using the available open resources, building analysts
-    should be able to seamlessly integrate and apply them in their own
-    M&V projects.
+*"A company wishes to sell their supply temperature reset control
+software package to a building owner and guarantees its effect in
+reducing building electricity usage. If the building owner decides to
+purchase the service, the company agrees to charge the service fee based
+on a fraction of the measured savings."*
+
+As required by the M&V scenario, we assessed the performance of both the
+conventional and the novel randomized M&V methods by estimating the
+intervention energy savings for all valid buildings in the dataset. By
+conducting such analysis, we hope to:
+
+1.  Demonstrate the implementation of the proposed randomized M&V method
+    using a public available dataset. We ensured the reproducibility of
+    the M&V method by making the analysis code open source including
+    randomized schedule generation, sequential statistical analysis,
+    energy modeling and normalized saving calculation. Using the
+    available open resources, building analysts should be able to
+    seamlessly integrate and apply them in their own M&V projects.
 
@@
 other scientific research fields. We outlined the methodology of the
-study in figure and outlined several key steps in this section.
+study in Figure \@ref(fig:flowchart) and outlined several key components
+in this section.
 
-```{r flowchart}
+```{r flowchart, fig.cap = "Workflow summary of the methodology of this paper", out.width="50%"}
+
+knitr::include_graphics(paste0(fig_path, "flowchart.png"))
+```
+
+## Building filtering
+
+### 'Tidy' subset
+
+In this study, we extracted the electricity measurements from the BDG2
+dataset. On a first pass, we filtered out buildings with less noise
+based on the following criteria:
+
+1.  Missing values \< 1000: given the hourly resolution of all
+    measurements, this is equivalent to more than a month of missing
+    days.
+
+2.  Mean electricity usage \> 0 kWh: target buildings should have active
+    electricity usage.
+
+3.  No statistical significant difference (P-value \> 0.05) between the
+    two-year electricity usage: target buildings should have no change
+    in the electricity usage between the two years.
+
+4.  Target buildings should have known site location: buildings with
+    anonymous location can not request typical meteorological weather.
+
+5.  Warehouse and parking types are excluded: target buildings that are
+    likely to have non-regular electricity and chilled water usage
+    should be excluded.
+
+As a summary, the resulting subset contains all buildings with 'tidy'
+measurements. Figure \@ref(fig:tidysite) shows in total, the subset
+contains `r nrow(all_names_tidy)` buildings in `r nrow(all_types_tidy)`
+types from `r nrow(all_sites_tidy)` different climate zones.
+
+```{r tidysite, fig.cap = "Site summary of the tidy building subset", fig.width=8, fig.height=8}
+
+set3 <- colorRampPalette(brewer.pal('Set3',n=12))
+type_colors <- setNames(set3(13), all_types_messy$type)
+
+p1 <- df_energy_tidy %>%
+    group_by(type) %>%
+    distinct(name) %>%
+    summarise(n = n()) %>%
+    ungroup() %>%
+    mutate(type = fct_reorder(type, n, .desc = F)) %>%
+    ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
+    geom_col() +
+    geom_text(aes(label = ifelse(n > 10, as.character(n), "")), color = "black", position = position_stack(vjust = 0.5)) +
+    scale_y_continuous(expand = c(0, 0),
+                       breaks = breaks_pretty(n = 4)) +
+    scale_x_discrete(expand = c(0, 0.1)) +
+    scale_fill_manual(values = type_colors) +
+    labs(x = NULL,
+         y = NULL,
+         subtitle = "Across all sites",
+         fill = NULL) +
+    theme(axis.text.x = element_blank(),
+          panel.grid.major.y = element_line(color = "grey80"),
+          legend.direction = "horizontal",
+          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+  p2 <- df_energy_tidy %>%
+    group_by(site, type) %>%
+    distinct(name) %>%
+    summarise(n = n()) %>%
+    mutate(proportion = n / sum(n),
+           total = sum(n),
+           ymax = cumsum(proportion),
+           ymin = c(0, head(ymax, n = -1))) %>%
+    mutate(label_pos = (ymax + ymin) / 2) %>%
+    ungroup() %>%
+    group_by(type) %>%
+    mutate(order = sum(n)) %>%
+    ungroup() %>%
+    mutate(type = fct_reorder(type, order, .desc = F)) %>%
+    ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = type)) +
+    geom_rect() +
+    coord_polar(theta = "y") +
+    xlim(c(2, 4)) +
+    facet_wrap(~ site, nrow = 3) +
+    labs(x = NULL,
+         y = NULL,
+         subtitle = "For each site",
+         fill = NULL) +
+    scale_fill_manual(values = type_colors) +
+    geom_text(aes(x = 3.5, y = label_pos, label = n)) +
+    geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black") +
+    theme(legend.direction = "horizontal",
+          axis.text = element_blank(),
+          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+  ggarrange(p1, p2,
+            ncol = 2, nrow = 1,
+            widths = c(0.2, 1),
+            common.legend = TRUE,
+            legend = "bottom") +
+    plot_annotation(title = "Case study building type summary")
+```
+
+### 'Messy' subset
+
+The 'tidy' subset might be presentative to the measurements collected
+among the existing building stock since in reality, whole-building
+electricity measurements collected over two years normally exhibit
+larger changes. Those measured changes can be caused by sensor itself
+such as lack of calibration, or inherent changes of the building, which
+illustrated before, can bias the result of a M&V. Therefore, to compare
+the robustness of the two M&V method more realistically, we included an
+additional 'messy' subset which first exclude the 'tidy' subset and then
+re-apply the filtering rule with one amendment:
+
+3.  Absolute mean difference of the two-year electricity usage \< 25%:
+    any increase or decrease of building electricity usage in the second
+    year should be less than 25% of that in th first year.
+
+Figure \@ref(fig:messysite) shows the summary of the 'messy' dataset. As
+this is less aggressive, it contains `r nrow(all_names_messy)` buildings
+in `r nrow(all_types_messy)` types from `r nrow(all_sites_messy)`
+different climate zones.
+
+```{r messysite, fig.cap = "Site summary of the messy building subset", fig.width=8, fig.height=8}
+
+p1 <- df_energy_messy %>%
+    group_by(type) %>%
+    distinct(name) %>%
+    summarise(n = n()) %>%
+    ungroup() %>%
+    mutate(type = fct_reorder(type, n, .desc = F)) %>%
+    ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
+    geom_col() +
+    geom_text(aes(label = ifelse(n > 10, as.character(n), "")), color = "black", position = position_stack(vjust = 0.5)) +
+    scale_y_continuous(expand = c(0, 0),
+                       breaks = breaks_pretty(n = 4)) +
+    scale_x_discrete(expand = c(0, 0.1)) +
+    scale_fill_manual(values = type_colors) +
+    labs(x = NULL,
+         y = NULL,
+         subtitle = "Across all sites",
+         fill = NULL) +
+    theme(axis.text.x = element_blank(),
+          panel.grid.major.y = element_line(color = "grey80"),
+          legend.direction = "horizontal",
+          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+  p2 <- df_energy_messy %>%
+    group_by(site, type) %>%
+    distinct(name) %>%
+    summarise(n = n()) %>%
+    mutate(proportion = n / sum(n),
+           total = sum(n),
+           ymax = cumsum(proportion),
+           ymin = c(0, head(ymax, n = -1))) %>%
+    mutate(label_pos = (ymax + ymin) / 2) %>%
+    ungroup() %>%
+    group_by(type) %>%
+    mutate(order = sum(n)) %>%
+    ungroup() %>%
+    mutate(type = fct_reorder(type, order, .desc = F)) %>%
+    ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = type)) +
+    geom_rect() +
+    coord_polar(theta = "y") +
+    xlim(c(2, 4)) +
+    facet_wrap(~ site, nrow = 3) +
+    labs(x = NULL,
+         y = NULL,
+         subtitle = "For each site",
+         fill = NULL) +
+    scale_fill_manual(values = type_colors) +
+    geom_text(aes(x = 3.5, y = label_pos, label = n)) +
+    geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black") +
+    theme(legend.direction = "horizontal",
+          axis.text = element_blank(),
+          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+  ggarrange(p1, p2,
+            ncol = 2, nrow = 1,
+            widths = c(0.2, 1),
+            common.legend = TRUE,
+            legend = "bottom") +
+    plot_annotation(title = "Case study building type summary")
+```
+
+## Apply control intervention
+
+Figure \@ref(fig:chwst) shows the algorithm for the proposed control
+intervention that reset the chiller supply temperature based on the
+outdoor weather conditions. Under both strategies, we assume the chiller
+is enabled when outdoor temperature is higher than 10°C. The baseline,
+which assumed to be the existing measurements from the dataset, runs at
+a constant water supply temperature. The intervention as the figure
+shows reset from 7 °C to 12 °C.
+
+```{r chwst, fig.cap = "Proposed chilled water supply temperature reset based on outdoor temperature", fig.width=6, fig.height=4}
+
+data.frame(
+  OAT = c(10, 15, 25, 30),  
+  Baseline = c(6, 6, 6, 6),      
+  Intervention = c(12, 12, 7, 7)
+  ) %>% 
+  pivot_longer(cols = c("Baseline", "Intervention"), names_to = "strategy", values_to = "value") %>% 
+  ggplot(aes(x = OAT, y = value, color = strategy)) +
+  geom_line() +
+  scale_color_manual(values = ls_colors) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = c(4, 6, 8, 10, 12),
+                     limits = c(5.5, 12.5),
+                     labels = number_format(suffix = " °C")) +
+  scale_x_continuous(expand = c(0, 0), 
+                     breaks = c(10, 15, 20, 25, 30), 
+                     labels = number_format(suffix = " °C")) +
+  labs(x = "Outdoor temperature", 
+       y = "Chilled water supply temperature", 
+       title = "Chilled water supply temperature reset on outdoor temperature") +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.direction = "horizontal",
+        legend.position = "bottom",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
 
@@
 
-## Data filtering process
+We map the chilled water supply temperature reset to the electrical
+energy savings as: $$
+\text{savings} = mean(E_{CHW}) \times 25\% \times \text{perc_savings}
+$$
 
-### 'Tidy' dataset - buildings with the same two-year electricity usage
+We assume in general, HVAC uses around 50% of whole-building electricity
+and the chilled water plant further uses 50% electricity of the HVAC
+system. This assumption largely simplifies the unique characteristics of
+electricity usage in all types of commercial buildings, but for the
+scope of this paper, we therefore assume 25% of whole-building
+electricity is consumed by the chilled water plant ($E_{CHW}$).
+<!--# TODO: Add reference -->In most cases, savings from an intervention
+is unproportional to the building hourly electricity usage, which
+renders M&V challenging. Therefore, we map the resulting electricity
+savings as a percentage of the mean electricity consumption of the plant
+determined by outdoor temperature ($OAT$), intervention supply water
+temperature ($T_{interv}$), baseline supply water temperature
+($T_{base}$) and hour of the day ($\delta_{occ}$, binary indicator
+whether during peak hours from 9 AM to 4 PM).
 
-### 'Messy' dataset - buildings with reasonable change
+$$
+\text{perc_savings = }
+\begin{cases}
+0, & \text{ if } OAT < 10°C \\
+\left[ \beta_{\text{temp}} \times (\text{T}_{\text{interv}} - \text{T}_{\text{base}}) \right] \times 
+\left[ \beta_{\text{occ}} \times \delta_{occ} + \beta_{\text{unocc}} \times (1 - \delta_{occ})\right], & \text{ otherwise}
+\end{cases}
+$$
 
-## Control intervention - chilled water supply temperature reset
+Parameters and their pre-defined values are summarized in the table
+below. Those parameters were not rigorously calibrated for each
+building, and were used uniformly across the dataset. Although this
+calibration process should be included, it is considered tangential to
+the study scope outlined in this paper.
 
-## Building Energy modeling
+| Parameter              | Description                                                      | Value |
+|----------------|----------------------------------------|----------------|
+| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
+| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
+| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |
+
+: Parameters for calculating the intervention savings
+
+## Apply example non-routine event
+
+To demonstrate the influence of non-routine events that are not properly
+adjusted, we developed three scenarios to quantify the energy
+consumption change associated with a hypothetical occupancy change. In
+summary, we assume an increase in occupancy in 2016 (i.e. refered as the
+'baseline year' if using conventional method) resulted in 20% increase
+in whole-building measured electricity usage. The three scenarios
+specifies the increase happened during January to April (S1); May to
+August (S2); and September to December (S3). We applied such change only
+to the buildings in the 'tidy' subset as originally they are more likely
+to have no non-routine events.
+
+## Run M&V methods
+
+### Conventional M&V
+
+```{r convmethod, fig.cap="Flow chart showing the conventional M&V process for an energy-saving intervention", out.width="50%"}
+
+knitr::include_graphics(paste0(fig_path, "convmethod.png"))
+```
+
+As described in Section \@ref(introduction) and Figure
+\@ref(fig:convmethod), conventional method for M&V is time-consuming and
+unable to separate non-routine event impact from measured savings. If
+using conventional method to estimate intervention energy savings, the
+result (after 24 months) is calculated by the difference between
+measured intervention and projected baseline in the post-retrofit
+period. In this study, we leveraged a piece-wise linear regression
+considering time-of-week and outdoor temperature (TOWT) as independent
+variables for projection and normalization on typical meteorological
+year <!--# TODO: Add reference -->.
+
+### Randomized M&V
+
+```{r randmethod, fig.cap="Flow chart showing the novel randomized M&V process for an energy-saving intervention", out.width="50%"}
+
+knitr::include_graphics(paste0(fig_path, "randmethod.png"))
+```
+
+Compared to the conventional method, the randomized M&V can provide an
+estimation more rapid and reliable. Section \@ref(introduction)
+describes the application of randomization techniques and Figure
+\@ref(fig:randmethod) specifically outlined the process of conducting a
+randomized switchback experiment for control retrofit projects. The
+method sequentially evaluate the intervention effect and returns a
+saving estimation with 95% confidence interval when all stopping
+criteria are satisfied. <!--# TODO: Add paper DOI --> describes all
+stopping criteria and recommened threshold to consider in more details
+and we breifly summarized the thresholds used in this paper here:
+
+-   The HVAC system operates from 06:00 to 22:00 each day, so we use a
+    daily sampling interval with the sampling time at midnight each day.
+
+-   Block by day of the week with a block period of 12 weeks.
+
+Stopping criteria are:
+
+-   A minimum and maximum of 12 and 108 weeks respectively. The
+    randomized schedule covers the entire two-year period but stopping
+    criteria enables an early stop at the end of satisfied blocking
+    period.
+
+-   At least 80% of the drybulb temperature range in the annual TMY data
+    sampled by both strategies.
+
+-   Test for no carryover effect using a t-test with a p-value not
+    exceeding a defined significance threshold of 0.05.
+
+-   90% confidence that energy savings exceed or do not exceed 0% using
+    the SPRT test. Medium effect size (d = 0.5) quantified by cohen's d
+    and calculated SPRT statistics either falls below the lower
+    threshold or exceeds the upper threshold.
+
+-   As no baseline data is available, test with an equal sampling ratio
+    (50% baseline, 50% intervention).
+
+To remain consistency, savings normalization on typical meteorological
+year is also modeled through TOWT.
 
@@
 
-## Randomized M&V results
+## M&V methods comparison
 
-### Saving estimation accuracy comparison ('tidy' dataset results)
+In this section, we compare the performance of two M&V methods. The key
+aspects of the assessment include: 1) time required to reach a saving
+estimation: in most cases, a shorter M&V timeline reduces associated
+cost and interruption for the building owner; and 2) saving estimation
+accuracy: this is particularly important for any software-as-a-service
+company to set a reasonable price with their customers.
 
-### Required estimation time
+### Savings estimation time
 
-### Continue sampling results
+The timeline for the conventional M&V method is outlined in previous
+sections. Although ASHRAE Guideline 14 offers minimum requirements for
+each path, the required baseline period is either a full range of all
+independent variables (typically outdoor weather conditions) under
+normal facility operation or a 12-month worth of continuous
+measurements. The same requirements also applies to intervention
+installed in the post-retrofit period. Thus, the conventional M&V method
+is likely to run over 24 months or even longer due to missing data. For
+example if an analyst is asked to determine the energy savings of a
+chilled water plant retrofit but missed most of the cooling season due
+to delay in retrofit deployment, he/she needs to wait until the next
+cooling season to measure the savings.
 
-## Influence of non-routine events
+If the randomized M&V method is applicable, even if there is a delay in
+the retrofit deployment, the building owner faces less risk and impact
+since random sampling allows measuring the full range of all independent
+variables simultaneously among all control strategies or modes. In
+addition, sampling at an equal probability (i.e. 50%/50% between
+baseline/intervention) effectively balances the level of independent
+variables measured. In other words, unless one strategy is sampled with
+a large amount of consecutive days (e.g. 7 days or more), it is rare
+that the independent variable, such as outdoor weather condition, is
+measured significantly different among all sampled control strategies.
+Figure \@ref(fig:timeline) shows the average estimated timeline for each
+building across all sites and climate conditions if using randomized M&V
+for the applied chilled water supply temperature reset intervention. The
+timeline figures for each individual building are attached in the
+supplementary material. The figure shows for all climate zones, covering
+a sufficient range outdoor weather condition is the most stringent
+criterion. However, most buildings can satisfy the 80% range by 6 months
+and all buildings can quantitatively determine the savings with
+associated uncertainty by 9 months, which is even shorter than the
+length of baseline measurement required by the conventional method.
 
-### Example non-routine event: occupancy change
+```{r timeline, fig.cap="Average randomized M&V timeline summary for all buildings at each site", fig.width=8, fig.height=4}
 
-### Saving estimation accuracy comparison ('messy' dataset results)
+df_sprt_all_messy %>% 
+  filter(seq == "eob" | seq == "temp") %>% 
+  bind_rows(df_sprt_all_tidy %>% filter(seq == "eob" | seq == "temp")) %>% 
+  group_by(site, seq) %>% 
+  summarise(n_weeks = mean(n_weeks)) %>% 
+  ungroup() %>% 
+  mutate(seq = as.factor(seq), 
+         seq = recode_factor(seq, "temp" = "80% temperature range satisfied", "eob" = "remaining weeks till finish")) %>% 
+  ggplot() +
+  geom_col(aes(x = site, y = n_weeks, fill = seq), position = "identity") +
+  labs(x = "Site", 
+       y = "Number of weeks", 
+       fill = NULL, 
+       title = "Average finishing time for all buildings using randomized M&V") +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = seq(0, 36, by = 12)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.direction = "horizontal",
+        legend.position = "bottom",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+```
+
+### Savings estimation accuracy
+
+In this study, we define the overall target savings (i.e. the ground
+truth) that an M&V analyst wishes to detect, either in normalized
+savings, fractional savings or simply the measured difference between
+the baseline and the intervention measurements, is the mean electricity
+energy reduction over the two-year period from implementing the chilled
+water supply temperature reset control. We estimate the M&V savings by
+the two methods following the process given in Figure
+\@ref(fig:convmethod) and Figure \@ref(fig:randmethod). In orther words,
+we compare the conventional M&V savings at the end of the two-year
+period with the randomized M&V savings after all criteria are satisfied.
+
+Figure \@ref(fig:norm) shows the overall comparison of the savings
+estimation from all buildings normalized on the typical meteorological
+year weather conditions of each site. In subplot a), the narrower range
+of the true savings indicates the dependence of the intervention effect
+on the outdoor weather condition, which is intended as shown in Figure
+\@ref(fig:chwst). Sites with mild climate all year round such as
+locations in California shows higher savings potential above 10%, while
+locations with more extreme climate such as Washington DC shows only 6%
+savings annually. In addition, although the box plot indicates some
+savings estimation are 'outliers', they should not be interpreted as
+such since the distribution can be biased by the unbalanced sample sizes
+across different site shown in Figure \@ref(fig:tidysite) and Figure
+\@ref(fig:messysite). In subplot b), the results show that the
+conventional M&V method tends to estimate savings with a greater
+uncertainty and we will demonstrate in the following sections that
+non-routine events can have a significant impact. For the randomized
+method, we displayed two scenarios, the estimation at the end of
+two-year period with 50%/50% sampling throughout and the estimation
+provided after all stopping criteria satisfied (i.e. stops early at 24
+weeks or 36 weeks). Since we found no significant difference between the
+two distributions as shown in the figure, we demonstrated that an early
+stop has no impact on the estimation accuracy but can significantly
+reduces the cost associated with M&V.
+
+```{r norm, fig.cap="Comparison of normalized annual fractional savings on site TMY between conventional and randomized method", fig.width=8, fig.height=4}
+
+p1 <- df_FS_tidy %>% 
+  bind_rows(df_FS_messy) %>% 
+  filter(method == "true") %>% 
+  mutate(method = "true savings") %>% 
+  ggplot(aes(x = method, y = savings, fill = method)) +
+  stat_boxplot(geom ='errorbar', width = 0.5) +
+  geom_boxplot(outlier.size = 1) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 3), 
+                     limits = c(0, 30), 
+                     labels = number_format(suffix = "%")) +
+  labs(x = NULL, 
+       y = NULL, 
+       fill = NULL) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.direction = "horizontal",
+        legend.position = "bottom",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+p2 <- df_FS_tidy %>% 
+  bind_rows(df_FS_messy) %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  left_join(df_seq_FS_tidy %>% 
+              bind_rows(df_seq_FS_messy) %>% 
+              filter(seq == "eob"), by = c("name", "site")) %>% 
+  pivot_longer(c(conv, rand, FS), names_to = "parameter", values_to = "value") %>% 
+  mutate(parameter = as.factor(parameter), 
+         parameter = recode_factor(parameter, 
+                                   "conv" = "Conventional M&V", 
+                                   "rand" = "Randomized M&V", 
+                                   "FS" = "Randomized M&V at early stop")) %>% 
+  ggplot(aes(x = parameter, y = value, fill = parameter)) +
+  stat_boxplot(geom ='errorbar', width = 0.5) +
+  geom_boxplot(outlier.shape = NA) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 3), 
+                     limits = c(0, 30)) +
+  labs(x = NULL, 
+       y = NULL, 
+       fill = NULL) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.direction = "horizontal",
+        legend.position = "bottom",
+        axis.text.y = element_blank(),
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+  ggarrange(p1, p2,
+            labels = c("a)", "b)"),
+            ncol = 2, nrow = 1,
+            widths = c(0.35, 1),
+            legend = "bottom") +
+    plot_annotation(title = "Normalized annual fractional savings estimation comparison", 
+                    subtitle = "between convention and randomized M&V (with early stop)")
+```
+
+To assess the accuracy on individual building, we calculate the mean
+absolute difference between the true savings and estimated savings.
+Figure \@ref(fig:tidyfrac) shows the comparison between the two methods
+using the tidy subset. Because of the large variation in electricity
+energy usage, we express the savings as a normalized value which is a
+fraction of energy reduction compared to baseline. Furthermore, we focus
+on the median and 95% threshold of the mean absolute difference
+distribution in the comparitive assessment to filter out outliers. As a
+result, the conventional method performs better than the randomized
+method. One hypothesis is that the tidy subset only includes buildings
+exposed to similar usage throughout the two years, meaning the
+measurement used for regression model fitting resembles the model
+prediction results. Although the results indicate the conventional M&V
+method is preferable if a building can achieve long-term consistency in
+electricity usage, such condition is rarely garrenteed in reality.
+
+```{r tidyfrac, fig.cap="Absolute deviation of M&V estimated fractional savings from true target savings using tidy subset", fig.width=8, fig.height=4}
+
+plot_data <- df_seq_FS_tidy %>% 
+  filter(seq == "eob") %>% 
+  left_join(df_FS_tidy %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(abs_diff = abs(savings - FS), 
+         plot_max = max(abs_diff))
+
+median_rand <- median(plot_data$abs_diff)
+upper_rand <- quantile(plot_data$abs_diff, probs = 0.95)
+
+p1 <- plot_data %>% 
+  arrange(abs_diff) %>% 
+  ggplot() +
+  geom_bar(aes(x = seq(1, nrow(all_names_tidy), by = 1), y = abs_diff), 
+           stat = "identity", alpha = 0.5) +
+  geom_hline(yintercept = median_rand, 
+             linetype = "dashed", 
+             color = "red") +
+  annotate(y = median_rand + 0.5, 
+           x = 20, 
+           geom = "text", 
+           label = paste0("(", "50% buildings: < ", round(median_rand, digits = 1), " %)"),
+           size = 4) +
+  geom_hline(yintercept = upper_rand, 
+             linetype = "dashed", 
+             color = "red") +
+  annotate(y = upper_rand + 0.5, 
+           x = 20, 
+           geom = "text", 
+           label = paste0("(", "95% buildings: < ", round(upper_rand, digits = 1), " %)"),
+           size = 4) +
+  scale_x_continuous(expand = c(0.02, 0), 
+                     breaks = breaks_pretty(n = 3)) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 3), 
+                     labels = number_format(suffix = " %"),
+                     limits = c(0, max(plot_data$plot_max) + 0.5)) +
+  labs(x = "Number of buildings", 
+       y = "Absolute difference in measured savings", 
+       subtitle = "Randomized M&V at early stop") +
+  coord_cartesian(ylim = c(0, 10)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.direction = "horizontal",
+        legend.position = "bottom",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+plot_data <- df_FS_tidy %>% 
+  filter(method != "rand") %>% 
+  group_by(name) %>% 
+  summarise(abs_diff = abs(diff(savings))) %>% 
+  ungroup() %>% 
+  mutate(plot_max = max(abs_diff))
+
+median_conv <- median(plot_data$abs_diff)
+upper_conv <- quantile(plot_data$abs_diff, probs = 0.95)
+
+p2 <- plot_data %>% 
+  arrange(abs_diff) %>% 
+  ggplot() +
+  geom_bar(aes(x = seq(1, nrow(all_names_tidy), by = 1), y = abs_diff), 
+           stat = "identity", alpha = 0.5) +
+  geom_hline(yintercept = median_conv, 
+             linetype = "dashed", 
+             color = "red") +
+  annotate(y = median_conv + 0.5,
+           x = 20, 
+           geom = "text", 
+           label = paste0("(", "50% buildings: < ", round(median_conv, digits = 1), " %)"),
+           size = 4) +
+  geom_hline(yintercept = upper_conv, 
+             linetype = "dashed", 
+             color = "red") +
+  annotate(y = upper_conv + 0.5, 
+           x = 20,
+           geom = "text", 
+           label = paste0("(", "95% buildings: < ", round(upper_conv, digits = 1), " %)"),
+           size = 4) +
+  scale_x_continuous(expand = c(0.02, 0), 
+                     breaks = breaks_pretty(n = 3)) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 3), 
+                     labels = number_format(suffix = " %"),
+                     limits = c(0, max(plot_data$plot_max) + 0.1)) +
+  labs(x = "Number of buildings", 
+       y = NULL, 
+       subtitle = "Conventional M&V") +
+  coord_cartesian(ylim = c(0, 10)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.direction = "horizontal",
+        legend.position = "bottom",
+        axis.text.y = element_blank(),
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 0, unit = "mm"))
+
+ggarrange(p1, p2,
+          ncol=2, nrow=1,
+          labels = c("a)", "b)"),
+          align = "hv",
+          legend="bottom") +
+  plot_annotation(title = str_glue("M&V method accuracy comparison"), 
+                  subtitle = "fractional savings in tidy subset")
+
+```
+
+A more realistic condition is shown in the messy subset in Figure
+\@ref(fig:messyfrac). The randomized method demonstrates consistent
+accuracy in savings estimation, even in the presence of significant
+measurement noise, indicating strong robustness. On the contrary, the
+conventional method shows approximately in 50% of all cases, the
+deviations of M&V estimation are more than 5%. We showed the
+non-normalized mean absolute difference comparison plots in the
+supplementary material. The poor performance shows if the there exist a
+change in electricity usage, baseline projection based on regression
+models are unreliable. Those changes are normally referred as
+non-routine events, which summarizes all the influential factors that
+are not considered in the regression model either because they are
+impractical to measure, such as hourly occupancy rate, or too random
+with no expected patterns to quantify, such as a sudden change of use
+from an office to a warehouse. We will demonstrate this in more details
+in the following sections.
+
+```{r messyfrac, fig.cap="Absolute deviation of M&V estimated fractional savings from true target savings using messy subset", fig.width=8, fig.height=4}
+
+plot_data <- df_seq_FS_messy %>% 
+  filter(seq == "eob") %>% 
+  left_join(df_FS_messy %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(abs_diff = abs(savings - FS), 
+         plot_max = max(abs_diff))
+
+median_rand <- median(plot_data$abs_diff)
+upper_rand <- quantile(plot_data$abs_diff, probs = 0.95)
+
+p1 <- plot_data %>% 
+  arrange(abs_diff) %>% 
+  ggplot() +
+  geom_bar(aes(x = seq(1, nrow(all_names_messy), by = 1), y = abs_diff), 
+           stat = "identity", alpha = 0.5) +
+  geom_hline(yintercept = median_rand, 
+             linetype = "dashed", 
+             color = "red") +
+  annotate(y = median_rand + 1, 
+           x = 220, 
+           geom = "text", 
+           label = paste0("(", "50% buildings: < ", round(median_rand, digits = 1), " %)"),
+           size = 4) +
+  geom_hline(yintercept = upper_rand, 
+             linetype = "dashed", 
+             color = "red") +
+  annotate(y = upper_rand + 1, 
+           x = 220, 
+           geom = "text", 
+           label = paste0("(", "95% buildings: < ", round(upper_rand, digits = 1), " %)"),
+           size = 4) +
+  scale_x_continuous(expand = c(0.02, 0), 
+                     breaks = breaks_pretty(n = 3)) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 3), 
+                     labels = number_format(suffix = " %"),
+                     limits = c(0, max(plot_data$plot_max) + 0.5)) +
+  labs(x = "Number of buildings", 
+       y = "Absolute difference in measured savings", 
+       subtitle = "Randomized M&V at early stop") +
+  coord_cartesian(ylim = c(0, 25)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.direction = "horizontal",
+        legend.position = "bottom",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+plot_data <- df_FS_messy %>% 
+  filter(method != "rand") %>% 
+  group_by(name) %>% 
+  summarise(abs_diff = abs(diff(savings))) %>% 
+  ungroup() %>% 
+  mutate(plot_max = max(abs_diff))
+
+median_conv <- median(plot_data$abs_diff)
+upper_conv <- quantile(plot_data$abs_diff, probs = 0.95)
+
+p2 <- plot_data %>% 
+  arrange(abs_diff) %>% 
+  ggplot() +
+  geom_bar(aes(x = seq(1, nrow(all_names_messy), by = 1), y = abs_diff), 
+           stat = "identity", alpha = 0.5) +
+  geom_hline(yintercept = median_conv, 
+             linetype = "dashed", 
+             color = "red") +
+  annotate(y = median_conv + 1,
+           x = 220, 
+           geom = "text", 
+           label = paste0("(", "50% buildings: < ", round(median_conv, digits = 1), " %)"),
+           size = 4) +
+  geom_hline(yintercept = upper_conv, 
+             linetype = "dashed", 
+             color = "red") +
+  annotate(y = upper_conv + 1, 
+           x = 220,
+           geom = "text", 
+           label = paste0("(", "95% buildings: < ", round(upper_conv, digits = 1), " %)"),
+           size = 4) +
+  scale_x_continuous(expand = c(0.02, 0), 
+                     breaks = breaks_pretty(n = 3)) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 3), 
+                     labels = number_format(suffix = " %"),
+                     limits = c(0, max(plot_data$plot_max) + 0.1)) +
+  labs(x = "Number of buildings", 
+       y = NULL, 
+       subtitle = "Conventional M&V") +
+  coord_cartesian(ylim = c(0, 25)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.direction = "horizontal",
+        legend.position = "bottom",
+        axis.text.y = element_blank(),
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 0, unit = "mm"))
+
+ggarrange(p1, p2,
+          ncol=2, nrow=1,
+          labels = c("a)", "b)"),
+          align = "hv",
+          legend="bottom") +
+  plot_annotation(title = str_glue("M&V method accuracy comparison"), 
+                  subtitle = "fractional savings in messy subset")
+
+```
+
+## Non-routine events impact on savings estimation
+
+In this section, we demonstrate the influence of non-routine events
+through two scenarios. In the first scenario, we quantified how much
+occupancy would influence the electricity usage and added to the tidy
+measurement set as they originally indicates no such change. In the
+second scenario, we simply run the two M&V methods on the messy
+measurement sets before adding the chilled water supply temperature
+reset. Since no intervention is added, the more reliable method should
+detect a 'saving' closer to 0.
+
+### Occupancy change
+
+Although occupancy can be approximated in several ways in commercial
+buildings, such as through counting check-in, or WIFI connections, or
+monitoring indoor CO2 concentration as a proxy. However, it is not a
+cost-efficient measures to add to the routine operation for most of the
+buildings and there might be privacy concerns associated. The scenario
+proposed hypothesize that during the baseline measurement period of the
+target building, a floor of tenants moved out leaving the space
+unoccupied for four months before new tenants moved in and this led to a
+20% electricity decrease as a result. The M&V protocol only requires the
+M&V analyst to fit a TOWT model so he/she only measures outdoor
+temperature. Figure \@ref(fig:occchange) shows one example
+
+```{r occ}
+
+n = 6
+
+site_info <- df_energy_tidy %>%
+    filter(name == all_names_tidy$name[n]) %>%
+    select(site, type) %>%
+    distinct()
+
+site <- site_info$site
+
+site_weather <- df_weather_tidy %>%
+    filter(site == site_info$site) %>%
+    select(timestamp, t_out) %>%
+    group_by(timestamp) %>%
+    summarise(t_out = mean(t_out)) %>%
+    ungroup()
+
+df_all <- df_energy_tidy %>%
+    filter(name == all_names_tidy$name[n]) %>%
+    select(timestamp, eload) %>%
+    left_join(site_weather, by = "timestamp")
+
+# Linear interpolation of baseline
+df_all <- df_all %>%
+  run_interpo()
+
+plot_scale <- get_scale(df_all$base_eload)
+
+df_hourly_conv <- df_all %>%
+  run_reset()
+
+df_base_conv <- df_hourly_conv %>%
+    select(datetime,
+           eload = base_eload,
+           t_out) %>% 
+    drop_na()
+
+df_interv_conv <- df_hourly_conv %>%
+    select(datetime,
+           eload = interv_eload,
+           t_out) %>% 
+    drop_na()
+
+change_start <- occ_params$change_start
+change_end <- occ_params$change_end
+change <- occ_params$change
+
+base_pre_meas <- df_base_conv %>%
+  filter(datetime < as.Date("2017-01-01")) %>%
+  mutate(eload = ifelse(month(datetime) >= month(change_start) & month(datetime) <= month(change_end), eload * (1 - change / 100), eload))
+
+# Baseline projection
+towt_base <- base_pre_meas %>%
+  model_fit()
+
+df_towt <- df_interv_conv %>%
+  filter(datetime >= as.Date("2017-01-01")) %>%
+  select(time = datetime,
+         temp = t_out,
+         eload)
+
+base_pos_proj <- model_pred(df_towt, towt_base) %>%
+  rename("datetime" = "time") %>%
+  select(datetime, towt, eload)
+
+base_pos_true <- df_base_conv %>%
+  filter(datetime >= as.Date("2017-01-01"))
+
+interv_pos_meas <- df_interv_conv %>%
+  filter(datetime >= as.Date("2017-01-01"))
+```
+
+```{r occchange, fig.cap="Occupancy change impact on TOWT model fitting and baseline projection", fig.width=8, fig.height=4}
+
+p1 <- ggplot() +
+    geom_point(data = base_pre_meas %>%
+                 slice_sample(n = 500),
+               aes(x = datetime, y = eload, color = "Measured baseline"),
+               size = 0.5,
+               alpha = 0.3) +
+    geom_smooth(data = base_pre_meas,
+                aes(x = datetime, y = eload, color = "Measured baseline"), 
+                formula = y ~ x, method = "loess") +
+    scale_x_datetime(date_breaks = "2 months",
+                     date_labels = "%b")  +
+    scale_y_continuous(expand = c(0, 0),
+                       breaks = breaks_pretty(n = 3),
+                       labels = number_format(suffix = " kW")) +
+    scale_color_manual(values = ls_colors) +
+    coord_cartesian(ylim = plot_scale) +
+    labs(x = NULL,
+         y = NULL,
+         color = NULL,
+         title = NULL,
+         subtitle = "Pre-retrofit period") +
+    theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+          legend.direction = "horizontal",
+          legend.position = "bottom",
+          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+p2 <- ggplot() +
+  geom_point(data = base_pos_proj %>%
+               slice_sample(n = 500),
+             aes(x = datetime, y = towt, color = "Projected baseline\n(no change)"),
+             size = 0.5,
+             alpha = 0.3) +
+  geom_smooth(data = base_pos_proj,
+              aes(x = datetime, y = towt, color = "Projected baseline\n(no change)"), 
+              formula = y ~ x, method = "loess") +
+  geom_point(data = base_pos_true %>%
+               slice_sample(n = 500),
+             aes(x = datetime, y = eload, color = "Adjusted baseline"),
+             size = 0.5,
+             alpha = 0.3) +
+  geom_smooth(data = base_pos_true,
+              aes(x = datetime, y = eload, color = "Adjusted baseline"), 
+              formula = y ~ x, method = "loess") +
+  geom_point(data = interv_pos_meas %>%
+               slice_sample(n = 500),
+             aes(x = datetime, y = eload, color = "Measured interv"),
+             size = 0.5,
+             alpha = 0.3) +
+  geom_smooth(data = interv_pos_meas,
+              aes(x = datetime, y = eload, color = "Measured interv"), 
+              formula = y ~ x, method = "loess") +
+  scale_x_datetime(date_breaks = "2 months",
+                   date_labels = "%b")  +
+  scale_y_continuous(expand = c(0, 0),
+                     breaks = breaks_pretty(n = 3),
+                     labels = number_format(suffix = " kW")) +
+  scale_color_manual(values = ls_colors) +
+  coord_cartesian(ylim = plot_scale) +
+  labs(x = NULL,
+       y = NULL,
+       color = NULL,
+       title = NULL,
+       subtitle = "post-retrofit period") +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        axis.text.y = element_blank(),
+        legend.direction = "horizontal",
+        legend.position = "bottom",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+ggarrange(p1, p2,
+          ncol=2, nrow=1,
+          labels = c("a)", "b)"),
+          widths = c(1, 1),
+          align = "v",
+          legend="bottom") +
+  plot_annotation(title = "Building energy consumption with added occupancy change")
+```
+
+### No-saving detection
 
@@
 
+## TOWT modeling accuracy
+
+## Sampling ratio for randomized M&V
+
+## Limitations and future study
+
 # Conclusion

### 2024-09-21 code/manuscript.Rmd
Updated result section

@@
 
-```{r tidysite, fig.cap = "Site summary of the tidy building subset", fig.width=8, fig.height=8}
+```{r tidysite, fig.cap = "Site summary of the tidy building subset", fig.width=8, fig.height=6}
 
@@
 
-```{r chwst, fig.cap = "Proposed chilled water supply temperature reset based on outdoor temperature", fig.width=6, fig.height=4}
+```{r chwst, fig.cap = "Proposed chilled water supply temperature reset based on outdoor temperature", fig.width=8, fig.height=4}
 
@@
 | Parameter              | Description                                                      | Value |
-|----------------|----------------------------------------|----------------|
+|-----------------|--------------------------------------|-----------------|
 | $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
@@
 M&V analyst to fit a TOWT model so he/she only measures outdoor
-temperature. Figure \@ref(fig:occchange) shows one example
+temperature. We showed one example in Figure \@ref(fig:occchange) where
+from May 1st 2016 to August 31st 2016, the target building electricity
+decreased 20% due to reduced occupancy. Thus subplot (a) shows measured
+baseline with such change, which are then used for TOWT model fitting.
+Starting from 2017, the occupancy returns normal and only intervention
+strategy was measured and the projected baseline is the prediction
+results from the fitted model. To demonstrate, we also plotted the
+original measurements before adding the intervention effect as 'adjusted
+baseline'. This assumes that we can accurately adjust the baseline in
+the post-retrofit period on back-to-normal occupancy. Subplot (b) shows
+that the fitted regression underestimate the true baseline condition
+leading to underestimated savings.
 
@@
 
+Figure \@ref(fig:occsavings) shows the normalized fractional savings
+estimation accuracy comparison between the two M&V methods. To compare
+which method is more robust to the applied occupancy change, a
+difference-in-difference value is plotted in the figure. Each bar in the
+plot represent the difference in the deviation of savings estimated by
+the convention method and the randomized method:
+
+$$
+|S_{conv} - S_{true}| - |S_{rand} - S_{true}|
+$$
+
+In other words, a positive value indicate the randomized method provides
+an estimation more aligned to the true target saving. Thus the plot
+shows the randomized method shows uniformly superior robustness for all
+sites. The red-dotted line shows the absolute deviation in savings
+estimation by the randomized method as a reference for scaling. The
+deviation is consistent with the distribution shown in Figure
+\@ref(fig:tidysite) and thus further reinforce its robustness to
+non-routine events.
+
+```{r occsavings, fig.cap="Savings estimation accuracy comparison between two M&V methods with added occupancy change (with the site average difference-in-difference displayed at the bottom of each site).", fig.width=8, fig.height=4}
+s <- 'S4'
+dev_conv <- df_NRE_occ %>%
+  filter(scenario == s) %>%
+  filter(method != "rand") %>%
+  group_by(name, site, scenario) %>%
+  summarise(conv = abs(diff(savings))) %>%
+  ungroup()
+
+dev_rand <- df_NRE_occ %>%
+  filter(scenario == s) %>%
+  filter(method != "conv") %>%
+  group_by(name, site, scenario) %>%
+  summarise(rand = abs(diff(savings))) %>%
+  ungroup()
+
+dev_FS <- dev_rand %>%
+  left_join(dev_conv, by = c("name", "site", "scenario")) %>% 
+  mutate(diff_in_diff = conv - rand)
+
+mean_diff <- dev_FS %>%
+  group_by(site) %>%
+  summarise(mean = round(mean(diff_in_diff), digits = 1),
+            pos = round(n() / 2) + 1,
+            .groups = 'keep')
+
+dev_FS %>%
+  ggplot(aes(group = site)) +
+  geom_col(aes(x = name, y = diff_in_diff), position = "identity", alpha = 0.3) +
+  facet_wrap(~site, nrow = 1, scales = "free_x") +
+  scale_y_continuous(expand = c(0.1, 0),
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = "%")) +
+  geom_text(data = mean_diff,
+            aes(x = pos, y = -.5, group = site, label = paste0(mean, "%"))) +
+  geom_line(aes(x = name, y = rand, color = "Absolute deviation of randomized method"), alpha = 0.5) +
+  geom_point(aes(x = name, y = rand, color = "Absolute deviation of randomized method"), alpha = 0.5) +
+  labs(x = NULL,
+       fill = NULL,
+       y = NULL,
+       color = NULL, 
+       title = "Difference-in-difference of fractional savings") +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        axis.text.x = element_blank(),
+        legend.direction = "horizontal",
+        legend.position = "bottom",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+```
+
 ### No-saving detection

### 2024-09-24 code/manuscript.Rmd
Updated manuscript for running cont savings

@@
   bookdown::word_document2:
-    reference_docx: "../paper/TF_Template_Word_Mac_2011.dotx"
+    reference_docx: "../paper/template.docx"
 editor_options: 
@@
 df_cont_FS_tidy <- read_rds(paste0(readfile_tidy, "df_cont_FS.rds"))
+df_model_acc_tidy <- read_rds(paste0(readfile_tidy, "df_model_acc.rds"))
 
@@
 df_seq_FS_nsprt <- read_rds(paste0(readfile_messy, "df_seq_FS_nsprt.rds"))
+df_model_acc_messy <- read_rds(paste0(readfile_messy, "df_model_acc.rds"))
 
@@
 ## Literature review
+
+### Measurement and verification
+
+### Randomized experimental design
 
@@
 | Parameter              | Description                                                      | Value |
-|-----------------|--------------------------------------|-----------------|
+|------------------|------------------------------------|------------------|
 | $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
@@
 
+The outbreak of pandemic in 2020 followed by working-from-home policy is
+a practical example of building energy consumption influenced by
+non-routine events. During unoccupied months, building energy
+consumption dropped compared to the same time period in the previous
+years. If an energy-efficient intervention was deployed onsite starting
+at 2020, given by the whole-building electricity measurements, the large
+difference in measured electricity usage in 2020 and projected baseline
+usage in 2019 can hardly indicate the retrofit intervention effect.
+
+In other words, if there is no intervention applied or the intervention
+effect is known to be null (e.g. constantly being overriden by the
+baseline), an ideal M&V method should detect no savings. To test this,
+we kept the randomized schedule as sampled earlier but removed added
+chilled water plant intervention effect. Therefore, all sampled
+intervention days are considered additional sampled baseline days.
+Earlier we mentioned the 'messy' dataset contains some change in the
+two-year electricity measurement, which is purely due to various
+non-routine events such as occupancy change or even other measures
+implemented by the building manager but irrelevant to the chilled water
+set point reset intervention. Thus, when the proposed chilled water
+reset intervention was removed, a more reliable M&V method should
+overcome the influence of those confounding factors and inform no
+savings.
+
+```{r nosavings, fig.cap="Distribution of savings estimation results by the convention M&V method and randomized method.", fig.width=8, fig.height=4}
+
+df_FS_nsprt %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  left_join(df_seq_FS_nsprt %>% filter(seq == "eob"), by = c("name", "site")) %>% 
+  pivot_longer(c(true, conv, rand, FS), names_to = "parameter", values_to = "value") %>% 
+  filter(parameter != "true") %>% 
+  mutate(parameter = as.factor(parameter), 
+         parameter = recode_factor(parameter, 
+                                   "conv" = "Conventional M&V", 
+                                   "rand" = "Randomized M&V", 
+                                   "FS" = "Randomized M&V at early stop")) %>% 
+  ggplot(aes(x = parameter, y = value, fill = parameter)) +
+  stat_boxplot(geom ='errorbar', width = 0.2) +
+  geom_boxplot(outlier.shape = NA, width = 0.5) +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "red", linewidth = 1.2) +
+  annotate(geom = "text", 
+           x = 0.5, 
+           y = 0.5, 
+           color = 'red',
+           label = "True savings = 0") +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 5), 
+                     limits = c(0, 16), 
+                     labels = number_format(suffix = "%")) +
+  labs(x = NULL, 
+       y = NULL, 
+       fill = NULL, 
+       title = "Savings estimation comparison", 
+       subtitle = "between convention and randomized M&V (with early stop)") +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.direction = "horizontal",
+        legend.position = "bottom",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+```
+
+Figure \@ref(fig:nosavings) shows that the conventional M&V method in
+average detected 5% savings when there should be none. The randomized
+method even when stops right after satisfying all criteria, the savings
+estimation is much closer to 0%. This is mostly because when there is a
+change or disturbance applied to the target building such as lighting
+retrofit or thermal leakage in the envelope, randomly sample the two
+strategies at 50%/50% ensures the resulting effect either as a decrease
+in lighting electricity or an increase in reheat electricity is balanced
+between the two implemented strategies.
+
 # Discussion
@@
 
+Although this paper argues that regression model prediction results can
+be largely biased by various non-routine events, we further clarify that
+this is not related to regression model fitting accuracy. Time-of-week
+temperature model is used here for its simplicity and convenience since
+the model only requires outdoor temperature measurements and time of
+week as independent variables. The model assumes a linear composition of
+building energy consumption as temperature-dependent and time-dependent
+load. The time-dependent component accounts for day-to-day variations
+and the temperature dependent component considers a piecewise linear
+relationship across a range of temperature intervals.
+<!--# TODO: add reference -->
+
+To assess modeling accuracy, we used the Coefficient of Variation of
+Root-Mean Squared Error, or CV(RMSE) as the error metric. Since this
+metric is calculated as a normalized value, it is useful to compare
+different model fitting results. Guideline 14 requires that
+whole-building baseline model fitting accuracy should maintain a
+CV(RMSE) lower than 30%. In addition One study focusing on the baseline
+energy data-driven model fitting indicates that TOWT performs as
+accurate as other more advanced machine learning models
+<!--# TODO: reference --> and the calculated CV(RMSE) distribution for a
+large sample of commercial buildings indicates a median of 20%. Figure
+\@ref(fig:towtacc) plots the distribution of the model fitting accuracy
+calculated separately for the two measurement sets. The result shows
+that the TOWT model performance is even better in this study compared to
+the literature and no significant difference between the two subsets.
+This is mostly related to low-quality measurement sets filtering in the
+pre-processing mentioned in Section \@ref(method).
+
+```{r towtacc, fig.cap="TOWT model fitting accuracy distribution on filtered measurement set", fig.width=8, fig.height=4}
+
+df_model_acc_messy %>% 
+  mutate(group = "messy") %>% 
+  rbind(df_model_acc_tidy %>% mutate(group = "tidy")) %>% 
+  mutate(group = as.factor(group)) %>% 
+  ggplot(aes(x = group, y = cvrmse, fill = group)) +
+  stat_boxplot(geom ='errorbar', width = 0.5) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_text(aes(x = group, y = median, label = paste0(round(median, digits = 0), " %")), 
+            data = . %>% 
+              group_by(group) %>% 
+              summarise(median = median(cvrmse)) %>% 
+              ungroup(), 
+            position = position_nudge(y = 1)) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 3), 
+                     labels = number_format(suffix = " %")) +
+  coord_cartesian(ylim = c(0, 38)) +
+  labs(x = NULL, 
+       y = NULL, 
+       fill = NULL, 
+       title = "TOWT model accuracy", 
+       subtitle = "CV(RMSE)") +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.direction = "horizontal",
+        legend.position = "bottom",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+```
+
+In general, the limitation is that despite the regression model tends to
+capture well on the mean building energy consumption but can
+underestimate and peak or overestimate the lower base load
+<!--# TODO: reference -->, which is not useful for assessing demand
+response. Additionally, a data-driven model can learn the training
+dataset very efficiently with high accuracy, but the well-trained model
+is challenging to generalize prediction outside the range of training
+set.
+
 ## Sampling ratio for randomized M&V
 
+Another advantage of using randomized M&V is the flexibility of changing
+sampling ratio after the M&V. For example, the building owner can
+continue sampling at 50%/50% between the baseline and the intervention
+to further reduce the uncertainty range associated with the savings. Or
+he/she can switch to 100% intervention to optimize energy savings, but
+whether the existing baseline becomes outdated is unknown. A compromised
+approach is to continue sampling at 80%/20% between the baseline and the
+intervention. Figure \@ref(fig:cont) shows after the analyst reports the
+randomized M&V results, a new schedule sampling at 80%/20% was
+implemented for another 36 weeks.
+
+```{r cont, fig.cap="Savings estimation accuracy for continue sampling at 80%/20% for 36 weeks after all stopping criteria are satisfied", fig.width=8, fig.height=4}
+
+plot_data <- df_cont_FS_tidy %>% 
+  rename(cont = FS) %>% 
+  left_join(df_FS_tidy %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(abs_diff = abs(savings - cont), 
+         plot_max = max(abs_diff))
+
+median_rand <- median(plot_data$abs_diff)
+upper_rand <- quantile(plot_data$abs_diff, probs = 0.95)
+
+p1 <- plot_data %>% 
+  arrange(abs_diff) %>% 
+  ggplot() +
+  geom_bar(aes(x = seq(1, nrow(all_names_tidy), by = 1), y = abs_diff), 
+           stat = "identity", alpha = 0.5) +
+  geom_hline(yintercept = median_rand, 
+             linetype = "dashed", 
+             color = "red") +
+  annotate(y = median_rand + 0.5, 
+           x = 20, 
+           geom = "text", 
+           label = paste0("(", "50% buildings: < ", round(median_rand, digits = 1), " %)"),
+           size = 4) +
+  geom_hline(yintercept = upper_rand, 
+             linetype = "dashed", 
+             color = "red") +
+  annotate(y = upper_rand + 0.5, 
+           x = 20, 
+           geom = "text", 
+           label = paste0("(", "95% buildings: < ", round(upper_rand, digits = 1), " %)"),
+           size = 4) +
+  scale_x_continuous(expand = c(0.02, 0), 
+                     breaks = breaks_pretty(n = 3)) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 3), 
+                     labels = number_format(suffix = " %"),
+                     limits = c(0, max(plot_data$plot_max) + 0.5)) +
+  labs(x = "Number of buildings", 
+       y = "Absolute difference in measured savings", 
+       subtitle = "Tidy subset") +
+  coord_cartesian(ylim = c(0, 25)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.direction = "horizontal",
+        legend.position = "bottom",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+plot_data <- df_cont_FS_messy %>% 
+  rename(cont = FS) %>% 
+  left_join(df_FS_messy %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(abs_diff = abs(savings - cont), 
+         plot_max = max(abs_diff))
+
+median_rand <- median(plot_data$abs_diff)
+upper_rand <- quantile(plot_data$abs_diff, probs = 0.95)
+
+p2 <- plot_data %>% 
+  arrange(abs_diff) %>% 
+  ggplot() +
+  geom_bar(aes(x = seq(1, nrow(all_names_messy), by = 1), y = abs_diff), 
+           stat = "identity", alpha = 0.5) +
+  geom_hline(yintercept = median_rand, 
+             linetype = "dashed", 
+             color = "red") +
+  annotate(y = median_rand + 0.5, 
+           x = 220, 
+           geom = "text", 
+           label = paste0("(", "50% buildings: < ", round(median_rand, digits = 1), " %)"),
+           size = 4) +
+  geom_hline(yintercept = upper_rand, 
+             linetype = "dashed", 
+             color = "red") +
+  annotate(y = upper_rand + 0.5, 
+           x = 220, 
+           geom = "text", 
+           label = paste0("(", "95% buildings: < ", round(upper_rand, digits = 1), " %)"),
+           size = 4) +
+  scale_x_continuous(expand = c(0.02, 0), 
+                     breaks = breaks_pretty(n = 3)) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 3), 
+                     labels = number_format(suffix = " %"),
+                     limits = c(0, max(plot_data$plot_max) + 0.1)) +
+  labs(x = "Number of buildings", 
+       y = NULL, 
+       subtitle = "Messy subset") +
+  coord_cartesian(ylim = c(0, 25)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.direction = "horizontal",
+        legend.position = "bottom",
+        axis.text.y = element_blank(),
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 0, unit = "mm"))
+
+ggarrange(p1, p2,
+          ncol=2, nrow=1,
+          labels = c("a)", "b)"),
+          align = "hv",
+          legend="bottom") +
+  plot_annotation(title = str_glue("Continue sampling accuracy comparison"), 
+                  subtitle = "with 80%/20% intervention/baseline")
+```
+
+The deviation remains consistent in the tidy subset but increases
+slightly in the messy subset, suggesting that changes in energy usage in
+the target building create a trade-off between measured savings and
+estimation accuracy when down-sampling the baseline. In this scenario,
+opting for an early stop at 24 or 36 weeks with 50% intervention and
+resampling 80% of the data for 36 weeks still allows the building owner
+to capture 65% of the full-range savings.
+
 ## Limitations and future study
 
+Considering the diversity of the measurement sets, the assumption made
+in this study is relative general. For example,
+
 # Conclusion

### 2024-09-24 code/1_analysis.R
verified for continue sampling

@@ -56,9 +56,9 @@ source(paste0(function_path, "cont_plot.R"))
 # section run control
 run_params <- list(type = "tidy", 
                    sprt = T, 
-                   sprt_cont = F, 
+                   sprt_cont = T, 
                    nre_occ = F, 
-                   nsprt = F)
+                   nsprt = T)
 
 # Adding intervention effect as advanced chiller operation
 ctr_params <- list(peak_hours = 10:16,                      # accounts for peak hours

### 2024-09-24 code/2_tidyvisual.R

@@ -40,7 +40,7 @@ ls_colors <- c("Baseline" = "#99d8c9",
 
 # define parameters
 run_params <- list(type = "tidy", 
-                   site = T)
+                   site = F)
 
 sprt_param <- list(baseline = "Baseline",
                    strategy = "Intervention",

### 2024-10-11 code/manuscript.Rmd
Added references and removed continue sampling from result sections

@@
 author:
-  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA] ^[Correspondence to aoyuzou@berkeley.edu], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Durate ^1^
+  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Durate ^1^
 abstract: Conventional measurement and verification (M&V) methods for whole-building energy savings estimation are both time-consuming and unreliable, especially when non-routine events occur during the M&V process.
@@
     wrap: 72
-# bibliography: references.bib
-# biblio-style: apalike
+bibliography: references.bib
+biblio-style: apalike
 
@@
 
-The conventional M&V process can be found in ASHRAE Guideline 14
-<!--# TODO:Add reference -->, which are summarized in Figure
-\@ref(fig:convmethod). The process normally starting with baseline
-measurement for a whole year before retrofit. After deploying the
-intervention, building performance measurement continues for another
-year. At the end of the two-year period,
+Measurement and Verification (M&V) is the process of quantifying energy
+savings from energy efficiency projects by comparing actual energy
+consumption against a baseline, adjusting for factors like weather and
+occupancy. This process ensures that improvements in energy performance
+are accurately evaluated. In the United States, practitioners often
+refer to ASHRAE Guideline 14, the International Performance Measurement
+and Verification Protocol (IPMVP), and the Federal Energy Management
+Program (FEMP) for standard guidelines [@doe_mv_2008,
+@efficiency_valuation_organisation_international_2007,
+@ashrae_ashrae_2014]. These guidelines outline standardized methods for
+quantifying energy savings, whether through calibrated simulations or
+monitored measurements for specific equipment or systems (isolation
+methods) or for entire buildings (whole-building methods). In this
+study, we will focus on the energy savings quantified at the
+whole-building level where the measurements are obtained from utility
+bills or whole-building meters. The corresponding M&V process required
+by ASHRAE Guideline 14, which we refer to as the convention method in
+this paper. Typically, the process begins with baseline measurements
+taken over a year before implementing any energy-efficiency retrofit,
+followed by the same measurement procedure during the intervention
+period. After collecting two years of data, an M&V analyst fits an
+energy prediction model, using variables such as outdoor temperature and
+time [@mathieu_quantifying_2011] to project baseline energy consumption
+in the post-retrofit period. The difference between the counter-factual
+baseline and the measured intervention represents the energy savings. A
+key drawback of this method is its reliance on a two-year timeline to
+quantify savings, during which baseline measurements can become outdated
+due to changes in building performance caused by non-routine events
+unrelated to the intervention. This limitation reduces the feasibility
+of rapid M&V and complicates the quantification of estimation
+uncertainty, thus impacting the accuracy and timeliness of savings
+assessments.
 
@@
 
-### Non-routine events
-
-A common non-routine event in an energy-saving M&V project is a change
-in occupancy or a substantial change in occupant behavior. Those changes
-have a significant impact on measured building energy consumption and
-are typically not caused by the intervention strategy. If the M&V
-analyst are not aware of the change and has no reasonable approach to
-make adjustment on the energy measurements, the saving estimation result
-is largely biased. For example, most commercial buildings are unoccupied
-during the pandemic in 2020 and thus building managers have observed a
-drastically decrease in monthly energy bills despite no energy-efficient
-measures were implemented. In addition, common adjustments by an
-building analyst only consider the variations of outdoor weather
-conditions such as by fitting a regression model (e.g. Time-Of-Week
-Temperature model).
+To address the limitations of conventional M&V methods and the
+challenges posed by non-routine events, we propose a novel M&V method
+that adopts the randomized crossover design, a concept borrowed from
+medical and agricultural studies. Another improvement is that we
+proposed a sequential evaluation framework and defined stopping criteria
+to end the M&V if the target effect is detected. This is to avoid
+unnecessary measurement collection over the full 2-year M&V cycle. The
+full framework is detailed in another study with all stopping criteria
+outlined <!--# TODO: Add references -->. In summary, this method
+provides M&V analysts with a randomized schedule that alternates between
+baseline and intervention implementation while ensuring balanced
+sampling across days of the week. For example, given a 10-week M&V
+period for 1 intervention, the balanced randomized schedule would
+equally sample 5 Mondays with the baseline and 5 Mondays with the
+intervention. The limitation of the randomized M&V is that it is only
+applicable to a subset of retrofit projects such as control
+interventions. However, for all applicable use cases, it allows analysts
+to detect energy savings sequentially as the study progresses meaning
+once the desired savings target is achieved, analysts can terminate the
+M&V. The key advantage of randomization, which is one of the study
+objectives, is that if control strategies are sampled with equal
+probability, the influence of non-routine events is likely to be evenly
+distributed between the baseline and intervention measurements. This
+means that the savings estimate, calculated as the difference between
+the two, effectively 'cancels out' the impact of these disturbances,
+leading to a more accurate and unbiased assessment of the intervention's
+impact.
 
@@
 
+The Building Genome Dataset 2 (BGD2) is an extensive open-access dataset
+designed to advance research and development in building energy
+efficiency and control strategies acting as a test-bed for modeling,
+simulation, and algorithm development [@miller_building_2020]. BGD2
+contains over 500 buildings' metadata and realistic operational
+information from across North America and Europe, making it one of the
+most comprehensive collections of building-related data available for
+scientific use. The dataset includes various commercial building types
+such as offices, education facilities, public, and retail buildings, and
+provides detailed information on their physical characteristics (e.g.
+enregy ratings, heating types and floor area) and hourly measurements of
+chilled and hot water, electricity, gas usage as well as site outdoor
+weather conditions.
+
+For the purpose of this study, we only queried whole-building hourly
+electricity usage. All building names and precise locations were erased
+but the climate zone and city name were provided, which are shown in
+Figure \@ref(fig:tidysite) and Figure \@ref(fig:messysite). In addition,
+the measurements were pre-processed by the authors with timestamp
+already converted to local time. We further described the data filtering
+process in Section \@ref(method).
+
 ## Literature review
 
-### Measurement and verification
-
-### Randomized experimental design
+### Measurement and verification energy-efficient measures
+
+Most research related to M&V for whole-building approach focuses on the
+accuracy of baseline modeling, exploring model performance from simple
+regression models to more complex machine learning techniques. One study
+reviewed various models suitable for M&V applications as well as
+selected input features [@alrobaie_review_2022] and another study
+provided a definitive methodology to apply machine learning models for
+M&V use cases [@gallagher_development_2018]. In addition, a few studies
+investigated the critical performance metrics to evaluate the developed
+baseline models [@granderson_automated_2015,
+@granderson_development_2014] and compared a variety of models using
+those metrics [@granderson_accuracy_2016]. These studies made
+significant contributions by emphasizing the uncertainty associated with
+the model-fitting process, a key factor in accurately determining energy
+savings. Furthermore, other researchers addressed this issue by
+leveraging statistical formulation and inference to improve baseline
+energy models [@burkhart_measurement_2014, @heo_calibration_2012,
+@walter_uncertainty_2014]. However, a gap still remains in the
+literature regarding the rigorous quantification of uncertainties
+directly associated with calculated savings, for instance, accounting
+for the potential bias that baseline model might deteriorate (i.e.
+becomes 'stale') over an extended period of pre- and post-analysis.
+
+### Impact of non-routine events on building energy usage
+
+A common non-routine event in energy-saving M&V projects is a change in
+occupancy or a significant shift in occupant behavior. These changes can
+greatly affect measured energy consumption in buildings and are
+typically unrelated to the intervention strategy. For instance, during
+the COVID-19 pandemic in 2020, most commercial buildings were
+unoccupied, leading to a noticeable drop in energy bills despite no
+energy-efficiency measures being implemented [@cai_impact_2024,
+@chihib_impact_2021, @gaspar_assessing_2022, @kang_changes_2021].
+Furthermore, subsequent research has shown that hybrid working modes,
+allowing employees to work remotely, have persisted after the outbreak
+of the pandemic [@aksoy_working_2022], adding further complexity to
+energy consumption patterns due to evolved occupant behaviors with
+higher flexibility [@gui_impact_2021, @mantesi_office_2022,
+@xie_does_2021]. One study realized the limitation of current M&V
+methods, which only consider adjusting for outdoor weather, is
+insufficient and emphasized the importance of requiring matched
+comparison groups to control for exogenous factors beyond weather
+differences when comparing between baseline and intervention
+[@demand_side_analytics_population_2022]. Another type of non-routin
+event is filter clogging in air handling units due to particle
+accumulation. This can cause supply fans to gradually consume more
+energy to maintain required duct static pressure [@feng_newly_2019,
+@zhai_full-scale_2017]. If M&V analysts are unaware of such changes and
+lack an appropriate adjustment method (e.g., replacing filters before
+the intervention begins), the savings could be underestimated as
+increased energy use is incorrectly attributed to the intervention
+rather than the mechanical issue.
+
+### Randomized experimental design in other scientific fields
+
+To address the research gap identified in existing M&V studies,
+particularly the challenges posed by non-routine events, we applied a
+randomized experimental design method---an approach commonly used in
+other scientific disciplines. In clinical trials, for instance,
+randomization is employed to determine the effect of a medical treatment
+(such as a new drug) by randomly assigning participants to either a
+control group (e.g., receiving a placebo) or a treatment group
+[@angus_adaptive_2019, @burger_importance_2021, @lim_randomization_2019,
+@wiley_crossover_2016]. The primary purpose of randomization is to block
+confounding variables such as age, gender, and socioeconomic status when
+evaluating treatment effects. In the context of buildings, a "treatment"
+refers to a control retrofit, while the "placebo" represents the
+existing baseline control. Unlike clinical trials, where participants
+are randomly assigned, our approach randomizes treatment assignment
+longitudinally for each individual building---an approach known as
+n-of-1 trials [@gabler2011]. Although the BDG2 dataset allows for
+randomization at a population level (e.g., across climate zones),
+building owners are generally more interested in understanding the
+effects of control retrofits specific to their own buildings, rather
+than at a generalized population level.
 
@@
 
-As mentioned, the goal of a M&V project is to determine the effect
-(normally savings) of an energy-efficient intervention. And in this
-study, we further limit the study scope to switchable interventions,
-which mostly encapsulates control retrofits. An example intervention of
-such type can be a control retrofit developed by a software-as-a-service
-company that enables the chilled water plant to reset its supply water
-temperature based on outdoor weather condition. Therefore, we defined
-the M&V scenario as follows:
-
-*"A company wishes to sell their supply temperature reset control
-software package to a building owner and guarantees its effect in
-reducing building electricity usage. If the building owner decides to
-purchase the service, the company agrees to charge the service fee based
-on a fraction of the measured savings."*
+As mentioned, the goal of an M&V project is to determine the
+effect---typically energy savings---of an energy-efficient intervention.
+In this study, we focus primarily on switchable interventions, which
+often involve control retrofits. An example of such an intervention is a
+control retrofit developed by a software-as-a-service company that
+adjusts the chilled water plant's supply water temperature based on
+outdoor weather conditions [@roa_field_2023,
+@lee_simulationoptimization_2012, @qiu_chilled_2022]. Therefore, we
+defined the M&V scenario as follows:
+
+*"A company aims to sell its supply temperature reset control software
+package to a customer, such as a building owner, with a guarantee that
+it will reduce the building's electricity usage. If the building owner
+decides to purchase the service, the company agrees to charge a service
+fee based on a percentage of the measured energy savings."*
 
@@
 intervention energy savings for all valid buildings in the dataset. By
-conducting such analysis, we hope to:
-
-1.  Demonstrate the implementation of the proposed randomized M&V method
+conducting such analysis, we aim to:
+
+1.  Compare the energy saving estimation accuracy between the
+    conventional and the randomized method. This study extends the
+    comparison to a large sample of buildings, covering a variety of
+    types and climate zones. The comparison metrics include both
+    estimation accuracy and M&V finishing timeline.
+
+2.  Verify the enhanced robustness of the randomized method. By using
+    realistic measurements from real-world buildings, which include
+    various sources of noise, we aim to reflect the challenges faced by
+    building analysts in real projects. As will be demonstrated in the
+    following sections, the randomized approach is less impacted by
+    non-routine events (i.e., measurement 'noise'), resulting in more
+    reliable energy savings estimates.
+
+3.  Demonstrate the implementation of the proposed randomized M&V method
     using a public available dataset. We ensured the reproducibility of
-    the M&V method by making the analysis code open source including
+    the method by making the analysis code open source including
     randomized schedule generation, sequential statistical analysis,
@@
 
-2.  Compare the energy saving estimation accuracy between the
-    conventional and the randomized method. In particular, this study
-    extends the comparison to large samples of buildings of various
-    types and across multiple climate zones.
-
-3.  Verify the superior robustness of the randomized method over the
-    conventional method. By using the realistic measurements from
-    real-world buildings, which contains various sources of noises,
-    could largely reflect the challenges that a building analyst would
-    be facing in any real project. Particularly, as we will demonstrate
-    in the following sections, non-routine events (i.e. 'noises' in the
-    measurements) have less impact in the energy saving estimation when
-    the analyst uses the randomized approach.
-
 # Method
 
-As mentioned, this study leverages a large public dataset to demonstrate
-the energy saving estimation results of a novel M&V method inspired by
-other scientific research fields. We outlined the methodology of the
-study in Figure \@ref(fig:flowchart) and outlined several key components
-in this section.
+We outlined the methodology of the study in Figure \@ref(fig:flowchart)
+and extended several key components in this section.
 
@@
 1.  Missing values \< 1000: given the hourly resolution of all
-    measurements, this is equivalent to more than a month of missing
-    days.
+    measurements, this is equivalent to 1.5 months of missing days.
 
@@
 4.  Target buildings should have known site location: buildings with
-    anonymous location can not request typical meteorological weather.
-
-5.  Warehouse and parking types are excluded: target buildings that are
-    likely to have non-regular electricity and chilled water usage
-    should be excluded.
-
-As a summary, the resulting subset contains all buildings with 'tidy'
+    anonymous location are excluded due to unavailable typical
+    meteorological weather.
+
+5.  Warehouse and parking types are excluded: target buildings have less
+    demand flexibility to implement a chilled water setpoint reset
+    control.
+
+Therefor the resulting subset contains all buildings with 'tidy'
 measurements. Figure \@ref(fig:tidysite) shows in total, the subset
@@
 
-The 'tidy' subset might be presentative to the measurements collected
-among the existing building stock since in reality, whole-building
-electricity measurements collected over two years normally exhibit
-larger changes. Those measured changes can be caused by sensor itself
-such as lack of calibration, or inherent changes of the building, which
-illustrated before, can bias the result of a M&V. Therefore, to compare
-the robustness of the two M&V method more realistically, we included an
-additional 'messy' subset which first exclude the 'tidy' subset and then
-re-apply the filtering rule with one amendment:
-
-3.  Absolute mean difference of the two-year electricity usage \< 25%:
-    any increase or decrease of building electricity usage in the second
-    year should be less than 25% of that in th first year.
-
-Figure \@ref(fig:messysite) shows the summary of the 'messy' dataset. As
-this is less aggressive, it contains `r nrow(all_names_messy)` buildings
-in `r nrow(all_types_messy)` types from `r nrow(all_sites_messy)`
-different climate zones.
+In reality, the 'tidy' subset is less representative of the measurements
+typically collected from existing building stock, as whole-building
+electricity measurements collected over two years often show more
+variability the the filtering critieron. These variations can stem from
+issues like sensor calibration errors or inherent changes in the
+building, which, as previously discussed, can bias M&V results.
+Therefore, to more realistically assess the robustness of the two M&V
+method, we included an additional 'messy' subset which first exclude the
+'tidy' subset and then re-apply the filtering rule with one amendment:
+
+3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
+    decrease of building electricity usage in the second year should be
+    less than 25% of that in th first year.
+
+Figure \@ref(fig:messysite) shows the summary of the 'messy' dataset,
+which contains `r nrow(all_names_messy)` buildings in
+`r nrow(all_types_messy)` types from `r nrow(all_sites_messy)` different
+climate zones.
 
@@
 intervention that reset the chiller supply temperature based on the
-outdoor weather conditions. Under both strategies, we assume the chiller
-is enabled when outdoor temperature is higher than 10°C. The baseline,
-which assumed to be the existing measurements from the dataset, runs at
-a constant water supply temperature. The intervention as the figure
-shows reset from 7 °C to 12 °C.
+outdoor weather conditions, which can be commonly found in the
+literature [@lee_chilled_2022, @congradac_recognition_2012]. For both
+strategies, we assume that the chiller is activated when the outdoor
+temperature exceeds 10°C. The baseline strategy, representing the
+existing measurements from the dataset, operates with a constant water
+supply temperature. The intervention strategy, as illustrated in the
+figure, adjusts the water supply temperature dynamically, resetting it
+from 7°C to 12°C.
 
@@
 
-We map the chilled water supply temperature reset to the electrical
+We mapped the chilled water supply temperature reset to the electrical
 energy savings as: $$
@@
 
-We assume in general, HVAC uses around 50% of whole-building electricity
-and the chilled water plant further uses 50% electricity of the HVAC
-system. This assumption largely simplifies the unique characteristics of
-electricity usage in all types of commercial buildings, but for the
-scope of this paper, we therefore assume 25% of whole-building
-electricity is consumed by the chilled water plant ($E_{CHW}$).
-<!--# TODO: Add reference -->In most cases, savings from an intervention
-is unproportional to the building hourly electricity usage, which
-renders M&V challenging. Therefore, we map the resulting electricity
-savings as a percentage of the mean electricity consumption of the plant
-determined by outdoor temperature ($OAT$), intervention supply water
-temperature ($T_{interv}$), baseline supply water temperature
-($T_{base}$) and hour of the day ($\delta_{occ}$, binary indicator
-whether during peak hours from 9 AM to 4 PM).
+We assume on average, HVAC systems account for approximately 50% of a
+building's total electricity consumption, and the chilled water plant
+further consumes around 50% of the HVAC electricity. While this
+assumption largely simplifies the diverse energy usage across various
+building types, for the scope of this paper, we assume that 25% of the
+total building electricity is used by the chilled water plant, $E_{CHW}$
+[@us2012commercial]. Typically, the savings from an intervention are not
+proportional to the building's hourly electricity usage, which is
+generally the challenge for M&V. To address this, we mapped the
+resulting electricity savings as a percentage of the plant's normal
+operation, calculated as its mean electricity usage over the two-year
+period. This percentage is influenced by factors such as outdoor
+temperature ($OAT$), intervention supply water temperature
+($T_{interv}$), baseline supply water temperature ($T_{base}$) and hour
+of the day ($\delta_{occ}$, binary indicator whether during peak hours
+from 9 AM to 4 PM).
 
@@
 Parameters and their pre-defined values are summarized in the table
-below. Those parameters were not rigorously calibrated for each
-building, and were used uniformly across the dataset. Although this
-calibration process should be included, it is considered tangential to
-the study scope outlined in this paper.
-
+below. For simplicity, those parameters were not rigorously calibrated
+for each building, and were used uniformly across the dataset.
+
+|                        |                                                                  |       |
+|------------------|------------------------------------|------------------|
 | Parameter              | Description                                                      | Value |
-|------------------|------------------------------------|------------------|
 | $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
@@
 
-: Parameters for calculating the intervention savings
+: Table 1. Parameters for calculating the intervention savings.
 
@@
 
-To demonstrate the influence of non-routine events that are not properly
-adjusted, we developed three scenarios to quantify the energy
-consumption change associated with a hypothetical occupancy change. In
-summary, we assume an increase in occupancy in 2016 (i.e. refered as the
-'baseline year' if using conventional method) resulted in 20% increase
-in whole-building measured electricity usage. The three scenarios
-specifies the increase happened during January to April (S1); May to
-August (S2); and September to December (S3). We applied such change only
-to the buildings in the 'tidy' subset as originally they are more likely
-to have no non-routine events.
+To illustrate the impact of non-routine events that are not properly
+adjusted for, we developed three scenarios to quantify the energy
+consumption changes associated with a hypothetical increase in
+occupancy. Specifically, we assume that a rise in occupancy in 2016
+(referred to as the "baseline year" when using the conventional method)
+led to a 20% increase in whole-building measured electricity usage. The
+three scenarios describe when this increase occurred: January to April
+(S1), May to August (S2), and September to December (S3). This
+hypothetical change was applied only to buildings in the "tidy" subset,
+as these buildings were initially more likely to be free from
+non-routine events.
 
@@
 
-```{r convmethod, fig.cap="Flow chart showing the conventional M&V process for an energy-saving intervention", out.width="50%"}
-
-knitr::include_graphics(paste0(fig_path, "convmethod.png"))
-```
-
-As described in Section \@ref(introduction) and Figure
-\@ref(fig:convmethod), conventional method for M&V is time-consuming and
-unable to separate non-routine event impact from measured savings. If
-using conventional method to estimate intervention energy savings, the
-result (after 24 months) is calculated by the difference between
-measured intervention and projected baseline in the post-retrofit
-period. In this study, we leveraged a piece-wise linear regression
-considering time-of-week and outdoor temperature (TOWT) as independent
-variables for projection and normalization on typical meteorological
-year <!--# TODO: Add reference -->.
+As described in Section \@ref(introduction), conventional method for M&V
+is time-consuming and unable to separate non-routine event impact from
+measured savings. If using conventional method to estimate intervention
+energy savings, the result (after 24 months) is calculated by the
+difference between measured intervention and projected baseline in the
+post-retrofit period. In this study, we leveraged a piece-wise linear
+regression considering time-of-week and outdoor temperature (TOWT) as
+independent variables for projection for the counter-factual baseline in
+the post-retrofit period and normalization on typical meteorological
+year [@mathieu_quantifying_2011].
 
@@
 
-```{r randmethod, fig.cap="Flow chart showing the novel randomized M&V process for an energy-saving intervention", out.width="50%"}
-
-knitr::include_graphics(paste0(fig_path, "randmethod.png"))
-```
-
-Compared to the conventional method, the randomized M&V can provide an
-estimation more rapid and reliable. Section \@ref(introduction)
-describes the application of randomization techniques and Figure
-\@ref(fig:randmethod) specifically outlined the process of conducting a
-randomized switchback experiment for control retrofit projects. The
-method sequentially evaluate the intervention effect and returns a
-saving estimation with 95% confidence interval when all stopping
-criteria are satisfied. <!--# TODO: Add paper DOI --> describes all
-stopping criteria and recommened threshold to consider in more details
-and we breifly summarized the thresholds used in this paper here:
+Compared to the conventional method, the randomized M&V approach offers
+a more rapid and reliable estimation of energy savings. To apply this
+method, analysts first define the target savings and stopping criteria.
+Then they design a randomized switchback schedule and perform sequential
+statistical tests, such as the sequential probability ratio test (SPRT),
+to monitor savings as data is collected. Once the target savings are
+detected, the analysts fit a prediction model (e.g. TOWT) to adjust for
+differences in outdoor temperature, following the same adjustment
+process as the conventional method. We provide example switchback
+experimental design and stopping criteria thresholds below:
 
@@
 
-The timeline for the conventional M&V method is outlined in previous
-sections. Although ASHRAE Guideline 14 offers minimum requirements for
-each path, the required baseline period is either a full range of all
-independent variables (typically outdoor weather conditions) under
-normal facility operation or a 12-month worth of continuous
+ASHRAE Guideline 14 offering minimum requirements for whole-building
+measurement path states that the baseline period is either a full range
+of all independent variables (typically outdoor weather conditions)
+under normal facility operation or a 12-month worth of continuous
 measurements. The same requirements also applies to intervention
@@
 to delay in retrofit deployment, he/she needs to wait until the next
-cooling season to measure the savings.
-
-If the randomized M&V method is applicable, even if there is a delay in
-the retrofit deployment, the building owner faces less risk and impact
-since random sampling allows measuring the full range of all independent
-variables simultaneously among all control strategies or modes. In
-addition, sampling at an equal probability (i.e. 50%/50% between
-baseline/intervention) effectively balances the level of independent
-variables measured. In other words, unless one strategy is sampled with
-a large amount of consecutive days (e.g. 7 days or more), it is rare
-that the independent variable, such as outdoor weather condition, is
-measured significantly different among all sampled control strategies.
+cooling season to measure the savings. If the randomized M&V method is
+applicable, delays in retrofit deployment pose less risk to the building
+owner since baseline are continuously monitored. Sampling at equal
+probabilities (e.g., 50%/50% between baseline and intervention) helps to
+balance the distribution of independent variables like outdoor weather
+conditions across all sampled control strategies. As long as one
+strategy is not sampled over a large number of consecutive days (e.g., 7
+days or more), it is unlikely that the measurements of independent
+variables will differ significantly between the control strategies.
+
 Figure \@ref(fig:timeline) shows the average estimated timeline for each
@@
 supplementary material. The figure shows for all climate zones, covering
-a sufficient range outdoor weather condition is the most stringent
-criterion. However, most buildings can satisfy the 80% range by 6 months
-and all buildings can quantitatively determine the savings with
-associated uncertainty by 9 months, which is even shorter than the
-length of baseline measurement required by the conventional method.
+a sufficient range of outdoor weather condition is the most stringent
+requirement. However, most buildings can achieve 80% of the required
+range within 6 months, and all buildings can determine the savings,
+including associated uncertainties, within 9 months---significantly
+shorter than the baseline measurement period required by the
+conventional method.
 
@@
 
-In this study, we define the overall target savings (i.e. the ground
-truth) that an M&V analyst wishes to detect, either in normalized
-savings, fractional savings or simply the measured difference between
-the baseline and the intervention measurements, is the mean electricity
-energy reduction over the two-year period from implementing the chilled
-water supply temperature reset control. We estimate the M&V savings by
-the two methods following the process given in Figure
-\@ref(fig:convmethod) and Figure \@ref(fig:randmethod). In orther words,
-we compare the conventional M&V savings at the end of the two-year
-period with the randomized M&V savings after all criteria are satisfied.
+In this study, we define the overall target savings (i.e., the ground
+truth) that an M&V analyst aims to detect as the mean reduction in
+electricity consumption over a two-year period resulting from the
+implementation of the chilled water supply temperature reset control.
+This target can be expressed in terms of normalized savings, fractional
+savings, or simply the measured difference between baseline and
+intervention measurements. Due to different timeline required by the two
+methods, we compare the conventional M&V savings at the end of the
+two-year period with the randomized M&V savings after all criteria are
+satisfied. Since those are the time in reality, an M&V analyst report
+estimated savings.
 
 Figure \@ref(fig:norm) shows the overall comparison of the savings
-estimation from all buildings normalized on the typical meteorological
-year weather conditions of each site. In subplot a), the narrower range
-of the true savings indicates the dependence of the intervention effect
-on the outdoor weather condition, which is intended as shown in Figure
-\@ref(fig:chwst). Sites with mild climate all year round such as
-locations in California shows higher savings potential above 10%, while
-locations with more extreme climate such as Washington DC shows only 6%
-savings annually. In addition, although the box plot indicates some
-savings estimation are 'outliers', they should not be interpreted as
-such since the distribution can be biased by the unbalanced sample sizes
-across different site shown in Figure \@ref(fig:tidysite) and Figure
-\@ref(fig:messysite). In subplot b), the results show that the
-conventional M&V method tends to estimate savings with a greater
-uncertainty and we will demonstrate in the following sections that
-non-routine events can have a significant impact. For the randomized
-method, we displayed two scenarios, the estimation at the end of
-two-year period with 50%/50% sampling throughout and the estimation
-provided after all stopping criteria satisfied (i.e. stops early at 24
-weeks or 36 weeks). Since we found no significant difference between the
-two distributions as shown in the figure, we demonstrated that an early
-stop has no impact on the estimation accuracy but can significantly
-reduces the cost associated with M&V.
+estimated for all buildings in both subsets normalized on the typical
+meteorological year weather conditions of each site. In subplot a), the
+narrower range and clustering of the true savings indicates the
+dependence of the intervention effect on the outdoor weather condition,
+which is intended after weather normalization. Sites with mild climate
+all year round such as locations in California shows higher savings
+potential above 10%, while locations with more extreme climate such as
+Washington DC shows only 6% savings annually. In subplot b), the results
+indicate that the conventional M&V method tends to estimate savings with
+greater uncertainty, and its distribution median deviates from the true
+savings. In comparison, the randomized method, which requires much less
+time, shows more accurate and precise estimation results.
 
@@
   ggplot(aes(x = method, y = savings, fill = method)) +
-  stat_boxplot(geom ='errorbar', width = 0.5) +
-  geom_boxplot(outlier.size = 1) +
+  geom_errorbar(aes(ymin = min(savings), ymax = max(savings)), width = 0.2) +
+  geom_boxplot(outlier.size = 2) +
+  geom_text(aes(x = 1.35, 
+                y = median(savings) + 1.5, 
+                label = paste0("median: ", round(median(savings), digits = 0), "%")), 
+            check_overlap = T) +
   scale_y_continuous(expand = c(0, 0), 
@@
   pivot_longer(c(conv, rand, FS), names_to = "parameter", values_to = "value") %>% 
+  filter(parameter != "rand") %>% 
   mutate(parameter = as.factor(parameter), 
@@
                                    "conv" = "Conventional M&V", 
-                                   "rand" = "Randomized M&V", 
-                                   "FS" = "Randomized M&V at early stop")) %>% 
+                                   "FS" = "Randomized M&V")) %>% 
   ggplot(aes(x = parameter, y = value, fill = parameter)) +
@@
             ncol = 2, nrow = 1,
-            widths = c(0.35, 1),
-            legend = "bottom") +
+            widths = c(0.5, 1),
+            legend = "none") +
     plot_annotation(title = "Normalized annual fractional savings estimation comparison", 
@@
 
-To assess the accuracy on individual building, we calculate the mean
-absolute difference between the true savings and estimated savings.
-Figure \@ref(fig:tidyfrac) shows the comparison between the two methods
-using the tidy subset. Because of the large variation in electricity
-energy usage, we express the savings as a normalized value which is a
-fraction of energy reduction compared to baseline. Furthermore, we focus
-on the median and 95% threshold of the mean absolute difference
-distribution in the comparitive assessment to filter out outliers. As a
-result, the conventional method performs better than the randomized
-method. One hypothesis is that the tidy subset only includes buildings
-exposed to similar usage throughout the two years, meaning the
-measurement used for regression model fitting resembles the model
-prediction results. Although the results indicate the conventional M&V
-method is preferable if a building can achieve long-term consistency in
-electricity usage, such condition is rarely garrenteed in reality.
+To assess accuracy at the individual building level, we calculated the
+mean absolute difference between the true savings and the estimated
+savings and normalized the results as a fraction of measured baseline
+for generalized comparison. Figure \@ref(fig:tidyfrac) shows the
+comparison between the two methods using the tidy subset. In this
+comparison, we focused on the median and the 95th percentile of the
+plotted distribution to minimize the influence of outliers. The results
+indicate that the conventional method outperforms the randomized method
+in this subset. This is mostly because the tidy subset includes
+buildings with consistent usage patterns throughout the two-year period,
+meaning the data used for regression model fitting closely aligns with
+the conditions during the model prediction phase. In our case, this
+justifies the TOWT model selection when there is no additional
+adjustment needed for such pre- and post-analysis.
 
@@
 
-A more realistic condition is shown in the messy subset in Figure
-\@ref(fig:messyfrac). The randomized method demonstrates consistent
-accuracy in savings estimation, even in the presence of significant
-measurement noise, indicating strong robustness. On the contrary, the
-conventional method shows approximately in 50% of all cases, the
-deviations of M&V estimation are more than 5%. We showed the
-non-normalized mean absolute difference comparison plots in the
-supplementary material. The poor performance shows if the there exist a
-change in electricity usage, baseline projection based on regression
-models are unreliable. Those changes are normally referred as
-non-routine events, which summarizes all the influential factors that
-are not considered in the regression model either because they are
-impractical to measure, such as hourly occupancy rate, or too random
-with no expected patterns to quantify, such as a sudden change of use
-from an office to a warehouse. We will demonstrate this in more details
-in the following sections.
+While these findings suggest that the conventional M&V method is
+advantageous when a building maintains stable electricity usage over
+time, such stability is rarely guaranteed in real-world scenarios. A
+more realistic condition is shown in the messy subset in Figure
+\@ref(fig:messyfrac). In this case, the randomized method demonstrates
+consistent accuracy in savings estimation, even in the presence of
+random measurement noise, implying strong robustness. In contrast, the
+conventional method shows deviations exceeding 5% in approximately 50%
+of all cases. Non-normalized mean absolute difference comparison plots
+are provided in the supplementary material. The much deteriorated
+performance of the conventional method highlights its vulnerability when
+changes in electricity usage occur. This is because baseline projections
+based on regression models become unreliable when non-routine events are
+present. Non-routine events encompass all influential factors not
+accounted for in the regression model, either due to their
+impracticality to measure---such as hourly occupancy rates---or their
+unpredictability---such as a sudden change in building use where an
+office being converted into a warehouse. We dedicate the following
+section to continue the discussion of the impact.
 
@@
 
-```{r nosavings, fig.cap="Distribution of savings estimation results by the convention M&V method and randomized method.", fig.width=8, fig.height=4}
+```{r nosavings, fig.cap="Distribution of savings estimation results by the convention M&V method and randomized method.", fig.width=6, fig.height=3}
 
@@
   pivot_longer(c(true, conv, rand, FS), names_to = "parameter", values_to = "value") %>% 
-  filter(parameter != "true") %>% 
+  filter(parameter != "true" & parameter != "rand") %>% 
   mutate(parameter = as.factor(parameter), 
@@
   annotate(geom = "text", 
-           x = 0.5, 
+           x = 0.7, 
            y = 0.5, 
@@
         legend.direction = "horizontal",
-        legend.position = "bottom",
+        legend.position = "none",
         plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
@@
 average detected 5% savings when there should be none. The randomized
-method even when stops right after satisfying all criteria, the savings
-estimation is much closer to 0%. This is mostly because when there is a
+method on the other hand is able to provide an estimation much closer to
+0% with reduced uncertainty. This is mostly because when there is a
 change or disturbance applied to the target building such as lighting
@@
 randomized M&V results, a new schedule sampling at 80%/20% was
-implemented for another 36 weeks.
+implemented till the end of 2016 and the plot shows the saving
+estimation accuracy calculated as difference-in-difference when sampled
+baseline reduced.
 
@@
 opting for an early stop at 24 or 36 weeks with 50% intervention and
-resampling 80% of the data for 36 weeks still allows the building owner
-to capture 65% of the full-range savings.
+re-sampling 80% till the year end (\~ another 20 weeks) still allows the
+building owner to capture 65% of the full-range savings. On the other
+hand, if the building owner follows the conventional method, there is no
+savings due to required baseline measurement.
 
@@
 
-Considering the diversity of the measurement sets, the assumption made
-in this study is relative general. For example,
+We identify two key limitations of this study: 1) the application of
+control intervention, and 2) the design of the randomized switchback
+experiment. While the primary focus is on accurately detecting
+intervention effects rather than validating their impact, the simulated
+intervention remains somewhat generic due to the diverse building types
+and climate zones in the BDG2 dataset. For simplicity, we applied the
+parameters listed in Table 1 uniformly across all buildings. However,
+the electricity savings from raising the water temperature by 1°C may be
+less than 8% in some cases, due to increased chilled water pump speeds.
+Similarly, the randomized schedule design is also simplified. We assumed
+a daily sampling interval would be sufficient for most commercial
+buildings, but exceptions exist, particularly for buildings with
+significant thermal lag in heating and cooling systems, such as those
+with heavy concrete construction, hot water tanks, or Thermally Active
+Building Systems (TABS). In such cases, the effect of a previously
+sampled strategy can carry over and influence subsequent measurements
+due to thermal storage, known as the carryover effect. For example, if
+the intervention pre-charges chilled water in thermal mass one day
+before switching back to the baseline, the analyst is likely to observe
+reduced energy usage in the following days. This study does not account
+for the carryover effect, but in practice, we recommend using a 3-day
+sampling interval and dropping non-consecutive days to 'wash out' any
+lingering effects from previous strategies.
+
+Our major goal for the future study is to further extend the application
+of the proposed randomized M&V. For example, a customer maybe interested
+in whether a Model Predictive Control (MPC) can reduce energy bill under
+dynamic pricing structure and what the estimated Return of Investment
+(ROI) for the retrofit is. Or whether a load shift control can save
+operational cost by shifting load from grid peak to off peak period.
+Furthermore, the novel method is generalizable beyond energy saving M&V
+cases. The statistical analytic framework is applicable to a variety of
+retrofit metrics such as indoor air quality, operational carbon
+emissions and thermal comfort.
 
 # Conclusion
+
+This research demonstrated the usage and assessed the performance of a
+novel whole-building measurement and verification method using a large
+open-source commericial dataset. The method leverages the randomized
+experimental design concept from other scientific research fields and
+statistical sequential inference techniques to inform whether a target
+savings are detected. We made an example of chilled water setpoint reset
+based on outdoor weather condition as a virtual control retrofit use
+case and applied to over 500 filtered commercial buildings. By comparing
+the savings estimation by the conventional method (outlined in ASHRAE
+Guideline 14) and the novel randomized method, we showed that the
+randomized method is able to provide a more rapid and robust saving
+estimation.
+
+We showed that throughout 7 climate zones we assessed, the randomized
+M&V can provide a saving estimation by 36 weeks (with the majority
+finishes by 24 weeks) once all stopping criteria satisfied. Meanwhile,
+the conventional method requires a full range measurement of both
+baseline and intervention performance under normal operation conditions,
+which normally translate to 6 - 9 months each. We further showed the
+estimated savings at early stop has acceptable accuracy when compared
+with the true calculated savings.
+
+We also assessed the impact of non-routine events on the proposed M&V
+method both by 1) adding a known change, such as occupancy-induced
+energy reduction, and 2) detecting removed intervention when no reset is
+applied to buildings with a marginal difference in energy consumption.
+We showed in both scenarios that baseline model fitting can be biased
+while the randomization can efficiently block those confounding effect
+to ensure the robustness of the saving estimation.
+
+Although the limitation of the method is that it only applies to a
+subset of all M&V use cases (i.e. only if the strategy can be switched
+on and off), we believe the true value lies in the usefulness for most
+control retrofit validation in the field test.
+
+# Acknowledgements
+
+# References

### 2024-10-13 code/manuscript.Rmd
Adjust for NRE re-simulation

@@
 |                        |                                                                  |       |
-|------------------|------------------------------------|------------------|
+|------------------|-----------------------------------|------------------|
 | Parameter              | Description                                                      | Value |
@@
 
-In this section, we demonstrate the influence of non-routine events
-through two scenarios. In the first scenario, we quantified how much
-occupancy would influence the electricity usage and added to the tidy
-measurement set as they originally indicates no such change. In the
-second scenario, we simply run the two M&V methods on the messy
-measurement sets before adding the chilled water supply temperature
-reset. Since no intervention is added, the more reliable method should
-detect a 'saving' closer to 0.
+We included two examples in this section to demonstrate the influence of
+non-routine events on M&V savings. In the first scenario, we manually
+subtracted a certain amount of electricity usage for a period of time
+quantified how much occupancy would influence the electricity usage and
+added to the tidy measurement set as they originally indicates no such
+change. In the second scenario, we simply run the two M&V methods on the
+messy measurement sets before adding the chilled water supply
+temperature reset. Since no intervention is added, the more reliable
+method should detect a 'saving' closer to 0.
 

### 2024-10-13 code/1_analysis.R

@@ -56,9 +56,9 @@ source(paste0(function_path, "cont_plot.R"))
 # section run control
 run_params <- list(type = "tidy", 
                    sprt = T, 
-                   sprt_cont = T, 
-                   nre_occ = F, 
-                   nsprt = T)
+                   sprt_cont = F, 
+                   nre_occ = T, 
+                   nsprt = F)
 
 # Adding intervention effect as advanced chiller operation
 ctr_params <- list(peak_hours = 10:16,                      # accounts for peak hours
@@ -95,7 +95,7 @@ cont_param <- list(baseline = "Baseline",
 # NRE: Occupancy change
 occ_params <- list(change_start = c(1, 5, 9),
                    change_end = c(4, 8, 12),
-                   change = c(10, 20))
+                   change = 20)
 
 
 
@@ -380,75 +380,75 @@ for (n in 1:(nrow(all_names))){
     run_reset()
   
   # power-temp plot
-  p1 <- df_hourly_conv %>%
-    pivot_longer(c(base_eload, interv_eload), names_to = "strategy", values_to = "eload") %>%
-    mutate(strategy = as.factor(strategy),
-           strategy = recode_factor(strategy, "base_eload" = "Baseline", "interv_eload" = "Intervention")) %>%
-    ggplot(aes(x = t_out, y = eload, color = strategy)) +
-    geom_point(data= .%>%
-                 group_by(strategy) %>%
-                 slice_sample(n = 1000),
-               size = 0.7,
-               alpha = 0.4,
-               shape = 16) +
-    geom_smooth(formula = y ~ x, method = "loess", linewidth = 1.25, alpha = 0.15) +
-    scale_x_continuous(expand = c(0, 0),
-                       breaks = breaks_pretty(n = 4),
-                       labels = number_format(suffix = " °C")) +
-    scale_y_continuous(expand = c(0, 0),
-                       breaks = breaks_pretty(n = 4),
-                       labels = number_format(suffix = " kW")) +
-    scale_color_manual(values = ls_colors) +
-    coord_cartesian(ylim = plot_scale) +
-    labs(x = NULL,
-         y = NULL,
-         color = NULL,
-         subtitle = "by outdoor drybulb temperature") +
-    theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-          legend.direction = "horizontal",
-          legend.position = "bottom",
[…]

### 2024-10-20 code/manuscript.Rmd
Changed site names and regrouped visualization + minor figure edits

--- 2024-10-11 code/manuscript.Rmd
@@
   - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Durate ^1^
-abstract: Conventional measurement and verification (M&V) methods for whole-building energy savings estimation are both time-consuming and unreliable, especially when non-routine events occur during the M&V process.
-Those events are unrelated to the proposed intervention strategy but have substaintial impacts on the building energy consumption.
-In this study, we argue that for switchable interventions (e.g.
-most of the control retrofits) can benefit from random sampling where the analyst randomly decide which strategy (i.e.
-baseline or intervention) to implement each day.
-We tested the novel randomized M&V method on a large public dataset which covers multiple climate zones and types of commercial buildings.
-We applied a virtual chilled water supply temperature reset based on outdoor weather as a control retrofit intervention.
-Our study shows that the new M&V method can estimate the savings accurately much quicker than the conventional method and most importantly, the estimation results are much more robust compared to the conventional method when non-routine events are present.
+abstract: Conventional measurement and verification (M&V) methods relying on pre- and post-  retrofit comparison for estimating whole-building energy savings are often time-consuming and unreliable, especially when non-routine events such as reduced occupancy occur during the M&V process.
+Those events are unrelated to the intervention but significantgly affect building energy consumption and thus when the analyst applies conventiona M&V suggested by industry guidelines, the results can be largely confounded.
+In this study, we demonstrated that switchable interventions, such as most of the control retrofits, can benefit from random sampling, where the analyst randomly selects whether to implement the baseline or the intervention strategy each day.
+We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention.
+Our results show that comparing to the conventional method, the randomized method provides faster and more accurate savings estimation.
+Additionaly we found that the randomized M&V approach estimates much closer to the true savings when non-routine events are present demonstrating much improved reliability.
 
@@
 # colors
-ls_colors <- c("Baseline" = "#a6bddb",
-               "Measured baseline" = "#a6bddb",
-               "Adjusted baseline" = "#a6bddb",
-               "Projected baseline\n(no change)" = "#2b8cbe",
+ls_colors <- c("Baseline" = "#c6dbef",
+               "Measured baseline" = "#c6dbef",
+               "Adjusted baseline" = "#c6dbef",
+               "Projected baseline" = "#2171b5",
                "Intervention" = "#fdbb84",
                "Measured interv" = "#fdbb84",
-               "True savings" = "black",
-               "True normalized savings" = "black")
+               "Conventional M&V" = "grey70", 
+               "Randomized M&V" = "#66c2a4", 
+               "Buildings satisfying all criteria" = "#66c2a4", 
+               "Buildings satisfying 80% TMY range" = "black", 
+               "Buildings satisfying SPRT" = "black")
 
@@
 #### READ DATA ####
+df_loc <- read.csv("../readfiles/loc_map.csv")
+
 # Tidy dataset
@@
 
-```{r flowchart, fig.cap = "Workflow summary of the methodology of this paper", out.width="50%"}
+```{r flowchart, fig.cap = "Flow chat showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="50%"}
 
@@
 measurements. Figure \@ref(fig:tidysite) shows in total, the subset
-contains `r nrow(all_names_tidy)` buildings in `r nrow(all_types_tidy)`
-types from `r nrow(all_sites_tidy)` different climate zones.
-
-```{r tidysite, fig.cap = "Site summary of the tidy building subset", fig.width=8, fig.height=6}
+contains `r nrow(all_names_tidy)` buildings of `r nrow(all_types_tidy)`
+types.
+
+```{r tidysite, fig.cap = "Site summary of the tidy building subset (counts < 2 are omitted for visualization; left: aggragated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}
 
@@
   geom_col() +
-    geom_text(aes(label = ifelse(n > 10, as.character(n), "")), color = "black", position = position_stack(vjust = 0.5)) +
+  geom_text(aes(label = ifelse(n > 2, as.character(n), "")), color = "black", position = position_stack(vjust = 0.5)) +
   scale_y_continuous(expand = c(0, 0),
@@
        y = NULL,
-         subtitle = "Across all sites",
+       subtitle = "Combining all locations",
        fill = NULL) +
@@
   p2 <- df_energy_tidy %>%
-    group_by(site, type) %>%
+    left_join(df_loc, by = "site") %>% 
+    group_by(location, type) %>%
     distinct(name) %>%
@@
     xlim(c(2, 4)) +
-    facet_wrap(~ site, nrow = 3) +
+    facet_wrap(~ location, nrow = 3) +
     labs(x = NULL,
          y = NULL,
-         subtitle = "For each site",
+         subtitle = "For each location",
          fill = NULL) +
     scale_fill_manual(values = type_colors) +
-    geom_text(aes(x = 3.5, y = label_pos, label = n)) +
+    geom_text(aes(x = 3.5, y = label_pos, label = ifelse(n > 2, as.character(n), ""))) +
     geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black") +
@@
             legend = "bottom") +
-    plot_annotation(title = "Case study building type summary")
+    plot_annotation(title = "Summary of building numbers and types for 'tidy' subset")
 ```
@@
 which contains `r nrow(all_names_messy)` buildings in
-`r nrow(all_types_messy)` types from `r nrow(all_sites_messy)` different
-climate zones.
-
-```{r messysite, fig.cap = "Site summary of the messy building subset", fig.width=8, fig.height=8}
+`r nrow(all_types_messy)` types.
+
+```{r messysite, fig.cap = "Site summary of the messy building subset (counts < 2 are omitted for visualization; left: aggragated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=8}
 
@@
          y = NULL,
-         subtitle = "Across all sites",
+         subtitle = "Combining all locations",
          fill = NULL) +
@@
   p2 <- df_energy_messy %>%
-    group_by(site, type) %>%
+    left_join(df_loc, by = "site") %>% 
+    group_by(location, type) %>%
     distinct(name) %>%
@@
     xlim(c(2, 4)) +
-    facet_wrap(~ site, nrow = 3) +
+    facet_wrap(~ location, nrow = 3) +
     labs(x = NULL,
          y = NULL,
-         subtitle = "For each site",
+         subtitle = "For each location",
          fill = NULL) +
     scale_fill_manual(values = type_colors) +
-    geom_text(aes(x = 3.5, y = label_pos, label = n)) +
+    geom_text(aes(x = 3.5, y = label_pos, label = ifelse(n > 1, as.character(n), ""))) +
     geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black") +
@@
             legend = "bottom") +
-    plot_annotation(title = "Case study building type summary")
+    plot_annotation(title = "Summary of building numbers and types for 'messy' subset")
 ```
@@
 
-```{r chwst, fig.cap = "Proposed chilled water supply temperature reset based on outdoor temperature", fig.width=8, fig.height=4}
+```{r chwst, fig.cap = "Proposed intervention strategy: chilled water supply temperature reset based on outdoor temperature", fig.width=8, fig.height=4}
 
@@
   ggplot(aes(x = OAT, y = value, color = strategy)) +
-  geom_line() +
+  geom_line(linewidth = 1.2) +
   scale_color_manual(values = ls_colors) +
@@
        y = "Chilled water supply temperature", 
-       title = "Chilled water supply temperature reset on outdoor temperature") +
+       color = NULL, 
+       title = "Intervention logic: chilled water supply temperature reset") +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
 |                        |                                                                  |       |
-|------------------|------------------------------------|------------------|
+|-------------------|----------------------------------|-------------------|
 | Parameter              | Description                                                      | Value |
@@
 timeline figures for each individual building are attached in the
-supplementary material. The figure shows for all climate zones, covering
-a sufficient range of outdoor weather condition is the most stringent
-requirement. However, most buildings can achieve 80% of the required
-range within 6 months, and all buildings can determine the savings,
-including associated uncertainties, within 9 months---significantly
-shorter than the baseline measurement period required by the
-conventional method.
-
-```{r timeline, fig.cap="Average randomized M&V timeline summary for all buildings at each site", fig.width=8, fig.height=4}
-
-df_sprt_all_messy %>% 
-  filter(seq == "eob" | seq == "temp") %>% 
-  bind_rows(df_sprt_all_tidy %>% filter(seq == "eob" | seq == "temp")) %>% 
-  group_by(site, seq) %>% 
-  summarise(n_weeks = mean(n_weeks)) %>% 
-  ungroup() %>% 
-  mutate(seq = as.factor(seq), 
-         seq = recode_factor(seq, "temp" = "80% temperature range satisfied", "eob" = "remaining weeks till finish")) %>% 
+supplementary material. The figure shows all buildings can detect a
+saving statistically from sampled measurements within 12 months, which
+is equivalent to the minimum test criterion. Similar to the findings
+from our pervious study, covering a sufficient range of outdoor weather
+condition is the most stringent requirement. If target buildings are
+located in a climate zone with more preferable outdoor weather
+conditions such as California, covering 80% of the TMY range should only
+take 3 \~ 4 months. In this case, the analyst can conclude the M&V
+simultaneously if the blocking period ends at the same time. However,
+given most buildings requires 6 months to achieve 80% of the TMY range,
+all buildings can determine the savings, including associated
+uncertainties, within 9 months---significantly shorter than the baseline
+measurement period required by the conventional method.
+
+```{r timeline, fig.cap="Randomized M&V timeline of satisfying key stopping criteria for all buildings", fig.width=8, fig.height=4}
+
+df_time <- df_sprt_all_messy %>% 
+  filter(seq != "final") %>% 
+  bind_rows(df_sprt_all_tidy %>% filter(seq != "final")) %>% 
+  select(name, seq, n_weeks)
+
+count <- list()
+
+for (i in seq(1, 36, by = 1)){
+
+  df_sprt <- df_time %>% 
+    filter(seq == "sprt",
+           n_weeks <= i)
+
+  df_eob <- df_time %>% 
+    filter(seq == "eob",
+           n_weeks <= i)
+
+  df_temp <- df_time %>% 
+    filter(seq == "temp", 
+           n_weeks <= i)
+
+  count[[i]] <- tibble("n_weeks" = i, 
+                       "eob" = nrow(df_eob), 
+                       "temp" = nrow(df_temp), 
+                       "sprt" = nrow(df_sprt))
+}
+
+count <- bind_rows(count)
+
+count %>% 
   ggplot() +
-  geom_col(aes(x = site, y = n_weeks, fill = seq), position = "identity") +
-  labs(x = "Site", 
-       y = "Number of weeks", 
+  geom_bar(data = . %>% filter(n_weeks == 24 | n_weeks == 36), 
+           aes(x = n_weeks, y = eob, fill = "Buildings satisfying all criteria"), 
+           stat = "identity", 
+           alpha = 0.8, 
+           width = 1) +
+  geom_line(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"),
+            alpha = 0.2) +
+  geom_point(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"), 
+             size = 1.5) +
+  geom_line(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"),
+            alpha = 0.2) +
+  geom_point(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"), 
+             size = 1.5, 
+             shape = 17) +
+  geom_vline(xintercept = seq(12, 36, by = 12), lty = "dashed", color = "grey80") +
+  annotate(geom = "text", 
+           x = seq(3, 33, by = 12), 
+           y = 700, 
+           label = paste0("12-week block"), 
+           alpha = 0.5) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = ls_colors) +
+  scale_x_continuous(expand = c(0, 0), 
+                     limits = c(0, 38),
+                     breaks = c(12, 24, 36), 
+                     labels = c("12\nweeks", "24\nweeks", "36\nweeks")) +
+  coord_cartesian(ylim = c(0, 720)) +
+  labs(x = NULL, 
+       y = "Number of buildings", 
        fill = NULL, 
-       title = "Average finishing time for all buildings using randomized M&V") +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = seq(0, 36, by = 12)) +
+       color = NULL, 
+       title = "Overall timeline for buildings using randomized M&V") +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
 
-```{r norm, fig.cap="Comparison of normalized annual fractional savings on site TMY between conventional and randomized method", fig.width=8, fig.height=4}
+```{r norm, fig.cap="Comparison of annual fractional savings normalized on TMY weather conditions between conventional and randomized M&V method", fig.width=8, fig.height=4}
 
@@
   filter(method == "true") %>% 
-  mutate(method = "true savings") %>% 
+  mutate(method = "True savings") %>% 
   ggplot(aes(x = method, y = savings, fill = method)) +
@@
                      breaks = breaks_pretty(n = 3), 
-                     limits = c(0, 30), 
+                     limits = c(0, 28), 
                      labels = number_format(suffix = "%")) +
@@
                      breaks = breaks_pretty(n = 3), 
-                     limits = c(0, 30)) +
+                     limits = c(0, 28)) +
+  scale_fill_manual(values = ls_colors) +
   labs(x = NULL, 
@@
             legend = "none") +
-    plot_annotation(title = "Normalized annual fractional savings estimation comparison", 
-                    subtitle = "between convention and randomized M&V (with early stop)")
+    plot_annotation(title = "Normalized annual fractional savings estimation comparison")
 ```
@@
 
-```{r tidyfrac, fig.cap="Absolute deviation of M&V estimated fractional savings from true target savings using tidy subset", fig.width=8, fig.height=4}
+```{r tidyfrac, fig.cap="Absolute deviation of fractional savings from true target savings shown for the 'tidy' subset", fig.width=8, fig.height=4}
 
@@
   ggplot() +
-  geom_bar(aes(x = seq(1, nrow(all_names_tidy), by = 1), y = abs_diff), 
+  geom_bar(aes(x = seq(1, nrow(all_names_tidy), by = 1), y = abs_diff, fill = "Randomized M&V"), 
            stat = "identity", alpha = 0.5) +
@@
              linetype = "dashed", 
-             color = "red") +
+             color = "black") +
   annotate(y = median_rand + 0.5, 
@@
              linetype = "dashed", 
-             color = "red") +
+             color = "black") +
   annotate(y = upper_rand + 0.5, 
@@
                      limits = c(0, max(plot_data$plot_max) + 0.5)) +
-  labs(x = "Number of buildings", 
-       y = "Absolute difference in measured savings", 
-       subtitle = "Randomized M&V at early stop") +
+  scale_fill_manual(values = ls_colors) +
+  labs(fill = NULL, 
+       x = "Number of buildings", 
+       y = "Absolute difference of fractional savings", 
+       subtitle = "Randomized M&V") +
   coord_cartesian(ylim = c(0, 10)) +
@@
   ggplot() +
-  geom_bar(aes(x = seq(1, nrow(all_names_tidy), by = 1), y = abs_diff), 
+  geom_bar(aes(x = seq(1, nrow(all_names_tidy), by = 1), y = abs_diff, fill = "Conventional M&V"), 
            stat = "identity", alpha = 0.5) +
@@
              linetype = "dashed", 
-             color = "red") +
+             color = "black") +
   annotate(y = median_conv + 0.5,
@@
              linetype = "dashed", 
-             color = "red") +
+             color = "black") +
   annotate(y = upper_conv + 0.5, 
@@
                      limits = c(0, max(plot_data$plot_max) + 0.1)) +
-  labs(x = "Number of buildings", 
+  scale_fill_manual(values = ls_colors) +
+  labs(fill = NULL, 
+       x = "Number of buildings", 
        y = NULL, 
@@
           align = "hv",
-          legend="bottom") +
-  plot_annotation(title = str_glue("M&V method accuracy comparison"), 
-                  subtitle = "fractional savings in tidy subset")
+          legend="none") +
+  plot_annotation(title = "Savings estimation accuracy comparison", 
+                  subtitle = "using 'tidy' subset")
 
@@
 
-```{r messyfrac, fig.cap="Absolute deviation of M&V estimated fractional savings from true target savings using messy subset", fig.width=8, fig.height=4}
+```{r messyfrac, fig.cap="Absolute deviation of fractional savings from true target savings shown for the 'messy' subset", fig.width=8, fig.height=4}
 
@@
   ggplot() +
-  geom_bar(aes(x = seq(1, nrow(all_names_messy), by = 1), y = abs_diff), 
+  geom_bar(aes(x = seq(1, nrow(all_names_messy), by = 1), y = abs_diff, fill = "Randomized M&V"), 
            stat = "identity", alpha = 0.5) +
@@
              linetype = "dashed", 
-             color = "red") +
+             color = "black") +
   annotate(y = median_rand + 1, 
@@
              linetype = "dashed", 
-             color = "red") +
+             color = "black") +
   annotate(y = upper_rand + 1, 
@@
                      limits = c(0, max(plot_data$plot_max) + 0.5)) +
+  scale_fill_manual(values = ls_colors) +
   labs(x = "Number of buildings", 
-       y = "Absolute difference in measured savings", 
-       subtitle = "Randomized M&V at early stop") +
+       y = "Absolute difference of fractional savings", 
+       fill = NULL, 
+       subtitle = "Randomized M&V") +
   coord_cartesian(ylim = c(0, 25)) +
@@
   ggplot() +
-  geom_bar(aes(x = seq(1, nrow(all_names_messy), by = 1), y = abs_diff), 
+  geom_bar(aes(x = seq(1, nrow(all_names_messy), by = 1), y = abs_diff, fill = "Conventional M&V"), 
            stat = "identity", alpha = 0.5) +
@@
              linetype = "dashed", 
-             color = "red") +
+             color = "black") +
   annotate(y = median_conv + 1,
@@
              linetype = "dashed", 
-             color = "red") +
+             color = "black") +
   annotate(y = upper_conv + 1, 
@@
                      limits = c(0, max(plot_data$plot_max) + 0.1)) +
+  scale_fill_manual(values = ls_colors) +
   labs(x = "Number of buildings", 
        y = NULL, 
+       fill = NULL, 
        subtitle = "Conventional M&V") +
@@
           align = "hv",
-          legend="bottom") +
-  plot_annotation(title = str_glue("M&V method accuracy comparison"), 
-                  subtitle = "fractional savings in messy subset")
+          legend="none") +
+  plot_annotation(title = "Savings estimation accuracy comparison", 
+                  subtitle = "using 'messy' subset")
 
@@
 
-In this section, we demonstrate the influence of non-routine events
-through two scenarios. In the first scenario, we quantified how much
-occupancy would influence the electricity usage and added to the tidy
-measurement set as they originally indicates no such change. In the
-second scenario, we simply run the two M&V methods on the messy
-measurement sets before adding the chilled water supply temperature
-reset. Since no intervention is added, the more reliable method should
-detect a 'saving' closer to 0.
+We included two examples in this section to illustrate the impact of
+non-routine events on M&V savings estimation. In the first scenario, we
+manually subtracted a portion of the electricity usage for a specific
+period in the tidy subset to simulate occupancy-confounded measurements,
+which were not originally present in the dataset. In the second
+scenario, focusing on the messy subset, which already contains
+underlying changes, we removed the intervention effect and applied both
+M&V methods to assess which method could more accurately detect the
+resulting zero 'savings.'
 
@@
 
-Although occupancy can be approximated in several ways in commercial
-buildings, such as through counting check-in, or WIFI connections, or
-monitoring indoor CO2 concentration as a proxy. However, it is not a
-cost-efficient measures to add to the routine operation for most of the
-buildings and there might be privacy concerns associated. The scenario
-proposed hypothesize that during the baseline measurement period of the
-target building, a floor of tenants moved out leaving the space
-unoccupied for four months before new tenants moved in and this led to a
-20% electricity decrease as a result. The M&V protocol only requires the
-M&V analyst to fit a TOWT model so he/she only measures outdoor
-temperature. We showed one example in Figure \@ref(fig:occchange) where
+Although occupancy can be estimated in various ways, such as counting
+check-ins, monitoring WiFi connections, or using indoor CO2
+concentrations as a proxy, these methods are often not cost-effective
+for routine operations in most buildings and may raise privacy concerns.
+In the proposed scenario, we hypothesize that during the baseline
+measurement period, one floor of tenants vacated the building, leaving
+the space unoccupied for four months. This led to a fixed reduction in
+electricity usage, approximated as 20% of the yearly average. We showed
+the impact on TOWT model fitting in Figure \@ref(fig:occchange) where
 from May 1st 2016 to August 31st 2016, the target building electricity
-decreased 20% due to reduced occupancy. Thus subplot (a) shows measured
-baseline with such change, which are then used for TOWT model fitting.
-Starting from 2017, the occupancy returns normal and only intervention
-strategy was measured and the projected baseline is the prediction
-results from the fitted model. To demonstrate, we also plotted the
-original measurements before adding the intervention effect as 'adjusted
-baseline'. This assumes that we can accurately adjust the baseline in
-the post-retrofit period on back-to-normal occupancy. Subplot (b) shows
-that the fitted regression underestimate the true baseline condition
-leading to underestimated savings.
+dropped. Subplot (a) displays the measured baseline data, including the
+occupancy-related change, used as input for model training. Subplot (b)
+shows the model's prediction results for the post-retrofit period. For
+reference, we also plotted the correctly adjusted baseline (labeled as
+'adjusted baseline') in the post-retrofit period. Since occupancy was
+not considered an independent variable in the model, the baseline change
+was incorrectly attributed to temperature and time variations shown as
+the 'projected baseline', leading to underestimated savings estimation.
 
@@
 change_end <- occ_params$change_end
-change <- occ_params$change
+change <- mean(df_base_conv$eload) * occ_params$change / 100
 
@@
   filter(datetime < as.Date("2017-01-01")) %>%
-  mutate(eload = ifelse(month(datetime) >= month(change_start) & month(datetime) <= month(change_end), eload * (1 - change / 100), eload))
+  mutate(eload = ifelse(month(datetime) >= month(change_start) & month(datetime) <= month(change_end), eload - change, eload))
 
@@
 
-```{r occchange, fig.cap="Occupancy change impact on TOWT model fitting and baseline projection", fig.width=8, fig.height=4}
+```{r occchange, fig.cap="Illustration of occupancy change impact on TOWT model fitting and baseline projection in the post-retrofit period", fig.width=9, fig.height=4}
 
@@
                slice_sample(n = 500),
-             aes(x = datetime, y = towt, color = "Projected baseline\n(no change)"),
+             aes(x = datetime, y = towt, color = "Projected baseline"),
              size = 0.5,
@@
   geom_smooth(data = base_pos_proj,
-              aes(x = datetime, y = towt, color = "Projected baseline\n(no change)"), 
+              aes(x = datetime, y = towt, color = "Projected baseline"), 
               formula = y ~ x, method = "loess") +
@@
 
-Figure \@ref(fig:occsavings) shows the normalized fractional savings
-estimation accuracy comparison between the two M&V methods. To compare
-which method is more robust to the applied occupancy change, a
-difference-in-difference value is plotted in the figure. Each bar in the
-plot represent the difference in the deviation of savings estimated by
-the convention method and the randomized method:
+Figure \@ref(fig:occsavings) shows the overall estimation accuracy
+between the two M&V methods where the accuracy is indicated as a
+difference-in-difference value:
 
 $$
-|S_{conv} - S_{true}| - |S_{rand} - S_{true}|
+|FS_{conv} - FS_{true}| - |FS_{rand} - FS_{true}|
 $$
 
-In other words, a positive value indicate the randomized method provides
-an estimation more aligned to the true target saving. Thus the plot
-shows the randomized method shows uniformly superior robustness for all
-sites. The red-dotted line shows the absolute deviation in savings
-estimation by the randomized method as a reference for scaling. The
-deviation is consistent with the distribution shown in Figure
-\@ref(fig:tidysite) and thus further reinforce its robustness to
-non-routine events.
-
-```{r occsavings, fig.cap="Savings estimation accuracy comparison between two M&V methods with added occupancy change (with the site average difference-in-difference displayed at the bottom of each site).", fig.width=8, fig.height=4}
-s <- 'S4'
+Therefore, a positive value indicates that the randomized method
+provides an estimation more closely aligned with the true target
+savings. The bar plot shows the difference-in-difference of the
+calculated fractional savings ($FS$) for each target building and the
+dotted line shows the absolute deviation of randomized method (i.e.
+$|FS_{rand} - FS_{true}|$ ) as a reference. In addition, we showed the
+averaged difference-in-difference metric for each site in the figure,
+quantifying how much more accurate the randomized method is compared to
+the conventional method. As a result, across all sites, the randomized
+method consistently outperforms the conventional method and the values
+of dotted line show similar results compared to Figure
+\@ref(fig:tidyfrac) further implying that the non-routine event has
+negligible influence on the randomized method.
+
+```{r occsavings, fig.cap="Savings estimation accuracy comparison between two M&V methods with added occupancy change (with the average difference-in-difference average on each location displayed at the bottom).", fig.width=9, fig.height=4}
+s <- 'S2'
 dev_conv <- df_NRE_occ %>%
+  left_join(df_loc, by = "site") %>% 
   filter(scenario == s) %>%
   filter(method != "rand") %>%
-  group_by(name, site, scenario) %>%
+  group_by(name, location, scenario) %>%
   summarise(conv = abs(diff(savings))) %>%
@@
 dev_rand <- df_NRE_occ %>%
+  left_join(df_loc, by = "site") %>% 
   filter(scenario == s) %>%
   filter(method != "conv") %>%
-  group_by(name, site, scenario) %>%
+  group_by(name, location, scenario) %>%
   summarise(rand = abs(diff(savings))) %>%
@@
 dev_FS <- dev_rand %>%
-  left_join(dev_conv, by = c("name", "site", "scenario")) %>% 
+  left_join(dev_conv, by = c("name", "location", "scenario")) %>% 
   mutate(diff_in_diff = conv - rand)
@@
 mean_diff <- dev_FS %>%
-  group_by(site) %>%
+  group_by(location) %>%
   summarise(mean = round(mean(diff_in_diff), digits = 1),
@@
 dev_FS %>%
-  ggplot(aes(group = site)) +
-  geom_col(aes(x = name, y = diff_in_diff), position = "identity", alpha = 0.3) +
-  facet_wrap(~site, nrow = 1, scales = "free_x") +
+  ggplot(aes(group = location)) +
+  geom_line(aes(x = name, y = rand, color = "Randomized M&V")) +
+  geom_point(aes(x = name, y = rand, color = "Randomized M&V"), alpha = 0.8) +
+  geom_col(aes(x = name, y = diff_in_diff), position = "identity", alpha = 0.2, fill = "#78c679") +
+  facet_wrap(~location, nrow = 1, scales = "free_x") +
   scale_y_continuous(expand = c(0.1, 0),
@@
   geom_text(data = mean_diff,
-            aes(x = pos, y = -.5, group = site, label = paste0(mean, "%"))) +
-  geom_line(aes(x = name, y = rand, color = "Absolute deviation of randomized method"), alpha = 0.5) +
-  geom_point(aes(x = name, y = rand, color = "Absolute deviation of randomized method"), alpha = 0.5) +
+            aes(x = pos, y = -.5, group = location, label = paste0(mean, "%"))) +
+  scale_color_manual(values = ls_colors) +
   labs(x = NULL,
        fill = NULL,
-       y = NULL,
+       y = "Difference-in-difference of fractional savings",
        color = NULL, 
-       title = "Difference-in-difference of fractional savings") +
+       title = "Savings estimation accuracy comparison for non-routine events", 
+       subtitle = "using 'tidy' subset") +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
 
-The outbreak of pandemic in 2020 followed by working-from-home policy is
-a practical example of building energy consumption influenced by
-non-routine events. During unoccupied months, building energy
-consumption dropped compared to the same time period in the previous
-years. If an energy-efficient intervention was deployed onsite starting
-at 2020, given by the whole-building electricity measurements, the large
-difference in measured electricity usage in 2020 and projected baseline
-usage in 2019 can hardly indicate the retrofit intervention effect.
-
-In other words, if there is no intervention applied or the intervention
-effect is known to be null (e.g. constantly being overriden by the
-baseline), an ideal M&V method should detect no savings. To test this,
-we kept the randomized schedule as sampled earlier but removed added
-chilled water plant intervention effect. Therefore, all sampled
-intervention days are considered additional sampled baseline days.
-Earlier we mentioned the 'messy' dataset contains some change in the
-two-year electricity measurement, which is purely due to various
-non-routine events such as occupancy change or even other measures
-implemented by the building manager but irrelevant to the chilled water
-set point reset intervention. Thus, when the proposed chilled water
-reset intervention was removed, a more reliable M&V method should
-overcome the influence of those confounding factors and inform no
-savings.
-
-```{r nosavings, fig.cap="Distribution of savings estimation results by the convention M&V method and randomized method.", fig.width=6, fig.height=3}
+As previously mentioned, the 'savings' detected in commercial buildings
+during the pandemic were largely confounded by null occupancy. In other
+words, if no intervention is applied or the intervention effect is known
+to be null (e.g., constantly overridden by the baseline), an ideal M&V
+method should detect no savings. Given that the messy dataset allows for
+up to a 25% annual usage change in the original dataset, a more reliable
+M&V method should inform the analyst that no savings occurred prior to
+the addition of the intervention. To test this, we removed the chilled
+water plant reset effect from the sampled days and assessed whether the
+M&V methods could correctly detect zero savings.
+
+```{r nosavings, fig.cap="Comparison of no intervention effect detection results (with 0% as true annual fractional savings) between conventional and randomized M&V method.", fig.width=6, fig.height=3}
 
@@
   geom_boxplot(outlier.shape = NA, width = 0.5) +
-  geom_hline(yintercept = 0, linetype = "dashed", color = "red", linewidth = 1.2) +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 1.2) +
   annotate(geom = "text", 
@@
            y = 0.5, 
-           color = 'red',
+           color = "black",
            label = "True savings = 0") +
@@
                      labels = number_format(suffix = "%")) +
+  scale_fill_manual(values = ls_colors) +
   labs(x = NULL, 
@@
        fill = NULL, 
-       title = "Savings estimation comparison", 
-       subtitle = "between convention and randomized M&V (with early stop)") +
+       title = "Normalized confounded annual fractional savings comparison") +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
 method on the other hand is able to provide an estimation much closer to
-0% with reduced uncertainty. This is mostly because when there is a
-change or disturbance applied to the target building such as lighting
-retrofit or thermal leakage in the envelope, randomly sample the two
-strategies at 50%/50% ensures the resulting effect either as a decrease
-in lighting electricity or an increase in reheat electricity is balanced
-between the two implemented strategies.
+0% with reduced uncertainty. This improvement is largely due to the
+randomized sampling approach, where the two strategies (baseline and
+intervention) are sampled at a 50%/50% ratio. This ensures that any
+changes or disturbances additional to the intervention implementation,
+such as concurrent lighting retrofits, are evenly distributed across
+both strategies, balancing their effects.
 
@@
 ## TOWT modeling accuracy
-
-Although this paper argues that regression model prediction results can
-be largely biased by various non-routine events, we further clarify that
-this is not related to regression model fitting accuracy. Time-of-week
-temperature model is used here for its simplicity and convenience since
-the model only requires outdoor temperature measurements and time of
-week as independent variables. The model assumes a linear composition of
-building energy consumption as temperature-dependent and time-dependent
-load. The time-dependent component accounts for day-to-day variations
-and the temperature dependent component considers a piecewise linear
-relationship across a range of temperature intervals.
-<!--# TODO: add reference -->
 
@@
 whole-building baseline model fitting accuracy should maintain a
-CV(RMSE) lower than 30%. In addition One study focusing on the baseline
-energy data-driven model fitting indicates that TOWT performs as
-accurate as other more advanced machine learning models
-<!--# TODO: reference --> and the calculated CV(RMSE) distribution for a
-large sample of commercial buildings indicates a median of 20%. Figure
-\@ref(fig:towtacc) plots the distribution of the model fitting accuracy
-calculated separately for the two measurement sets. The result shows
-that the TOWT model performance is even better in this study compared to
-the literature and no significant difference between the two subsets.
-This is mostly related to low-quality measurement sets filtering in the
-pre-processing mentioned in Section \@ref(method).
-
-```{r towtacc, fig.cap="TOWT model fitting accuracy distribution on filtered measurement set", fig.width=8, fig.height=4}
+CV(RMSE) lower than 30% [@ashrae_ashrae_2014]. In addition one study
+focusing on the baseline energy data-driven model fitting indicates that
+TOWT performs as accurate as other more advanced machine learning models
+[@granderson_accuracy_2016] and the calculated CV(RMSE) distribution for
+a large sample of commercial buildings indicates a median of 20%. Figure
+\@ref(fig:towtacc)shows the distribution of model fitting accuracy,
+calculated separately for the two measurement sets. The box plot
+provides a statistical summary of the accuracy data points, with error
+bars highlighting any points beyond the range, which should be
+considered outliers. The results show that the TOWT model's performance
+in this study aligns with, and is even slightly better than, the
+performance reported in the literature. The improved results can be
+attributed to relatively strict data filtering criteria for
+pre-processing described in Section \@ref(method).
+
+```{r towtacc, fig.cap="TOWT model fitting accuracy distribution for all buildings included (with each data point representing one building)", fig.width=8, fig.height=4}
 
 df_model_acc_messy %>% 
-  mutate(group = "messy") %>% 
-  rbind(df_model_acc_tidy %>% mutate(group = "tidy")) %>% 
+  mutate(group = "Messy subset") %>% 
+  rbind(df_model_acc_tidy %>% mutate(group = "Tidy subset")) %>% 
   mutate(group = as.factor(group)) %>% 
-  ggplot(aes(x = group, y = cvrmse, fill = group)) +
-  stat_boxplot(geom ='errorbar', width = 0.5) +
-  geom_boxplot(outlier.shape = NA) +
+  ggplot(aes(x = group, y = cvrmse)) +
+  stat_boxplot(geom ='errorbar', width = 0.1) +
+  geom_boxplot(outlier.shape = NA, width = 0.25) +
+  geom_jitter(position = position_jitter(0.1), alpha = 0.1, size = 1.2) +
   geom_text(aes(x = group, y = median, label = paste0(round(median, digits = 0), " %")), 
@@
               ungroup(), 
-            position = position_nudge(y = 1)) +
+            position = position_nudge(x = 0.25)) +
   scale_y_continuous(expand = c(0, 0), 
@@
                      labels = number_format(suffix = " %")) +
-  coord_cartesian(ylim = c(0, 38)) +
+  coord_cartesian(ylim = c(0, 50)) +
   labs(x = NULL, 
        y = NULL, 
-       fill = NULL, 
-       title = "TOWT model accuracy", 
+       color = NULL, 
+       title = "TOWT model fitting accuracy", 
        subtitle = "CV(RMSE)") +
@@
 
-In general, the limitation is that despite the regression model tends to
-capture well on the mean building energy consumption but can
-underestimate and peak or overestimate the lower base load
-<!--# TODO: reference -->, which is not useful for assessing demand
-response. Additionally, a data-driven model can learn the training
-dataset very efficiently with high accuracy, but the well-trained model
-is challenging to generalize prediction outside the range of training
-set.
+The limitation of regression models is that despite they capture well on
+the mean building energy consumption but can underestimate the 15-min
+daily peak load [@granderson2021], which is not useful for assessing
+savings for demand response events.
 
@@
 Another advantage of using randomized M&V is the flexibility of changing
-sampling ratio after the M&V. For example, the building owner can
-continue sampling at 50%/50% between the baseline and the intervention
-to further reduce the uncertainty range associated with the savings. Or
-he/she can switch to 100% intervention to optimize energy savings, but
-whether the existing baseline becomes outdated is unknown. A compromised
-approach is to continue sampling at 80%/20% between the baseline and the
-intervention. Figure \@ref(fig:cont) shows after the analyst reports the
-randomized M&V results, a new schedule sampling at 80%/20% was
-implemented till the end of 2016 and the plot shows the saving
-estimation accuracy calculated as difference-in-difference when sampled
-baseline reduced.
-
-```{r cont, fig.cap="Savings estimation accuracy for continue sampling at 80%/20% for 36 weeks after all stopping criteria are satisfied", fig.width=8, fig.height=4}
+sampling ratio after the target savings detected. For instance, the
+building owner can continue sampling at a 50%/50% ratio between the
+baseline and the intervention to further reduce the uncertainty
+associated with the savings. Alternatively, they could switch to 100%
+intervention to maximize savings on the utility bill, though this
+approach risks the baseline becoming outdated. A middle-ground approach
+would be to sample at an 20%/80% ratio between the baseline and the
+intervention. Through this way, the software-as-a-service company can
+accurately model future baseline changes and adjust customers' bill
+accordinly with minimum baseline days sampled. To demonstrate, figure
+\@ref(fig:cont) shows after the analyst reports the randomized M&V
+results indicated in Figure \@ref(fig:timeline), a new schedule sampling
+at 20%/80% was implemented till the end of the year. Similarly, we
+calculated the accuracy metric as the absolute deviation from the annual
+true savings.
+
+```{r cont, fig.cap="Randomized M&V absolute deviation of fractional savings from true target savings for a whole year with more intervention days sampled after all stopping criteria satisfied", fig.width=8, fig.height=4}
 
@@
                      limits = c(0, max(plot_data$plot_max) + 0.5)) +
+  scale_fill_manual(values = ls_colors) +
   labs(x = "Number of buildings", 
-       y = "Absolute difference in measured savings", 
+       y = "Absolute difference of fractional savings", 
+       fill = NULL, 
        subtitle = "Tidy subset") +
@@
                      limits = c(0, max(plot_data$plot_max) + 0.1)) +
+  scale_fill_manual(values = ls_colors) +
   labs(x = "Number of buildings", 
        y = NULL, 
+       fill = NULL, 
        subtitle = "Messy subset") +
@@
           align = "hv",
-          legend="bottom") +
-  plot_annotation(title = str_glue("Continue sampling accuracy comparison"), 
-                  subtitle = "with 80%/20% intervention/baseline")
+          legend="none") +
+  plot_annotation(title = "Randomized M&V savings estimation accuracy comparison with continue sampling included", 
+                  subtitle = "20%/80% baseline/intervention")
 ```
 
-The deviation remains consistent in the tidy subset but increases
-slightly in the messy subset, suggesting that changes in energy usage in
-the target building create a trade-off between measured savings and
-estimation accuracy when down-sampling the baseline. In this scenario,
-opting for an early stop at 24 or 36 weeks with 50% intervention and
-re-sampling 80% till the year end (\~ another 20 weeks) still allows the
-building owner to capture 65% of the full-range savings. On the other
-hand, if the building owner follows the conventional method, there is no
-savings due to required baseline measurement.
+The plot shows the deviation remains consistent in the tidy subset
+comapred to Figure \@ref(fig:tidyfrac) but increased slightly in the
+messy subset compare to to Figure \@ref(fig:messyfrac). With unbalanced
+sampling ratio, there is a trade-off for estimation accuracy. But
+compared with conventional method, this is still a preferred approach.
+Most importantly, this approach allows customers to realize a proportion
+of expected savings while documenting baseline measurement. For example,
+opting for an early stop at 24 or 36 weeks with 50% intervention,
+followed by re-sampling at 80% until the year's end (\~another 20
+weeks), would enable customers to capture about 65% of the full-range
+savings.
 
@@
 
-We identify two key limitations of this study: 1) the application of
-control intervention, and 2) the design of the randomized switchback
-experiment. While the primary focus is on accurately detecting
-intervention effects rather than validating their impact, the simulated
-intervention remains somewhat generic due to the diverse building types
-and climate zones in the BDG2 dataset. For simplicity, we applied the
-parameters listed in Table 1 uniformly across all buildings. However,
-the electricity savings from raising the water temperature by 1°C may be
-less than 8% in some cases, due to increased chilled water pump speeds.
-Similarly, the randomized schedule design is also simplified. We assumed
-a daily sampling interval would be sufficient for most commercial
-buildings, but exceptions exist, particularly for buildings with
-significant thermal lag in heating and cooling systems, such as those
-with heavy concrete construction, hot water tanks, or Thermally Active
-Building Systems (TABS). In such cases, the effect of a previously
-sampled strategy can carry over and influence subsequent measurements
-due to thermal storage, known as the carryover effect. For example, if
-the intervention pre-charges chilled water in thermal mass one day
-before switching back to the baseline, the analyst is likely to observe
-reduced energy usage in the following days. This study does not account
-for the carryover effect, but in practice, we recommend using a 3-day
-sampling interval and dropping non-consecutive days to 'wash out' any
-lingering effects from previous strategies.
-
-Our major goal for the future study is to further extend the application
-of the proposed randomized M&V. For example, a customer maybe interested
-in whether a Model Predictive Control (MPC) can reduce energy bill under
-dynamic pricing structure and what the estimated Return of Investment
-(ROI) for the retrofit is. Or whether a load shift control can save
-operational cost by shifting load from grid peak to off peak period.
-Furthermore, the novel method is generalizable beyond energy saving M&V
-cases. The statistical analytic framework is applicable to a variety of
-retrofit metrics such as indoor air quality, operational carbon
-emissions and thermal comfort.
+We identify two key limitations in this study:
+
+1.  Application of control intervention: Although the primary focus is
+    on accurately detecting intervention effects rather than validating
+    their broader impact, the simulated intervention remains somewhat
+    generic considering the diverse building types and climate zones in
+    the BDG2 dataset. For simplicity, we applied the parameters listed
+    in Table 1 uniformly across all buildings. Yet, in some cases,
+    raising the water temperature by 1°C might yield more or less than
+    8% in electricity savings due to target buildings' demand
+    flexibility.
+2.  Design of the randomized switchback experiment: In this study, we
+    assumed a daily sampling interval would suffice for most commercial
+    buildings, but exceptions exist. Buildings with significant thermal
+    lag, such as those with heavy concrete construction, hot water
+    tanks, or Thermally Active Building Systems (TABS), may experience
+    carryover effects. In these cases, the impact of one day's control
+    strategy can influence subsequent measurements due to thermal
+    storage. For example, if chilled water is pre-charged in the thermal
+    mass at the end of a sampled intervention day and the control
+    swicthes to baseline at 12 AM the following day, the analyst would
+    observe lower energy consumption in that sampled baseline day. While
+    this study does not account for carryover effects, we recommend
+    using a 3-day sampling interval and excluding non-consecutive days
+    in practice to 'wash out' residual effects from previous strategies.
+
+For future work, our main objective is to extend the application of the
+proposed randomized M&V approach. For instance, a customer may want to
+know whether Model Predictive Control (MPC) can reduce energy bills
+under a dynamic pricing structure or estimate the Return on Investment
+(ROI) for a retrofit. Or whether a load shift control can save
+operational cost by shifting load out of a given time window such grid
+peak price/emissions period. We also believe that this approach could be
+adapted to analyze the effects of other retrofit interventions at a
+population level. Additionally, we aim to demonstrate that the
+randomized method and sequential analysis framework can be applied to
+various performance metrics, including indoor air quality, operational
+carbon emissions, and thermal comfort.
 
@@
 
-This research demonstrated the usage and assessed the performance of a
-novel whole-building measurement and verification method using a large
-open-source commericial dataset. The method leverages the randomized
-experimental design concept from other scientific research fields and
-statistical sequential inference techniques to inform whether a target
-savings are detected. We made an example of chilled water setpoint reset
-based on outdoor weather condition as a virtual control retrofit use
-case and applied to over 500 filtered commercial buildings. By comparing
-the savings estimation by the conventional method (outlined in ASHRAE
-Guideline 14) and the novel randomized method, we showed that the
-randomized method is able to provide a more rapid and robust saving
-estimation.
-
-We showed that throughout 7 climate zones we assessed, the randomized
-M&V can provide a saving estimation by 36 weeks (with the majority
-finishes by 24 weeks) once all stopping criteria satisfied. Meanwhile,
-the conventional method requires a full range measurement of both
-baseline and intervention performance under normal operation conditions,
-which normally translate to 6 - 9 months each. We further showed the
-estimated savings at early stop has acceptable accuracy when compared
-with the true calculated savings.
-
-We also assessed the impact of non-routine events on the proposed M&V
-method both by 1) adding a known change, such as occupancy-induced
-energy reduction, and 2) detecting removed intervention when no reset is
-applied to buildings with a marginal difference in energy consumption.
-We showed in both scenarios that baseline model fitting can be biased
-while the randomization can efficiently block those confounding effect
-to ensure the robustness of the saving estimation.
+This research demonstrated the application of a novel whole-building
+measurement and verification (M&V) method, comparing its performance to
+the conventional approach outlined in ASHRAE Guideline 14 using a large,
+open-source commercial building dataset. The proposed M&V method
+leverages the randomized experimental design concept from other
+scientific fields, along with statistical sequential inference
+techniques, to determine when target savings are detected. We used a
+virtual control retrofit case---resetting the chilled water setpoint
+based on outdoor weather conditions---and applied it to over 500
+filtered commercial buildings. By comparing the savings estimations of
+the conventional method with the novel randomized method, we found that
+the randomized approach provides faster and more robust savings
+estimations.
+
+Specifically, we showed that throughout 7 climate zones assessed in this
+study, the randomized M&V can provide a saving estimation by 36 weeks
+(with the majority finishes by 24 weeks) once all stopping criteria
+satisfied. In contrast, the conventional method requires a full range of
+baseline and intervention measurements under normal operating
+conditions, typically taking 6-9 months for each phase. Most
+importantly, Furthermore, we verified that with reduced M&V timeline the
+randomized method can estimate savings more accurately than the two-year
+conventional method.
+
+We also evaluated the impact of non-routine events on the proposed M&V
+method by: 1) introducing a known change, such as occupancy-induced
+energy reductions, and 2) detecting when no intervention was applied in
+buildings with marginal energy consumption differences. In both cases,
+we demonstrated that baseline model fitting could be biased, while
+randomization effectively blocked confounding effects, ensuring the
+robustness of savings estimations.
 
 Although the limitation of the method is that it only applies to a
-subset of all M&V use cases (i.e. only if the strategy can be switched
-on and off), we believe the true value lies in the usefulness for most
-control retrofit validation in the field test.
+subset of all M&V use cases (i.e., strategies that can be switched on
+and off), we believe its true value lies in the usefulness and
+conviencience for most control retrofit validation in the field test.
 

### 2024-11-03 code/manuscript.Rmd
Added code for graphical abstract

@@
 |                        |                                                                  |       |
-|-------------------|----------------------------------|-------------------|
+|------------------------|------------------------------------------------------------------|-------|
 | Parameter              | Description                                                      | Value |

### 2024-11-03 code/graphical.R

@@ -1,0 +1,326 @@
+#### LIBRARIES ####
+require(pacman)
+
+# load packages using pacman
+pacman::p_load(tidyverse, lubridate, here, stats, zoo, scales, ggpubr, patchwork, RColorBrewer)
+
+# turn off scientific notation
+options(scipen = 999, digits = 15)
+
+# set directory
+here::i_am("manuscript.rmd")
+
+# set default theme for ggplot
+theme_set(theme_minimal())
+
+# define base ggplot theme
+theme_update(plot.title = element_text(size = 14, colour = "grey20", face = "bold", hjust = 0.5),
+             plot.subtitle = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
+             plot.caption = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5),
+             plot.background = element_rect(fill = "white", colour = NA),
+             panel.grid.minor = element_blank(),
+             panel.grid.major = element_blank(),
+             axis.text = element_text(size = 10),
+             strip.text = element_text(size = 10, color = "grey20", face = "bold"),
+             strip.background = element_blank())
+
+# colors
+ls_colors <- c("Baseline" = "#c6dbef",
+               "Measured baseline" = "#c6dbef",
+               "Adjusted baseline" = "#c6dbef",
+               "Projected baseline" = "#2171b5",
+               "Intervention" = "#fdbb84",
+               "Measured interv" = "#fdbb84",
+               "Conventional M&V" = "grey70", 
+               "Randomized M&V" = "#66c2a4", 
+               "Buildings satisfying all criteria" = "#66c2a4", 
+               "Buildings satisfying 80% TMY range" = "black", 
+               "Buildings satisfying SPRT" = "black")
+
+# parameters
+ctr_params <- list(peak_hours = 10:16,
+                   chwl_perc = 0.25,
+                   step_perc = 0.08,
+                   conv_swt = 6,
+                   weather_knots = c(15, 25),
+                   swt_knots = c(12, 7),
[…]

### 2025-02-01 code/manuscript.Rmd
changed interval dataset error and modified graph, manuscript

@@
 ---
-title: "Reliability Demonstration of A Novel Randomized Measurement and Verification Method for Switchable Control Retrofit Using Large-scale Public Dataset"
+title: "Reliability Demonstration of A Novel Randomized Measurement and Verification Method for Switchable Control Retrofit Using Large Open Source Dataset"
 author:
-  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Durate ^1^
+  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
 abstract: Conventional measurement and verification (M&V) methods relying on pre- and post-  retrofit comparison for estimating whole-building energy savings are often time-consuming and unreliable, especially when non-routine events such as reduced occupancy occur during the M&V process.
@@
 # load packages using pacman
-pacman::p_load(tidyverse, lubridate, here, stats, zoo, scales, ggpubr, patchwork, RColorBrewer)
+pacman::p_load(tidyverse, lubridate, here, stats, zoo, scales, lvplot, ggpubr, gridExtra, patchwork, RColorBrewer)
 
@@
                "Measured interv" = "#fdbb84",
-               "Conventional M&V" = "grey70", 
-               "Randomized M&V" = "#66c2a4", 
-               "Buildings satisfying all criteria" = "#66c2a4", 
+               "Conventional" = "grey70", 
+               "Randomized" = "#99d8c9", 
+               "Randomized\n(24 months)" = "#66c2a4",
+               "Randomized\n(50/50)" = "#66c2a4",
+               "Daily\nsampling" = "#66c2a4",
+               "2-day\nsampling" = "#66c2a4",
+               "3-day\nsampling" = "#66c2a4",
+               "7-day\nsampling" = "#66c2a4",
+               "Dropped" = "#41ae76",
+               "Kept" = "#006d2c",
+               "Randomized\n(20/80)" = "#ccece6",
+               "Buildings finishing randomized M&V" =  "#99d8c9",
                "Buildings satisfying 80% TMY range" = "black", 
@@
 
-# Tidy dataset
-readfile_tidy <- str_glue("../readfiles/tidy/")
+# stable dataset
+readfile_stable <- str_glue("../readfiles/stable/")
 fig_path = "../figs/manuscript/"
 
-df_energy_tidy <- read_rds(paste0(readfile_tidy, "df_energy.rds"))
-df_meta_tidy <- read_rds(paste0(readfile_tidy, "df_meta.rds"))
-df_weather_tidy <- read_rds(paste0(readfile_tidy, "df_weather.rds"))
-df_sprt_all_tidy <- read_rds(paste0(readfile_tidy, "df_sprt_all.rds"))
-df_seq_FS_tidy <- read_rds(paste0(readfile_tidy, "df_seq_FS.rds"))
-df_NRE_occ <- read_rds(paste0(readfile_tidy, "df_NRE_occ.rds"))
-df_MD_tidy <- read_rds(paste0(readfile_tidy, "df_MD.rds"))
-df_FS_tidy <- read_rds(paste0(readfile_tidy, "df_FS.rds"))
-df_eui_tidy <- read_rds(paste0(readfile_tidy, "df_eui.rds"))
-df_cont_MD_tidy <- read_rds(paste0(readfile_tidy, "df_cont_MD.rds"))
-df_cont_FS_tidy <- read_rds(paste0(readfile_tidy, "df_cont_FS.rds"))
-df_model_acc_tidy <- read_rds(paste0(readfile_tidy, "df_model_acc.rds"))
-
-all_sites_tidy <- df_energy_tidy %>%
+df_energy_stable <- read_rds(paste0(readfile_stable, "df_energy.rds"))
+df_weather_stable <- read_rds(paste0(readfile_stable, "df_weather.rds"))
+df_sprt_all_stable <- read_rds(paste0(readfile_stable, "df_sprt_all.rds"))
+df_seq_fs_stable <- read_rds(paste0(readfile_stable, "df_seq_fs.rds"))
+df_fs_stable <- read_rds(paste0(readfile_stable, "df_fs.rds"))
+df_cont_stable <- read_rds(paste0(readfile_stable, "df_cont.rds"))
+df_fs_null_stable <- read_rds(paste0(readfile_stable, "df_fs_null.rds"))
+df_seq_fs_null_stable <- read_rds(paste0(readfile_stable, "df_seq_fs_null.rds"))
+df_model_acc_stable <- read_rds(paste0(readfile_stable, "df_model_acc.rds"))
+df_fs_tmy_stable <- read_rds(paste0(readfile_stable, "df_fs_tmy.rds"))
+df_fs_tmy_null_stable <- read_rds(paste0(readfile_stable, "df_fs_tmy_null.rds"))
+df_interval_drop_stable <- read_rds(paste0(readfile_stable, "df_interval_drop.rds"))
+df_interval_keep_stable <- read_rds(paste0(readfile_stable, "df_interval_keep.rds"))
+df_seq_interval_fs_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_fs_drop.rds"))
+df_seq_interval_fs_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_fs_keep.rds"))
+df_seq_interval_nm_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_nm_drop.rds"))
+df_seq_interval_nm_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_nm_keep.rds"))
+df_seq_interval_tl_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_tl_drop.rds"))
+df_seq_interval_tl_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_tl_keep.rds"))
+
+all_sites_stable <- df_energy_stable %>%
   select(site) %>%
@@
 
-all_types_tidy <- df_energy_tidy %>%
+all_types_stable <- df_energy_stable %>%
   select(type) %>%
@@
 
-all_names_tidy <- df_energy_tidy %>%
+all_names_stable <- df_energy_stable %>%
   select(name) %>%
@@
 
-# Messy dataset
-readfile_messy <- str_glue("../readfiles/messy/")
-
-df_energy_messy <- read_rds(paste0(readfile_messy, "df_energy.rds"))
-df_meta_messy <- read_rds(paste0(readfile_messy, "df_meta.rds"))
-df_weather_messy <- read_rds(paste0(readfile_messy, "df_weather.rds"))
-df_sprt_all_messy <- read_rds(paste0(readfile_messy, "df_sprt_all.rds"))
-df_seq_FS_messy <- read_rds(paste0(readfile_messy, "df_seq_FS.rds"))
-df_MD_messy <- read_rds(paste0(readfile_messy, "df_MD.rds"))
-df_FS_messy <- read_rds(paste0(readfile_messy, "df_FS.rds"))
-df_eui_messy <- read_rds(paste0(readfile_messy, "df_eui.rds"))
-df_cont_MD_messy <- read_rds(paste0(readfile_messy, "df_cont_MD.rds"))
-df_cont_FS_messy <- read_rds(paste0(readfile_messy, "df_cont_FS.rds"))
-df_FS_nsprt <- read_rds(paste0(readfile_messy, "df_FS_nsprt.rds"))
-df_MD_nsprt <- read_rds(paste0(readfile_messy, "df_MD_nsprt.rds"))
-df_seq_FS_nsprt <- read_rds(paste0(readfile_messy, "df_seq_FS_nsprt.rds"))
-df_model_acc_messy <- read_rds(paste0(readfile_messy, "df_model_acc.rds"))
-
-all_sites_messy <- df_energy_messy %>%
+# variable dataset
+readfile_variable <- str_glue("../readfiles/variable/")
+
+df_energy_variable <- read_rds(paste0(readfile_variable, "df_energy.rds"))
+df_weather_variable <- read_rds(paste0(readfile_variable, "df_weather.rds"))
+df_sprt_all_variable <- read_rds(paste0(readfile_variable, "df_sprt_all.rds"))
+df_seq_fs_variable <- read_rds(paste0(readfile_variable, "df_seq_fs.rds"))
+df_fs_variable <- read_rds(paste0(readfile_variable, "df_fs.rds"))
+df_cont_variable <- read_rds(paste0(readfile_variable, "df_cont.rds"))
+df_fs_null_variable <- read_rds(paste0(readfile_variable, "df_fs_null.rds"))
+df_seq_fs_null_variable <- read_rds(paste0(readfile_variable, "df_seq_fs_null.rds"))
+df_model_acc_variable <- read_rds(paste0(readfile_variable, "df_model_acc.rds"))
+df_fs_tmy_variable <- read_rds(paste0(readfile_variable, "df_fs_tmy.rds"))
+df_fs_tmy_null_variable <- read_rds(paste0(readfile_variable, "df_fs_tmy_null.rds"))
+df_interval_drop_variable <- read_rds(paste0(readfile_variable, "df_interval_drop.rds"))
+df_interval_keep_variable <- read_rds(paste0(readfile_variable, "df_interval_keep.rds"))
+df_seq_interval_fs_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_fs_drop.rds"))
+df_seq_interval_fs_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_fs_keep.rds"))
+df_seq_interval_nm_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_nm_drop.rds"))
+df_seq_interval_nm_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_nm_keep.rds"))
+df_seq_interval_tl_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_tl_drop.rds"))
+df_seq_interval_tl_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_tl_keep.rds"))
+
+all_sites_variable <- df_energy_variable %>%
   select(site) %>%
@@
 
-all_types_messy <- df_energy_messy %>%
+all_types_variable <- df_energy_variable %>%
   select(type) %>%
@@
 
-all_names_messy <- df_energy_messy %>%
+all_names_variable <- df_energy_variable %>%
   select(name) %>%
@@
 source(paste0(function_path, "prepost_plot.R"))
+
+S_building <- nrow(df_fs_tmy_stable)
+V_building <- nrow(df_fs_tmy_variable)
+A_building <- S_building + V_building
 ```
@@
 
-For the purpose of this study, we only queried whole-building hourly
-electricity usage. All building names and precise locations were erased
-but the climate zone and city name were provided, which are shown in
-Figure \@ref(fig:tidysite) and Figure \@ref(fig:messysite). In addition,
-the measurements were pre-processed by the authors with timestamp
-already converted to local time. We further described the data filtering
-process in Section \@ref(method).
-
 ## Literature review
 
-### Measurement and verification energy-efficient measures
+### Whole building approach
 
@@
 
-### Impact of non-routine events on building energy usage
+### Non-routine events impact
 
@@
 
-### Randomized experimental design in other scientific fields
-
-To address the research gap identified in existing M&V studies,
-particularly the challenges posed by non-routine events, we applied a
-randomized experimental design method---an approach commonly used in
-other scientific disciplines. In clinical trials, for instance,
-randomization is employed to determine the effect of a medical treatment
-(such as a new drug) by randomly assigning participants to either a
-control group (e.g., receiving a placebo) or a treatment group
-[@angus_adaptive_2019, @burger_importance_2021, @lim_randomization_2019,
-@wiley_crossover_2016]. The primary purpose of randomization is to block
-confounding variables such as age, gender, and socioeconomic status when
-evaluating treatment effects. In the context of buildings, a "treatment"
-refers to a control retrofit, while the "placebo" represents the
-existing baseline control. Unlike clinical trials, where participants
-are randomly assigned, our approach randomizes treatment assignment
-longitudinally for each individual building---an approach known as
-n-of-1 trials [@gabler2011]. Although the BDG2 dataset allows for
-randomization at a population level (e.g., across climate zones),
-building owners are generally more interested in understanding the
-effects of control retrofits specific to their own buildings, rather
-than at a generalized population level.
-
 ## Objectives
@@
 
-## Building filtering
-
-### 'Tidy' subset
+## Data filtering
 
 In this study, we extracted the electricity measurements from the BDG2
-dataset. On a first pass, we filtered out buildings with less noise
-based on the following criteria:
+dataset and filter out all qualified buildings based on the following
+criteria:
 
@@
 
-3.  No statistical significant difference (P-value \> 0.05) between the
-    two-year electricity usage: target buildings should have no change
-    in the electricity usage between the two years.
-
-4.  Target buildings should have known site location: buildings with
-    anonymous location are excluded due to unavailable typical
-    meteorological weather.
-
-5.  Warehouse and parking types are excluded: target buildings have less
+3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
+    decrease of building electricity usage in the second year should be
+    less than 25% of that in the first year.
+
+4.  Warehouse and parking types are excluded: target buildings have less
     demand flexibility to implement a chilled water setpoint reset
@@
 
-Therefor the resulting subset contains all buildings with 'tidy'
-measurements. Figure \@ref(fig:tidysite) shows in total, the subset
-contains `r nrow(all_names_tidy)` buildings of `r nrow(all_types_tidy)`
-types.
-
-```{r tidysite, fig.cap = "Site summary of the tidy building subset (counts < 2 are omitted for visualization; left: aggragated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}
+```{r stablesite, fig.cap = "Site summary of the stable building subset (counts < 2 are omitted for visualization; left: aggragated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}
 
 set3 <- colorRampPalette(brewer.pal('Set3',n=12))
-type_colors <- setNames(set3(13), all_types_messy$type)
-
-p1 <- df_energy_tidy %>%
+type_colors <- setNames(set3(13), all_types_variable$type)
+
+stable_set <- df_energy_stable %>%
   group_by(type) %>%
   distinct(name) %>%
-  summarise(n = n()) %>%
+  summarise(stable = n()) %>%
+  ungroup()
+
+variable_set <- df_energy_variable %>%
+  group_by(type) %>%
+  distinct(name) %>%
+  summarise(variable = n()) %>%
+  ungroup()
+
+p1 <- variable_set %>% 
+  left_join(stable_set, by = "type") %>% 
+  mutate(stable = replace_na(stable, 0)) %>% 
+  group_by(type) %>% 
+  summarise(n = stable + variable) %>% 
   ungroup() %>% 
@@
   geom_col() +
-  geom_text(aes(label = ifelse(n > 2, as.character(n), "")), color = "black", position = position_stack(vjust = 0.5)) +
+  geom_text(aes(label = ifelse(n > 10, as.character(n), "")), color = "black", position = position_stack(vjust = 0.5)) +
   scale_y_continuous(expand = c(0, 0),
@@
 
-  p2 <- df_energy_tidy %>%
+stable_set <- df_energy_stable %>% 
+  group_by(type, site) %>%
+  distinct(name) %>%
+  summarise(stable = n()) %>%
+  ungroup()
+
+variable_set <- df_energy_variable %>% 
+  group_by(type, site) %>% 
+  distinct(name) %>%
+  summarise(variable = n()) %>%
+  ungroup()
+
+p2 <- variable_set %>% 
+  left_join(stable_set, by = c("type", "site")) %>% 
+  mutate(stable = replace_na(stable, 0)) %>% 
+  group_by(site, type) %>% 
+  summarise(n = stable + variable) %>% 
+  ungroup() %>% 
   left_join(df_loc, by = "site") %>% 
-    group_by(location, type) %>%
-    distinct(name) %>%
-    summarise(n = n()) %>%
+  group_by(location) %>%
   mutate(proportion = n / sum(n),
@@
   scale_fill_manual(values = type_colors) +
-    geom_text(aes(x = 3.5, y = label_pos, label = ifelse(n > 2, as.character(n), ""))) +
-    geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black") +
+  geom_text(aes(x = 3.5, y = label_pos, label = ifelse(n > 10, as.character(n), "")), check_overlap = T) +
+  geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black", check_overlap = T) +
   theme(legend.direction = "horizontal",
@@
           legend = "bottom") +
-    plot_annotation(title = "Summary of building numbers and types for 'tidy' subset")
+  plot_annotation(title = "Summary of building numbers and types")
 ```
 
-### 'Messy' subset
-
-In reality, the 'tidy' subset is less representative of the measurements
-typically collected from existing building stock, as whole-building
-electricity measurements collected over two years often show more
-variability the the filtering critieron. These variations can stem from
-issues like sensor calibration errors or inherent changes in the
-building, which, as previously discussed, can bias M&V results.
-Therefore, to more realistically assess the robustness of the two M&V
-method, we included an additional 'messy' subset which first exclude the
-'tidy' subset and then re-apply the filtering rule with one amendment:
-
-3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
-    decrease of building electricity usage in the second year should be
-    less than 25% of that in th first year.
-
-Figure \@ref(fig:messysite) shows the summary of the 'messy' dataset,
-which contains `r nrow(all_names_messy)` buildings in
-`r nrow(all_types_messy)` types.
-
-```{r messysite, fig.cap = "Site summary of the messy building subset (counts < 2 are omitted for visualization; left: aggragated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=8}
-
-p1 <- df_energy_messy %>%
-    group_by(type) %>%
-    distinct(name) %>%
-    summarise(n = n()) %>%
-    ungroup() %>%
-    mutate(type = fct_reorder(type, n, .desc = F)) %>%
-    ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
-    geom_col() +
-    geom_text(aes(label = ifelse(n > 10, as.character(n), "")), color = "black", position = position_stack(vjust = 0.5)) +
-    scale_y_continuous(expand = c(0, 0),
-                       breaks = breaks_pretty(n = 4)) +
-    scale_x_discrete(expand = c(0, 0.1)) +
-    scale_fill_manual(values = type_colors) +
-    labs(x = NULL,
-         y = NULL,
-         subtitle = "Combining all locations",
-         fill = NULL) +
-    theme(axis.text.x = element_blank(),
-          panel.grid.major.y = element_line(color = "grey80"),
-          legend.direction = "horizontal",
-          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-  p2 <- df_energy_messy %>%
-    left_join(df_loc, by = "site") %>% 
-    group_by(location, type) %>%
-    distinct(name) %>%
-    summarise(n = n()) %>%
-    mutate(proportion = n / sum(n),
-           total = sum(n),
-           ymax = cumsum(proportion),
-           ymin = c(0, head(ymax, n = -1))) %>%
-    mutate(label_pos = (ymax + ymin) / 2) %>%
-    ungroup() %>%
-    group_by(type) %>%
-    mutate(order = sum(n)) %>%
-    ungroup() %>%
-    mutate(type = fct_reorder(type, order, .desc = F)) %>%
-    ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = type)) +
-    geom_rect() +
-    coord_polar(theta = "y") +
-    xlim(c(2, 4)) +
-    facet_wrap(~ location, nrow = 3) +
-    labs(x = NULL,
-         y = NULL,
-         subtitle = "For each location",
-         fill = NULL) +
-    scale_fill_manual(values = type_colors) +
-    geom_text(aes(x = 3.5, y = label_pos, label = ifelse(n > 1, as.character(n), ""))) +
-    geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black") +
-    theme(legend.direction = "horizontal",
-          axis.text = element_blank(),
-          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-  ggarrange(p1, p2,
-            ncol = 2, nrow = 1,
-            widths = c(0.2, 1),
-            common.legend = TRUE,
-            legend = "bottom") +
-    plot_annotation(title = "Summary of building numbers and types for 'messy' subset")
-```
-
-## Apply control intervention
+For all qualified buildings, we further added one more stringent
+criterion to label all buildings with very stable electricity usage
+between the two years:
+
+1.  No statistical significant difference (P-value \> 0.05) between the
+    two-year electricity usage.
+
+As a results, there are only 66 buildings were labeled in this 'stable'
+subset (out of total of 600 buildings) implying that the assumption made
+by conventional M&V is overly simplified. As mentioned in Section
+\@ref(non-routine-events-impact), such variability is largely associated
+with non-routine events and according to the statistics, is more
+commonly observed in real applications.
+
+## Control intervention implementation
 
@@
 
-## Apply example non-routine event
-
-To illustrate the impact of non-routine events that are not properly
-adjusted for, we developed three scenarios to quantify the energy
-consumption changes associated with a hypothetical increase in
-occupancy. Specifically, we assume that a rise in occupancy in 2016
-(referred to as the "baseline year" when using the conventional method)
-led to a 20% increase in whole-building measured electricity usage. The
-three scenarios describe when this increase occurred: January to April
-(S1), May to August (S2), and September to December (S3). This
-hypothetical change was applied only to buildings in the "tidy" subset,
-as these buildings were initially more likely to be free from
-non-routine events.
-
-## Run M&V methods
-
-### Conventional M&V
-
-As described in Section \@ref(introduction), conventional method for M&V
-is time-consuming and unable to separate non-routine event impact from
-measured savings. If using conventional method to estimate intervention
-energy savings, the result (after 24 months) is calculated by the
-difference between measured intervention and projected baseline in the
-post-retrofit period. In this study, we leveraged a piece-wise linear
-regression considering time-of-week and outdoor temperature (TOWT) as
-independent variables for projection for the counter-factual baseline in
-the post-retrofit period and normalization on typical meteorological
-year [@mathieu_quantifying_2011].
-
-### Randomized M&V
-
-Compared to the conventional method, the randomized M&V approach offers
-a more rapid and reliable estimation of energy savings. To apply this
-method, analysts first define the target savings and stopping criteria.
-Then they design a randomized switchback schedule and perform sequential
-statistical tests, such as the sequential probability ratio test (SPRT),
-to monitor savings as data is collected. Once the target savings are
-detected, the analysts fit a prediction model (e.g. TOWT) to adjust for
-differences in outdoor temperature, following the same adjustment
-process as the conventional method. We provide example switchback
-experimental design and stopping criteria thresholds below:
+## M&V methods comparison
+
+We described more in detail the workflow of both conventional and
+randomized M&V methos in the previous study [@RAFTERY2024111134].
+
+In a similar style, we summarized the experimental design as:
 
@@
 
-Stopping criteria are:
+and stopping criteria:
 
@@
 
-To remain consistency, savings normalization on typical meteorological
-year is also modeled through TOWT.
-
 # Results
 
-## M&V methods comparison
-
-In this section, we compare the performance of two M&V methods. The key
-aspects of the assessment include: 1) time required to reach a saving
-estimation: in most cases, a shorter M&V timeline reduces associated
-cost and interruption for the building owner; and 2) saving estimation
-accuracy: this is particularly important for any software-as-a-service
-company to set a reasonable price with their customers.
-
-### Savings estimation time
-
-ASHRAE Guideline 14 offering minimum requirements for whole-building
-measurement path states that the baseline period is either a full range
-of all independent variables (typically outdoor weather conditions)
-under normal facility operation or a 12-month worth of continuous
-measurements. The same requirements also applies to intervention
-installed in the post-retrofit period. Thus, the conventional M&V method
-is likely to run over 24 months or even longer due to missing data. For
-example if an analyst is asked to determine the energy savings of a
-chilled water plant retrofit but missed most of the cooling season due
-to delay in retrofit deployment, he/she needs to wait until the next
-cooling season to measure the savings. If the randomized M&V method is
-applicable, delays in retrofit deployment pose less risk to the building
-owner since baseline are continuously monitored. Sampling at equal
-probabilities (e.g., 50%/50% between baseline and intervention) helps to
-balance the distribution of independent variables like outdoor weather
-conditions across all sampled control strategies. As long as one
-strategy is not sampled over a large number of consecutive days (e.g., 7
-days or more), it is unlikely that the measurements of independent
-variables will differ significantly between the control strategies.
-
-Figure \@ref(fig:timeline) shows the average estimated timeline for each
-building across all sites and climate conditions if using randomized M&V
-for the applied chilled water supply temperature reset intervention. The
-timeline figures for each individual building are attached in the
-supplementary material. The figure shows all buildings can detect a
-saving statistically from sampled measurements within 12 months, which
-is equivalent to the minimum test criterion. Similar to the findings
-from our pervious study, covering a sufficient range of outdoor weather
-condition is the most stringent requirement. If target buildings are
-located in a climate zone with more preferable outdoor weather
-conditions such as California, covering 80% of the TMY range should only
-take 3 \~ 4 months. In this case, the analyst can conclude the M&V
-simultaneously if the blocking period ends at the same time. However,
-given most buildings requires 6 months to achieve 80% of the TMY range,
-all buildings can determine the savings, including associated
-uncertainties, within 9 months---significantly shorter than the baseline
-measurement period required by the conventional method.
-
-```{r timeline, fig.cap="Randomized M&V timeline of satisfying key stopping criteria for all buildings", fig.width=8, fig.height=4}
-
-df_time <- df_sprt_all_messy %>% 
+```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both at the stopping criteria and over a two-year period).", fig.height=12, fig.width=10}
+
+#### GA ####
+rand_eob_S <- df_seq_fs_stable %>% 
+  filter(seq == "eob") %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - fs), 
+         method = "rand_eob") %>% 
+  select(name, method, diff)
+
+conv_S <- df_fs_stable %>% 
+  filter(method != "rand") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = abs(true - conv), 
+         method = "conv") %>% 
+  select(name, method, diff)
+
+rand_final_S <- df_fs_stable %>% 
+  filter(method != "conv") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = abs(true - rand), 
+         method = "rand_final") %>% 
+  select(name, method, diff)
+
+df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)
+
+# plot for the variable subset
+rand_eob_V <- df_seq_fs_variable %>% 
+  filter(seq == "eob") %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site"))%>% 
+  mutate(diff = abs(savings - fs), 
+         method = "rand_eob") %>% 
+  select(name, method, diff)
+
+conv_V <- df_fs_variable %>% 
+  filter(method != "rand") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = abs(true - conv), 
+         method = "conv") %>% 
+  select(name, method, diff)
+
+rand_final_V <- df_fs_variable %>% 
+  filter(method != "conv") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = abs(true - rand), 
+         method = "rand_final") %>% 
+  select(name, method, diff)
+
+df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)
+
+# plot for combined dataset
+rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
+  select(name, rand_eob_diff = diff)
+
+conv_A <- bind_rows(conv_V, conv_S) %>% 
+  select(name, conv_diff = diff)
+
+rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
+  select(name, rand_final_diff = diff)
+
+df_MW_A <- rand_eob_A %>% 
+  left_join(conv_A, by = "name") %>% 
+  left_join(rand_final_A, by = "name") %>% 
+  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")
+
+p_top <- df_MW_A %>% 
+  mutate(method = as.factor(method), 
+         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
+  ggplot(aes(x = method, y = diff, fill = method)) +
+  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
+  geom_lv(k = 4, outlier.shape = NA) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
+  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
+            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
+  labs(fill = NULL, 
+       color = NULL, 
+       x = NULL, 
+       y = "Absolute error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with measured weather")) +
+  coord_cartesian(ylim = c(0, 23)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "none",
+        axis.text.x = element_blank(),
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+# TMY
+# plot for the stable subset
+rand_eob_S <- df_sprt_all_stable %>% 
+  filter(seq == "eob") %>% 
+  select(-c(seq, n_weeks)) %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - annual), 
+         method = "rand_eob") %>% 
+  select(name, method, diff)
+
+conv_S <- df_fs_tmy_stable %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - conv), 
+         method = "conv") %>% 
+  select(name, method, diff)
+
+rand_final_S <- df_fs_tmy_stable %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - rand), 
+         method = "rand_final") %>% 
+  select(name, method, diff)
+
+df_TW_S <- rbind(rand_eob_S, conv_S, rand_final_S)
+
+# plot for the variable subset
+rand_eob_V <- df_sprt_all_variable %>% 
+  filter(seq == "eob") %>% 
+  select(-c(seq, n_weeks)) %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - annual), 
+         method = "rand_eob") %>% 
+  select(name, method, diff)
+
+conv_V <- df_fs_tmy_variable %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - conv), 
+         method = "conv") %>% 
+  select(name, method, diff)
+
+rand_final_V <- df_fs_tmy_variable %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - rand), 
+         method = "rand_final") %>% 
+  select(name, method, diff)
+
+df_TW_V <- rbind(rand_eob_V, conv_V, rand_final_V)
+
+# plot for combined dataset
+rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
+  select(name, rand_eob_diff = diff)
+
+conv_A <- bind_rows(conv_V, conv_S) %>% 
+  select(name, conv_diff = diff)
+
+rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
+  select(name, rand_final_diff = diff)
+
+df_TW_A <- rand_eob_A %>% 
+  left_join(conv_A, by = "name") %>% 
+  left_join(rand_final_A, by = "name") %>% 
+  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")
+
+p_middle <- df_TW_A %>% 
+  mutate(method = as.factor(method), 
+         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
+  ggplot(aes(x = method, y = diff, fill = method)) +
+  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
+  geom_lv(k = 4, outlier.shape = NA) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
+  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
+            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
+  labs(fill = NULL, 
+       x = NULL, 
+       y = "Absolute error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with TMY weather")) +
+  coord_cartesian(ylim = c(0, 23)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "none",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+# Timeline plot
+df_time <- df_sprt_all_variable %>% 
   filter(seq != "final") %>% 
-  bind_rows(df_sprt_all_tidy %>% filter(seq != "final")) %>% 
+  bind_rows(df_sprt_all_stable %>% filter(seq != "final")) %>% 
   select(name, seq, n_weeks)
@@
 count <- list()
-
-for (i in seq(1, 36, by = 1)){
+n <- 1
+for (i in seq(0, 36, by = 3)){
 
@@
 
-  count[[i]] <- tibble("n_weeks" = i, 
+  count[[n]] <- tibble("n_weeks" = i, 
                        "eob" = nrow(df_eob), 
@@
                        "sprt" = nrow(df_sprt))
+
+  n <- n + 1
 }
@@
 
-count %>% 
+p_bottom <- count %>% 
   ggplot() +
-  geom_bar(data = . %>% filter(n_weeks == 24 | n_weeks == 36), 
-           aes(x = n_weeks, y = eob, fill = "Buildings satisfying all criteria"), 
+  geom_bar(data = . %>% 
+             filter(n_weeks == 24 | n_weeks == 36),
+           aes(x = n_weeks, y = eob, fill = "Buildings finishing randomized M&V"), 
            stat = "identity", 
+           position = "stack",
            alpha = 0.8, 
-           width = 1) +
+           width = 3) +
   geom_line(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"),
@@
              shape = 17) +
-  geom_vline(xintercept = seq(12, 36, by = 12), lty = "dashed", color = "grey80") +
+  geom_segment(aes(x = 36.5, y = max(eob), xend = 95.5, yend = max(eob)),
+               arrow = arrow(length = unit(0.25, "in")),   
+               linewidth = 1.1,  
+               color = "#fb8072") + 
   annotate(geom = "text", 
-           x = seq(3, 33, by = 12), 
-           y = 700, 
-           label = paste0("12-week block"), 
+           x = 66, 
+           y = max(count$eob) - 30, 
+           size = 5,
+           label = "Excess time by conventional M&V") +
+  geom_vline(xintercept = c(12, 24, 36, 48, 96), lty = "dashed", color = "grey80") +
+  annotate(geom = "text", 
+           x = seq(6, 48, by = 12), 
+           y = 200, 
+           label = paste0("12-week\nblock"), 
            alpha = 0.5) +
@@
   scale_x_continuous(expand = c(0, 0), 
-                     limits = c(0, 38),
-                     breaks = c(12, 24, 36), 
-                     labels = c("12\nweeks", "24\nweeks", "36\nweeks")) +
-  coord_cartesian(ylim = c(0, 720)) +
+                     limits = c(0, 100),
+                     breaks = c(12, 24, 36, 48, 96), 
+                     labels = c("12 weeks", "24 weeks", "36 weeks", "1 year", "2 years")) +
+  coord_cartesian(ylim = c(0, 650)) +
   labs(x = NULL, 
@@
        color = NULL, 
-       title = "Overall timeline for buildings using randomized M&V") +
+       subtitle = "timeline comparison") +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
 
+ggarrange(p_top, p_middle, p_bottom, 
+          ncol = 1, nrow = 3,
+          heights = c(1.2, 1.2, 1),
+          labels = c("a)", "b)", "c)"),
+          common.legend = T, 
+          legend = "bottom") +
+  plot_annotation(title = "Overall comparison of conventional and randomized M&V")
 ```
 
-### Savings estimation accuracy
-
-In this study, we define the overall target savings (i.e., the ground
-truth) that an M&V analyst aims to detect as the mean reduction in
-electricity consumption over a two-year period resulting from the
-implementation of the chilled water supply temperature reset control.
-This target can be expressed in terms of normalized savings, fractional
-savings, or simply the measured difference between baseline and
-intervention measurements. Due to different timeline required by the two
-methods, we compare the conventional M&V savings at the end of the
-two-year period with the randomized M&V savings after all criteria are
-satisfied. Since those are the time in reality, an M&V analyst report
-estimated savings.
-
-Figure \@ref(fig:norm) shows the overall comparison of the savings
-estimated for all buildings in both subsets normalized on the typical
-meteorological year weather conditions of each site. In subplot a), the
-narrower range and clustering of the true savings indicates the
-dependence of the intervention effect on the outdoor weather condition,
-which is intended after weather normalization. Sites with mild climate
-all year round such as locations in California shows higher savings
-potential above 10%, while locations with more extreme climate such as
-Washington DC shows only 6% savings annually. In subplot b), the results
-indicate that the conventional M&V method tends to estimate savings with
-greater uncertainty, and its distribution median deviates from the true
-savings. In comparison, the randomized method, which requires much less
-time, shows more accurate and precise estimation results.
-
-```{r norm, fig.cap="Comparison of annual fractional savings normalized on TMY weather conditions between conventional and randomized M&V method", fig.width=8, fig.height=4}
-
-p1 <- df_FS_tidy %>% 
-  bind_rows(df_FS_messy) %>% 
-  filter(method == "true") %>% 
-  mutate(method = "True savings") %>% 
-  ggplot(aes(x = method, y = savings, fill = method)) +
-  geom_errorbar(aes(ymin = min(savings), ymax = max(savings)), width = 0.2) +
-  geom_boxplot(outlier.size = 2) +
-  geom_text(aes(x = 1.35, 
-                y = median(savings) + 1.5, 
-                label = paste0("median: ", round(median(savings), digits = 0), "%")), 
-            check_overlap = T) +
+Figure \@ref(fig:abstract) shows the overall results of M&V methods
+comparison. The estimation error is calculated as the absolute deviation
+from the true savings. Subplot a) shows the savings estimated from
+measured weather condition and subplot b) shows the savings normalized
+on typical meteorological weather file using TOWT model prediction. In
+average, the conventional method could deviate 6% in saving estimation
+while the the deviation from our proposed randomized method is much
+smaller. If the analyst stops right after satisfying all stopping
+criteria, then the deviation is around 2% and it further decreases to 1%
+when the M&V extends over the same period of the convention method. By
+comparing subplot a) and b), we noticed a slight improvement in
+estimation accuracy in model fitting, which is less than 1% on average.
+This implies that baseline projection to post-retrofit period through
+data-driven models is inefficient when the actual baseline changes.
+Subplot c) shows the overall timeline of the two methods. It is clear
+that the all buildings satisfied all stopping criteria in 36 weeks and
+given by the difference in weather conditions, a large portion of the
+buildings can get an accurate M&V results in 24 weeks. Therefore, on
+average a building analyst should expect 1% accuracy improvement over 1
+year extension of the randomized M&V method. However, the red arrow also
+implies the excess time required by the conventional M&V method that
+only leads to more unreliable results.
+
+## Non-routine events impact
+
+```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect", fig.height=9, fig.width=12}
+
+#### MEAN-NULL ####
+rand_eob_S <- df_seq_fs_null_stable %>% 
+  filter(seq == "eob") %>% 
+  mutate(diff = 0 - fs, 
+         method = "rand_eob") %>% 
+  select(name, method, diff)
+
+conv_S <- df_fs_null_stable %>% 
+  filter(method != "rand") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = true - conv, 
+         method = "conv") %>% 
+  select(name, method, diff)
+
+rand_final_S <- df_fs_null_stable %>% 
+  filter(method != "conv") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = true - rand, 
+         method = "rand_final") %>% 
+  select(name, method, diff)
+
+df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)
+
+# plot for the variable subset
+rand_eob_V <- df_seq_fs_null_variable %>% 
+  filter(seq == "eob") %>% 
+  mutate(diff = 0 - fs, 
+         method = "rand_eob") %>% 
+  select(name, method, diff)
+
+conv_V <- df_fs_null_variable %>% 
+  filter(method != "rand") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = true - conv, 
+         method = "conv") %>% 
+  select(name, method, diff)
+
+rand_final_V <- df_fs_null_variable %>% 
+  filter(method != "conv") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = true - rand, 
+         method = "rand_final") %>% 
+  select(name, method, diff)
+
+df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)
+
+# plot for combined dataset
+rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
+  select(name, rand_eob_diff = diff)
+
+conv_A <- bind_rows(conv_V, conv_S) %>% 
+  select(name, conv_diff = diff)
+
+rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
+  select(name, rand_final_diff = diff)
+
+df_MW_A <- rand_eob_A %>% 
+  left_join(conv_A, by = "name") %>% 
+  left_join(rand_final_A, by = "name") %>% 
+  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")
+
+df_MW_A %>% 
+  mutate(method = as.factor(method), 
+         method = recode_factor(method, 
+                                "conv_diff" = "Conventional", 
+                                "rand_eob_diff" = "Randomized", 
+                                "rand_final_diff" = "Randomized\n(24 months)")) %>% 
+  ggplot(aes(x = method, y = diff, fill = method)) +
+  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
+  geom_lv(k = 4, outlier.shape = NA) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
+            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
   scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 3), 
-                     limits = c(0, 28), 
-                     labels = number_format(suffix = "%")) +
-  labs(x = NULL, 
-       y = NULL, 
-       fill = NULL) +
+                     breaks = breaks_pretty(n = 6), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
+  labs(fill = NULL, 
+       x = NULL, 
+       y = "Mean effect in fractional savings",
+       title = "Mean effect of non-routine events when using different M&V methods", 
+       subtitle = str_glue("All {A_building} buildings with measured weather")) +
+  coord_cartesian(ylim = c(-18, 18)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.direction = "horizontal",
-        legend.position = "bottom",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-p2 <- df_FS_tidy %>% 
-  bind_rows(df_FS_messy) %>% 
-  pivot_wider(names_from = method, values_from = savings) %>% 
-  left_join(df_seq_FS_tidy %>% 
-              bind_rows(df_seq_FS_messy) %>% 
-              filter(seq == "eob"), by = c("name", "site")) %>% 
-  pivot_longer(c(conv, rand, FS), names_to = "parameter", values_to = "value") %>% 
-  filter(parameter != "rand") %>% 
-  mutate(parameter = as.factor(parameter), 
-         parameter = recode_factor(parameter, 
-                                   "conv" = "Conventional M&V", 
-                                   "FS" = "Randomized M&V")) %>% 
-  ggplot(aes(x = parameter, y = value, fill = parameter)) +
-  stat_boxplot(geom ='errorbar', width = 0.5) +
-  geom_boxplot(outlier.shape = NA) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 3), 
-                     limits = c(0, 28)) +
-  scale_fill_manual(values = ls_colors) +
-  labs(x = NULL, 
-       y = NULL, 
-       fill = NULL) +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.direction = "horizontal",
-        legend.position = "bottom",
-        axis.text.y = element_blank(),
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-  ggarrange(p1, p2,
-            labels = c("a)", "b)"),
-            ncol = 2, nrow = 1,
-            widths = c(0.5, 1),
-            legend = "none") +
-    plot_annotation(title = "Normalized annual fractional savings estimation comparison")
-```
-
-To assess accuracy at the individual building level, we calculated the
-mean absolute difference between the true savings and the estimated
-savings and normalized the results as a fraction of measured baseline
-for generalized comparison. Figure \@ref(fig:tidyfrac) shows the
-comparison between the two methods using the tidy subset. In this
-comparison, we focused on the median and the 95th percentile of the
-plotted distribution to minimize the influence of outliers. The results
-indicate that the conventional method outperforms the randomized method
-in this subset. This is mostly because the tidy subset includes
-buildings with consistent usage patterns throughout the two-year period,
-meaning the data used for regression model fitting closely aligns with
-the conditions during the model prediction phase. In our case, this
-justifies the TOWT model selection when there is no additional
-adjustment needed for such pre- and post-analysis.
-
-```{r tidyfrac, fig.cap="Absolute deviation of fractional savings from true target savings shown for the 'tidy' subset", fig.width=8, fig.height=4}
-
-plot_data <- df_seq_FS_tidy %>% 
-  filter(seq == "eob") %>% 
-  left_join(df_FS_tidy %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(abs_diff = abs(savings - FS), 
-         plot_max = max(abs_diff))
-
-median_rand <- median(plot_data$abs_diff)
-upper_rand <- quantile(plot_data$abs_diff, probs = 0.95)
-
-p1 <- plot_data %>% 
-  arrange(abs_diff) %>% 
-  ggplot() +
-  geom_bar(aes(x = seq(1, nrow(all_names_tidy), by = 1), y = abs_diff, fill = "Randomized M&V"), 
-           stat = "identity", alpha = 0.5) +
-  geom_hline(yintercept = median_rand, 
-             linetype = "dashed", 
-             color = "black") +
-  annotate(y = median_rand + 0.5, 
-           x = 20, 
-           geom = "text", 
-           label = paste0("(", "50% buildings: < ", round(median_rand, digits = 1), " %)"),
-           size = 4) +
-  geom_hline(yintercept = upper_rand, 
-             linetype = "dashed", 
-             color = "black") +
-  annotate(y = upper_rand + 0.5, 
-           x = 20, 
-           geom = "text", 
-           label = paste0("(", "95% buildings: < ", round(upper_rand, digits = 1), " %)"),
-           size = 4) +
-  scale_x_continuous(expand = c(0.02, 0), 
-                     breaks = breaks_pretty(n = 3)) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 3), 
-                     labels = number_format(suffix = " %"),
-                     limits = c(0, max(plot_data$plot_max) + 0.5)) +
-  scale_fill_manual(values = ls_colors) +
-  labs(fill = NULL, 
-       x = "Number of buildings", 
-       y = "Absolute difference of fractional savings", 
-       subtitle = "Randomized M&V") +
-  coord_cartesian(ylim = c(0, 10)) +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.direction = "horizontal",
-        legend.position = "bottom",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-plot_data <- df_FS_tidy %>% 
-  filter(method != "rand") %>% 
-  group_by(name) %>% 
-  summarise(abs_diff = abs(diff(savings))) %>% 
-  ungroup() %>% 
-  mutate(plot_max = max(abs_diff))
-
-median_conv <- median(plot_data$abs_diff)
-upper_conv <- quantile(plot_data$abs_diff, probs = 0.95)
-
-p2 <- plot_data %>% 
-  arrange(abs_diff) %>% 
-  ggplot() +
-  geom_bar(aes(x = seq(1, nrow(all_names_tidy), by = 1), y = abs_diff, fill = "Conventional M&V"), 
-           stat = "identity", alpha = 0.5) +
-  geom_hline(yintercept = median_conv, 
-             linetype = "dashed", 
-             color = "black") +
-  annotate(y = median_conv + 0.5,
-           x = 20, 
-           geom = "text", 
-           label = paste0("(", "50% buildings: < ", round(median_conv, digits = 1), " %)"),
-           size = 4) +
-  geom_hline(yintercept = upper_conv, 
-             linetype = "dashed", 
-             color = "black") +
-  annotate(y = upper_conv + 0.5, 
-           x = 20,
-           geom = "text", 
-           label = paste0("(", "95% buildings: < ", round(upper_conv, digits = 1), " %)"),
-           size = 4) +
-  scale_x_continuous(expand = c(0.02, 0), 
-                     breaks = breaks_pretty(n = 3)) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 3), 
-                     labels = number_format(suffix = " %"),
-                     limits = c(0, max(plot_data$plot_max) + 0.1)) +
-  scale_fill_manual(values = ls_colors) +
-  labs(fill = NULL, 
-       x = "Number of buildings", 
-       y = NULL, 
-       subtitle = "Conventional M&V") +
-  coord_cartesian(ylim = c(0, 10)) +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.direction = "horizontal",
-        legend.position = "bottom",
-        axis.text.y = element_blank(),
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 0, unit = "mm"))
-
-ggarrange(p1, p2,
-          ncol=2, nrow=1,
-          labels = c("a)", "b)"),
-          align = "hv",
-          legend="none") +
-  plot_annotation(title = "Savings estimation accuracy comparison", 
-                  subtitle = "using 'tidy' subset")
-
-```
-
-While these findings suggest that the conventional M&V method is
-advantageous when a building maintains stable electricity usage over
-time, such stability is rarely guaranteed in real-world scenarios. A
-more realistic condition is shown in the messy subset in Figure
-\@ref(fig:messyfrac). In this case, the randomized method demonstrates
-consistent accuracy in savings estimation, even in the presence of
-random measurement noise, implying strong robustness. In contrast, the
-conventional method shows deviations exceeding 5% in approximately 50%
-of all cases. Non-normalized mean absolute difference comparison plots
-are provided in the supplementary material. The much deteriorated
-performance of the conventional method highlights its vulnerability when
-changes in electricity usage occur. This is because baseline projections
-based on regression models become unreliable when non-routine events are
-present. Non-routine events encompass all influential factors not
-accounted for in the regression model, either due to their
-impracticality to measure---such as hourly occupancy rates---or their
-unpredictability---such as a sudden change in building use where an
-office being converted into a warehouse. We dedicate the following
-section to continue the discussion of the impact.
-
-```{r messyfrac, fig.cap="Absolute deviation of fractional savings from true target savings shown for the 'messy' subset", fig.width=8, fig.height=4}
-
-plot_data <- df_seq_FS_messy %>% 
-  filter(seq == "eob") %>% 
-  left_join(df_FS_messy %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(abs_diff = abs(savings - FS), 
-         plot_max = max(abs_diff))
-
-median_rand <- median(plot_data$abs_diff)
-upper_rand <- quantile(plot_data$abs_diff, probs = 0.95)
-
-p1 <- plot_data %>% 
-  arrange(abs_diff) %>% 
-  ggplot() +
-  geom_bar(aes(x = seq(1, nrow(all_names_messy), by = 1), y = abs_diff, fill = "Randomized M&V"), 
-           stat = "identity", alpha = 0.5) +
-  geom_hline(yintercept = median_rand, 
-             linetype = "dashed", 
-             color = "black") +
-  annotate(y = median_rand + 1, 
-           x = 220, 
-           geom = "text", 
-           label = paste0("(", "50% buildings: < ", round(median_rand, digits = 1), " %)"),
-           size = 4) +
-  geom_hline(yintercept = upper_rand, 
-             linetype = "dashed", 
-             color = "black") +
-  annotate(y = upper_rand + 1, 
-           x = 220, 
-           geom = "text", 
-           label = paste0("(", "95% buildings: < ", round(upper_rand, digits = 1), " %)"),
-           size = 4) +
-  scale_x_continuous(expand = c(0.02, 0), 
-                     breaks = breaks_pretty(n = 3)) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 3), 
-                     labels = number_format(suffix = " %"),
-                     limits = c(0, max(plot_data$plot_max) + 0.5)) +
-  scale_fill_manual(values = ls_colors) +
-  labs(x = "Number of buildings", 
-       y = "Absolute difference of fractional savings", 
-       fill = NULL, 
-       subtitle = "Randomized M&V") +
-  coord_cartesian(ylim = c(0, 25)) +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.direction = "horizontal",
-        legend.position = "bottom",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-plot_data <- df_FS_messy %>% 
-  filter(method != "rand") %>% 
-  group_by(name) %>% 
-  summarise(abs_diff = abs(diff(savings))) %>% 
-  ungroup() %>% 
-  mutate(plot_max = max(abs_diff))
-
-median_conv <- median(plot_data$abs_diff)
-upper_conv <- quantile(plot_data$abs_diff, probs = 0.95)
-
-p2 <- plot_data %>% 
-  arrange(abs_diff) %>% 
-  ggplot() +
-  geom_bar(aes(x = seq(1, nrow(all_names_messy), by = 1), y = abs_diff, fill = "Conventional M&V"), 
-           stat = "identity", alpha = 0.5) +
-  geom_hline(yintercept = median_conv, 
-             linetype = "dashed", 
-             color = "black") +
-  annotate(y = median_conv + 1,
-           x = 220, 
-           geom = "text", 
-           label = paste0("(", "50% buildings: < ", round(median_conv, digits = 1), " %)"),
-           size = 4) +
-  geom_hline(yintercept = upper_conv, 
-             linetype = "dashed", 
-             color = "black") +
-  annotate(y = upper_conv + 1, 
-           x = 220,
-           geom = "text", 
-           label = paste0("(", "95% buildings: < ", round(upper_conv, digits = 1), " %)"),
-           size = 4) +
-  scale_x_continuous(expand = c(0.02, 0), 
-                     breaks = breaks_pretty(n = 3)) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 3), 
-                     labels = number_format(suffix = " %"),
-                     limits = c(0, max(plot_data$plot_max) + 0.1)) +
-  scale_fill_manual(values = ls_colors) +
-  labs(x = "Number of buildings", 
-       y = NULL, 
-       fill = NULL, 
-       subtitle = "Conventional M&V") +
-  coord_cartesian(ylim = c(0, 25)) +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.direction = "horizontal",
-        legend.position = "bottom",
-        axis.text.y = element_blank(),
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 0, unit = "mm"))
-
-ggarrange(p1, p2,
-          ncol=2, nrow=1,
-          labels = c("a)", "b)"),
-          align = "hv",
-          legend="none") +
-  plot_annotation(title = "Savings estimation accuracy comparison", 
-                  subtitle = "using 'messy' subset")
-
-```
-
-## Non-routine events impact on savings estimation
-
-We included two examples in this section to illustrate the impact of
-non-routine events on M&V savings estimation. In the first scenario, we
-manually subtracted a portion of the electricity usage for a specific
-period in the tidy subset to simulate occupancy-confounded measurements,
-which were not originally present in the dataset. In the second
-scenario, focusing on the messy subset, which already contains
-underlying changes, we removed the intervention effect and applied both
-M&V methods to assess which method could more accurately detect the
-resulting zero 'savings.'
-
-### Occupancy change
-
-Although occupancy can be estimated in various ways, such as counting
-check-ins, monitoring WiFi connections, or using indoor CO2
-concentrations as a proxy, these methods are often not cost-effective
-for routine operations in most buildings and may raise privacy concerns.
-In the proposed scenario, we hypothesize that during the baseline
-measurement period, one floor of tenants vacated the building, leaving
-the space unoccupied for four months. This led to a fixed reduction in
-electricity usage, approximated as 20% of the yearly average. We showed
-the impact on TOWT model fitting in Figure \@ref(fig:occchange) where
-from May 1st 2016 to August 31st 2016, the target building electricity
-dropped. Subplot (a) displays the measured baseline data, including the
-occupancy-related change, used as input for model training. Subplot (b)
-shows the model's prediction results for the post-retrofit period. For
-reference, we also plotted the correctly adjusted baseline (labeled as
-'adjusted baseline') in the post-retrofit period. Since occupancy was
-not considered an independent variable in the model, the baseline change
-was incorrectly attributed to temperature and time variations shown as
-the 'projected baseline', leading to underestimated savings estimation.
-
-```{r occ}
-
-n = 6
-
-site_info <- df_energy_tidy %>%
-    filter(name == all_names_tidy$name[n]) %>%
-    select(site, type) %>%
-    distinct()
-
-site <- site_info$site
-
-site_weather <- df_weather_tidy %>%
-    filter(site == site_info$site) %>%
-    select(timestamp, t_out) %>%
-    group_by(timestamp) %>%
-    summarise(t_out = mean(t_out)) %>%
-    ungroup()
-
-df_all <- df_energy_tidy %>%
-    filter(name == all_names_tidy$name[n]) %>%
-    select(timestamp, eload) %>%
-    left_join(site_weather, by = "timestamp")
-
-# Linear interpolation of baseline
-df_all <- df_all %>%
-  run_interpo()
-
-plot_scale <- get_scale(df_all$base_eload)
-
-df_hourly_conv <- df_all %>%
-  run_reset()
-
-df_base_conv <- df_hourly_conv %>%
-    select(datetime,
-           eload = base_eload,
-           t_out) %>% 
-    drop_na()
-
-df_interv_conv <- df_hourly_conv %>%
-    select(datetime,
-           eload = interv_eload,
-           t_out) %>% 
-    drop_na()
-
-change_start <- occ_params$change_start
-change_end <- occ_params$change_end
-change <- mean(df_base_conv$eload) * occ_params$change / 100
-
-base_pre_meas <- df_base_conv %>%
-  filter(datetime < as.Date("2017-01-01")) %>%
-  mutate(eload = ifelse(month(datetime) >= month(change_start) & month(datetime) <= month(change_end), eload - change, eload))
-
-# Baseline projection
-towt_base <- base_pre_meas %>%
-  model_fit()
-
-df_towt <- df_interv_conv %>%
-  filter(datetime >= as.Date("2017-01-01")) %>%
-  select(time = datetime,
-         temp = t_out,
-         eload)
-
-base_pos_proj <- model_pred(df_towt, towt_base) %>%
-  rename("datetime" = "time") %>%
-  select(datetime, towt, eload)
-
-base_pos_true <- df_base_conv %>%
-  filter(datetime >= as.Date("2017-01-01"))
-
-interv_pos_meas <- df_interv_conv %>%
-  filter(datetime >= as.Date("2017-01-01"))
-```
-
-```{r occchange, fig.cap="Illustration of occupancy change impact on TOWT model fitting and baseline projection in the post-retrofit period", fig.width=9, fig.height=4}
-
-p1 <- ggplot() +
-    geom_point(data = base_pre_meas %>%
-                 slice_sample(n = 500),
-               aes(x = datetime, y = eload, color = "Measured baseline"),
-               size = 0.5,
-               alpha = 0.3) +
-    geom_smooth(data = base_pre_meas,
-                aes(x = datetime, y = eload, color = "Measured baseline"), 
-                formula = y ~ x, method = "loess") +
-    scale_x_datetime(date_breaks = "2 months",
-                     date_labels = "%b")  +
-    scale_y_continuous(expand = c(0, 0),
-                       breaks = breaks_pretty(n = 3),
-                       labels = number_format(suffix = " kW")) +
-    scale_color_manual(values = ls_colors) +
-    coord_cartesian(ylim = plot_scale) +
-    labs(x = NULL,
-         y = NULL,
-         color = NULL,
-         title = NULL,
-         subtitle = "Pre-retrofit period") +
-    theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-          legend.direction = "horizontal",
-          legend.position = "bottom",
-          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-p2 <- ggplot() +
-  geom_point(data = base_pos_proj %>%
-               slice_sample(n = 500),
-             aes(x = datetime, y = towt, color = "Projected baseline"),
-             size = 0.5,
-             alpha = 0.3) +
-  geom_smooth(data = base_pos_proj,
-              aes(x = datetime, y = towt, color = "Projected baseline"), 
-              formula = y ~ x, method = "loess") +
-  geom_point(data = base_pos_true %>%
-               slice_sample(n = 500),
-             aes(x = datetime, y = eload, color = "Adjusted baseline"),
-             size = 0.5,
-             alpha = 0.3) +
-  geom_smooth(data = base_pos_true,
-              aes(x = datetime, y = eload, color = "Adjusted baseline"), 
-              formula = y ~ x, method = "loess") +
-  geom_point(data = interv_pos_meas %>%
-               slice_sample(n = 500),
-             aes(x = datetime, y = eload, color = "Measured interv"),
-             size = 0.5,
-             alpha = 0.3) +
-  geom_smooth(data = interv_pos_meas,
-              aes(x = datetime, y = eload, color = "Measured interv"), 
-              formula = y ~ x, method = "loess") +
-  scale_x_datetime(date_breaks = "2 months",
-                   date_labels = "%b")  +
-  scale_y_continuous(expand = c(0, 0),
-                     breaks = breaks_pretty(n = 3),
-                     labels = number_format(suffix = " kW")) +
-  scale_color_manual(values = ls_colors) +
-  coord_cartesian(ylim = plot_scale) +
-  labs(x = NULL,
-       y = NULL,
-       color = NULL,
-       title = NULL,
-       subtitle = "post-retrofit period") +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        axis.text.y = element_blank(),
-        legend.direction = "horizontal",
-        legend.position = "bottom",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-ggarrange(p1, p2,
-          ncol=2, nrow=1,
-          labels = c("a)", "b)"),
-          widths = c(1, 1),
-          align = "v",
-          legend="bottom") +
-  plot_annotation(title = "Building energy consumption with added occupancy change")
-```
-
-Figure \@ref(fig:occsavings) shows the overall estimation accuracy
-between the two M&V methods where the accuracy is indicated as a
-difference-in-difference value:
-
-$$
-|FS_{conv} - FS_{true}| - |FS_{rand} - FS_{true}|
-$$
-
-Therefore, a positive value indicates that the randomized method
-provides an estimation more closely aligned with the true target
-savings. The bar plot shows the difference-in-difference of the
-calculated fractional savings ($FS$) for each target building and the
-dotted line shows the absolute deviation of randomized method (i.e.
-$|FS_{rand} - FS_{true}|$ ) as a reference. In addition, we showed the
-averaged difference-in-difference metric for each site in the figure,
-quantifying how much more accurate the randomized method is compared to
-the conventional method. As a result, across all sites, the randomized
-method consistently outperforms the conventional method and the values
-of dotted line show similar results compared to Figure
-\@ref(fig:tidyfrac) further implying that the non-routine event has
-negligible influence on the randomized method.
-
-```{r occsavings, fig.cap="Savings estimation accuracy comparison between two M&V methods with added occupancy change (with the average difference-in-difference average on each location displayed at the bottom).", fig.width=9, fig.height=4}
-s <- 'S2'
-dev_conv <- df_NRE_occ %>%
-  left_join(df_loc, by = "site") %>% 
-  filter(scenario == s) %>%
-  filter(method != "rand") %>%
-  group_by(name, location, scenario) %>%
-  summarise(conv = abs(diff(savings))) %>%
-  ungroup()
-
-dev_rand <- df_NRE_occ %>%
-  left_join(df_loc, by = "site") %>% 
-  filter(scenario == s) %>%
-  filter(method != "conv") %>%
-  group_by(name, location, scenario) %>%
-  summarise(rand = abs(diff(savings))) %>%
-  ungroup()
-
-dev_FS <- dev_rand %>%
-  left_join(dev_conv, by = c("name", "location", "scenario")) %>% 
-  mutate(diff_in_diff = conv - rand)
-
-mean_diff <- dev_FS %>%
-  group_by(location) %>%
-  summarise(mean = round(mean(diff_in_diff), digits = 1),
-            pos = round(n() / 2) + 1,
-            .groups = 'keep')
-
-dev_FS %>%
-  ggplot(aes(group = location)) +
-  geom_line(aes(x = name, y = rand, color = "Randomized M&V")) +
-  geom_point(aes(x = name, y = rand, color = "Randomized M&V"), alpha = 0.8) +
-  geom_col(aes(x = name, y = diff_in_diff), position = "identity", alpha = 0.2, fill = "#78c679") +
-  facet_wrap(~location, nrow = 1, scales = "free_x") +
-  scale_y_continuous(expand = c(0.1, 0),
-                     breaks = breaks_pretty(n = 4), 
-                     labels = number_format(suffix = "%")) +
-  geom_text(data = mean_diff,
-            aes(x = pos, y = -.5, group = location, label = paste0(mean, "%"))) +
-  scale_color_manual(values = ls_colors) +
-  labs(x = NULL,
-       fill = NULL,
-       y = "Difference-in-difference of fractional savings",
-       color = NULL, 
-       title = "Savings estimation accuracy comparison for non-routine events", 
-       subtitle = "using 'tidy' subset") +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        axis.text.x = element_blank(),
-        legend.direction = "horizontal",
-        legend.position = "bottom",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-```
-
-### No-saving detection
-
-As previously mentioned, the 'savings' detected in commercial buildings
-during the pandemic were largely confounded by null occupancy. In other
-words, if no intervention is applied or the intervention effect is known
-to be null (e.g., constantly overridden by the baseline), an ideal M&V
-method should detect no savings. Given that the messy dataset allows for
-up to a 25% annual usage change in the original dataset, a more reliable
-M&V method should inform the analyst that no savings occurred prior to
-the addition of the intervention. To test this, we removed the chilled
-water plant reset effect from the sampled days and assessed whether the
-M&V methods could correctly detect zero savings.
-
-```{r nosavings, fig.cap="Comparison of no intervention effect detection results (with 0% as true annual fractional savings) between conventional and randomized M&V method.", fig.width=6, fig.height=3}
-
-df_FS_nsprt %>% 
-  pivot_wider(names_from = method, values_from = savings) %>% 
-  left_join(df_seq_FS_nsprt %>% filter(seq == "eob"), by = c("name", "site")) %>% 
-  pivot_longer(c(true, conv, rand, FS), names_to = "parameter", values_to = "value") %>% 
-  filter(parameter != "true" & parameter != "rand") %>% 
-  mutate(parameter = as.factor(parameter), 
-         parameter = recode_factor(parameter, 
-                                   "conv" = "Conventional M&V", 
-                                   "FS" = "Randomized M&V")) %>% 
-  ggplot(aes(x = parameter, y = value, fill = parameter)) +
-  stat_boxplot(geom ='errorbar', width = 0.2) +
-  geom_boxplot(outlier.shape = NA, width = 0.5) +
-  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 1.2) +
-  annotate(geom = "text", 
-           x = 0.7, 
-           y = 0.5, 
-           color = "black",
-           label = "True savings = 0") +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 5), 
-                     limits = c(0, 16), 
-                     labels = number_format(suffix = "%")) +
-  scale_fill_manual(values = ls_colors) +
-  labs(x = NULL, 
-       y = NULL, 
-       fill = NULL, 
-       title = "Normalized confounded annual fractional savings comparison") +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.direction = "horizontal",
         legend.position = "none",
@@
 
-Figure \@ref(fig:nosavings) shows that the conventional M&V method in
-average detected 5% savings when there should be none. The randomized
-method on the other hand is able to provide an estimation much closer to
-0% with reduced uncertainty. This improvement is largely due to the
-randomized sampling approach, where the two strategies (baseline and
-intervention) are sampled at a 50%/50% ratio. This ensures that any
-changes or disturbances additional to the intervention implementation,
-such as concurrent lighting retrofits, are evenly distributed across
-both strategies, balancing their effects.
+As mentioned in Section \@ref(data-filtering), the majority of the
+selected buildings are affected by non-routine events. To assess the
+reliability of the two M&V methods, we repeated the M&V procedure on the
+dataset before applying the reset control. In other words, the
+intervention has no effect on building electricity usage and our
+objective is to compare which method can provide an estimation closer to
+0 savings. Figure \@ref(fig:nosaving) shows the randomized M&V method
+(both at the time of satisfying all stopping criteria and at the end of
+two-year period) produced consistent results. We included the absolute
+deviation in the appendix. Similarly, stopping early can lead to larger
+uncertainty similar to Figure \@ref(fig:abstract). As for conventional
+method, the uncertainty range is much larger than the randomized method
+and it shows on average the deviation is 1.3% meaning a building analyst
+can claim 1.3% energy savings were detected when there isn't any.
 
@@
 
-df_model_acc_messy %>% 
-  mutate(group = "Messy subset") %>% 
-  rbind(df_model_acc_tidy %>% mutate(group = "Tidy subset")) %>% 
+df_model_acc_variable %>% 
+  mutate(group = "variable subset") %>% 
+  rbind(df_model_acc_stable %>% mutate(group = "stable subset")) %>% 
   mutate(group = as.factor(group)) %>% 
@@
 
-## Sampling ratio for randomized M&V
+## Sampling interval impact
+
+```{r absinterv2, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy", fig.height=9, fig.width=12}
+
+#### ABS-INT ####
+# dropped version
+df_interval_keep_stable %<>% 
+  left_join(df_fs_stable %>% filter(method == "rand"), by = c("name", "site")) %>% 
+  select(-c("scenario", "method")) %>% 
+  rename("interval_fs_1" = "savings") %>% 
+  left_join(df_fs_tmy_stable, by = c("name", "site")) %>% 
+  select(-conv) %>% 
+  rename("interval_tmy_1" = "rand")
+
+df_interval_keep_variable %<>% 
+  left_join(df_fs_variable %>% filter(method == "rand"), by = c("name", "site")) %>% 
+  select(-c("scenario", "method")) %>% 
+  rename("interval_fs_1" = "savings") %>% 
+  left_join(df_fs_tmy_variable, by = c("name", "site")) %>% 
+  select(-conv) %>% 
+  rename("interval_tmy_1" = "rand")
+
+df_drop <- df_interval_drop_stable %>% 
+  bind_rows(df_interval_drop_variable) %>% 
+  select(name, contains("fs")) %>% 
+  rename("drop_1" = "interval_fs_1", 
+         "drop_2" = "interval_fs_2", 
+         "drop_3" = "interval_fs_3", 
+         "drop_7" = "interval_fs_7")
+
+df_keep <- df_interval_keep_stable %>% 
+  bind_rows(df_interval_keep_variable) %>% 
+  select(name, contains("fs")) %>% 
+  rename("keep_1" = "interval_fs_1",
+         "keep_2" = "interval_fs_2", 
+         "keep_3" = "interval_fs_3", 
+         "keep_7" = "interval_fs_7")
+
+df_MW_A <- df_drop %>% 
+  left_join(df_keep, by = c("name")) %>%
+  left_join(rbind(df_fs_stable %>% 
+              filter(scenario == "ref" & method == "true"), 
+              df_fs_variable %>% 
+              filter(scenario == "ref" & method == "true")), by = "name") %>% 
+  mutate(drop_1 = abs(drop_1 - savings),
+         drop_2 = abs(drop_2 - savings), 
+         drop_3 = abs(drop_3 - savings), 
+         drop_7 = abs(drop_7 - savings), 
+         keep_1 = abs(keep_1 - savings),
+         keep_2 = abs(keep_2 - savings), 
+         keep_3 = abs(keep_3 - savings), 
+         keep_7 = abs(keep_7 - savings)) %>% 
+  pivot_longer(cols = contains("drop") | contains("keep"),  
+               names_to = c("group", "interval"),
+               names_pattern = "(drop|keep)_(\\d+)",
+               values_to = "diff")
+
+p_top <- df_MW_A %>% 
+  mutate(interval = as.factor(interval), 
+         interval = recode_factor(interval, 
+                                  "1" = "1-day\nsampling",
+                                  "2" = "2-day\nsampling", 
+                                  "3" = "3-day\nsampling", 
+                                  "7" = "7-day\nsampling"), 
+         group = as.factor(group), 
+         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
+  ggplot(aes(x = interval, y = diff, fill = group)) +
+  geom_jitter(alpha = 0.8, 
+              size = 0.5, 
+              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
+  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% group_by(group, interval) %>% summarise(mean = mean(diff, na.rm = T)) %>% ungroup(), 
+            aes(x = interval, y = mean, group = group,
+                label = paste0(round(mean, digits = 1), " %")), 
+            position = position_dodge(width = 0.75)) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80")) + 
+  labs(fill = NULL, 
+       color = NULL, 
+       x = NULL, 
+       y = "Absolute error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with measured weather")) +
+  coord_cartesian(ylim = c(0, 10)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "bottom",
+        axis.text.x = element_blank(),
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+# TMY version
+df_drop <- df_interval_drop_stable %>% 
+  bind_rows(df_interval_drop_variable) %>% 
+  select(name, contains("tmy")) %>% 
+  rename("drop_1" = "interval_tmy_1",
+         "drop_2" = "interval_tmy_2", 
+         "drop_3" = "interval_tmy_3", 
+         "drop_7" = "interval_tmy_7")
+
+df_keep <- df_interval_keep_stable %>% 
+  bind_rows(df_interval_keep_variable) %>% 
+  select(name, contains("tmy")) %>% 
+  rename("keep_1" = "interval_tmy_1",
+         "keep_2" = "interval_tmy_2", 
+         "keep_3" = "interval_tmy_3", 
+         "keep_7" = "interval_tmy_7")
+
+df_TW_A <- df_drop %>% 
+  left_join(df_keep, by = c("name")) %>%
+  left_join(rbind(df_fs_stable %>% 
+              filter(scenario == "ref" & method == "true"), 
+              df_fs_variable %>% 
+              filter(scenario == "ref" & method == "true")), by = "name") %>% 
+  mutate(drop_1 = abs(drop_1 - savings), 
+         drop_2 = abs(drop_2 - savings), 
+         drop_3 = abs(drop_3 - savings), 
+         drop_7 = abs(drop_7 - savings), 
+         keep_1 = abs(keep_1 - savings), 
+         keep_2 = abs(keep_2 - savings), 
+         keep_3 = abs(keep_3 - savings), 
+         keep_7 = abs(keep_7 - savings)) %>% 
+  pivot_longer(cols = contains("drop") | contains("keep"),  
+               names_to = c("group", "interval"),
+               names_pattern = "(drop|keep)_(\\d+)",
+               values_to = "diff")
+
+p_bottom <- df_TW_A %>% 
+  mutate(interval = as.factor(interval), 
+         interval = recode_factor(interval, 
+                                  "1" = "1-day\nsampling",
+                                  "2" = "2-day\nsampling", 
+                                  "3" = "3-day\nsampling", 
+                                  "7" = "7-day\nsampling"), 
+         group = as.factor(group), 
+         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
+  ggplot(aes(x = interval, y = diff, fill = group)) +
+  geom_jitter(alpha = 0.8, 
+              size = 0.5, 
+              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
+  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% group_by(group, interval) %>% summarise(mean = mean(diff, na.rm = T)) %>% ungroup(), 
+            aes(x = interval, y = mean, group = group,
+                label = paste0(round(mean, digits = 1), " %")), 
+            position = position_dodge(width = 0.75)) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80")) + 
+  labs(fill = NULL, 
+       color = NULL, 
+       x = NULL, 
+       y = "Absolute error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with TMY weather")) +
+  coord_cartesian(ylim = c(0, 10)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "bottom",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+ggarrange(p_top, p_bottom, 
+          ncol = 1, nrow = 2, 
+          align = "hv", 
+          common.legend = T, 
+          legend = "bottom") +
+  plot_annotation(title = "Absolute error in savings estimation of ranodmized M&V by different sampling intervals")
+
+```
+
+```{r absinterv, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy", fig.height=9, fig.width=12}
+
+# sprt sequence plot for different sampling intervals
+df_seq_interval_fs_keep_stable %<>% 
+  select(-c(sprt, temp, final)) %>% 
+  bind_rows(df_seq_fs_stable %>% 
+              filter(seq == "eob") %>% 
+              select(-seq) %>% 
+              mutate(interval = 1) %>% 
+              rename(eob = "fs"))
+
+df_seq_interval_fs_keep_variable %<>% 
+  select(-c(sprt, temp, final)) %>% 
+  bind_rows(df_seq_fs_variable %>% 
+              filter(seq == "eob") %>% 
+              select(-seq) %>% 
+              mutate(interval = 1) %>% 
+              rename(eob = "fs"))
+
+df_drop <- df_seq_interval_fs_drop_stable %>% 
+  bind_rows(df_seq_interval_fs_drop_variable) %>% 
+  select(name, eob, interval) %>% 
+  mutate(group = "drop")
+
+df_keep <- df_seq_interval_fs_keep_stable %>% 
+  bind_rows(df_seq_interval_fs_keep_variable) %>% 
+  select(name, eob, interval) %>% 
+  mutate(group = "keep")
+
+df_MW_A <- df_drop %>% 
+  bind_rows(df_keep) %>% 
+  left_join(rbind(df_fs_stable %>% 
+                    filter(scenario == "ref" & method == "true"), 
+                  df_fs_variable %>% 
+                    filter(scenario == "ref" & method == "true")), by = "name") %>% 
+  mutate(diff = abs(eob - savings)) %>% 
+  filter(is.finite(diff))
+
+p_top <- df_MW_A %>% 
+  mutate(interval = as.factor(interval), 
+         interval = recode_factor(interval, 
+                                  "1" = "1-day\nsampling",
+                                  "2" = "2-day\nsampling", 
+                                  "3" = "3-day\nsampling", 
+                                  "7" = "7-day\nsampling"), 
+         group = as.factor(group), 
+         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
+   ggplot(aes(x = interval, y = diff, fill = group)) +
+  geom_jitter(alpha = 0.8, 
+              size = 0.5, 
+              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
+  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% group_by(group, interval) %>% summarise(mean = mean(diff, inf.rm = T)) %>% ungroup(), 
+            aes(x = interval, y = mean, group = group,
+                label = paste0(round(mean, digits = 1), " %")), 
+            position = position_dodge(width = 0.75)) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80")) + 
+  labs(fill = NULL, 
+       color = NULL, 
+       x = NULL, 
+       y = "Absolute error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with measured weather")) +
+  coord_cartesian(ylim = c(0, 20)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "bottom",
+        axis.text.x = element_blank(),
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+# with tmy version
+df_seq_interval_nm_keep_stable %<>% 
+  select(-c(sprt, temp, final)) %>% 
+  bind_rows(df_sprt_all_stable %>% 
+             filter(seq == "eob") %>% 
+              select(-c(seq, n_weeks)) %>% 
+              mutate(interval = 1) %>% 
+              rename(eob = "annual"))
+
+df_seq_interval_nm_keep_variable %<>% 
+  select(-c(sprt, temp, final)) %>% 
+  bind_rows(df_sprt_all_variable %>% 
+              filter(seq == "eob") %>% 
+              select(-c(seq, n_weeks)) %>% 
+              mutate(interval = 1) %>% 
+              rename(eob = "annual"))
+
+df_drop <- df_seq_interval_nm_drop_stable %>% 
+  bind_rows(df_seq_interval_nm_drop_variable) %>% 
+  select(name, eob, interval) %>% 
+  mutate(group = "drop")
+
+df_keep <- df_seq_interval_nm_keep_stable %>% 
+  bind_rows(df_seq_interval_nm_keep_variable) %>% 
+  select(name, eob, interval) %>% 
+  mutate(group = "keep")
+
+df_TW_A <- df_drop %>% 
+  bind_rows(df_keep) %>% 
+  left_join(rbind(df_fs_stable %>% 
+                    filter(scenario == "ref" & method == "true"), 
+                  df_fs_variable %>% 
+                    filter(scenario == "ref" & method == "true")), by = "name") %>% 
+  mutate(diff = abs(eob - savings)) %>% 
+  filter(is.finite(diff))
+
+p_bottom <- df_TW_A %>% 
+  mutate(interval = as.factor(interval), 
+         interval = recode_factor(interval, 
+                                  "1" = "1-day\nsampling",
+                                  "2" = "2-day\nsampling", 
+                                  "3" = "3-day\nsampling", 
+                                  "7" = "7-day\nsampling"), 
+         group = as.factor(group), 
+         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
+  ggplot(aes(x = interval, y = diff, fill = group)) +
+  geom_jitter(alpha = 0.8, 
+              size = 0.5, 
+              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
+  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% group_by(group, interval) %>% summarise(mean = mean(diff, na.rm = T)) %>% ungroup(), 
+            aes(x = interval, y = mean, group = group,
+                label = paste0(round(mean, digits = 1), " %")), 
+            position = position_dodge(width = 0.75)) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80")) + 
+  labs(fill = NULL, 
+       color = NULL, 
+       x = NULL, 
+       y = "Absolute error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with TMY weather")) +
+  coord_cartesian(ylim = c(0, 20)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "bottom",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+ggarrange(p_top, p_bottom, 
+          ncol = 1, nrow = 2, 
+          align = "hv", 
+          common.legend = T, 
+          legend = "bottom") +
+  plot_annotation(title = "Absolute error in savings estimation of randomized M&V after satisfying all stopping criteria")
+```
+
+## Sampling ratio impact
 
@@
 
-```{r cont, fig.cap="Randomized M&V absolute deviation of fractional savings from true target savings for a whole year with more intervention days sampled after all stopping criteria satisfied", fig.width=8, fig.height=4}
-
-plot_data <- df_cont_FS_tidy %>% 
-  rename(cont = FS) %>% 
-  left_join(df_FS_tidy %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(abs_diff = abs(savings - cont), 
-         plot_max = max(abs_diff))
-
-median_rand <- median(plot_data$abs_diff)
-upper_rand <- quantile(plot_data$abs_diff, probs = 0.95)
-
-p1 <- plot_data %>% 
-  arrange(abs_diff) %>% 
-  ggplot() +
-  geom_bar(aes(x = seq(1, nrow(all_names_tidy), by = 1), y = abs_diff, fill = "Randomized M&V"), 
-           stat = "identity", alpha = 0.5) +
-  geom_hline(yintercept = median_rand, 
-             linetype = "dashed", 
-             color = "black") +
-  annotate(y = median_rand + 0.5, 
-           x = 20, 
-           geom = "text", 
-           label = paste0("(", "50% buildings: < ", round(median_rand, digits = 1), " %)"),
-           size = 4) +
-  geom_hline(yintercept = upper_rand, 
-             linetype = "dashed", 
-             color = "black") +
-  annotate(y = upper_rand + 0.5, 
-           x = 20, 
-           geom = "text", 
-           label = paste0("(", "95% buildings: < ", round(upper_rand, digits = 1), " %)"),
-           size = 4) +
-  scale_x_continuous(expand = c(0.02, 0), 
-                     breaks = breaks_pretty(n = 3)) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 3), 
-                     labels = number_format(suffix = " %"),
-                     limits = c(0, max(plot_data$plot_max) + 0.5)) +
-  scale_fill_manual(values = ls_colors) +
-  labs(x = "Number of buildings", 
-       y = "Absolute difference of fractional savings", 
-       fill = NULL, 
-       subtitle = "Tidy subset") +
-  coord_cartesian(ylim = c(0, 25)) +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.direction = "horizontal",
-        legend.position = "bottom",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-plot_data <- df_cont_FS_messy %>% 
-  rename(cont = FS) %>% 
-  left_join(df_FS_messy %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(abs_diff = abs(savings - cont), 
-         plot_max = max(abs_diff))
-
-median_rand <- median(plot_data$abs_diff)
-upper_rand <- quantile(plot_data$abs_diff, probs = 0.95)
-
-p2 <- plot_data %>% 
-  arrange(abs_diff) %>% 
-  ggplot() +
-  geom_bar(aes(x = seq(1, nrow(all_names_messy), by = 1), y = abs_diff, fill = "Randomized M&V"), 
-           stat = "identity", alpha = 0.5) +
-  geom_hline(yintercept = median_rand, 
-             linetype = "dashed", 
-             color = "black") +
-  annotate(y = median_rand + 0.5, 
-           x = 220, 
-           geom = "text", 
-           label = paste0("(", "50% buildings: < ", round(median_rand, digits = 1), " %)"),
-           size = 4) +
-  geom_hline(yintercept = upper_rand, 
-             linetype = "dashed", 
-             color = "black") +
-  annotate(y = upper_rand + 0.5, 
-           x = 220, 
-           geom = "text", 
-           label = paste0("(", "95% buildings: < ", round(upper_rand, digits = 1), " %)"),
-           size = 4) +
-  scale_x_continuous(expand = c(0.02, 0), 
-                     breaks = breaks_pretty(n = 3)) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 3), 
-                     labels = number_format(suffix = " %"),
-                     limits = c(0, max(plot_data$plot_max) + 0.1)) +
-  scale_fill_manual(values = ls_colors) +
-  labs(x = "Number of buildings", 
-       y = NULL, 
-       fill = NULL, 
-       subtitle = "Messy subset") +
-  coord_cartesian(ylim = c(0, 25)) +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.direction = "horizontal",
-        legend.position = "bottom",
-        axis.text.y = element_blank(),
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 0, unit = "mm"))
-
-ggarrange(p1, p2,
-          ncol=2, nrow=1,
-          labels = c("a)", "b)"),
-          align = "hv",
-          legend="none") +
-  plot_annotation(title = "Randomized M&V savings estimation accuracy comparison with continue sampling included", 
-                  subtitle = "20%/80% baseline/intervention")
-```
-
-The plot shows the deviation remains consistent in the tidy subset
-comapred to Figure \@ref(fig:tidyfrac) but increased slightly in the
-messy subset compare to to Figure \@ref(fig:messyfrac). With unbalanced
-sampling ratio, there is a trade-off for estimation accuracy. But
-compared with conventional method, this is still a preferred approach.
-Most importantly, this approach allows customers to realize a proportion
-of expected savings while documenting baseline measurement. For example,
-opting for an early stop at 24 or 36 weeks with 50% intervention,
-followed by re-sampling at 80% until the year's end (\~another 20
-weeks), would enable customers to capture about 65% of the full-range
-savings.
-
-## Limitations and future study
+The plot shows the deviation remains consistent in the stable subset
+comapred to Figure \@ref(fig:stablefrac) but increased slightly in the
+variable subset compare to to Figure \@ref(fig:variablefrac). With
+unbalanced sampling ratio, there is a trade-off for estimation accuracy.
+But compared with conventional method, this is still a preferred
+approach. Most importantly, this approach allows customers to realize a
+proportion of expected savings while documenting baseline measurement.
+For example, opting for an early stop at 24 or 36 weeks with 50%
+intervention, followed by re-sampling at 80% until the year's end
+(\~another 20 weeks), would enable customers to capture about 65% of the
+full-range savings.
+
+## Limitations
 
@@
 
-For future work, our main objective is to extend the application of the
-proposed randomized M&V approach. For instance, a customer may want to
-know whether Model Predictive Control (MPC) can reduce energy bills
-under a dynamic pricing structure or estimate the Return on Investment
-(ROI) for a retrofit. Or whether a load shift control can save
-operational cost by shifting load out of a given time window such grid
-peak price/emissions period. We also believe that this approach could be
-adapted to analyze the effects of other retrofit interventions at a
-population level. Additionally, we aim to demonstrate that the
-randomized method and sequential analysis framework can be applied to
-various performance metrics, including indoor air quality, operational
-carbon emissions, and thermal comfort.
-
 # Conclusion
@@
 and off), we believe its true value lies in the usefulness and
-conviencience for most control retrofit validation in the field test.
+convenience for most control retrofit validation in the field test.
 

### 2025-02-01 code/1_analysis.R

@@ -58,8 +58,8 @@ source(paste0(function_path, "rand_seq.R"))
 run_params <- list(type = "stable", 
                    sprt = T, 
                    sprt_cont = F, 
-                   interval = F, 
-                   null = T)
+                   interval = T, 
+                   null = F)
 
 # Adding intervention effect as advanced chiller operation
 ctr_params <- list(peak_hours = 10:16,                      # accounts for peak hours

### 2025-02-01 code/graphical.R

@@ -172,9 +172,9 @@ df_interval_keep_stable <- read_rds(paste0(readfile_stable, "df_interval_keep.rd
 df_seq_interval_fs_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_fs_drop.rds"))
 df_seq_interval_fs_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_fs_keep.rds"))
 df_seq_interval_nm_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_nm_drop.rds"))
-df_seq_interval_nm_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_nm_drop.rds"))
+df_seq_interval_nm_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_nm_keep.rds"))
 df_seq_interval_tl_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_tl_drop.rds"))
-df_seq_interval_tl_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_tl_drop.rds"))
+df_seq_interval_tl_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_tl_keep.rds"))
 
 
 all_sites_stable <- df_energy_stable %>%
@@ -871,6 +871,14 @@ df_interval_keep_stable %<>%
   select(-conv) %>% 
   rename("interval_tmy_1" = "rand")
 
+df_interval_keep_variable %<>% 
+  left_join(df_fs_variable %>% filter(method == "rand"), by = c("name", "site")) %>% 
+  select(-c("scenario", "method")) %>% 
+  rename("interval_fs_1" = "savings") %>% 
+  left_join(df_fs_tmy_variable, by = c("name", "site")) %>% 
+  select(-conv) %>% 
+  rename("interval_tmy_1" = "rand")
+
 df_drop <- df_interval_drop_stable %>% 
   bind_rows(df_interval_drop_variable) %>% 
   select(name, contains("fs")) %>% 
@@ -889,7 +897,10 @@ df_keep <- df_interval_keep_stable %>%
 
 df_MW_A <- df_drop %>% 
   left_join(df_keep, by = c("name")) %>%
-  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = "name") %>% 
+  left_join(rbind(df_fs_stable %>% 
+                    filter(scenario == "ref" & method == "true"), 
[…]

### 2025-03-01 code/manuscript.Rmd
push for second draft review

@@
 ---
-title: "Reliability Demonstration of A Novel Randomized Measurement and Verification Method for Switchable Control Retrofit Using Large Open Source Dataset"
+title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
 author:
   - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
-abstract: Conventional measurement and verification (M&V) methods relying on pre- and post-  retrofit comparison for estimating whole-building energy savings are often time-consuming and unreliable, especially when non-routine events such as reduced occupancy occur during the M&V process.
-Those events are unrelated to the intervention but significantgly affect building energy consumption and thus when the analyst applies conventiona M&V suggested by industry guidelines, the results can be largely confounded.
-In this study, we demonstrated that switchable interventions, such as most of the control retrofits, can benefit from random sampling, where the analyst randomly selects whether to implement the baseline or the intervention strategy each day.
+abstract: Conventional measurement and verification (M&V) methods rely on pre- and post- retrofit comparison to estimate energy savings.
+They are often time-consuming and unreliable, especially when non-routine events such as reduced occupancy, or equipment failures occur during the M&V process.
+Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies conventional M&V suggested by industry guidelines, the results can be largely confounded.
+In this study, we demonstrated that switchable interventions, such as most of the control retrofits, can benefit from a new M&V method which randomly samples whether to implement the baseline or the intervention strategy each day.
 We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention.
-Our results show that comparing to the conventional method, the randomized method provides faster and more accurate savings estimation.
-Additionaly we found that the randomized M&V approach estimates much closer to the true savings when non-routine events are present demonstrating much improved reliability.
+Our results show that compared to the conventional method, the randomized method provides faster and more accurate savings estimation.
+Additionally we found that when non-routine events are present, the randomized M&V approach estimates savings that are much closer to the ground-truth savings than the conventional M&V method, demonstrating much improved reliability.
 
@@
 
-## Background
-
-### Conventional M&V
+## Conventional M&V
 
@@
 and Verification Protocol (IPMVP), and the Federal Energy Management
-Program (FEMP) for standard guidelines [@doe_mv_2008,
-@efficiency_valuation_organisation_international_2007,
-@ashrae_ashrae_2014]. These guidelines outline standardized methods for
+Program (FEMP) for standard guidelines [@doe_mv_2008;
+@efficiency_valuation_organisation_international_2012;
+@ashrae_ashrae_2023]. These guidelines outline standardized methods for
 quantifying energy savings, whether through calibrated simulations or
@@
 whole-building level where the measurements are obtained from utility
-bills or whole-building meters. The corresponding M&V process required
-by ASHRAE Guideline 14, which we refer to as the convention method in
-this paper. Typically, the process begins with baseline measurements
-taken over a year before implementing any energy-efficiency retrofit,
-followed by the same measurement procedure during the intervention
-period. After collecting two years of data, an M&V analyst fits an
-energy prediction model, using variables such as outdoor temperature and
-time [@mathieu_quantifying_2011] to project baseline energy consumption
-in the post-retrofit period. The difference between the counter-factual
-baseline and the measured intervention represents the energy savings. A
-key drawback of this method is its reliance on a two-year timeline to
-quantify savings, during which baseline measurements can become outdated
-due to changes in building performance caused by non-routine events
-unrelated to the intervention. This limitation reduces the feasibility
-of rapid M&V and complicates the quantification of estimation
-uncertainty, thus impacting the accuracy and timeliness of savings
-assessments.
-
-### Randomized M&V
+bills or whole-building meters, however the overall conclusions would
+also apply to isolation methods. Typically, the process begins with
+baseline measurements taken during the year before implementing the 
+energy-efficiency retrofit, followed by the same measurement procedure
+during the year after the intervention. After collecting two years of
+data, an M&V analyst fits an energy prediction model, using variables
+such as outdoor temperature and time [@mathieu_quantifying_2011] to
+project baseline energy consumption in the post-retrofit period. The
+difference between the counter-factual baseline and the measured
+intervention period energy consumption represents the energy savings.
+One of the key drawbacks of this method is its reliance on a two-year
+timeline to quantify savings, during which baseline measurements can
+become outdated due to changes in building performance caused by
+non-routine events unrelated to the intervention. This limitation
+reduces the feasibility of rapid M&V and complicates the quantification
+of estimation uncertainty, thus impacting the accuracy and timeliness of
+savings assessments.
+
+## Randomized M&V
 
@@
 challenges posed by non-routine events, we propose a novel M&V method
-that adopts the randomized crossover design, a concept borrowed from
-medical and agricultural studies. Another improvement is that we
+that adopts the randomized crossover design, a gold standard from
+medical and agricultural studies [@blackston2019; @duan2013; @gupta2019;
+@mie2022; @munro; @raseduzzaman2017]. Another improvement is that we
 proposed a sequential evaluation framework and defined stopping criteria
@@
 unnecessary measurement collection over the full 2-year M&V cycle. The
-full framework is detailed in another study with all stopping criteria
-outlined <!--# TODO: Add references -->. In summary, this method
-provides M&V analysts with a randomized schedule that alternates between
-baseline and intervention implementation while ensuring balanced
-sampling across days of the week. For example, given a 10-week M&V
-period for 1 intervention, the balanced randomized schedule would
-equally sample 5 Mondays with the baseline and 5 Mondays with the
-intervention. The limitation of the randomized M&V is that it is only
-applicable to a subset of retrofit projects such as control
-interventions. However, for all applicable use cases, it allows analysts
-to detect energy savings sequentially as the study progresses meaning
-once the desired savings target is achieved, analysts can terminate the
-M&V. The key advantage of randomization, which is one of the study
-objectives, is that if control strategies are sampled with equal
-probability, the influence of non-routine events is likely to be evenly
-distributed between the baseline and intervention measurements. This
-means that the savings estimate, calculated as the difference between
-the two, effectively 'cancels out' the impact of these disturbances,
-leading to a more accurate and unbiased assessment of the intervention's
-impact.
-
-### BDG2 dataset
+full framework is detailed in a previous published study with all
+stopping criteria and example usecases outlined [@raftery2024]. In
+summary, this method provides M&V analysts with a randomized schedule
+that alternates between baseline and intervention implementation while
+ensuring balanced sampling across days of the week and seasons. For
+example, given a 10-week M&V period for 1 intervention, the balanced
+randomized schedule would equally sample 5 Mondays with the baseline
+strategy in operation and 5 Mondays with the intervention strategy in
+operation. The limitation of this method is that it is only applicable
+to a subset of retrofit projects where interventions can be easily
+switched on and off. However, for all applicable use cases, it allows
+analysts to detect energy savings sequentially right after test begins.
+In addition, once the desired savings target is achieved, analysts can
+terminate the M&V and switch to 100% intervention. The key advantage of
+randomization is that if control strategies are sampled with equal
+probability, the influence of other confounding factors such as
+occupancy change (or commonly known as non-routine events) is likely to
+be evenly distributed among measurements, leading to a more accurate and
+unbiased assessment of the intervention's impact.
+
+## BDG2 dataset
 
@@
 simulation, and algorithm development [@miller_building_2020]. BGD2
-contains over 500 buildings' metadata and realistic operational
-information from across North America and Europe, making it one of the
-most comprehensive collections of building-related data available for
-scientific use. The dataset includes various commercial building types
-such as offices, education facilities, public, and retail buildings, and
-provides detailed information on their physical characteristics (e.g.
-enregy ratings, heating types and floor area) and hourly measurements of
-chilled and hot water, electricity, gas usage as well as site outdoor
-weather conditions.
+contains over 1000 buildings' metadata and realistic operational
+measurements between 2016 and 2017 from across North America and Europe,
+making it one of the most comprehensive collections of building-related
+data available for scientific use. The dataset includes various
+commercial building types such as offices, education facilities, public,
+and retail buildings, and provides detailed information on their
+physical characteristics (e.g. energy ratings, heating types and floor
+area) and hourly measurements of chilled and hot water, electricity, gas
+usage as well as site outdoor weather conditions. In this study, we
+mostly use the metered electricity and outdoor weather measurements for
+running the M&V methods.
 
@@
 
-Most research related to M&V for whole-building approach focuses on the
-accuracy of baseline modeling, exploring model performance from simple
-regression models to more complex machine learning techniques. One study
-reviewed various models suitable for M&V applications as well as
-selected input features [@alrobaie_review_2022] and another study
+ASHRAE Guidelinge 14 and IPMVP provides options for whole building M&V
+use cases. For code compliance using prescriptive option, 12 months of
+baseline and 12 months of postretrofit measurements are required. In
+addition, the expected savings should be larger than 10% and the
+baseline model fitting accuracy should have CV(RMSE) lower than 20%.
+Thus, most research related to M&V for whole-building approach focuses
+on the accuracy of baseline modeling, exploring model performance from
+simple regression models to more complex machine learning techniques.
+One study reviewed various models suitable for M&V applications as well
+as selected input features [@alrobaie_review_2022] and another study
 provided a definitive methodology to apply machine learning models for
@@
 investigated the critical performance metrics to evaluate the developed
-baseline models [@granderson_automated_2015,
+baseline models [@granderson_automated_2015;
 @granderson_development_2014] and compared a variety of models using
@@
 leveraging statistical formulation and inference to improve baseline
-energy models [@burkhart_measurement_2014, @heo_calibration_2012,
+energy models [@burkhart_measurement_2014; @heo_calibration_2012;
 @walter_uncertainty_2014]. However, a gap still remains in the
@@
 
-A common non-routine event in energy-saving M&V projects is a change in
-occupancy or a significant shift in occupant behavior. These changes can
-greatly affect measured energy consumption in buildings and are
-typically unrelated to the intervention strategy. For instance, during
-the COVID-19 pandemic in 2020, most commercial buildings were
-unoccupied, leading to a noticeable drop in energy bills despite no
-energy-efficiency measures being implemented [@cai_impact_2024,
-@chihib_impact_2021, @gaspar_assessing_2022, @kang_changes_2021].
-Furthermore, subsequent research has shown that hybrid working modes,
-allowing employees to work remotely, have persisted after the outbreak
-of the pandemic [@aksoy_working_2022], adding further complexity to
-energy consumption patterns due to evolved occupant behaviors with
-higher flexibility [@gui_impact_2021, @mantesi_office_2022,
-@xie_does_2021]. One study realized the limitation of current M&V
-methods, which only consider adjusting for outdoor weather, is
-insufficient and emphasized the importance of requiring matched
-comparison groups to control for exogenous factors beyond weather
-differences when comparing between baseline and intervention
-[@demand_side_analytics_population_2022]. Another type of non-routin
-event is filter clogging in air handling units due to particle
-accumulation. This can cause supply fans to gradually consume more
-energy to maintain required duct static pressure [@feng_newly_2019,
-@zhai_full-scale_2017]. If M&V analysts are unaware of such changes and
-lack an appropriate adjustment method (e.g., replacing filters before
-the intervention begins), the savings could be underestimated as
-increased energy use is incorrectly attributed to the intervention
-rather than the mechanical issue.
+Non-routine events during M&V commonly refers to unexpected changes in a
+building that influences its energy usage. These changes can greatly
+affect measured energy consumption in buildings and are typically
+unrelated to the intervention strategy. Therefore, those changes are
+considered 'static factors' respect to measured independent variables. A
+common non-routine event in energy-saving M&V projects is a change in
+occupancy or a significant shift in occupant behavior, equipment run
+time, and operating conditions (e.g. set points, lighting and
+ventilation levels). However, current standards or guidelines only
+provides general approach for consideration. For example, IPMVP requires
+facility owner and the M&V analyst periodically perform inspections of
+all equipment and operations during reporting period, which is labor
+intensive and error-prone. ASHRAE guideline 14 recommends performing
+engineering calculations or computer software simulations to adjust
+post-retrofit baseline. Additional, it is very rare to have access to
+all measurements such as occupancy, and thus the analyst normally assume
+those factors remain unchanged throughout the study. Specifically,
+studies investigating the effects of demand response on building energy
+efficiency commonly uses linear interpolation to estimate
+counter-factual baseline [@beil_round-trip_2015;
+@keskar_commercial_2020], and one study points out that it is inaccurate
+to assume no change in the operating conditions during the test period
+[@huang_experimental_2023]. One likely encountered non-routine event due
+to operation condition change is filter clogging in air handling units
+due to particle accumulation. This can cause supply fans to gradually
+consume more energy to maintain required duct static pressure
+[@feng_newly_2019; @zhai_full-scale_2017]. If M&V analysts are unaware
+of such changes and lack an appropriate adjustment method (e.g.,
+replacing filters before the intervention begins), the savings could be
+underestimated as increased energy use is incorrectly attributed to the
+intervention rather than the mechanical issue. Some studies realized the
+limitation of current M&V methods, which only consider adjusting for
+outdoor weather, is insufficient and emphasized the importance of
+requiring matched comparison groups to control for exogenous factors
+beyond weather differences when comparing between baseline and
+intervention [@demand_side_analytics_population_2022;
+@huang_experimental_2023]. However, those methods are still unable to
+accurately quantify uncertainties due to changing baseline.
 
@@
 As mentioned, the goal of an M&V project is to determine the
-effect---typically energy savings---of an energy-efficient intervention.
-In this study, we focus primarily on switchable interventions, which
-often involve control retrofits. An example of such an intervention is a
+effect---typically energy savings---of an switchable energy-efficient
+intervention. The goal is to determine, for a large sample of buildings,
+how much more accurately the novel M&V method would estimate the savings
+of such an intervention compared to conventional M&V, and how much more
+quickly it would reach a result. An example of such an intervention is a
 control retrofit developed by a software-as-a-service company that
-adjusts the chilled water plant's supply water temperature based on
-outdoor weather conditions [@roa_field_2023,
-@lee_simulationoptimization_2012, @qiu_chilled_2022]. Therefore, we
-defined the M&V scenario as follows:
-
-*"A company aims to sell its supply temperature reset control software
-package to a customer, such as a building owner, with a guarantee that
-it will reduce the building's electricity usage. If the building owner
-decides to purchase the service, the company agrees to charge a service
-fee based on a percentage of the measured energy savings."*
+adjusts the chilled water plant's supply water temperature, which can be
+commonly found in the literature [@duarte_field_2023;
+@lee_simulationoptimization_2012; @qiu_chilled_2022; @jin2007;
+@taylor_optimizing_2012]. In our case, we make it even simpler by
+adjusting setpoint based on outdoor weather conditions with more
+description shown in Section \@ref(apply-virtue-intervention).
+Therefore, we defined the M&V scenario as follows:
+
+"A company aims to sell its supply temperature reset control software
+package to a customer, in this case, the building owner, with a
+promotion that it will reduce the building's electricity usage. If the
+building owner decides to purchase the service, the company agrees to
+charge a service fee based on a percentage of the measured energy
+savings."
 
@@
 conventional and the novel randomized M&V methods by estimating the
-intervention energy savings for all valid buildings in the dataset. By
-conducting such analysis, we aim to:
+intervention energy savings for all valid buildings in the BG2 dataset.
+By conducting such analysis, we aim to:
 
@@
     types and climate zones. The comparison metrics include both
-    estimation accuracy and M&V finishing timeline.
+    estimation accuracy and M&V completion timeline.
 
@@
     various sources of noise, we aim to reflect the challenges faced by
-    building analysts in real projects. As will be demonstrated in the
-    following sections, the randomized approach is less impacted by
-    non-routine events (i.e., measurement 'noise'), resulting in more
-    reliable energy savings estimates.
-
-3.  Demonstrate the implementation of the proposed randomized M&V method
-    using a public available dataset. We ensured the reproducibility of
-    the method by making the analysis code open source including
-    randomized schedule generation, sequential statistical analysis,
-    energy modeling and normalized saving calculation. Using the
+    building analysts in real projects. As we will perform numercial
+    simulation on the existing dataset, the ground-truth savings can be
+    calculated as the reference for comparison.
+
+3.  Open-source implementation of the proposed randomized M&V method
+    using a public dataset. We ensured the reproducibility of the method
+    by making the analysis code open source including randomized
+    schedule generation, sequential statistical analysis, energy
+    modeling and normalized saving calculation. In addition, we also
+    included codes for extended use cases of the randomized method, such
+    as changing sampling ratio and sampling intervals. Using the
     available open resources, building analysts should be able to
@@
 
-```{r flowchart, fig.cap = "Flow chat showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="50%"}
+```{r flowchart, fig.cap = "Flow chat showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="75%"}
 
@@
 
-## Data filtering
+## Filter and clean dataset
 
@@
 
-4.  Warehouse and parking types are excluded: target buildings have less
-    demand flexibility to implement a chilled water setpoint reset
+4.  Electric EUI \< 750 kWh/$m^2$: excludes buildings at the top 5%
+    electrical energy usage intensity according to the statistics
+    provided by the Building Performance Database
+    [@lawrenceberkeleynationallab; @mathew_big-data_2015].
+
+5.  Warehouse and parking types are excluded: target buildings have less
+    demand flexibility to implement a chilled water set point reset
     control.
 
-```{r stablesite, fig.cap = "Site summary of the stable building subset (counts < 2 are omitted for visualization; left: aggragated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}
+```{r stablesite, fig.cap = "Site summary of the stable building subset (counts < 10 are omitted for visualization; left: aggragated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}
 
@@
   geom_col() +
-  geom_text(aes(label = ifelse(n > 10, as.character(n), "")), color = "black", position = position_stack(vjust = 0.5)) +
+  geom_text(aes(label = ifelse(n > 10, n, "")), color = "black", position = position_stack(vjust = 0.5)) +
   scale_y_continuous(expand = c(0, 0),
@@
 stable_set <- df_energy_stable %>% 
-  group_by(type, site) %>%
+  left_join(df_loc, by = "site") %>% 
+  group_by(type, location) %>%
   distinct(name) %>%
@@
 variable_set <- df_energy_variable %>% 
-  group_by(type, site) %>% 
+   left_join(df_loc, by = "site") %>% 
+  group_by(type, location) %>%
   distinct(name) %>%
@@
 p2 <- variable_set %>% 
-  left_join(stable_set, by = c("type", "site")) %>% 
+  left_join(stable_set, by = c("type", "location")) %>% 
   mutate(stable = replace_na(stable, 0)) %>% 
-  group_by(site, type) %>% 
+  group_by(location, type) %>% 
   summarise(n = stable + variable) %>% 
   ungroup() %>% 
-  left_join(df_loc, by = "site") %>% 
   group_by(location) %>%
@@
   mutate(type = fct_reorder(type, order, .desc = F)) %>%
-  ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = type)) +
+  ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3.2, fill = type)) +
   geom_rect() +
@@
   scale_fill_manual(values = type_colors) +
-  geom_text(aes(x = 3.5, y = label_pos, label = ifelse(n > 10, as.character(n), "")), check_overlap = T) +
+  geom_text(aes(x = 3.7, y = label_pos, label = ifelse(n > 10, n, ""))) +
   geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black", check_overlap = T) +
@@
 For all qualified buildings, we further added one more stringent
-criterion to label all buildings with very stable electricity usage
+criterion to explore all buildings with very stable electricity usage
 between the two years:
@@
 1.  No statistical significant difference (P-value \> 0.05) between the
-    two-year electricity usage.
+    two-year electricity usage, which are sometimes assumed to be the
+    case for whole-building measurement and verification.
 
@@
 \@ref(non-routine-events-impact), such variability is largely associated
-with non-routine events and according to the statistics, is more
+with non-routine events and according to the statistics shown here, is
 commonly observed in real applications.
 
-## Control intervention implementation
+## Apply virtue intervention
 
@@
 outdoor weather conditions, which can be commonly found in the
-literature [@lee_chilled_2022, @congradac_recognition_2012]. For both
+literature [@lee_chilled_2022; @congradac_recognition_2012]. For both
 strategies, we assume that the chiller is activated when the outdoor
@@
                      breaks = c(4, 6, 8, 10, 12),
-                     limits = c(5.5, 12.5),
+                     limits = c(5, 12.5),
                      labels = number_format(suffix = " °C")) +
   scale_x_continuous(expand = c(0, 0), 
-                     breaks = c(10, 15, 20, 25, 30), 
+                     breaks = c(5, 10, 15, 20, 25, 30), 
                      labels = number_format(suffix = " °C")) +
+  geom_vline(xintercept = 10, lty = "dashed", color = "red") +
+  annotate(geom = "text", x = 7.5, y = 8, label = "No chiller operation\nbelow 10 °C") +
+  coord_cartesian(xlim = c(5, 30)) +
   labs(x = "Outdoor temperature", 
@@
        color = NULL, 
-       title = "Intervention logic: chilled water supply temperature reset") +
+       title = "Chilled water supply temperature reset intervention") +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
 below. For simplicity, those parameters were not rigorously calibrated
-for each building, and were used uniformly across the dataset.
+for each building and were applied uniformly in the filtered dataset.
+Interested readers can also change the parameters in our open-source
+code to simulate different scenarios.
 
 |                        |                                                                  |       |
-|------------------------|------------------------------------------------------------------|-------|
+|--------------|--------------------------------------------|--------------|
 | Parameter              | Description                                                      | Value |
@@
 
-## M&V methods comparison
+## Run M&V methods
 
 We described more in detail the workflow of both conventional and
-randomized M&V methos in the previous study [@RAFTERY2024111134].
-
-In a similar style, we summarized the experimental design as:
-
--   The HVAC system operates from 06:00 to 22:00 each day, so we use a
-    daily sampling interval with the sampling time at midnight each day.
+randomized M&V methods in the previous study [@raftery2024]. The
+conventional method is a simple pre- and post- comparison on a 12-month
+baseline and 12-month intervention timeline. The randomized method
+defines sampling requirements as:
+
+-   Use a daily sampling interval with the sampling time at midnight
+    each day.
 
@@
 
-and stopping criteria:
+and stopping criteria as:
 
@@
 
--   At least 80% of the drybulb temperature range in the annual TMY data
-    sampled by both strategies.
+-   At least 80% of the dry-bulb temperature range in the annual TMY
+    data sampled by both strategies.
 
@@
 
--   As no baseline data is available, test with an equal sampling ratio
-    (50% baseline, 50% intervention).
+-   test with an equal sampling ratio (50% baseline, 50% intervention).
 
@@
 
-```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both at the stopping criteria and over a two-year period).", fig.height=12, fig.width=10}
+The accuracy ($Acc$) is calculated as: $$
+Acc = \frac{E_{baseline} - E_{intervention}}{E_{baseline}}-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
+$$where $E$ indicates true electricity usage and $E'$ indicates the
+estimated electricity usage either through the conventional method or
+the new randomized method. We present the distribution of savings
+estimation accuracy using boxen plots, also known as letter plots, which
+is an advanced variation of the box plot designed to extend beyond the
+interquartile range (IQR) by progressively dividing the data into
+smaller percentiles, revealing more detail in the tails of the
+distribution. We set the division parameter to k = 4, meaning the entire
+box area represents the data distribution from 6.25% to 93.75%. As the
+steps move closer to the center line (50% median), the distribution
+range progressively narrows with the next step represents 12.5% to
+87.5%, followed by 25% to 75%. In addition, for each boxen plot, we also
+showed the mean value of the distribution on top.
+
+```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both at the stopping criteria and over a two-year period).", fig.height=12, fig.width=12}
 
@@
   geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
-  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
+  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff, na.rm = T)) %>% ungroup(), 
             aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
@@
 
+p_accuracy <- ggarrange(p_top, p_middle, 
+                        ncol = 1, nrow = 2,
+                        labels = c("a)", "b)"),
+                        common.legend = T, 
+                        legend = "bottom")
+
 # Timeline plot
@@
 
-p_bottom <- count %>% 
+p_timeline <- count %>% 
   ggplot() +
@@
 
-ggarrange(p_top, p_middle, p_bottom, 
-          ncol = 1, nrow = 3,
-          heights = c(1.2, 1.2, 1),
-          labels = c("a)", "b)", "c)"),
-          common.legend = T, 
+ggarrange(p_accuracy, p_timeline, 
+          ncol = 1, nrow = 2,
+          heights = c(2.2, 1),
+          labels = c("", "c)"),
+          common.legend = F, 
           legend = "bottom") +
@@
 Figure \@ref(fig:abstract) shows the overall results of M&V methods
-comparison. The estimation error is calculated as the absolute deviation
-from the true savings. Subplot a) shows the savings estimated from
-measured weather condition and subplot b) shows the savings normalized
-on typical meteorological weather file using TOWT model prediction. In
-average, the conventional method could deviate 6% in saving estimation
-while the the deviation from our proposed randomized method is much
-smaller. If the analyst stops right after satisfying all stopping
-criteria, then the deviation is around 2% and it further decreases to 1%
-when the M&V extends over the same period of the convention method. By
-comparing subplot a) and b), we noticed a slight improvement in
-estimation accuracy in model fitting, which is less than 1% on average.
-This implies that baseline projection to post-retrofit period through
-data-driven models is inefficient when the actual baseline changes.
-Subplot c) shows the overall timeline of the two methods. It is clear
-that the all buildings satisfied all stopping criteria in 36 weeks and
-given by the difference in weather conditions, a large portion of the
-buildings can get an accurate M&V results in 24 weeks. Therefore, on
-average a building analyst should expect 1% accuracy improvement over 1
-year extension of the randomized M&V method. However, the red arrow also
-implies the excess time required by the conventional M&V method that
-only leads to more unreliable results.
+comparison. Subplots a) and b) calculate the savings estimation error as
+the absolute deviation from the true savings (i.e. $|Acc|$) and shows
+the results distribution using conventional M&V method in the first
+column, the randomized M&V method that stops after satisfying all
+stopping criteria in the second column, and the randomized M&V method
+continues 50%/50% sampling throughout 2 years in the third column. In
+addition, subplot a) shows the savings estimated from measured weather
+condition and subplot b) shows the savings normalized on typical
+meteorological weather file after using TOWT model. As shown clearly in
+the plots, the conventional M&V method exhibits an average deviation of
+6% in savings estimation, whereas the proposed randomized method
+demonstrates significantly smaller deviations. If the analyst stops
+immediately after satisfying all stopping criteria, the deviation is
+reduced to approximately 2%. Extending the M&V period to match the
+conventional method further improves accuracy to 1%. Furthermore, by
+comparing subplot (a) and (b), we observe a minor improvement (less than
+1%) in estimation accuracy due to model fitting. This is because the
+weather condition over the two years for all selected locations are
+similar. Thus the effect of adjusting temperature as an independent
+variable in the model is not significant. Overall the results suggest
+that projecting the baseline into the post-retrofit period that only
+considers weather variation is inefficient. Subplot (c) presents the
+overall timeline comparison of the two M&V methods. The results indicate
+that a significant portion achieve accurate M&V results within 24 weeks
+and all buildings satisfy the stopping criteria within 36 weeks, mostly
+influenced by weather variability. Meanwhile, the red arrow highlights
+the excess time required by the conventional M&V method, which only
+leads to less reliable results.
+
+We also plotted the mean error in Figure \@ref(fig:meanerr) and we
+noticed that the uncertainty range associated with the conventional
+method is significantly larger compared to the randomized method that
+only takes 36 weeks according to subplot c) in the Figure
+\@ref(fig:abstract).
+
+```{r meanerr, fig.cap="Distribution of mean deviation from the ground-truth savings calculated by the two M&V methods", fig.height=6, fig.width=12}
+# plot for the stable subset
+rand_eob_S <- df_seq_fs_stable %>% 
+  filter(seq == "eob") %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = savings - fs, 
+         method = "rand_eob") %>% 
+  select(name, method, diff)
+
+conv_S <- df_fs_stable %>% 
+  filter(method != "rand") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = true - conv, 
+         method = "conv") %>% 
+  select(name, method, diff)
+
+rand_final_S <- df_fs_stable %>% 
+  filter(method != "conv") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = true - rand, 
+         method = "rand_final") %>% 
+  select(name, method, diff)
+
+df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)
+
+# plot for the variable subset
+rand_eob_V <- df_seq_fs_variable %>% 
+  filter(seq == "eob") %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site"))%>% 
+  mutate(diff = savings - fs, 
+         method = "rand_eob") %>% 
+  select(name, method, diff)
+
+conv_V <- df_fs_variable %>% 
+  filter(method != "rand") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = true - conv, 
+         method = "conv") %>% 
+  select(name, method, diff)
+
+rand_final_V <- df_fs_variable %>% 
+  filter(method != "conv") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = true - rand, 
+         method = "rand_final") %>% 
+  select(name, method, diff)
+
+df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)
+
+# plot for combined dataset
+rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
+  select(name, rand_eob_diff = diff)
+
+conv_A <- bind_rows(conv_V, conv_S) %>% 
+  select(name, conv_diff = diff)
+
+rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
+  select(name, rand_final_diff = diff)
+
+df_MW_A <- rand_eob_A %>% 
+  left_join(conv_A, by = "name") %>% 
+  left_join(rand_final_A, by = "name") %>% 
+  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")
+
+p_top <- df_MW_A %>% 
+  mutate(method = as.factor(method), 
+         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
+  ggplot(aes(x = method, y = diff, fill = method)) +
+  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
+  geom_lv(k = 4, outlier.shape = NA) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
+  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
+            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
+  labs(fill = NULL, 
+       color = NULL, 
+       x = NULL, 
+       y = "Mean error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with measured weather")) +
+  coord_cartesian(ylim = c(-20, 20)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "none",
+        axis.text.x = element_blank(),
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+# TMY
+# plot for the stable subset
+rand_eob_S <- df_sprt_all_stable %>% 
+  filter(seq == "eob") %>% 
+  select(-c(seq, n_weeks)) %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = savings - annual, 
+         method = "rand_eob") %>% 
+  select(name, method, diff)
+
+conv_S <- df_fs_tmy_stable %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = savings - conv, 
+         method = "conv") %>% 
+  select(name, method, diff)
+
+rand_final_S <- df_fs_tmy_stable %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = savings - rand, 
+         method = "rand_final") %>% 
+  select(name, method, diff)
+
+df_TW_S <- rbind(rand_eob_S, conv_S, rand_final_S)
+
+# plot for the variable subset
+rand_eob_V <- df_sprt_all_variable %>% 
+  filter(seq == "eob") %>% 
+  select(-c(seq, n_weeks)) %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = savings - annual, 
+         method = "rand_eob") %>% 
+  select(name, method, diff)
+
+conv_V <- df_fs_tmy_variable %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = savings - conv, 
+         method = "conv") %>% 
+  select(name, method, diff)
+
+rand_final_V <- df_fs_tmy_variable %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = savings - rand, 
+         method = "rand_final") %>% 
+  select(name, method, diff)
+
+df_TW_V <- rbind(rand_eob_V, conv_V, rand_final_V)
+
+# plot for combined dataset
+rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
+  select(name, rand_eob_diff = diff)
+
+conv_A <- bind_rows(conv_V, conv_S) %>% 
+  select(name, conv_diff = diff)
+
+rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
+  select(name, rand_final_diff = diff)
+
+df_TW_A <- rand_eob_A %>% 
+  left_join(conv_A, by = "name") %>% 
+  left_join(rand_final_A, by = "name") %>% 
+  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")
+
+p_middle <- df_TW_A %>% 
+  mutate(method = as.factor(method), 
+         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
+  ggplot(aes(x = method, y = diff, fill = method)) +
+  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
+  geom_lv(k = 4, outlier.shape = NA) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
+  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
+            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
+  labs(fill = NULL, 
+       x = NULL, 
+       y = "Mean error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
+  coord_cartesian(ylim = c(-20, 20)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "none",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+ggarrange(p_top, p_middle, 
+          ncol = 1, nrow = 2,
+          labels = c("a)", "b)")) +
+  plot_annotation(title = "Mean error in savings estimation by different M&V methods")
+
+percentiles <- df_TW_A %>%
+  filter(method == "conv_diff") %>% 
+  summarise(
+    Q1     = round(quantile(diff, 0.25, na.rm = TRUE), digits = 2),
+    Median = round(quantile(diff, 0.50, na.rm = TRUE), digits = 2),
+    Q3     = round(quantile(diff, 0.75, na.rm = TRUE), digits = 2)
+  )
+```
+
+Additionally in the literature we found another M&V method study using
+over 500+ commercial buildings, but their study scope was to compare a
+variety of M&V baseline models from simple week mean to more complex
+machine learning methods using the normalized mean bias error
+(NMBE)[@granderson_accuracy_2016]. Since we also calculated the
+normalized savings estimation error, we compared our normalized accuracy
+of 600+ buildings with TOWT model and TMY weather with their TOWT model
+prediction assessment in the following table. As each percentile is
+closely matched, our results are verified.
+
+|                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
+|-----------------|--------------------------------------|-----------------|
+| 25th percentile | -5.85                                               | `r percentiles$Q1`     |
+| 50th percentile | -1.25                                               | `r percentiles$Median` |
+| 75th percentile | 3.86                                                | `r percentiles$Q3`     |
+
+: Table 1. Savings accuracy comparison with a similar literature
 
@@
 
-```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect", fig.height=9, fig.width=12}
+```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect", fig.height=6, fig.width=12}
 
@@
 
-As mentioned in Section \@ref(data-filtering), the majority of the
-selected buildings are affected by non-routine events. To assess the
-reliability of the two M&V methods, we repeated the M&V procedure on the
-dataset before applying the reset control. In other words, the
-intervention has no effect on building electricity usage and our
-objective is to compare which method can provide an estimation closer to
-0 savings. Figure \@ref(fig:nosaving) shows the randomized M&V method
-(both at the time of satisfying all stopping criteria and at the end of
-two-year period) produced consistent results. We included the absolute
-deviation in the appendix. Similarly, stopping early can lead to larger
-uncertainty similar to Figure \@ref(fig:abstract). As for conventional
-method, the uncertainty range is much larger than the randomized method
-and it shows on average the deviation is 1.3% meaning a building analyst
-can claim 1.3% energy savings were detected when there isn't any.
+As mentioned in Section \@ref(filter-and-clean-dataset), the majority of
+the selected buildings are affected by non-routine events, which is the
+reason why we observe baseline electricity usage changes throughout the
+two year. To evaluate the reliability of the two M&V methods in blocking
+the effect of non-routine events, we repeated the M&V run on the
+original dataset prior to the application of the reset control. In other
+words, as there is no intervention applied or the intervention is known
+to have no impact on whole-building electricity usage, a more reliable
+M&V method should detect closer to 0 kW savings using the measurements.
+Therefore, the $Acc$ here are calculated as:
+
+$$
+Acc = 0-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
+$$
+
+Figure \@ref(fig:nosaving) shows the randomized M&V method (both at the
+time of satisfying all stopping criteria and at the end of two-year
+period) produced consistent results. We included the absolute deviation
+($|ACC|$) in the appendix. Similarly, stopping early can lead to
+slightly larger uncertainty similar to Figure \@ref(fig:abstract).
+However, as for conventional method, the uncertainty range is much
+larger than the randomized method and it shows on average the deviation
+is 1.3% meaning a building analyst can 'detect' 1.3% energy savings when
+there isn't any using the conventional M&V method.
 
@@
 
-To assess modeling accuracy, we used the Coefficient of Variation of
-Root-Mean Squared Error, or CV(RMSE) as the error metric. Since this
-metric is calculated as a normalized value, it is useful to compare
-different model fitting results. Guideline 14 requires that
-whole-building baseline model fitting accuracy should maintain a
-CV(RMSE) lower than 30% [@ashrae_ashrae_2014]. In addition one study
-focusing on the baseline energy data-driven model fitting indicates that
-TOWT performs as accurate as other more advanced machine learning models
-[@granderson_accuracy_2016] and the calculated CV(RMSE) distribution for
-a large sample of commercial buildings indicates a median of 20%. Figure
-\@ref(fig:towtacc)shows the distribution of model fitting accuracy,
-calculated separately for the two measurement sets. The box plot
-provides a statistical summary of the accuracy data points, with error
-bars highlighting any points beyond the range, which should be
-considered outliers. The results show that the TOWT model's performance
-in this study aligns with, and is even slightly better than, the
-performance reported in the literature. The improved results can be
-attributed to relatively strict data filtering criteria for
-pre-processing described in Section \@ref(method).
+To evaluate modeling accuracy, we used the Coefficient of Variation of
+Root-Mean Squared Error (CV(RMSE)) as the primary error metric. Since
+CV(RMSE) is a normalized measure, it enables direct comparison across
+different model fitting results. According to ASHRAE Guideline 14,
+whole-building baseline models should achieve a CV(RMSE) below 25% for
+computing savings or 30% when less than 12 months' of measurements are
+available [@ashrae_ashrae_2023]. In our study, Figure \@ref(fig:towtacc)
+compares the model fitting accuracy of the two M&V methods and it shows
+around 87.5% of the models fitted in the randomized M&V method are code
+compliance while 80% for the conventional M&V method. As the quantity of
+the training set is the same, this highlights the impact of data
+sampling on model fitting accuracy, and for TOWT, the distribution of
+outdoor weather conditions sampled in the training dataset is
+influential. The conventional method trains the model using the entire
+first-year dataset, if the location is heating or cooling dominant, the
+data-driven model is trained using colder or warmer outdoor conditions
+leading to biased model coefficients. Although the random sampling in
+our study does not specifically block the outdoor temperature factor,
+the CV(RMSE) distribution shown in the figure indicates a significant
+improvement in model fitting accuracy.
 
@@
 
-df_model_acc_variable %>% 
-  mutate(group = "variable subset") %>% 
-  rbind(df_model_acc_stable %>% mutate(group = "stable subset")) %>% 
-  mutate(group = as.factor(group)) %>% 
-  ggplot(aes(x = group, y = cvrmse)) +
-  stat_boxplot(geom ='errorbar', width = 0.1) +
-  geom_boxplot(outlier.shape = NA, width = 0.25) +
-  geom_jitter(position = position_jitter(0.1), alpha = 0.1, size = 1.2) +
-  geom_text(aes(x = group, y = median, label = paste0(round(median, digits = 0), " %")), 
-            data = . %>% 
-              group_by(group) %>% 
-              summarise(median = median(cvrmse)) %>% 
+cvrmse <- df_model_acc_stable %>% 
+  bind_rows(df_model_acc_variable) %>% 
+  pivot_longer(c(conv, rand), 
+               names_to = "method", 
+               values_to = "cvrmse") %>% 
+  mutate(method = as.factor(method), 
+         method = recode_factor(method, 
+                              "rand" = "Randomized", 
+                              "conv" = "Conventional"))
+
+result <- t.test(filter(cvrmse, method == "Randomized")$cvrmse, 
+                 filter(cvrmse, method == "Conventional")$cvrmse, 
+                 exact = TRUE,
+                 alternative = "less",
+                 mu = 0,
+                 conf.int = TRUE, 
+                 conf.level = 0.90,
+                 var.equal = TRUE)$p.value
+
+cvrmse %>% 
+  ggplot(aes(x = method, y = cvrmse, fill = method)) +
+  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
+  geom_lv(k = 4, outlier.shape = NA) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
+  annotate(geom = "text", 
+           x = 1.5, 
+           y = 40, 
+           size = 4, 
+           label = paste0("P value = ", round(result, digits = 3))) +
+  geom_errorbarh(aes(xmin = 1.2, xmax = 1.8, y = 38), height = 2, linewidth = 0.25) +
+  geom_text(data = . %>% 
+              group_by(method) %>% 
+              summarise(mean = mean(cvrmse, na.rm = T)) %>% 
               ungroup(), 
-            position = position_nudge(x = 0.25)) +
+            aes(x = method, y = mean, label = paste0(round(mean, digits = 0), " %"))) +
+  geom_hline(yintercept = 25, color = "red", lty = "dashed") +
+  annotate(geom = "text", 
+           x = 0.5, 
+           y = 27, 
+           size = 4, 
+           color = "red", 
+           label = paste0(25, " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80")) + 
   scale_y_continuous(expand = c(0, 0), 
@@
 
-The limitation of regression models is that despite they capture well on
-the mean building energy consumption but can underestimate the 15-min
-daily peak load [@granderson2021], which is not useful for assessing
-savings for demand response events.
+Additionally, we also noticed in our study that despite the regression
+models capture well on the mean energy consumption of the building but
+the tend to underestimate the 15-min daily peak load [@granderson2021],
+which is not useful for assessing savings for demand response events.
 
@@
 
-```{r absinterv2, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy", fig.height=9, fig.width=12}
+In certain situations, particularly those involving large thermal mass
+in active thermal energy storage systems, there is a known time-lagged
+effect associated with the intervention. This means that the effect of
+one strategy may persist after switching to another. To mitigate such
+carryover effects, building analysts may choose to increase the sampling
+interval and drop non-consecutive days in the switchback experiment. For
+example, with a hot water tank that is charged by an intervention
+strategy, residual heat may still be present when the next strategy
+begins, influencing subsequent measurements, particularly the
+performance measured on the first day (i.e. a non-consecutive day). In
+such cases, increasing the sampling interval (e.g., sample strategies
+every 2 or 3 days instead of daily) and discard measurements from those
+non-consecutive days can eliminate potential carryover effect. However,
+this approach reduces the total data collected: for instance, sampling
+every 3 days would drop 1/6 of measurements and sampling every 2 day
+means would remove 1/4 of measurements. Therefore, by increasing the
+sampling interval, there is a reduction in statistical power due to less
+randomness, and when decreasing the sampling interval, there is a
+accuracy penalty from dropping more non-consecutive days. To quantify
+such trade-off, \@ref(fig:absinterval) shows the accuracy of different
+sampling intervals and consequence of dropping the non-consecutive days
+in the dataset over the entire two-year period.
+
+```{r absinterval, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy (dropped: all non-consecutive days were dropped; kept: all measurements were kept)", fig.height=9, fig.width=12}
 
@@
   geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
-  geom_text(data = . %>% group_by(group, interval) %>% summarise(mean = mean(diff, na.rm = T)) %>% ungroup(), 
+  geom_text(data = . %>% 
+              group_by(group, interval) %>%
+              summarise(mean = mean(diff, na.rm = T)) %>% 
+              ungroup(), 
             aes(x = interval, y = mean, group = group,
                 label = paste0(round(mean, digits = 1), " %")), 
-            position = position_dodge(width = 0.75)) +
+            position = position_dodge(width = 0.75), 
+            color = "white") +
   scale_y_continuous(expand = c(0, 0), 
@@
               filter(scenario == "ref" & method == "true")), by = "name") %>% 
-  mutate(drop_1 = abs(drop_1 - savings), 
-         drop_2 = abs(drop_2 - savings), 
-         drop_3 = abs(drop_3 - savings), 
-         drop_7 = abs(drop_7 - savings), 
-         keep_1 = abs(keep_1 - savings), 
-         keep_2 = abs(keep_2 - savings), 
-         keep_3 = abs(keep_3 - savings), 
-         keep_7 = abs(keep_7 - savings)) %>% 
+  mutate(drop_1 = abs(savings - drop_1),
+         drop_2 = abs(savings - drop_2), 
+         drop_3 = abs(savings - drop_3), 
+         drop_7 = abs(savings - drop_7), 
+         keep_1 = abs(savings - keep_1),
+         keep_2 = abs(savings - keep_2), 
+         keep_3 = abs(savings - keep_3), 
+         keep_7 = abs(savings - keep_7)) %>% 
   pivot_longer(cols = contains("drop") | contains("keep"),  
@@
   geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
-  geom_text(data = . %>% group_by(group, interval) %>% summarise(mean = mean(diff, na.rm = T)) %>% ungroup(), 
+  geom_text(data = . %>% 
+              group_by(group, interval) %>% 
+              summarise(mean = mean(diff, na.rm = T)) %>% 
+              ungroup(), 
             aes(x = interval, y = mean, group = group,
                 label = paste0(round(mean, digits = 1), " %")), 
-            position = position_dodge(width = 0.75)) +
+            position = position_dodge(width = 0.75), 
+            color = "white") +
   scale_y_continuous(expand = c(0, 0), 
@@
 
-```{r absinterv, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy", fig.height=9, fig.width=12}
-
-# sprt sequence plot for different sampling intervals
-df_seq_interval_fs_keep_stable %<>% 
-  select(-c(sprt, temp, final)) %>% 
-  bind_rows(df_seq_fs_stable %>% 
-              filter(seq == "eob") %>% 
-              select(-seq) %>% 
-              mutate(interval = 1) %>% 
-              rename(eob = "fs"))
-
-df_seq_interval_fs_keep_variable %<>% 
-  select(-c(sprt, temp, final)) %>% 
-  bind_rows(df_seq_fs_variable %>% 
-              filter(seq == "eob") %>% 
-              select(-seq) %>% 
-              mutate(interval = 1) %>% 
-              rename(eob = "fs"))
-
-df_drop <- df_seq_interval_fs_drop_stable %>% 
-  bind_rows(df_seq_interval_fs_drop_variable) %>% 
-  select(name, eob, interval) %>% 
-  mutate(group = "drop")
-
-df_keep <- df_seq_interval_fs_keep_stable %>% 
-  bind_rows(df_seq_interval_fs_keep_variable) %>% 
-  select(name, eob, interval) %>% 
-  mutate(group = "keep")
-
-df_MW_A <- df_drop %>% 
-  bind_rows(df_keep) %>% 
-  left_join(rbind(df_fs_stable %>% 
-                    filter(scenario == "ref" & method == "true"), 
-                  df_fs_variable %>% 
-                    filter(scenario == "ref" & method == "true")), by = "name") %>% 
-  mutate(diff = abs(eob - savings)) %>% 
-  filter(is.finite(diff))
-
-p_top <- df_MW_A %>% 
-  mutate(interval = as.factor(interval), 
-         interval = recode_factor(interval, 
-                                  "1" = "1-day\nsampling",
-                                  "2" = "2-day\nsampling", 
-                                  "3" = "3-day\nsampling", 
-                                  "7" = "7-day\nsampling"), 
-         group = as.factor(group), 
-         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
-   ggplot(aes(x = interval, y = diff, fill = group)) +
-  geom_jitter(alpha = 0.8, 
-              size = 0.5, 
-              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
-  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8) +
-  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
-  geom_text(data = . %>% group_by(group, interval) %>% summarise(mean = mean(diff, inf.rm = T)) %>% ungroup(), 
-            aes(x = interval, y = mean, group = group,
-                label = paste0(round(mean, digits = 1), " %")), 
-            position = position_dodge(width = 0.75)) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 4), 
-                     labels = number_format(suffix = " %")) +
-  scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = c("grey80", "grey80")) + 
-  labs(fill = NULL, 
-       color = NULL, 
-       x = NULL, 
-       y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with measured weather")) +
-  coord_cartesian(ylim = c(0, 20)) +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.position = "bottom",
-        axis.text.x = element_blank(),
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-# with tmy version
-df_seq_interval_nm_keep_stable %<>% 
-  select(-c(sprt, temp, final)) %>% 
-  bind_rows(df_sprt_all_stable %>% 
-             filter(seq == "eob") %>% 
-              select(-c(seq, n_weeks)) %>% 
-              mutate(interval = 1) %>% 
-              rename(eob = "annual"))
-
-df_seq_interval_nm_keep_variable %<>% 
-  select(-c(sprt, temp, final)) %>% 
-  bind_rows(df_sprt_all_variable %>% 
-              filter(seq == "eob") %>% 
-              select(-c(seq, n_weeks)) %>% 
-              mutate(interval = 1) %>% 
-              rename(eob = "annual"))
-
-df_drop <- df_seq_interval_nm_drop_stable %>% 
-  bind_rows(df_seq_interval_nm_drop_variable) %>% 
-  select(name, eob, interval) %>% 
-  mutate(group = "drop")
-
-df_keep <- df_seq_interval_nm_keep_stable %>% 
-  bind_rows(df_seq_interval_nm_keep_variable) %>% 
-  select(name, eob, interval) %>% 
-  mutate(group = "keep")
-
-df_TW_A <- df_drop %>% 
-  bind_rows(df_keep) %>% 
-  left_join(rbind(df_fs_stable %>% 
-                    filter(scenario == "ref" & method == "true"), 
-                  df_fs_variable %>% 
-                    filter(scenario == "ref" & method == "true")), by = "name") %>% 
-  mutate(diff = abs(eob - savings)) %>% 
-  filter(is.finite(diff))
-
-p_bottom <- df_TW_A %>% 
-  mutate(interval = as.factor(interval), 
-         interval = recode_factor(interval, 
-                                  "1" = "1-day\nsampling",
-                                  "2" = "2-day\nsampling", 
-                                  "3" = "3-day\nsampling", 
-                                  "7" = "7-day\nsampling"), 
-         group = as.factor(group), 
-         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
-  ggplot(aes(x = interval, y = diff, fill = group)) +
-  geom_jitter(alpha = 0.8, 
-              size = 0.5, 
-              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
-  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8) +
-  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
-  geom_text(data = . %>% group_by(group, interval) %>% summarise(mean = mean(diff, na.rm = T)) %>% ungroup(), 
-            aes(x = interval, y = mean, group = group,
-                label = paste0(round(mean, digits = 1), " %")), 
-            position = position_dodge(width = 0.75)) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 4), 
-                     labels = number_format(suffix = " %")) +
-  scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = c("grey80", "grey80")) + 
-  labs(fill = NULL, 
-       color = NULL, 
-       x = NULL, 
-       y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with TMY weather")) +
-  coord_cartesian(ylim = c(0, 20)) +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.position = "bottom",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-ggarrange(p_top, p_bottom, 
-          ncol = 1, nrow = 2, 
-          align = "hv", 
-          common.legend = T, 
-          legend = "bottom") +
-  plot_annotation(title = "Absolute error in savings estimation of randomized M&V after satisfying all stopping criteria")
-```
+To quantify the impact of reduced randomness, all measurements are kept
+when calculating the distribution shown in the darker colored set. The
+results suggests a gradual increase in error from 1% to 1.8% when
+sampling from daily to weekly intervals. To quantify the combined impact
+of data loss due to preventing carryover effect, we generated the
+lighter colored set, which drops all non-consecutive days. As a result,
+we noticed a significant increase in estimaion error when sampling
+daily, which lead to a similar accuracy compared to weekly intervals.
+Therefore, considering this trade-off, we recommend using a three-day
+sampling interval if the carryover effect is likely but expected to last
+less than one day (as more sampling is required for a two-day
+swithback), and then normalizing the estimation via energy modeling. We
+also included the mean deviation distribution and also the accuracy
+distribution calculated after all stopping criteria satisfied in the
+supplementary material and the results shows the same indication.
 
@@
 associated with the savings. Alternatively, they could switch to 100%
-intervention to maximize savings on the utility bill, though this
-approach risks the baseline becoming outdated. A middle-ground approach
-would be to sample at an 20%/80% ratio between the baseline and the
-intervention. Through this way, the software-as-a-service company can
-accurately model future baseline changes and adjust customers' bill
-accordinly with minimum baseline days sampled. To demonstrate, figure
-\@ref(fig:cont) shows after the analyst reports the randomized M&V
-results indicated in Figure \@ref(fig:timeline), a new schedule sampling
-at 20%/80% was implemented till the end of the year. Similarly, we
-calculated the accuracy metric as the absolute deviation from the annual
-true savings.
-
-The plot shows the deviation remains consistent in the stable subset
-comapred to Figure \@ref(fig:stablefrac) but increased slightly in the
-variable subset compare to to Figure \@ref(fig:variablefrac). With
-unbalanced sampling ratio, there is a trade-off for estimation accuracy.
-But compared with conventional method, this is still a preferred
-approach. Most importantly, this approach allows customers to realize a
-proportion of expected savings while documenting baseline measurement.
-For example, opting for an early stop at 24 or 36 weeks with 50%
-intervention, followed by re-sampling at 80% until the year's end
-(\~another 20 weeks), would enable customers to capture about 65% of the
-full-range savings.
+intervention to maximize cost savings, though this approach risks the
+baseline becoming outdated, which is another trade-off to consider. A
+middle-ground approach would be to sample at an 20%/80% ratio or 10%/90%
+between the baseline and the intervention. To compare those possible
+choices to a building owner, figure \@ref(fig:abscont) shows the results
+of changing the original 50%/50% sampling ratio after satisfying all
+stopping criteria and continuing a new ratio till the end of the
+two-year period. We also provided the reference case when the building
+owner continues with no change. The results shows as the sampled
+baseline days and intervention days become unbalanced, the savings
+estimation error grows from 1% to 5%. This is reasonable since sampling
+at 10%/90% is similar to the conventional M&V method. We included the
+mean deviation version in the supplementary material.
+
+```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy", fig.height=6, fig.width=12}
+
+#### ABS-CONT ####
+rand_cont_S <- df_cont_stable %>% 
+  select(-cont_tmy) %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - cont_fs)) %>% 
+  select(name, diff, ratio)
+
+rand_final_S <- df_fs_stable %>% 
+  filter(method != "conv") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = abs(true - rand), 
+         ratio = 0) %>% 
+  select(name, diff, ratio)
+
+df_MW_S <- rbind(rand_cont_S, rand_final_S)
+
+# plot for the variable subset
+rand_cont_V <- df_cont_variable %>% 
+  select(-cont_tmy) %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - cont_fs)) %>% 
+  select(name, diff, ratio)
+
+rand_final_V <- df_fs_variable %>% 
+  filter(method != "conv") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = abs(true - rand), 
+         ratio = 0) %>% 
+  select(name, diff, ratio)
+
+df_MW_V <- rbind(rand_cont_V, rand_final_V)
+
+# plot for combined dataset
+rand_cont_A <- bind_rows(rand_cont_V, rand_cont_S) %>% 
+  select(name, rand_cont_diff = diff)
+
+rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
+  select(name, rand_final_diff = diff)
+
+df_MW_A <- rbind(df_MW_S, df_MW_V)
+
+df_MW_A %>% 
+  mutate(ratio = as.factor(ratio), 
+         ratio = recode_factor(ratio, "0" = "Randomized\n(50/50)", "1" = "Randomized\n(20/80)", "2" = "Randomized\n(10/90)")) %>% 
+  ggplot(aes(x = ratio, y = diff, fill = "Randomized")) +
+  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
+  geom_lv(k = 4, outlier.shape = NA) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio)) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% 
+              group_by(ratio) %>% 
+              summarise(mean = mean(diff)) %>% 
+              ungroup(), 
+            aes(x = ratio, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
+  labs(fill = NULL, 
+       x = NULL, 
+       y = "Absolute error in fractional savings", 
+       title = "Absolute error in savings estimation by different sampling ratio",
+       subtitle = str_glue("All {A_building} buildings with measured weather conditions")) +
+  coord_cartesian(ylim = c(0, 15)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "none",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+```
 
@@
 
-1.  Application of control intervention: Although the primary focus is
-    on accurately detecting intervention effects rather than validating
-    their broader impact, the simulated intervention remains somewhat
-    generic considering the diverse building types and climate zones in
-    the BDG2 dataset. For simplicity, we applied the parameters listed
-    in Table 1 uniformly across all buildings. Yet, in some cases,
-    raising the water temperature by 1°C might yield more or less than
-    8% in electricity savings due to target buildings' demand
-    flexibility.
-2.  Design of the randomized switchback experiment: In this study, we
-    assumed a daily sampling interval would suffice for most commercial
-    buildings, but exceptions exist. Buildings with significant thermal
-    lag, such as those with heavy concrete construction, hot water
-    tanks, or Thermally Active Building Systems (TABS), may experience
-    carryover effects. In these cases, the impact of one day's control
-    strategy can influence subsequent measurements due to thermal
-    storage. For example, if chilled water is pre-charged in the thermal
-    mass at the end of a sampled intervention day and the control
-    swicthes to baseline at 12 AM the following day, the analyst would
-    observe lower energy consumption in that sampled baseline day. While
-    this study does not account for carryover effects, we recommend
-    using a 3-day sampling interval and excluding non-consecutive days
-    in practice to 'wash out' residual effects from previous strategies.
+1.  The simulated intervention remains somewhat generic and simplistic
+    considering the diverse building types and climate zones in the BDG2
+    dataset. As the primary focus is to accurately detect the
+    intervention (or any intervention) effect so for simplicity, we
+    applied the parameters listed in Table 1 uniformly across all
+    buildings. Yet, in some cases, raising the water temperature by 1°C
+    might yield more or less than 8% in electricity savings due to
+    target buildings' demand flexibility. For example, one field
+    validation of HVAC optimization shows additional savings might be
+    limited if other best practice controls are already implemented
+    [@granderson2018].
+2.  For the main analysis, we assumed no carryover effect and a daily
+    sampling interval would suffice for most commercial buildings, but
+    exceptions exist. Based on the measurement from the dataset, we have
+    limited knowledge of how builings thermal mass respond to the
+    proposed chilled water reset strategy. However, in the discussion,
+    we addressed such concern by comparing different sampling strategies
+    and the impact of dropping measurements from non-consecutive days.
 
@@
 virtual control retrofit case---resetting the chilled water setpoint
-based on outdoor weather conditions---and applied it to over 500
+based on outdoor weather conditions---and applied it to over 600
 filtered commercial buildings. By comparing the savings estimations of
@@
 
-Specifically, we showed that throughout 7 climate zones assessed in this
-study, the randomized M&V can provide a saving estimation by 36 weeks
-(with the majority finishes by 24 weeks) once all stopping criteria
-satisfied. In contrast, the conventional method requires a full range of
-baseline and intervention measurements under normal operating
+Specifically, we showed that throughout 11 different locations assessed
+in this study, the randomized M&V can provide a saving estimation by 36
+weeks (with the majority finishes by 24 weeks) once all stopping
+criteria satisfied. In contrast, the conventional method requires a full
+range of baseline and intervention measurements under normal operating
 conditions, typically taking 6-9 months for each phase. Most
 importantly, Furthermore, we verified that with reduced M&V timeline the
-randomized method can estimate savings more accurately than the two-year
-conventional method.
-
-We also evaluated the impact of non-routine events on the proposed M&V
-method by: 1) introducing a known change, such as occupancy-induced
-energy reductions, and 2) detecting when no intervention was applied in
-buildings with marginal energy consumption differences. In both cases,
-we demonstrated that baseline model fitting could be biased, while
-randomization effectively blocked confounding effects, ensuring the
-robustness of savings estimations.
-
-Although the limitation of the method is that it only applies to a
-subset of all M&V use cases (i.e., strategies that can be switched on
-and off), we believe its true value lies in the usefulness and
-convenience for most control retrofit validation in the field test.
+randomized method can estimate savings more accurately by showing that
+the absolute error is only 1 - 2 % while for the two-year conventional
+method, the estimation error could be 6% in a typical building. We also
+evaluated the impact of non-routine events on the proposed M&V method
+and the results show that baseline changes in the post retrofit period
+can deviate savings estimated using the conventional method. As a
+comparison, we found those events have very negligible impact on the
+savings estimated using the randomized method, demonstrating ideal
+reliability for realistic use cases.
+
+We also discussed the impact of model fitting accuracy when using the
+two methods and we found a significant improvement in our method due to
+random sampling. We also provide detailed assessment on using different
+sampling interval and sampling ratio, those results provide useful
+considerations to building analysts when generating randomized schedules
+for a variety of use cases.
+
+# CRediT authorship contribution statement
+
+**Aoyu Zou**: Conceptualization, Data curation, Formal analysis,
+Methodology, Investigation, Software, Writing - original draft,
+Writing - review & editing. **Paul Raftery**: Conceptualization, Formal
+analysis, Methodology, Investigation, Project administration,
+Supervision, Writing - review & editing. **Stefano Schiavon**:
+Conceptualization, Methodology, Investigation, Supervision, Writing -
+review & editing. **Carlos Duarte**: Methodology, Investigation,
+Supervision, Writing - review & editing. **Gail Brager**: Supervision,
+Writing - review & editing.
+
+# Reproducibility
+
+Reproducible example with analysis code is available (MIT license) at
+<https://github.com/CenterForTheBuiltEnvironment/genome_mnv>.
+
+# Declaration of competing interest
+
+The authors declare that they have no known competing financial
+interests or personal relationships that could have appeared to
+influence the work reported in this paper.
 

### 2025-03-01 code/1_analysis.R

@@ -55,10 +55,10 @@ source(paste0(function_path, "rand_seq.R"))
 
 # define parameters
 # section run control
-run_params <- list(type = "stable", 
+run_params <- list(type = "variable", 
                    sprt = T, 
                    sprt_cont = F, 
-                   interval = T, 
+                   interval = F, 
                    null = F)
 
 # Adding intervention effect as advanced chiller operation
@@ -315,7 +315,8 @@ df_tmy <- get_tmy(all_sites$site)
 #### INDIVIDUAL ####
 # storing results
 fs_ref <- list()
-model_acc <- list()
+model_acc_rand <- list()
+model_acc_conv <- list()
 seq_timeline <- list()
 seq_frsaving <- list()
 seq_nmsaving <- list()
@@ -396,7 +397,7 @@ for (n in 1:(nrow(all_names))){
     model_fit()
   
   df_towt <- df_base_conv %>%
-    filter(datetime < as.Date("2017-01-01")) %>%
+    filter(datetime >= as.Date("2017-01-01")) %>%
     select(time = datetime,
            temp = t_out,
            eload)
@@ -407,9 +408,9 @@ for (n in 1:(nrow(all_names))){
   
   cv_rmse <- mean(sqrt(base_proj$error ^ 2)) / mean(base_proj$eload) * 100
   
-  model_acc[[n]] <- tibble("name" = name,
-                           "site" = site, 
-                           "cvrmse" = cv_rmse)
+  model_acc_conv[[n]] <- tibble("name" = name,
+                                "site" = site, 
+                                "conv" = cv_rmse)
   
   # base_proj %>%
   #   ggplot() +
@@ -452,6 +453,7 @@ for (n in 1:(nrow(all_names))){
   base_proj <- model_pred(df_towt, towt_base) %>%
     rename("datetime" = "time")
   
+  fs_conv <- (mean(base_proj$towt) - mean(base_proj$eload)) / mean(base_proj$towt) * 100
   
   
   
@@ -482,9 +484,35 @@ for (n in 1:(nrow(all_names))){
     select(-eload_type) %>% 
     drop_na()
   
+  # separate baseline and intervention
+  towt_base <- df_rand %>% 
+    filter(strategy == 1) %>%
+    select(-strategy) %>%
+    model_fit()
+  
+  df_towt <- df_hourly_conv %>% 
+    left_join(df_schedule, by = "datetime") %>%
[…]

### 2025-03-01 code/graphical.R

@@ -410,12 +410,18 @@ p_middle <- df_TW_A %>%
   labs(fill = NULL, 
        x = NULL, 
        y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with TMY weather")) +
+       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
   coord_cartesian(ylim = c(0, 23)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
         legend.position = "none",
         plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
 
+p_accuracy <- ggarrange(p_top, p_middle, 
+                        ncol = 1, nrow = 2,
+                        labels = c("a)", "b)"),
+                        common.legend = T, 
+                        legend = "bottom")
+  
 # Timeline plot
 df_time <- df_sprt_all_variable %>% 
   filter(seq != "final") %>% 
@@ -448,7 +454,7 @@ for (i in seq(0, 36, by = 3)){
 
 count <- bind_rows(count) 
 
-p_bottom <- count %>% 
+p_timeline <- count %>% 
   ggplot() +
   geom_bar(data = . %>% 
              filter(n_weeks == 24 | n_weeks == 36),
@@ -498,15 +504,15 @@ p_bottom <- count %>%
         legend.position = "bottom",
         plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
 
-ggarrange(p_top, p_middle, p_bottom, 
-          ncol = 1, nrow = 3,
-          heights = c(1.2, 1.2, 1),
-          labels = c("a)", "b)", "c)"),
-          common.legend = T, 
+ggarrange(p_accuracy, p_timeline, 
+          ncol = 1, nrow = 2,
+          heights = c(2.2, 1),
+          labels = c("", "c)"),
+          common.legend = F, 
           legend = "bottom") +
   plot_annotation(title = "Overall comparison of conventional and randomized M&V")
 
-ggsave(filename = "abstract.png", path = fig_path, units = "in", height = 12, width = 10, dpi = 300)
+ggsave(filename = "abstract.png", path = fig_path, units = "in", height = 10, width = 12, dpi = 300)
 
 
 
@@ -807,7 +813,7 @@ df_MW_S <- rbind(rand_cont_S, rand_final_S)
 # plot for the variable subset
[…]

### 2025-03-22 code/manuscript.Rmd
Uploaded revised manuscript, added sampling interval timeline

@@
 abstract: Conventional measurement and verification (M&V) methods rely on pre- and post- retrofit comparison to estimate energy savings.
-They are often time-consuming and unreliable, especially when non-routine events such as reduced occupancy, or equipment failures occur during the M&V process.
+They are often time-consuming and unreliable, especially when non-routine events including static changes and gradual changes denoted as operational drift occurs during the M&V process.
 Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies conventional M&V suggested by industry guidelines, the results can be largely confounded.
-In this study, we demonstrated that switchable interventions, such as most of the control retrofits, can benefit from a new M&V method which randomly samples whether to implement the baseline or the intervention strategy each day.
+In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method which randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily).
 We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention.
-Our results show that compared to the conventional method, the randomized method provides faster and more accurate savings estimation.
+Our results show that compared to the conventional method, the randomized method provides faster and more accurate savings estimations.
 Additionally we found that when non-routine events are present, the randomized M&V approach estimates savings that are much closer to the ground-truth savings than the conventional M&V method, demonstrating much improved reliability.
+We also assessed the impact of dropping samples affected by carryover when switching between strategies, and identify the optimal sampling interval for that scenario using this large dataset.
 
@@
 # colors
-ls_colors <- c("Baseline" = "#c6dbef",
-               "Measured baseline" = "#c6dbef",
-               "Adjusted baseline" = "#c6dbef",
-               "Projected baseline" = "#2171b5",
+ls_colors <- c("Baseline" = "grey10",
                "Intervention" = "#fdbb84",
@@
                "Randomized" = "#99d8c9", 
-               "Randomized\n(24 months)" = "#66c2a4",
-               "Randomized\n(50/50)" = "#66c2a4",
+               "Randomized\n(24 months)" = "#6baed6",
+               "Randomized\n(50/50)" = "#6baed6",
+               "Randomized\n(20/80)" = "#6baed6", 
+               "Randomized\n(10/90)" = "#6baed6", 
                "Daily\nsampling" = "#66c2a4",
@@
                "Kept" = "#006d2c",
-               "Randomized\n(20/80)" = "#ccece6",
+               "Dropped (24 months)" = "#4292c6",
+               "Kept (24 months)" = "#08519c",
                "Buildings finishing randomized M&V" =  "#99d8c9",
@@
 become outdated due to changes in building performance caused by
-non-routine events unrelated to the intervention. This limitation
-reduces the feasibility of rapid M&V and complicates the quantification
-of estimation uncertainty, thus impacting the accuracy and timeliness of
-savings assessments.
+non-routine events unrelated to the intervention. Specifically, an M&V
+analyst should account for two types of baseline changes: (1) static
+changes, which are known building operational changes: such as
+renovation, equipment addition or removal; and (2) gradual changes,
+which often occur subtly and unknown, such as the incremental adoption
+of LED lighting or more efficient plug loads. Identifying and
+quantifying these changes are particularly labor-intensive and
+challenging. Consequently, those limitations reduce the feasibility of
+M&V and complicates the quantification of estimation uncertainty, thus
+impacting the accuracy and timeliness of savings assessments.
 
@@
 To address the limitations of conventional M&V methods and the
-challenges posed by non-routine events, we propose a novel M&V method
-that adopts the randomized crossover design, a gold standard from
-medical and agricultural studies [@blackston2019; @duan2013; @gupta2019;
-@mie2022; @munro; @raseduzzaman2017]. Another improvement is that we
-proposed a sequential evaluation framework and defined stopping criteria
-to end the M&V if the target effect is detected. This is to avoid
-unnecessary measurement collection over the full 2-year M&V cycle. The
-full framework is detailed in a previous published study with all
-stopping criteria and example usecases outlined [@raftery2024]. In
-summary, this method provides M&V analysts with a randomized schedule
+challenges posed by non-routine events, we proposed a M&V method that
+adopts the randomized crossover design [@raftery2024], a gold standard
+from medical, agricultural and online controlled experimental (i.e. A/B
+testing) studies [@blackston2019; @duan2013; @gupta2019; @mie2022;
+@munro; @raseduzzaman2017]. An early application of this randomized
+design to the field of M&V in buildings was first used in
+[@raftery2018], which compared the energy performance on two different
+supply air temperature reset strategies in a building. Subsequent
+improvements to the M&V method include adding a sequential evaluation
+framework and defined stopping criteria to end the M&V period early
+after reaching a target level of uncertainty, among other improvements,
+such as using blocked randomization by weekday. The full framework is
+detailed in a previously published study [@raftery2024] along with a
+case study application to a real building, and other example use cases.
+In summary, this method provides M&V analysts with a randomized schedule
 that alternates between baseline and intervention implementation while
@@
 switched on and off. However, for all applicable use cases, it allows
-analysts to detect energy savings sequentially right after test begins.
-In addition, once the desired savings target is achieved, analysts can
-terminate the M&V and switch to 100% intervention. The key advantage of
-randomization is that if control strategies are sampled with equal
-probability, the influence of other confounding factors such as
-occupancy change (or commonly known as non-routine events) is likely to
-be evenly distributed among measurements, leading to a more accurate and
-unbiased assessment of the intervention's impact.
+analysts to detect energy savings sequentially shortly after the test
+begins. In addition, once the desired estimate uncertainty target is
+achieved, analysts can terminate the M&V study and switch to 100%
+intervention. The key advantage of randomization is that if control
+strategies are sampled with equal probability, the influence of other
+confounding factors such as occupancy change (commonly known as one of
+the non-routine events) and other more subtle long term changes in
+building energy use are likely to be evenly distributed among
+measurements, leading to a more accurate and unbiased assessment of the
+intervention effect.
 
@@
 
-The Building Genome Dataset 2 (BGD2) is an extensive open-access dataset
-designed to advance research and development in building energy
+To demonstrate and differences between the two M&V methods, the Building
+Genome Dataset 2 (BGD2) is used, which is an extensive open-access
+dataset designed to advance research and development in building energy
 efficiency and control strategies acting as a test-bed for modeling,
@@
 
-ASHRAE Guidelinge 14 and IPMVP provides options for whole building M&V
-use cases. For code compliance using prescriptive option, 12 months of
-baseline and 12 months of postretrofit measurements are required. In
+ASHRAE Guideline 14 and the IPMVP provide options for whole-building M&V
+use cases. For code compliance using the prescriptive option, 12 months
+of baseline and 12 months of post-retrofit measurements are required. In
 addition, the expected savings should be larger than 10% and the
@@
 directly associated with calculated savings, for instance, accounting
-for the potential bias that baseline model might deteriorate (i.e.
-becomes 'stale') over an extended period of pre- and post-analysis.
+for the potential bias that the baseline model might deteriorate (i.e.
+become 'stale') over the extended period of pre- and post-analysis.
 
@@
 
-Non-routine events during M&V commonly refers to unexpected changes in a
-building that influences its energy usage. These changes can greatly
-affect measured energy consumption in buildings and are typically
-unrelated to the intervention strategy. Therefore, those changes are
-considered 'static factors' respect to measured independent variables. A
-common non-routine event in energy-saving M&V projects is a change in
-occupancy or a significant shift in occupant behavior, equipment run
-time, and operating conditions (e.g. set points, lighting and
-ventilation levels). However, current standards or guidelines only
-provides general approach for consideration. For example, IPMVP requires
-facility owner and the M&V analyst periodically perform inspections of
-all equipment and operations during reporting period, which is labor
-intensive and error-prone. ASHRAE guideline 14 recommends performing
-engineering calculations or computer software simulations to adjust
-post-retrofit baseline. Additional, it is very rare to have access to
-all measurements such as occupancy, and thus the analyst normally assume
-those factors remain unchanged throughout the study. Specifically,
-studies investigating the effects of demand response on building energy
-efficiency commonly uses linear interpolation to estimate
-counter-factual baseline [@beil_round-trip_2015;
-@keskar_commercial_2020], and one study points out that it is inaccurate
-to assume no change in the operating conditions during the test period
-[@huang_experimental_2023]. One likely encountered non-routine event due
-to operation condition change is filter clogging in air handling units
-due to particle accumulation. This can cause supply fans to gradually
-consume more energy to maintain required duct static pressure
-[@feng_newly_2019; @zhai_full-scale_2017]. If M&V analysts are unaware
-of such changes and lack an appropriate adjustment method (e.g.,
-replacing filters before the intervention begins), the savings could be
-underestimated as increased energy use is incorrectly attributed to the
-intervention rather than the mechanical issue. Some studies realized the
-limitation of current M&V methods, which only consider adjusting for
-outdoor weather, is insufficient and emphasized the importance of
-requiring matched comparison groups to control for exogenous factors
-beyond weather differences when comparing between baseline and
-intervention [@demand_side_analytics_population_2022;
-@huang_experimental_2023]. However, those methods are still unable to
-accurately quantify uncertainties due to changing baseline.
+Non-routine events in M&V commonly refer to unexpected changes in a
+building that influence a building's energy usage and are typically
+unrelated to the intervention strategy. Current guidelines define
+non-routine events as 'static factors' that needed to be adjusted after
+projecting baseline in the post-retrofit period [@ashrae_ashrae_2023;
+@efficiency_valuation_organisation_international_2012]. A common
+non-routine event in energy-saving M&V projects is a change in occupancy
+or a significant shift in occupant behavior, equipment run time, and
+operating conditions (e.g. set points, lighting, and ventilation
+levels). However, current standards or guidelines only provide a general
+approach for consideration. For example, IPMVP requires the facility
+owner and the M&V analyst to periodically perform inspections of all
+equipment and operations during the reporting period, which is
+labor-intensive and error-prone. ASHRAE guideline 14 recommends
+performing engineering calculations or computer software simulations to
+adjust the post-retrofit baseline. Additionally, it is relatively rare
+to have access to all measurements needed for adjustments such as
+occupancy, and thus the analyst normally assumes those factors remain
+unchanged throughout the study. For example, some studies uses linear
+interpolation to estimate counter-factual baseline for demand response
+program [@beil_round-trip_2015; @keskar_commercial_2020]. Consequently,
+one study points out that it is inaccurate to assume no change in the
+operating conditions during the response period
+[@huang_experimental_2023]. Other studies realized such limitation in
+field study, and emphasized the importance of requiring matched groups
+to control for exogenous factors beyond weather differences when
+comparing between baseline and intervention
+[@demand_side_analytics_population_2022; @huang_experimental_2023].
+
+Overall, those non-routine events mostly refer to clearly observable
+changes, but there are also more gradual and subtle changes in
+buildings, which are hardly noticed. We define those gradual changes as
+'operational drift' in this study. One likely encountered operational
+drift situation is the filter clogging in air handling units due to
+particle accumulation. This causes supply fans to gradually consume more
+energy to maintain the required duct static pressure [@feng_newly_2019;
+@zhai_full-scale_2017]. If M&V analysts are unaware of such changes and
+lack an appropriate adjustment method (e.g., replacing filters before
+the intervention begins), the savings could be overestimated as the
+decreased energy use is incorrectly attributed to the intervention
+rather than the filter replacement. An operational drift has not been
+formally defined in the literature or standards, but in this study, we
+consider it as a special type of non-routine events as it also
+represents a change in facility operation but in a longer term.
 
@@
 
-As mentioned, the goal of an M&V project is to determine the
-effect---typically energy savings---of an switchable energy-efficient
-intervention. The goal is to determine, for a large sample of buildings,
-how much more accurately the novel M&V method would estimate the savings
-of such an intervention compared to conventional M&V, and how much more
-quickly it would reach a result. An example of such an intervention is a
-control retrofit developed by a software-as-a-service company that
-adjusts the chilled water plant's supply water temperature, which can be
-commonly found in the literature [@duarte_field_2023;
-@lee_simulationoptimization_2012; @qiu_chilled_2022; @jin2007;
-@taylor_optimizing_2012]. In our case, we make it even simpler by
-adjusting setpoint based on outdoor weather conditions with more
-description shown in Section \@ref(apply-virtue-intervention).
-Therefore, we defined the M&V scenario as follows:
-
-"A company aims to sell its supply temperature reset control software
-package to a customer, in this case, the building owner, with a
-promotion that it will reduce the building's electricity usage. If the
-building owner decides to purchase the service, the company agrees to
-charge a service fee based on a percentage of the measured energy
-savings."
-
-As required by the M&V scenario, we assessed the performance of both the
-conventional and the novel randomized M&V methods by estimating the
-intervention energy savings for all valid buildings in the BG2 dataset.
-By conducting such analysis, we aim to:
-
-1.  Compare the energy saving estimation accuracy between the
-    conventional and the randomized method. This study extends the
-    comparison to a large sample of buildings, covering a variety of
-    types and climate zones. The comparison metrics include both
-    estimation accuracy and M&V completion timeline.
-
-2.  Verify the enhanced robustness of the randomized method. By using
-    realistic measurements from real-world buildings, which include
-    various sources of noise, we aim to reflect the challenges faced by
-    building analysts in real projects. As we will perform numercial
-    simulation on the existing dataset, the ground-truth savings can be
-    calculated as the reference for comparison.
-
-3.  Open-source implementation of the proposed randomized M&V method
-    using a public dataset. We ensured the reproducibility of the method
-    by making the analysis code open source including randomized
-    schedule generation, sequential statistical analysis, energy
-    modeling and normalized saving calculation. In addition, we also
-    included codes for extended use cases of the randomized method, such
-    as changing sampling ratio and sampling intervals. Using the
-    available open resources, building analysts should be able to
-    seamlessly integrate and apply them in their own M&V projects.
+As mentioned, the goal of an M&V project is to determine the effect,
+typically energy savings, of some intervention in the building. In this
+study, we focus on switchable interventions and an example of such an
+intervention is a control strategy that adjusts the chilled water
+plant's supply water temperature, which can be commonly found in the
+literature [@duarte_field_2023; @lee_simulationoptimization_2012;
+@qiu_chilled_2022; @jin2007; @taylor_optimizing_2012]. In our case, we
+make it even simpler by adjusting the setpoint based on outdoor weather
+conditions with more description shown in Section
+\@ref(apply-virtue-intervention). By conducting such an analysis, we aim
+to:
+
+1.  Determine, using a large sample of buildings, how much more
+    accurately the randomized M&V method would estimate the savings of
+    the proposed intervention compared to the conventional M&V, and how
+    much more quickly it would reach a result.
+
+2.  Determine how much more robust the randomized method is compared to
+    the conventional method when static non-routine events and
+    operational drift are present in the dataset. The BDG2 dataset
+    contains real-world building energy usage data collected over two
+    years, inherently contains varying degrees of usage changes. By
+    virtually introducing intervention effects into the existing
+    dataset, we can calculate the ground-truth energy savings, which
+    serve as a reference point for method comparison.
+
+3.  Open-source implementation of the proposed randomized M&V method. We
+    ensured the reproducibility of the method by making the analysis
+    code open source including randomized schedule generation,
+    sequential statistical analysis, energy modeling and normalized
+    saving calculation. In addition, we also included code for extended
+    use cases of the randomized method, such as changing sampling ratio
+    and sampling intervals. Using the available open resources, building
+    analysts should be able to seamlessly integrate and apply them in
+    their own M&V projects.
 
@@
 
-```{r flowchart, fig.cap = "Flow chat showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="75%"}
+```{r flowchart, fig.cap = "Flow chart showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="75%"}
 
@@
 
-In this study, we extracted the electricity measurements from the BDG2
-dataset and filter out all qualified buildings based on the following
-criteria:
+We extracted the electricity measurements from the BDG2 dataset and
+filter out all qualified buildings based on the following criteria:
 
@@
 3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
-    decrease of building electricity usage in the second year should be
+    decrease in building electricity usage in the second year should be
     less than 25% of that in the first year.
 
-4.  Electric EUI \< 750 kWh/$m^2$: excludes buildings at the top 5%
+4.  Electric EUI \< 750 kWh/$m^2$: excludes buildings at the top 5% of
     electrical energy usage intensity according to the statistics
@@
 
-```{r stablesite, fig.cap = "Site summary of the stable building subset (counts < 10 are omitted for visualization; left: aggragated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}
+```{r stablesite, fig.cap = "Site summary of the filtered buildings from BDG2 dataset (counts < 10 are omitted for visualization; left: aggregated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}
 
@@
           legend = "bottom") +
-  plot_annotation(title = "Summary of building numbers and types")
+  plot_annotation(title = "Summary of filtered buildings and types")
 ```
@@
 
-1.  No statistical significant difference (P-value \> 0.05) between the
-    two-year electricity usage, which are sometimes assumed to be the
-    case for whole-building measurement and verification.
-
-As a results, there are only 66 buildings were labeled in this 'stable'
-subset (out of total of 600 buildings) implying that the assumption made
-by conventional M&V is overly simplified. As mentioned in Section
+1.  No statistically significant difference (P-value \> 0.05) between
+    the two-year electricity usage, which are sometimes assumed to be
+    the case for whole-building measurement and verification.
+
+As a result, there are only 66 buildings were labeled in this 'stable'
+subset (out of total of 600 buildings) implying that the assumption that
+buildings are typically stable over the timeframe involved in
+conventional M&V is rarely true. As mentioned in Section
 \@ref(non-routine-events-impact), such variability is largely associated
-with non-routine events and according to the statistics shown here, is
-commonly observed in real applications.
+with non-routine events, especially operational drifts since the
+assessment last for two years. And according to the statistics shown
+here, it is a typical case observed in real buildings applications.
 
@@
 literature [@lee_chilled_2022; @congradac_recognition_2012]. For both
-strategies, we assume that the chiller is activated when the outdoor
+strategies, we assume that the chiller operates when the outdoor
 temperature exceeds 10°C. The baseline strategy, representing the
@@
 
-```{r chwst, fig.cap = "Proposed intervention strategy: chilled water supply temperature reset based on outdoor temperature", fig.width=8, fig.height=4}
+```{r chwst, fig.cap = "Proposed intervention strategy: chilled water supply temperature reset based on outdoor temperature.", fig.width=8, fig.height=4}
 
@@
 
-We mapped the chilled water supply temperature reset to the electrical
-energy savings as: $$
+We used a simple relationship to map the chilled water supply
+temperature reset to the electrical energy savings: $$
 \text{savings} = mean(E_{CHW}) \times 25\% \times \text{perc_savings}
@@
 |                        |                                                                  |       |
-|--------------|--------------------------------------------|--------------|
+|------------------|------------------------------------|------------------|
 | Parameter              | Description                                                      | Value |
@@
 
-We described more in detail the workflow of both conventional and
-randomized M&V methods in the previous study [@raftery2024]. The
-conventional method is a simple pre- and post- comparison on a 12-month
-baseline and 12-month intervention timeline. The randomized method
-defines sampling requirements as:
+We described in detail the workflow of both conventional and randomized
+M&V methods in the previous study [@raftery2024]. The conventional
+method is a pre- and post- comparison on a 12-month baseline and
+12-month intervention timeline. The randomized method defines sampling
+requirements as:
 
@@
     randomized schedule covers the entire two-year period but stopping
-    criteria enables an early stop at the end of satisfied blocking
+    criteria enables an early stop at the end of the satisfied blocking
     period.
@@
 
--   Test for no carryover effect using a t-test with a p-value not
-    exceeding a defined significance threshold of 0.05.
+-   No carryover effect using a t-test with a p-value not exceeding a
+    defined significance threshold of 0.05.
 
@@
 
--   test with an equal sampling ratio (50% baseline, 50% intervention).
+-   Sample strategies at an equal sampling ratio (50% baseline, 50%
+    intervention).
 
@@
 steps move closer to the center line (50% median), the distribution
-range progressively narrows with the next step represents 12.5% to
+range progressively narrows with the next step representing 12.5% to
 87.5%, followed by 25% to 75%. In addition, for each boxen plot, we also
-showed the mean value of the distribution on top.
-
-```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both at the stopping criteria and over a two-year period).", fig.height=12, fig.width=12}
+showed the mean value of the distribution on top for reference.
+
+```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period).", fig.height=12, fig.width=12}
 
@@
        y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with measured weather")) +
+       subtitle = str_glue("All {A_building} buildings with no model normalization on weather")) +
   coord_cartesian(ylim = c(0, 23)) +
@@
        y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with TMY weather")) +
+       subtitle = str_glue("All {A_building} buildings with TOWT model normalized on TMY weather")) +
   coord_cartesian(ylim = c(0, 23)) +
@@
        color = NULL, 
-       subtitle = "timeline comparison") +
+       subtitle = "timeline comparison of satisfying all stopping criteria") +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
           legend = "bottom") +
-  plot_annotation(title = "Overall comparison of conventional and randomized M&V")
+  plot_annotation(title = "Overall accuracy and time required for savings estimation by different M&V methods")
 ```
@@
 Figure \@ref(fig:abstract) shows the overall results of M&V methods
-comparison. Subplots a) and b) calculate the savings estimation error as
-the absolute deviation from the true savings (i.e. $|Acc|$) and shows
-the results distribution using conventional M&V method in the first
+comparison and for clarity, we plotted conventional M&V results in grey,
+randomized M&V results after satisfying all stopping criteria in green,
+and the randomized M&V results sampled for the entire 24 months in blue.
+Subplots (a) and (b) calculate the savings estimation error as the
+absolute deviation from the true savings (i.e. $|Acc|$) and show the
+results distribution using the conventional M&V method in the first
 column, the randomized M&V method that stops after satisfying all
@@
 continues 50%/50% sampling throughout 2 years in the third column. In
-addition, subplot a) shows the savings estimated from measured weather
-condition and subplot b) shows the savings normalized on typical
-meteorological weather file after using TOWT model. As shown clearly in
-the plots, the conventional M&V method exhibits an average deviation of
-6% in savings estimation, whereas the proposed randomized method
+addition, subplot (a) shows the savings estimated from measured weather
+conditions, and subplot b) shows the savings normalized on the typical
+meteorological weather of the building location after using TOWT model.
+As a result, the conventional M&V method exhibits an average deviation
+of 6% in savings estimation, whereas the proposed randomized method
 demonstrates significantly smaller deviations. If the analyst stops
@@
 reduced to approximately 2%. Extending the M&V period to match the
-conventional method further improves accuracy to 1%. Furthermore, by
-comparing subplot (a) and (b), we observe a minor improvement (less than
-1%) in estimation accuracy due to model fitting. This is because the
-weather condition over the two years for all selected locations are
-similar. Thus the effect of adjusting temperature as an independent
-variable in the model is not significant. Overall the results suggest
-that projecting the baseline into the post-retrofit period that only
-considers weather variation is inefficient. Subplot (c) presents the
-overall timeline comparison of the two M&V methods. The results indicate
-that a significant portion achieve accurate M&V results within 24 weeks
-and all buildings satisfy the stopping criteria within 36 weeks, mostly
-influenced by weather variability. Meanwhile, the red arrow highlights
-the excess time required by the conventional M&V method, which only
-leads to less reliable results.
+length of the conventional method (24 months) further improves accuracy
+to 1%. Furthermore, by comparing subplots (a) and (b), we observe a
+minor improvement in estimation accuracy when a model is fitted to the
+data to account for weather differences. The model fitting is most
+beneficial for the shorter dataset (randomized and stopping when all
+criteria are met, typically 24-36 weeks of data), where it improves
+estimation accuracy by 0.6%. This is because the weather conditions
+between the first and second year for each of the locations are
+relatively similar on average over this time period, and thus the effect
+of adjusting for temperature as an independent variable in the model is
+relatively small. Subplot (c) compares the overall timeline of the two
+M&V methods. The results show that almost all buildings achieve accurate
+M&V results and meet the stopping criteria within 24 weeks (for this
+intervention), with the remainder meeting all stopping criteria within
+36 weeks. The additional 12 weeks is usually driven by a need to span
+sufficient weather conditions representative of a full year of data.
+Meanwhile, the red arrow highlights the excess time required by the
+conventional M&V method (24 months). Thus, the randomized M&V method
+finds more accurate results in approximately a quarter of the time
+required for conventional M&V. Even if it was the case that an existing
+12-month baseline data was available at the start of the M&V process,
+randomized M&V would still obtain a more accurate result, in less time,
+than obtaining the 12-month retrofit dataset required by conventional
+M&V.
 
@@
 method is significantly larger compared to the randomized method that
-only takes 36 weeks according to subplot c) in the Figure
+only takes 36 weeks according to subplot (c) in Figure
 \@ref(fig:abstract).
@@
        y = "Mean error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with measured weather")) +
+       subtitle = str_glue("All {A_building} buildings with no model normalization on weather")) +
   coord_cartesian(ylim = c(-20, 20)) +
@@
        y = "Mean error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
+       subtitle = str_glue("All {A_building} buildings with TOWT model normalized on TMY weather")) +
   coord_cartesian(ylim = c(-20, 20)) +
@@
 
-Additionally in the literature we found another M&V method study using
-over 500+ commercial buildings, but their study scope was to compare a
-variety of M&V baseline models from simple week mean to more complex
-machine learning methods using the normalized mean bias error
-(NMBE)[@granderson_accuracy_2016]. Since we also calculated the
-normalized savings estimation error, we compared our normalized accuracy
-of 600+ buildings with TOWT model and TMY weather with their TOWT model
-prediction assessment in the following table. As each percentile is
-closely matched, our results are verified.
+Additionally in the literature, we found another M&V study using over
+500+ commercial buildings, but their study scope was to compare the
+predictive performance of a variety of M&V baseline models from simple
+weekly mean to more complex machine learning methods using the
+normalized mean bias error (NMBE) [@granderson_accuracy_2016]. Since we
+also calculated the normalized savings estimation error, we compared our
+normalized accuracy for 600+ buildings with the TOWT model and TMY
+weather with their TOWT model prediction assessment in the following
+table, highlighting that the model and overall building datasets are
+performing similarly.
 
 |                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
-|-----------------|--------------------------------------|-----------------|
+|-------------------|-----------------------------------|-------------------|
 | 25th percentile | -5.85                                               | `r percentiles$Q1`     |
@@
 
-```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect", fig.height=6, fig.width=12}
+```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect when buildings are subject to static non-rountine events and gradual operational drift.", fig.height=6, fig.width=12}
 
@@
        title = "Mean effect of non-routine events when using different M&V methods", 
-       subtitle = str_glue("All {A_building} buildings with measured weather")) +
+       subtitle = str_glue("All {A_building} buildings with no model normalization on weather")) +
   coord_cartesian(ylim = c(-18, 18)) +
@@
 As mentioned in Section \@ref(filter-and-clean-dataset), the majority of
-the selected buildings are affected by non-routine events, which is the
+the selected buildings are affected by operational drift, which is the
 reason why we observe baseline electricity usage changes throughout the
-two year. To evaluate the reliability of the two M&V methods in blocking
-the effect of non-routine events, we repeated the M&V run on the
-original dataset prior to the application of the reset control. In other
-words, as there is no intervention applied or the intervention is known
-to have no impact on whole-building electricity usage, a more reliable
-M&V method should detect closer to 0 kW savings using the measurements.
-Therefore, the $Acc$ here are calculated as:
+two year period. To evaluate the robustness of the two M&V methods to
+the effect of operational drift and non-routine events, we repeated the
+M&V run on the original dataset prior to the application of the reset
+control. In other words, there is no intervention applied here, so the
+most reliable M&V method will be the one that detects closer to 0 kW
+savings using the measurements. Therefore, the $Acc$ here are calculated
+as:
 
@@
 Figure \@ref(fig:nosaving) shows the randomized M&V method (both at the
-time of satisfying all stopping criteria and at the end of two-year
+time of satisfying all stopping criteria and at the end of a two-year
 period) produced consistent results. We included the absolute deviation
-($|ACC|$) in the appendix. Similarly, stopping early can lead to
-slightly larger uncertainty similar to Figure \@ref(fig:abstract).
-However, as for conventional method, the uncertainty range is much
-larger than the randomized method and it shows on average the deviation
-is 1.3% meaning a building analyst can 'detect' 1.3% energy savings when
-there isn't any using the conventional M&V method.
+($|ACC|$) figure in the supplementary material. Similarly to Figure
+\@ref(fig:abstract), stopping when meeting all stopping criteria
+generally yields slightly larger uncertainty than continuing for 24
+months. However, it still yields a more accurate result than the
+conventional method. Interestingly, similar to
+[@granderson_accuracy_2016] where they obtained a median normalized
+deviation around -1.25%, our analysis on conventional method shows that
+the typical deviation is -1.3%, meaning that the typical building is
+decreasing energy use slightly in this dataset in the absence of any
+known intervention. Thus, if a building analyst were to apply an
+intervention to this building and use conventional M&V to measure the
+effect, they would 'detect' 1.3% energy savings even when the
+intervention had no actual impact on energy performance.
 
@@
 CV(RMSE) is a normalized measure, it enables direct comparison across
-different model fitting results. According to ASHRAE Guideline 14,
+different model-fitting results. According to ASHRAE Guideline 14,
 whole-building baseline models should achieve a CV(RMSE) below 25% for
-computing savings or 30% when less than 12 months' of measurements are
+computing savings or 30% when less than 12 months of measurements are
 available [@ashrae_ashrae_2023]. In our study, Figure \@ref(fig:towtacc)
@@
 around 87.5% of the models fitted in the randomized M&V method are code
-compliance while 80% for the conventional M&V method. As the quantity of
+compliant while 80% for the conventional M&V method. As the quantity of
 the training set is the same, this highlights the impact of data
@@
          method = recode_factor(method, 
-                              "rand" = "Randomized", 
+                              "rand" = "Randomized\n(24 months)", 
                               "conv" = "Conventional"))
 
-result <- t.test(filter(cvrmse, method == "Randomized")$cvrmse, 
+result <- t.test(filter(cvrmse, method == "Randomized\n(24 months)")$cvrmse, 
                  filter(cvrmse, method == "Conventional")$cvrmse, 
@@
        color = NULL, 
-       title = "TOWT model fitting accuracy", 
+       title = "TOWT model fitting accuracy by different M&V methods", 
        subtitle = "CV(RMSE)") +
@@
 
-Additionally, we also noticed in our study that despite the regression
-models capture well on the mean energy consumption of the building but
-the tend to underestimate the 15-min daily peak load [@granderson2021],
-which is not useful for assessing savings for demand response events.
-
 ## Sampling interval impact
 
-In certain situations, particularly those involving large thermal mass
-in active thermal energy storage systems, there is a known time-lagged
-effect associated with the intervention. This means that the effect of
-one strategy may persist after switching to another. To mitigate such
-carryover effects, building analysts may choose to increase the sampling
-interval and drop non-consecutive days in the switchback experiment. For
-example, with a hot water tank that is charged by an intervention
-strategy, residual heat may still be present when the next strategy
-begins, influencing subsequent measurements, particularly the
-performance measured on the first day (i.e. a non-consecutive day). In
-such cases, increasing the sampling interval (e.g., sample strategies
-every 2 or 3 days instead of daily) and discard measurements from those
-non-consecutive days can eliminate potential carryover effect. However,
-this approach reduces the total data collected: for instance, sampling
-every 3 days would drop 1/6 of measurements and sampling every 2 day
-means would remove 1/4 of measurements. Therefore, by increasing the
-sampling interval, there is a reduction in statistical power due to less
-randomness, and when decreasing the sampling interval, there is a
+In some cases, particularly those involving large thermal mass and
+active thermal energy storage systems or operating continuously, there
+can be a substantial time-lagged effect associated with interventions.
+This means that the effect of one control strategy may persist after
+switching to another. To mitigate such carryover effects, building
+analysts may choose to drop non-consecutive days in the switchback
+experiment. For example, consider a heat pump hot water heater where the
+tank is charged at very different times of the day by the intervention
+strategy compared to the baseline strategy. When switching strategy, the
+tank may be much warmer or cooler than typical for that strategy, thus
+influencing subsequent measurements, particularly the performance
+measured on the first day (i.e., a non-consecutive day). In such cases,
+increasing the sampling interval (e.g., sample strategies every 2 or 3
+days instead of daily) and discarding measurements from those
+non-consecutive days can eliminate carryover. However, this approach
+reduces the total data collected: for instance, sampling every 3 days
+would drop 1/6 of measurements and sampling every 2 day means would
+remove 1/4 of measurements. Therefore, by increasing the sampling
+interval, there is an accuracy penalty due to a reduction in how often
+samples occur, and when decreasing the sampling interval, there is an
 accuracy penalty from dropping more non-consecutive days. To quantify
 such trade-off, \@ref(fig:absinterval) shows the accuracy of different
-sampling intervals and consequence of dropping the non-consecutive days
-in the dataset over the entire two-year period.
+sampling intervals and the consequence of dropping the non-consecutive
+days in the dataset at the time when randomized M&V produces a result
+(after satisfying all stopping criteria).
 
@@
 #### ABS-INT ####
-# dropped version
-df_interval_keep_stable %<>% 
-  left_join(df_fs_stable %>% filter(method == "rand"), by = c("name", "site")) %>% 
-  select(-c("scenario", "method")) %>% 
-  rename("interval_fs_1" = "savings") %>% 
-  left_join(df_fs_tmy_stable, by = c("name", "site")) %>% 
-  select(-conv) %>% 
-  rename("interval_tmy_1" = "rand")
-
-df_interval_keep_variable %<>% 
-  left_join(df_fs_variable %>% filter(method == "rand"), by = c("name", "site")) %>% 
-  select(-c("scenario", "method")) %>% 
-  rename("interval_fs_1" = "savings") %>% 
-  left_join(df_fs_tmy_variable, by = c("name", "site")) %>% 
-  select(-conv) %>% 
-  rename("interval_tmy_1" = "rand")
-
-df_drop <- df_interval_drop_stable %>% 
-  bind_rows(df_interval_drop_variable) %>% 
-  select(name, contains("fs")) %>% 
-  rename("drop_1" = "interval_fs_1", 
-         "drop_2" = "interval_fs_2", 
-         "drop_3" = "interval_fs_3", 
-         "drop_7" = "interval_fs_7")
-
-df_keep <- df_interval_keep_stable %>% 
-  bind_rows(df_interval_keep_variable) %>% 
-  select(name, contains("fs")) %>% 
-  rename("keep_1" = "interval_fs_1",
-         "keep_2" = "interval_fs_2", 
-         "keep_3" = "interval_fs_3", 
-         "keep_7" = "interval_fs_7")
+df_seq_interval_fs_keep_stable %<>% 
+  select(-c(sprt, temp, final)) %>% 
+  bind_rows(df_seq_fs_stable %>% 
+              filter(seq == "eob") %>% 
+              select(-seq) %>% 
+              mutate(interval = 1) %>% 
+              rename(eob = "fs"))
+
+df_seq_interval_fs_keep_variable %<>% 
+  select(-c(sprt, temp, final)) %>% 
+  bind_rows(df_seq_fs_variable %>% 
+              filter(seq == "eob") %>% 
+              select(-seq) %>% 
+              mutate(interval = 1) %>% 
+              rename(eob = "fs"))
+
+df_drop <- df_seq_interval_fs_drop_stable %>% 
+  bind_rows(df_seq_interval_fs_drop_variable) %>% 
+  select(name, eob, interval) %>% 
+  mutate(group = "drop")
+
+df_keep <- df_seq_interval_fs_keep_stable %>% 
+  bind_rows(df_seq_interval_fs_keep_variable) %>% 
+  select(name, eob, interval) %>% 
+  mutate(group = "keep")
 
 df_MW_A <- df_drop %>% 
-  left_join(df_keep, by = c("name")) %>%
+  bind_rows(df_keep) %>% 
   left_join(rbind(df_fs_stable %>% 
@@
                     filter(scenario == "ref" & method == "true")), by = "name") %>% 
-  mutate(drop_1 = abs(drop_1 - savings),
-         drop_2 = abs(drop_2 - savings), 
-         drop_3 = abs(drop_3 - savings), 
-         drop_7 = abs(drop_7 - savings), 
-         keep_1 = abs(keep_1 - savings),
-         keep_2 = abs(keep_2 - savings), 
-         keep_3 = abs(keep_3 - savings), 
-         keep_7 = abs(keep_7 - savings)) %>% 
-  pivot_longer(cols = contains("drop") | contains("keep"),  
-               names_to = c("group", "interval"),
-               names_pattern = "(drop|keep)_(\\d+)",
-               values_to = "diff")
+  mutate(diff = abs(savings - eob)) %>% 
+  filter(is.finite(diff))
 
@@
               group_by(group, interval) %>% 
-              summarise(mean = mean(diff, na.rm = T)) %>% 
+              summarise(mean = mean(diff, inf.rm = T)) %>% 
               ungroup(), 
-            aes(x = interval, y = mean, group = group,
+            aes(x = interval, 
+                y = mean, group = group,
                 label = paste0(round(mean, digits = 1), " %")), 
-            position = position_dodge(width = 0.75), 
-            color = "white") +
+            color = "white",
+            position = position_dodge(width = 0.75)) +
   scale_y_continuous(expand = c(0, 0), 
@@
   scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = c("grey80", "grey80")) + 
+  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
   labs(fill = NULL, 
@@
        y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with measured weather")) +
-  coord_cartesian(ylim = c(0, 10)) +
+       subtitle = str_glue("All {A_building} buildings with no model normalization on weather")) +
+  coord_cartesian(ylim = c(0, 15)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
 
-# TMY version
-df_drop <- df_interval_drop_stable %>% 
-  bind_rows(df_interval_drop_variable) %>% 
-  select(name, contains("tmy")) %>% 
-  rename("drop_1" = "interval_tmy_1",
-         "drop_2" = "interval_tmy_2", 
-         "drop_3" = "interval_tmy_3", 
-         "drop_7" = "interval_tmy_7")
-
-df_keep <- df_interval_keep_stable %>% 
-  bind_rows(df_interval_keep_variable) %>% 
-  select(name, contains("tmy")) %>% 
-  rename("keep_1" = "interval_tmy_1",
-         "keep_2" = "interval_tmy_2", 
-         "keep_3" = "interval_tmy_3", 
-         "keep_7" = "interval_tmy_7")
+# with tmy version
+df_seq_interval_nm_keep_stable %<>% 
+  select(-c(sprt, temp, final)) %>% 
+  bind_rows(df_sprt_all_stable %>% 
+              filter(seq == "eob") %>% 
+              select(-c(seq, n_weeks)) %>% 
+              mutate(interval = 1) %>% 
+              rename(eob = "annual"))
+
+df_seq_interval_nm_keep_variable %<>% 
+  select(-c(sprt, temp, final)) %>% 
+  bind_rows(df_sprt_all_variable %>% 
+              filter(seq == "eob") %>% 
+              select(-c(seq, n_weeks)) %>% 
+              mutate(interval = 1) %>% 
+              rename(eob = "annual"))
+
+df_drop <- df_seq_interval_nm_drop_stable %>% 
+  bind_rows(df_seq_interval_nm_drop_variable) %>% 
+  select(name, eob, interval) %>% 
+  mutate(group = "drop")
+
+df_keep <- df_seq_interval_nm_keep_stable %>% 
+  bind_rows(df_seq_interval_nm_keep_variable) %>% 
+  select(name, eob, interval) %>% 
+  mutate(group = "keep")
 
 df_TW_A <- df_drop %>% 
-  left_join(df_keep, by = c("name")) %>%
+  bind_rows(df_keep) %>% 
   left_join(rbind(df_fs_stable %>% 
@@
                     filter(scenario == "ref" & method == "true")), by = "name") %>% 
-  mutate(drop_1 = abs(savings - drop_1),
-         drop_2 = abs(savings - drop_2), 
-         drop_3 = abs(savings - drop_3), 
-         drop_7 = abs(savings - drop_7), 
-         keep_1 = abs(savings - keep_1),
-         keep_2 = abs(savings - keep_2), 
-         keep_3 = abs(savings - keep_3), 
-         keep_7 = abs(savings - keep_7)) %>% 
-  pivot_longer(cols = contains("drop") | contains("keep"),  
-               names_to = c("group", "interval"),
-               names_pattern = "(drop|keep)_(\\d+)",
-               values_to = "diff")
+  mutate(diff = abs(savings - eob)) %>% 
+  filter(is.finite(diff))
 
@@
               ungroup(), 
-            aes(x = interval, y = mean, group = group,
+            aes(x = interval, 
+                y = mean, 
+                group = group,
                 label = paste0(round(mean, digits = 1), " %")),
-            position = position_dodge(width = 0.75), 
-            color = "white") +
+            color = "white",
+            position = position_dodge(width = 0.75)) +
   scale_y_continuous(expand = c(0, 0), 
@@
        y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with TMY weather")) +
-  coord_cartesian(ylim = c(0, 10)) +
+       subtitle = str_glue("All {A_building} buildings with TOWT model normalized on TMY weather")) +
+  coord_cartesian(ylim = c(0, 15)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
           align = "hv", 
+          labels = c("a)", "b)"),
           common.legend = T, 
           legend = "bottom") +
-  plot_annotation(title = "Absolute error in savings estimation of ranodmized M&V by different sampling intervals")
+  plot_annotation(title = "Absolute error in savings estimation of randomized M&V after satisfying all stopping criteria")
 
@@
 To quantify the impact of reduced randomness, all measurements are kept
-when calculating the distribution shown in the darker colored set. The
-results suggests a gradual increase in error from 1% to 1.8% when
-sampling from daily to weekly intervals. To quantify the combined impact
-of data loss due to preventing carryover effect, we generated the
-lighter colored set, which drops all non-consecutive days. As a result,
-we noticed a significant increase in estimaion error when sampling
-daily, which lead to a similar accuracy compared to weekly intervals.
-Therefore, considering this trade-off, we recommend using a three-day
-sampling interval if the carryover effect is likely but expected to last
-less than one day (as more sampling is required for a two-day
-swithback), and then normalizing the estimation via energy modeling. We
-also included the mean deviation distribution and also the accuracy
-distribution calculated after all stopping criteria satisfied in the
-supplementary material and the results shows the same indication.
+when calculating the distribution shown in the darker-colored set. The
+results suggest a gradual increase in error from 2.4% to 3.7% when the
+sampling interval increases from daily to weekly intervals. This is due
+to the fact that increasing the sample interval means reducing
+randomization. To quantify the combined impact of data loss due to
+preventing the carryover effect, we generated the lighter-colored set,
+which drops all non-consecutive days. As a result, we noticed a
+significant increase in estimation error when sampling daily, which led
+to a similar accuracy compared to weekly intervals. Therefore,
+considering this trade-off, we recommend using a two-day sampling
+interval if the carryover effect is likely but expected to last less
+than one day, and then normalizing the estimation via energy modeling.
+We also included the mean deviation distribution and the accuracy
+distribution calculated after all stopping criteria were satisfied in
+the supplementary material and the results show the same indication.
+
+```{r intervaltime, fig.cap="Comparison of time required to satisfy all stopping criteria when considering dropping non-consecutive days (or not)", fig.height=6, fig.width=12}
+# timeline to finish sequential test
+df_drop <- bind_rows(df_seq_interval_tl_drop_stable, df_seq_interval_tl_drop_variable) %>%
+  mutate(group = "drop", 
+         temp = ifelse(base_temp >= interv_temp, base_temp, interv_temp)) %>% 
+  pivot_longer(c(sprt, temp, eob, final), names_to = "seq", values_to = "n_weeks") %>% 
+  select(name, seq, n_weeks, interval, group)
+
+interval_1_keep <- bind_rows(df_sprt_all_stable, df_sprt_all_variable) %>%
+  select(c(name, seq, n_weeks)) %>%
+  filter(seq != "final") %>%
+  mutate(seq = as.factor(seq)) %>%
+  mutate(interval = 1, 
+         group = "keep")
+
+df_keep <- bind_rows(df_seq_interval_tl_keep_stable, df_seq_interval_tl_keep_variable) %>%
+  mutate(group = "keep", 
+         temp = ifelse(base_temp >= interv_temp, base_temp, interv_temp)) %>% 
+  pivot_longer(c(sprt, temp, eob, final), names_to = "seq", values_to = "n_weeks") %>% 
+  select(name, seq, n_weeks, interval, group) %>% 
+  bind_rows(interval_1_keep)
+
+df_MW_A <- bind_rows(df_keep, df_drop) %>% 
+  filter(seq != "final") %>% 
+  filter(seq == "eob") %>% 
+  mutate(seq = as.factor(seq), 
+         seq = recode_factor(seq, 
+                             "eob" = "Buildings finishing all criteria"), 
+         interval = as.factor(interval), 
+         interval = recode_factor(interval, 
+                                  "1" = "1-day\nsampling",
+                                  "2" = "2-day\nsampling", 
+                                  "3" = "3-day\nsampling", 
+                                  "7" = "7-day\nsampling"), 
+         group = as.factor(group), 
+         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept"), 
+         n_weeks = as.factor(n_weeks), 
+         n_weeks = recode_factor(n_weeks, 
+                                 "12" = "12\nweeks", 
+                                 "24" = "24\nweeks", 
+                                 "36" = "36\nweeks", 
+                                 "48" = "48\nweeks")) %>% 
+  group_by(n_weeks, group, interval) %>% 
+  summarise(n = n()) %>%
+  ungroup()
+
+levA <- levels(df_MW_A$n_weeks)
+levB <- levels(df_MW_A$group)
+levC <- levels(df_MW_A$interval)
+
+# 2. Expand to all combinations of factorA and factorB
+all_combinations <- expand_grid(
+  n_weeks = levA,
+  group = levB, 
+  interval = levC, 
+)
+
+df_MW_A <- all_combinations %>%
+  left_join(df_MW_A, by = c("n_weeks", "group", "interval")) %>% 
+  mutate(n = ifelse(is.na(n), 0, n)) %>% 
+  group_by(group, interval) %>% 
+  mutate(n = ifelse(n > sum(n) * 0.01, n, 0), 
+         perc = n / sum(n) * 100) %>% 
+  ungroup()
+
+df_MW_A %>% 
+  ggplot(aes(x = n_weeks, y = perc, fill = group)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  facet_wrap(~interval, nrow = 1) +
+  geom_text(aes(x = n_weeks, y = perc + 2, label = paste0(round(perc, digits = 0), "%"), group = group), position = position_dodge(1)) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = ls_colors) +
+  scale_y_continuous(breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  labs(x = NULL,
+       y = NULL,
+       fill = NULL,
+       title = "Percentage of buildings using randomized M&V when satisfying all stopping critiera", 
+       subtitle = "Counted at the end of each 12-week blocking period") +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "bottom",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+```
+
+Sampling at different intervals and excluding non-consecutive days also
+impact other stopping criteria, such as the time needed to satisfy
+statistical uncertainty and weather conditions thresholds. In Figure
+\@ref(fig:intervaltime), we compared the percentage of buildings
+satisfying all stopping criteria under different sampling intervals,
+with or without dropping non-consecutive days. The results indicate that
+randomized M&V concludes more quickly when all data are retained,
+particularly at shorter intervals. For instance, when sampling at a
+one-day interval, 90% of buildings meet all criteria within 24 weeks if
+no data are dropped, but this number decreases significantly to 52% when
+non-consecutive days (account for 1/3 of the data) are excluded. As the
+sampling interval increases, this difference gradually diminishes: when
+sampling at a weekly interval, removing non-consecutive days has no
+impact on the timeline of meeting stopping criteria.
 
@@
 Another advantage of using randomized M&V is the flexibility of changing
-sampling ratio after the target savings detected. For instance, the
-building owner can continue sampling at a 50%/50% ratio between the
+the sampling ratio after the target savings are detected. For instance,
+the building owner can continue sampling at a 50%/50% ratio between the
 baseline and the intervention to further reduce the uncertainty
@@
 intervention to maximize cost savings, though this approach risks the
-baseline becoming outdated, which is another trade-off to consider. A
-middle-ground approach would be to sample at an 20%/80% ratio or 10%/90%
-between the baseline and the intervention. To compare those possible
-choices to a building owner, figure \@ref(fig:abscont) shows the results
-of changing the original 50%/50% sampling ratio after satisfying all
-stopping criteria and continuing a new ratio till the end of the
-two-year period. We also provided the reference case when the building
-owner continues with no change. The results shows as the sampled
-baseline days and intervention days become unbalanced, the savings
-estimation error grows from 1% to 5%. This is reasonable since sampling
-at 10%/90% is similar to the conventional M&V method. We included the
-mean deviation version in the supplementary material.
-
-```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy", fig.height=6, fig.width=12}
+baseline becoming outdated over time, which is another trade-off to
+consider. A middle-ground approach may be to continue sampling at a
+20%/80% ratio or 10%/90% between the baseline and the intervention. To
+compare those possible choices to a building owner, figure
+\@ref(fig:abscont) shows the results of changing the original 50%/50%
+sampling ratio after satisfying all stopping criteria and continuing a
+new ratio till the end of the two-year period. Those results were
+calculated without dropping non-consecutive days and include the initial
+50%/50% sampled data. On the far left, we also provided the case when
+the building owner continues with no change in sampling ratio. The
+results show that as the sampled baseline days and intervention days
+become unbalanced, the savings estimation error grows from 1% to 5%.
+This is reasonable since sampling at 10%/90% is similar to the
+conventional M&V method. And on the far right, we show again the
+accuracy of the conventional M&V method, which is still worse then
+sampling at 10%/90%. Also, We include the mean deviation version in the
+supplementary material.
+
+```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months", fig.height=6, fig.width=12}
 
 #### ABS-CONT ####
+conv_S <- df_fs_stable %>% 
+  filter(method != "rand") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = abs(true - conv), 
+         ratio = "conv") %>% 
+  select(name, ratio, diff)
+
 rand_cont_S <- df_cont_stable %>% 
@@
 
-df_MW_S <- rbind(rand_cont_S, rand_final_S)
+df_MW_S <- rbind(rand_cont_S, rand_final_S, conv_S)
 
 # plot for the variable subset
+conv_V <- df_fs_variable %>% 
+  filter(method != "rand") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = abs(true - conv), 
+         ratio = "conv") %>% 
+  select(name, ratio, diff)
+
 rand_cont_V <- df_cont_variable %>% 
@@
 
-df_MW_V <- rbind(rand_cont_V, rand_final_V)
-
-# plot for combined dataset
-rand_cont_A <- bind_rows(rand_cont_V, rand_cont_S) %>% 
-  select(name, rand_cont_diff = diff)
-
-rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
-  select(name, rand_final_diff = diff)
+df_MW_V <- rbind(rand_cont_V, rand_final_V, conv_V)
 
@@
   mutate(ratio = as.factor(ratio), 
-         ratio = recode_factor(ratio, "0" = "Randomized\n(50/50)", "1" = "Randomized\n(20/80)", "2" = "Randomized\n(10/90)")) %>% 
-  ggplot(aes(x = ratio, y = diff, fill = "Randomized")) +
+         ratio = recode_factor(ratio, "0" = "Randomized\n(50/50)", "1" = "Randomized\n(20/80)", "2" = "Randomized\n(10/90)", "conv" = "Conventional")) %>% 
+  ggplot(aes(x = ratio, y = diff, fill = ratio)) +
   geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
@@
   scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
+  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80")) + 
   labs(fill = NULL, 
@@
        y = "Absolute error in fractional savings", 
-       title = "Absolute error in savings estimation by different sampling ratio",
-       subtitle = str_glue("All {A_building} buildings with measured weather conditions")) +
-  coord_cartesian(ylim = c(0, 15)) +
+       title = "Absolute error in savings estimation fo randomized M&V using different sampling ratio",
+       subtitle = str_glue("All {A_building} buildings with with no model normalization on weather")) +
+  coord_cartesian(ylim = c(0, 20)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
 
-We identify two key limitations in this study:
+Aside from the general limitation that this method only applies to the
+subset of interventions in buildings that are easily switchable, we
+identify two key limitations in this study:
 
@@
     buildings. Yet, in some cases, raising the water temperature by 1°C
-    might yield more or less than 8% in electricity savings due to
-    target buildings' demand flexibility. For example, one field
+    might yield more or less than the specified 8% electricity savings
+    due to target buildings' demand flexibility. For example, one field
     validation of HVAC optimization shows additional savings might be
@@
     sampling interval would suffice for most commercial buildings, but
-    exceptions exist. Based on the measurement from the dataset, we have
-    limited knowledge of how builings thermal mass respond to the
-    proposed chilled water reset strategy. However, in the discussion,
-    we addressed such concern by comparing different sampling strategies
-    and the impact of dropping measurements from non-consecutive days.
+    certainly exceptions exist. However, in the discussion, we addressed
+    such concern by comparing different sampling strategies and the
+    impact of dropping measurements from non-consecutive days.
 
@@
 
-This research demonstrated the application of a novel whole-building
-measurement and verification (M&V) method, comparing its performance to
-the conventional approach outlined in ASHRAE Guideline 14 using a large,
-open-source commercial building dataset. The proposed M&V method
-leverages the randomized experimental design concept from other
-scientific fields, along with statistical sequential inference
-techniques, to determine when target savings are detected. We used a
-virtual control retrofit case---resetting the chilled water setpoint
-based on outdoor weather conditions---and applied it to over 600
-filtered commercial buildings. By comparing the savings estimations of
-the conventional method with the novel randomized method, we found that
-the randomized approach provides faster and more robust savings
-estimations.
+This research demonstrated the application of a randomized
+whole-building measurement and verification (M&V) method, comparing its
+performance to the conventional approach described in the IPMVP and in
+ASHRAE Guideline 14, using a large, open-source commercial building
+dataset. The porposed M&V method leverages the randomized experimental
+design concept from other scientific fields, along with statistical
+sequential inference techniques, to determine when target savings are
+detected. We used a virtual control retrofit case---resetting the
+chilled water setpoint based on outdoor weather conditions---and applied
+it to over 600 filtered commercial buildings. By comparing the savings
+estimations of the conventional method with the novel randomized method,
+we found that the randomized approach provides faster and more robust
+savings estimations.
 
@@
 in this study, the randomized M&V can provide a saving estimation by 36
-weeks (with the majority finishes by 24 weeks) once all stopping
-criteria satisfied. In contrast, the conventional method requires a full
-range of baseline and intervention measurements under normal operating
-conditions, typically taking 6-9 months for each phase. Most
-importantly, Furthermore, we verified that with reduced M&V timeline the
+weeks (with the majority finished by 24 weeks) once all stopping
+criteria are satisfied. In contrast, the conventional method requires a
+full range of baseline and intervention measurements under normal
+operating conditions, typically taking 6-9 months for each phase. Most
+importantly, we verified that with a reduced M&V timeline, the
 randomized method can estimate savings more accurately by showing that
-the absolute error is only 1 - 2 % while for the two-year conventional
-method, the estimation error could be 6% in a typical building. We also
-evaluated the impact of non-routine events on the proposed M&V method
-and the results show that baseline changes in the post retrofit period
-can deviate savings estimated using the conventional method. As a
-comparison, we found those events have very negligible impact on the
-savings estimated using the randomized method, demonstrating ideal
-reliability for realistic use cases.
+the absolute error is only 1 - 2 % for a typical building. Whereas for
+the much longer two-year conventional method, the estimation error was
+approximately 6% in a typical building. We also evaluated the impact of
+non-routine events especially more gradual and subtle operational drift
+on the proposed M&V method and the results show that baseline changes in
+the post-retrofit period can deviate savings estimated using the
+conventional method. As a comparison, we found those events have a very
+negligible impact on the savings estimated using the randomized method,
+demonstrating robustness to this issue. Lastly, we consider scenarios
+where there is a known carryover effect from switching between
+strategies in the building and assess the impact of dropping samples
+where the strategy is non-consecutive and varying the sampling interval.
+Through that process, we showed that for the typical building, if a
+carryover is likely to be present, then using a 3-day sampling interval
+and dropping non-consecutive samples yields the optimal design.
 
@@
 two methods and we found a significant improvement in our method due to
-random sampling. We also provide detailed assessment on using different
-sampling interval and sampling ratio, those results provide useful
-considerations to building analysts when generating randomized schedules
-for a variety of use cases.
+random sampling. We also provide a detailed assessment of using
+different sampling ratios, those results provide useful considerations
+to building analysts when generating randomized schedules for a variety
+of use cases.
 
@@
 Methodology, Investigation, Software, Writing - original draft,
-Writing - review & editing. **Paul Raftery**: Conceptualization, Formal
-analysis, Methodology, Investigation, Project administration,
-Supervision, Writing - review & editing. **Stefano Schiavon**:
-Conceptualization, Methodology, Investigation, Supervision, Writing -
-review & editing. **Carlos Duarte**: Methodology, Investigation,
-Supervision, Writing - review & editing. **Gail Brager**: Supervision,
-Writing - review & editing.
+Writing - review & editing. **Paul Raftery**: Conceptualization, Funding
+acquisition, Formal analysis, Methodology, Investigation, Project
+administration, Supervision, Writing - review & editing. **Stefano
+Schiavon**: Conceptualization, Methodology, Investigation, Supervision,
+Writing - review & editing. **Carlos Duarte**: Methodology,
+Investigation, Supervision, Writing - review & editing. **Gail Brager**:
+Supervision, Writing - review & editing.
 
@@
 
-Reproducible example with analysis code is available (MIT license) at
+A reproducible example with analysis code is available (MIT license) at
 <https://github.com/CenterForTheBuiltEnvironment/genome_mnv>.

### 2025-03-22 code/graphical.R

@@ -1854,45 +1854,7 @@ df_keep <- bind_rows(df_seq_interval_tl_keep_stable, df_seq_interval_tl_keep_var
   bind_rows(interval_1_keep)
 
 df_MW_A <- bind_rows(df_keep, df_drop) %>% 
-  filter(seq != "final")
-
-p_top <- df_MW_A %>% 
-  filter(seq != "eob") %>% 
-  mutate(seq = as.factor(seq), 
-         seq = recode_factor(seq, 
-                             "sprt" = "Buildings satisfying SPRT", 
-                             "temp" = "Buildings satisfying temperature range", 
-                             "eob" = "Buildings finishing all criteria"), 
-         interval = as.factor(interval), 
-         interval = recode_factor(interval, 
-                                  "1" = "1-day\nsampling",
-                                  "2" = "2-day\nsampling", 
-                                  "3" = "3-day\nsampling", 
-                                  "7" = "7-day\nsampling"), 
-         group = as.factor(group), 
-         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
-  ggplot(aes(x = interval, y = n_weeks, fill = group)) +
-  geom_jitter(alpha = 0.8,
-              size = 0.5,
-              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
-  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
-  geom_boxplot(outlier.shape = NA, coef = 0, color = "grey50", width = 0.8) +
-  scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = ls_colors) +
-  scale_y_continuous(expand = c(0, 0),
-                     breaks = seq(0, 50, by = 12), 
-                     labels = number_format(suffix = " weeks")) +
-  coord_cartesian(ylim = c(0, 50)) +
-  facet_wrap(~seq, nrow = 1) +
-  labs(x = NULL,
-       y = NULL,
-       fill = NULL,
-       subtitle = "Satisfying EACH stopping criterion") +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.position = "bottom",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-p_bottom <- df_MW_A %>% 
+  filter(seq != "final") %>% 
   filter(seq == "eob") %>% 
[…]

### 2025-04-24 code/manuscript.Rmd
Modified blocking period for new starting time

@@
 |                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
-|-------------------|-----------------------------------|-------------------|
+|-------------------|----------------------------------|-------------------|
 | 25th percentile | -5.85                                               | `r percentiles$Q1`     |
@@
 
+## Blocking period length impact
+
+Selecting different blocking periods affects the timing of evaluating
+stopping criteria, thus influencing the duration of the measurement and
+verification (M&V) process. Here, we summarize the advantages and
+disadvantages of shorter versus longer blocking periods. As discussed in
+our previous study, the second step of this randomized M&V method
+involves choosing an appropriate blocking period to minimize potential
+biases caused by 'bad' randomization. For most practical scenarios, we
+recommend a blocking period of approximately 12 to 16 weeks, as it
+typically covers the duration of a season, ensuring a representative
+sampling of operational conditions. Opting for a shorter blocking
+period, such as 6 weeks, has the potential benefit of allowing the
+stopping criteria to be met earlier. However, this also means that the
+analyst should perform the hypothesis tests more frequently. To avoid
+false positive due to frequent testing, adjusting the significance level
+using methods such as the Bonferroni correction can control for
+increased type-I error risk [@vickerstaff_methods_2019]. Additionally, a
+shorter blocking period can reduce the adverse impact of missing or
+problematic data. For example, if the intervention requires fine-tuning
+after initial implementation, removing data from the first blocking
+period would have less damage compared to when the switchback experiment
+was designed with longer blocking period. On the other hand, longer
+blocking periods are advantageous when evaluating multiple intervention
+strategies or using sampling intervals greater than one day. In such
+cases, a longer block can more easily ensure a balanced distribution of
+both day-of-week and outdoor weather conditions.
+
 ## Limitations

### 2025-04-24 code/0_schedule.R

@@ -48,8 +48,8 @@ source(paste0(function_path, "rand_seq.R"))
 # adding random sampling schedules
 block_params <- list(start_date = "2016-03-01",
                      n_weeks = 96,
-                     n_seasons = 8, 
-                     block_unit = 12, 
+                     n_seasons = 16, 
+                     block_unit = 6, 
                      pool = 20)
 
 # schedule generation

### 2025-04-24 code/1_analysis.R

@@ -55,7 +55,7 @@ source(paste0(function_path, "rand_seq.R"))
 
 # define parameters
 # section run control
-run_params <- list(type = "stable", 
+run_params <- list(type = "variable", 
                    start_time = "2016-03-01", 
                    sprt = T, 
                    interval = T)
@@ -75,8 +75,8 @@ ctr_params <- list(peak_hours = 10:16,                      # accounts for peak 
 # adding random sampling schedules
 block_params <- list(start_date = run_params$start_time,
                      n_weeks = 96,
-                     n_seasons = 8, 
-                     block_unit = 12)
+                     n_seasons = 16, 
+                     block_unit = 6)
 
 # adding sprt criteria
 sprt_param <- list(baseline = "Baseline",
@@ -914,6 +914,7 @@ for (n in 1:(nrow(all_names))){
 
 
 
+
 #### BIND ####
 # savings calculation
 df_seq_fs <- bind_rows(seq_frsaving) %>%
@@ -958,3 +959,4 @@ if (run_params$interval){
   write_rds(seq_nmsaving_interval_keep, paste0(readfile_path, "df_seq_interval_nm_keep.rds"), compress = "gz")
   
 }
+

### 2025-04-24 code/graphical.R

@@ -57,10 +57,6 @@ ctr_params <- list(peak_hours = 10:16,
                    coe_off = 1.2,
                    enable_temp = 8)
 
-occ_params <- list(change_start = 5,
-                   change_end = 8,
-                   change = 20)
-
 
 
 
@@ -430,7 +426,7 @@ df_time <- df_sprt_all_variable %>%
 
 count <- list()
 n <- 1
-for (i in seq(0, 48, by = 3)){
+for (i in seq(0, 60, by = 3)){
   
   df_sprt <- df_time %>% 
     filter(seq == "sprt",
@@ -457,7 +453,7 @@ count <- bind_rows(count)
 p_timeline <- count %>% 
   ggplot() +
   geom_bar(data = . %>% 
-             filter(n_weeks %in% c(12, 24, 36, 48)),
+             filter(n_weeks %in% seq(12, 60, by = 6)),
            aes(x = n_weeks, y = eob, fill = "Buildings finishing randomized M&V"), 
            stat = "identity", 
            position = "stack",
@@ -472,27 +468,27 @@ p_timeline <- count %>%
   geom_point(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"), 
              size = 1.5, 
              shape = 17) +
-  geom_segment(aes(x = 48.5, y = max(eob), xend = 95.5, yend = max(eob)),
+  geom_segment(aes(x = 60.5, y = max(eob), xend = 104.5, yend = max(eob)),
                arrow = arrow(length = unit(0.25, "in")),   
                linewidth = 1.1,  
                color = "#fb8072") + 
   annotate(geom = "text", 
-           x = 70, 
+           x = 80, 
            y = max(count$eob) - 30, 
            size = 5,
            label = "Excess time by conventional M&V") +
-  geom_vline(xintercept = c(12, 24, 36, 48, 96), lty = "dashed", color = "grey80") +
+  geom_vline(xintercept = seq(6, 60, by = 6), lty = "dashed", color = "grey80") +
   annotate(geom = "text", 
-           x = seq(6, 48, by = 12), 
+           x = 3, 
            y = 200, 
-           label = paste0("12-week\nblock"), 
+           label = paste0("6-week\nblock"), 
            alpha = 0.5) +
   scale_fill_manual(values = ls_colors) +
   scale_color_manual(values = ls_colors) +
   scale_x_continuous(expand = c(0, 0), 
-                     limits = c(0, 100),
[…]

### 2025-04-26 code/1_analysis.R
modifed timeline calculation error and reruned analysis

@@ -55,7 +55,7 @@ source(paste0(function_path, "rand_seq.R"))
 
 # define parameters
 # section run control
-run_params <- list(type = "variable", 
+run_params <- list(type = "stable", 
                    start_time = "2016-03-01", 
                    sprt = T, 
                    interval = T)
@@ -188,8 +188,19 @@ run_interpo <- function(df_all){
 get_eob <- function(sprt_res, sprt_overlap_base, sprt_overlap_interv){
   
   sprt_check <- sprt_res %>% filter(flag == 1) %>% slice(1) %>% .$n_weeks
-  bt_check <- sprt_overlap_base %>% filter(flag == 1) %>% .$n_weeks
-  it_check <- sprt_overlap_interv %>% filter(flag == 1) %>% .$n_weeks
+
+  if (identical(sprt_overlap_base %>% filter(flag == 1) %>% .$n_weeks, integer(0))) {
+    bt_check <- sprt_param$n_weeks
+  } else {
+    bt_check <- sprt_overlap_base %>% filter(flag == 1) %>% .$n_weeks
+  }
+  
+  if (identical(sprt_overlap_interv %>% filter(flag == 1) %>% .$n_weeks, integer(0))) {
+    it_check <- sprt_param$n_weeks
+  } else {
+    it_check <- sprt_overlap_interv %>% filter(flag == 1) %>% .$n_weeks
+  }
+  
   eob <- seq(block_params$block_unit, sprt_param$n_weeks, by = block_params$block_unit)
   
   return(eob[eob >= max(c(sprt_check, bt_check, it_check))][1])
@@ -201,17 +212,17 @@ get_timeline <- function(sprt_res, sprt_overlap_base, sprt_overlap_interv){
   sprt <- sprt_res %>% filter(flag == 1) %>% slice(1) %>% .$n_weeks
   
   if (identical(sprt, integer(0))) {
-    sprt <- 48
+    sprt <- sprt_param$n_weeks
   } 
   
   if (identical(sprt_overlap_base %>% filter(flag == 1) %>% .$n_weeks, integer(0))) {
-    base <- 48
+    base <- sprt_param$n_weeks
   } else {
     base <- sprt_overlap_base %>% filter(flag == 1) %>% .$n_weeks
   }
   
   if (identical(sprt_overlap_interv %>% filter(flag == 1) %>% .$n_weeks, integer(0))) {
-    interv <- 48
+    interv <- sprt_param$n_weeks
[…]

### 2025-04-27 code/1_analysis.R
Modified timeline and reruned analysis

@@ -55,7 +55,7 @@ source(paste0(function_path, "rand_seq.R"))
 
 # define parameters
 # section run control
-run_params <- list(type = "variable", 
+run_params <- list(type = "stable", 
                    sprt = T, 
                    sprt_cont = F, 
                    interval = F, 
@@ -189,8 +189,19 @@ run_interpo <- function(df_all){
 get_eob <- function(sprt_res, sprt_overlap_base, sprt_overlap_interv){
   
   sprt_check <- sprt_res %>% filter(flag == 1) %>% slice(1) %>% .$n_weeks
-  bt_check <- sprt_overlap_base %>% filter(flag == 1) %>% .$n_weeks
-  it_check <- sprt_overlap_interv %>% filter(flag == 1) %>% .$n_weeks
+  
+  if (identical(sprt_overlap_base %>% filter(flag == 1) %>% .$n_weeks, integer(0))) {
+    bt_check <- sprt_param$n_weeks
+  } else {
+    bt_check <- sprt_overlap_base %>% filter(flag == 1) %>% .$n_weeks
+  }
+  
+  if (identical(sprt_overlap_interv %>% filter(flag == 1) %>% .$n_weeks, integer(0))) {
+    it_check <- sprt_param$n_weeks
+  } else {
+    it_check <- sprt_overlap_interv %>% filter(flag == 1) %>% .$n_weeks
+  }
+  
   eob <- seq(block_params$block_unit, sprt_param$n_weeks, by = block_params$block_unit)
   
   return(eob[eob >= max(c(sprt_check, bt_check, it_check))][1])

### 2025-05-13 code/graphical.R
updated graphical timeline and median calculation

@@ -419,7 +419,7 @@ p_accuracy <- ggarrange(p_top, p_middle,
                         labels = c("a)", "b)"),
                         common.legend = T, 
                         legend = "bottom")
-  
+
 # Timeline plot
 df_time <- df_sprt_all_variable %>% 
   filter(seq != "final") %>% 
@@ -695,6 +695,8 @@ ggarrange(p_top, p_bottom,
   plot_annotation(title = "Absolute error in savings estimation by different M&V methods")
 
 ggsave(filename = "abs.png", path = fig_path, units = "in", height = 9, width = 12, dpi = 300)
+
+
 
 
 
@@ -1495,7 +1497,6 @@ ggsave(filename = "mean_interval_sprt.png", path = fig_path, units = "in", heigh
 
 
 
-
 #### TIME-INT ####
 # timeline to finish sequential test
 df_drop <- bind_rows(df_seq_interval_tl_drop_stable, df_seq_interval_tl_drop_variable) %>%
@@ -1510,7 +1511,7 @@ interval_1_keep <- bind_rows(df_sprt_all_stable, df_sprt_all_variable) %>%
   mutate(seq = as.factor(seq)) %>%
   mutate(interval = 1, 
          group = "keep")
-  
+
 df_keep <- bind_rows(df_seq_interval_tl_keep_stable, df_seq_interval_tl_keep_variable) %>%
   mutate(group = "keep", 
          temp = ifelse(base_temp >= interv_temp, base_temp, interv_temp)) %>% 
@@ -1522,12 +1523,14 @@ df_MW_A <- bind_rows(df_keep, df_drop) %>%
   filter(seq != "final") %>% 
   filter(seq == "eob") %>% 
   mutate(category = ifelse(n_weeks <= 12, 12, 
-                           ifelse(n_weeks > 12 & n_weeks <= 18, 18, 
-                                  ifelse(n_weeks > 18 & n_weeks <= 24, 24, 
-                                         ifelse(n_weeks > 24 & n_weeks <= 30, 30,
-                                                ifelse(n_weeks > 30 & n_weeks <= 36, 36,
-                                                       ifelse(n_weeks > 36 & n_weeks <= 42, 42, 
-                                                              ifelse(n_weeks > 42, 48, n_weeks)))))))) %>% 
+                           ifelse(n_weeks > 12 & n_weeks <= 18, 18,
+                                  ifelse(n_weeks > 18 & n_weeks <= 24, 24,
[…]

### 2025-05-13 code/graphical.R
updated graphical timeline plot for 6-week blocking period

@@ -1865,9 +1865,12 @@ df_MW_A <- bind_rows(df_keep, df_drop) %>%
   filter(seq != "final") %>% 
   filter(seq == "eob") %>% 
   mutate(category = ifelse(n_weeks <= 12, 12, 
-                           ifelse(n_weeks > 12 & n_weeks <= 24, 24, 
-                                  ifelse(n_weeks > 24 & n_weeks <= 48, 48, 
-                                         ifelse(n_weeks > 48, 60, n_weeks))))) %>% 
+                           ifelse(n_weeks > 12 & n_weeks <= 18, 18,
+                                  ifelse(n_weeks > 18 & n_weeks <= 24, 24,
+                                         ifelse(n_weeks > 24 & n_weeks <= 30, 30,
+                                                ifelse(n_weeks > 30 & n_weeks <= 36, 36, 
+                                                       ifelse(n_weeks > 36 & n_weeks <= 42, 42, 
+                                                              ifelse(n_weeks > 42, 48, n_weeks)))))))) %>% 
   select(-n_weeks) %>% 
   rename(n_weeks = category) %>% 
   mutate(seq = as.factor(seq), 
@@ -1884,10 +1887,12 @@ df_MW_A <- bind_rows(df_keep, df_drop) %>%
          n_weeks = as.factor(n_weeks), 
          n_weeks = recode_factor(n_weeks, 
                                  "12" = "12\nweeks", 
-                                 "24" = "24\nweeks", 
+                                 "18" = "18\nweeks",
+                                 "24" = "24\nweeks",
+                                 "30" = "30\nweeks",
                                  "36" = "36\nweeks", 
-                                 "48" = "48\nweeks", 
-                                 "60" = "60\nweeks")) %>% 
+                                 "42" = "42\nweeks", 
+                                 "48" = "48\nweeks")) %>% 
   group_by(n_weeks, group, interval) %>% 
   summarise(n = n()) %>%
   ungroup()
@@ -1914,7 +1919,7 @@ df_MW_A <- all_combinations %>%
 df_MW_A %>% 
   ggplot(aes(x = n_weeks, y = perc, fill = group)) +
   geom_bar(stat = "identity", position = "dodge") +
-  facet_wrap(~interval, nrow = 1) +
+  facet_wrap(~interval, nrow = 2) +
[…]

### 2025-05-14 code/manuscript.Rmd
updated manuscript for discussion

--- 2025-03-22 code/manuscript.Rmd
@@
 |                        |                                                                  |       |
-|------------------|------------------------------------|------------------|
+|-------------------|------------------------------------|-------------------|
 | Parameter              | Description                                                      | Value |
@@
 
+## Blocking period impact
+
+Selecting different blocking periods affects how frequently the a M&V
+analyst evaluats stopping criteria, thus influencing the duration of the
+measurement and verification (M&V) process. Here, we summarize the
+advantages and disadvantages of shorter versus longer blocking periods.
+As discussed in our previous study, the second step of this randomized
+M&V method involves choosing an appropriate blocking period to minimize
+potential biases caused by incomplete randomization. For most practical
+scenarios, we recommend a blocking period of approximately 12 to 16
+weeks, as it typically covers the duration of a season, ensuring a
+representative sampling of operational conditions. In general, longer
+blocking periods are advantageous when evaluating multiple intervention
+strategies or using sampling intervals greater than one day. In such
+cases, extended blocking ensures a balanced distribution of both
+day-of-week and outdoor weather conditions.
+
+Opting for a shorter blocking period, such as 6 weeks, has the potential
+benefit of allowing the stopping criteria to be met more rapidly. We
+plotted the percentage of buildings finishing all stopping criteria when
+the blocking period is 6 weeks in Figure \@ref(fig:6weekblock). As a
+result, when compared to Figure \@ref(fig:intervaltime), the only
+difference we found is when sampling at daily interval and not dropping
+non-consecutive days, 22% of buildings can get an estimate earlier. This
+indicates most buildings require 9 months to meet outdoor weather
+conditions starting from January.
+
+```{r 6weekblock, fig.cap = "Time required to satisfy all stopping criteria when using a 6-week blocking period"}
+
+knitr::include_graphics(paste0(fig_path, "6week_time.png"))
+```
+
+Sampling at a shorter blocking period also means that the analyst must
+perform the hypothesis tests more frequently, which can necessitate
+adjusting the significance level using methods such as the Bonferroni
+correction to control for increased type-I error risk. On the plus side,
+a shorter blocking period can reduce the adverse impact of missing or
+problematic data. For example, if an intervention requires fine-tuning
+after initial implementation, removing data from the first blocking
+period would be less detrimental compared to longer blocking periods,
+preserving the overall integrity of the experiment.
+
+## M&V starting time impact
+
+In this study, we consistently set the baseline period to begin on
+January 1st, which offers the advantage of capturing colder outdoor
+weather conditions early in the M&V process. However, when the baseline
+measurements misses the initial winter period, meeting the required
+weather condition criteria can potentially take a longer time. To assess
+this impact, we repeated the randomized M&V analysis for all buildings
+using a new start date in March. As shown in Figure  \@ref(fig:startT),
+the new M&V process results in a noticeable delay in satisfying the
+stopping criteria. Compared to Figure  \@ref(fig:intervaltime), where a
+January start allowed most buildings to finish within 36 weeks, a March
+start extended the required duration to 48 weeks for the majority of
+buildings. Furthermore, when a longer sampling interval is used or
+non-consecutive days are excluded (to avoid carryover effects), some
+buildings may require up to 60 weeks to complete the randomized M&V. As
+expected, to collect sufficient data representing winter conditions, the
+sampling may need to continue until the next January. However, as the
+overall data collection period extends, the influence of dropping
+non-consecutive days diminishes and becomes less of a concern.
+
+```{r startT, fig.cap = "Time required to satisfy all stopping criteria starting from March"}
+
+knitr::include_graphics(paste0(fig_path, "startT.png"))
+```
+
+Similarly, we evaluated the time required to satisfy all stopping
+criteria when using a shorter blocking period, as shown in Figure
+ \@ref(fig:6weekstartT). Compared to Figure  \@ref(fig:startT), a
+significant proportion of buildings completed the randomized M&V by week
+42 (mid-December), which is one blocking period earlier than having a
+12-week blocking period. This result further highlights that the overall
+duration of the randomized M&V process is strongly driven by outdoor
+weather. Accordingly, selecting an optimal blocking period should
+consider historical or typical meteorological weather file to
+effectively minimize the time required.
+
+```{r 6weekstartT, fig.cap = "Time required to satisfy all stopping criteria starting from March and using a shorter blocking period of 6 weeks"}
+
+knitr::include_graphics(paste0(fig_path, "6week_start_time.png"))
+```
+
 ## Limitations

### 2025-05-14 code/graphical.R

@@ -1899,7 +1899,7 @@ df_MW_A <- all_combinations %>%
 df_MW_A %>% 
   ggplot(aes(x = n_weeks, y = perc, fill = group)) +
   geom_bar(stat = "identity", position = "dodge") +
-  facet_wrap(~interval, nrow = 1) +
+  facet_wrap(~interval, nrow = 2) +
   geom_text(aes(x = n_weeks, y = perc + 2, label = paste0(round(perc, digits = 0), "%"), group = group), position = position_dodge(1)) +
   scale_fill_manual(values = ls_colors) +
   scale_color_manual(values = ls_colors) +
@@ -1914,4 +1914,4 @@ df_MW_A %>%
         legend.position = "bottom",
         plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
 
-ggsave(filename = "timeline_interval_sprt.png", path = fig_path, units = "in", height = 6, width = 9, dpi = 300)
+ggsave(filename = "timeline_interval_sprt.png", path = fig_path, units = "in", height = 8, width = 9, dpi = 300)

### 2025-05-14 code/graphical.R
adjusted timeline plot

@@ -1524,7 +1524,8 @@ df_MW_A <- bind_rows(df_keep, df_drop) %>%
                                  "12" = "12\nweeks", 
                                  "24" = "24\nweeks", 
                                  "36" = "36\nweeks", 
-                                 "48" = "48\nweeks")) %>% 
+                                 "48" = "48\nweeks", 
+                                 "60" = "60\nweeks")) %>% 
   group_by(n_weeks, group, interval) %>% 
   summarise(n = n()) %>%
   ungroup()
@@ -1551,7 +1552,7 @@ df_MW_A <- all_combinations %>%
 df_MW_A %>% 
   ggplot(aes(x = n_weeks, y = perc, fill = group)) +
   geom_bar(stat = "identity", position = "dodge") +
-  facet_wrap(~interval, nrow = 1) +
+  facet_wrap(~interval, nrow = 2) +
   geom_text(aes(x = n_weeks, y = perc + 2, label = paste0(round(perc, digits = 0), "%"), group = group), position = position_dodge(1)) +
   scale_fill_manual(values = ls_colors) +
   scale_color_manual(values = ls_colors) +
@@ -1566,4 +1567,4 @@ df_MW_A %>%
         legend.position = "bottom",
         plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
 
-ggsave(filename = "timeline_interval_sprt.png", path = fig_path, units = "in", height = 6, width = 12, dpi = 300)
+ggsave(filename = "timeline_interval_sprt.png", path = fig_path, units = "in", height = 8, width = 12, dpi = 300)

### 2025-06-08 code/manuscript.Rmd
Updated continous sampling ratio and manuscript

@@
 abstract: Conventional measurement and verification (M&V) methods rely on pre- and post- retrofit comparison to estimate energy savings.
-They are often time-consuming and unreliable, especially when non-routine events including static changes and gradual changes denoted as operational drift occurs during the M&V process.
+They are often time-consuming and unreliable, especially when non-routine events which are static changes or more gradual changes denoted as operational drift occur during the M&V process.
 Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies conventional M&V suggested by industry guidelines, the results can be largely confounded.
@@
 Additionally we found that when non-routine events are present, the randomized M&V approach estimates savings that are much closer to the ground-truth savings than the conventional M&V method, demonstrating much improved reliability.
-We also assessed the impact of dropping samples affected by carryover when switching between strategies, and identify the optimal sampling interval for that scenario using this large dataset.
+We also assessed the impact of starting the M&V at different time of the year, continuing randomization with different sampling ratio after satisfying all criteria, dropping samples affected by carryover when switching between strategies, and identify the optimal sampling interval for that scenario using this large dataset.
 
@@
                "Randomized\n(10/90)" = "#6baed6", 
+               "Continued\n(20/80)" = "#6baed6", 
+               "Continued\n(10/90)" = "#6baed6",
                "Daily\nsampling" = "#66c2a4",
@@
 To address the limitations of conventional M&V methods and the
-challenges posed by non-routine events, we proposed a M&V method that
-adopts the randomized crossover design [@raftery2024], a gold standard
-from medical, agricultural and online controlled experimental (i.e. A/B
-testing) studies [@blackston2019; @duan2013; @gupta2019; @mie2022;
-@munro; @raseduzzaman2017]. An early application of this randomized
-design to the field of M&V in buildings was first used in
+challenges posed by building baseline changes, we proposed a M&V method
+that adopts the randomized crossover design [@raftery2024], a gold
+standard from medical, agricultural and online controlled experimental
+(i.e. A/B testing) studies [@blackston2019; @duan2013; @gupta2019;
+@mie2022; @munro; @raseduzzaman2017]. An early application of this
+randomized design to the field of M&V in buildings was first used in
 [@raftery2018], which compared the energy performance on two different
@@
 usage as well as site outdoor weather conditions. In this study, we
-mostly use the metered electricity and outdoor weather measurements for
+mostly used the metered electricity and outdoor weather measurements for
 running the M&V methods.
@@
 addition, the expected savings should be larger than 10% and the
-baseline model fitting accuracy should have CV(RMSE) lower than 20%.
-Thus, most research related to M&V for whole-building approach focuses
-on the accuracy of baseline modeling, exploring model performance from
-simple regression models to more complex machine learning techniques.
-One study reviewed various models suitable for M&V applications as well
-as selected input features [@alrobaie_review_2022] and another study
-provided a definitive methodology to apply machine learning models for
-M&V use cases [@gallagher_development_2018]. In addition, a few studies
-investigated the critical performance metrics to evaluate the developed
-baseline models [@granderson_automated_2015;
-@granderson_development_2014] and compared a variety of models using
-those metrics [@granderson_accuracy_2016]. These studies made
-significant contributions by emphasizing the uncertainty associated with
-the model-fitting process, a key factor in accurately determining energy
+baseline energy model fitting accuracy should have CV(RMSE) lower than
+25% for 12-month data collection. Thus, most research related to M&V for
+whole-building approach focuses on the accuracy of baseline modeling,
+exploring model performance from simple regression models to more
+complex machine learning techniques. One study reviewed various models
+suitable for M&V applications as well as selected input features
+[@alrobaie_review_2022] and another study provided a definitive
+methodology to apply machine learning models for M&V use cases
+[@gallagher_development_2018]. In addition, a few studies investigated
+the critical performance metrics to evaluate the developed baseline
+models [@granderson_automated_2015; @granderson_development_2014] and
+compared a variety of models using those metrics
+[@granderson_accuracy_2016]. These studies made significant
+contributions by emphasizing the uncertainty associated with the
+model-fitting process, a key factor in accurately determining energy
 savings. Furthermore, other researchers addressed this issue by
@@
 
-### Non-routine events impact
-
-Non-routine events in M&V commonly refer to unexpected changes in a
-building that influence a building's energy usage and are typically
-unrelated to the intervention strategy. Current guidelines define
-non-routine events as 'static factors' that needed to be adjusted after
-projecting baseline in the post-retrofit period [@ashrae_ashrae_2023;
+### Changes in building baseline measurements
+
+Changes in a building that influences its energy usage are commonly
+refered to as non-routine events in M&V and are typically unrelated to
+the intervention strategy. Current guidelines define non-routine events
+as 'static factors' that needed to be adjusted after projecting baseline
+in the post-retrofit period [@ashrae_ashrae_2023;
 @efficiency_valuation_organisation_international_2012]. A common
@@
 @qiu_chilled_2022; @jin2007; @taylor_optimizing_2012]. In our case, we
-make it even simpler by adjusting the setpoint based on outdoor weather
+made it even simpler by adjusting the setpoint based on outdoor weather
 conditions with more description shown in Section
@@
 
-As a result, there are only 66 buildings were labeled in this 'stable'
-subset (out of total of 600 buildings) implying that the assumption that
-buildings are typically stable over the timeframe involved in
-conventional M&V is rarely true. As mentioned in Section
-\@ref(non-routine-events-impact), such variability is largely associated
-with non-routine events, especially operational drifts since the
-assessment last for two years. And according to the statistics shown
+As a result, only 66 buildings are labeled in this 'stable' subset (out
+of total of 600 buildings) implying the assumption that buildings are
+typically stable over the timeframe involved in conventional M&V is
+rarely true. As mentioned in Section \@ref(non-routine-events-impact),
+such variability is largely associated with non-routine events,
+especially operational drifts since the assessment, even under ideal
+situations, last for two years. And according to the statistics shown
 here, it is a typical case observed in real buildings applications.
@@
 Figure \@ref(fig:chwst) shows the algorithm for the proposed control
-intervention that reset the chiller supply temperature based on the
+intervention that resets the chiller supply temperature based on the
 outdoor weather conditions, which can be commonly found in the
@@
 |                        |                                                                  |       |
-|-------------------|------------------------------------|-------------------|
+|-------------------|----------------------------------|-------------------|
 | Parameter              | Description                                                      | Value |
@@
     data sampled by both strategies.
-
--   No carryover effect using a t-test with a p-value not exceeding a
-    defined significance threshold of 0.05.
 
@@
 
-```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period).", fig.height=12, fig.width=12}
+```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period).", fig.height=11, fig.width=12}
 
@@
   geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
-  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff, na.rm = T)) %>% ungroup(), 
-            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
+  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
+            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
+            size = 4) +
   scale_y_continuous(expand = c(0, 0), 
@@
        y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with no model normalization on weather")) +
+       subtitle = str_glue("All {A_building} buildings with measured weather")) +
   coord_cartesian(ylim = c(0, 23)) +
@@
   geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
-  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
-            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
+  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
+            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
+            size = 4) +
   scale_y_continuous(expand = c(0, 0), 
@@
        y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with TOWT model normalized on TMY weather")) +
+       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
   coord_cartesian(ylim = c(0, 23)) +
@@
        color = NULL, 
-       subtitle = "timeline comparison of satisfying all stopping criteria") +
+       subtitle = "timeline comparison") +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
 meteorological weather of the building location after using TOWT model.
-As a result, the conventional M&V method exhibits an average deviation
-of 6% in savings estimation, whereas the proposed randomized method
+As a result, the conventional M&V method exhibits a median deviation of
+5% in savings estimation, whereas the proposed randomized method
 demonstrates significantly smaller deviations. If the analyst stops
 immediately after satisfying all stopping criteria, the deviation is
-reduced to approximately 2%. Extending the M&V period to match the
+reduced to approximately 1.6%. Extending the M&V period to match the
 length of the conventional method (24 months) further improves accuracy
-to 1%. Furthermore, by comparing subplots (a) and (b), we observe a
-minor improvement in estimation accuracy when a model is fitted to the
-data to account for weather differences. The model fitting is most
-beneficial for the shorter dataset (randomized and stopping when all
-criteria are met, typically 24-36 weeks of data), where it improves
-estimation accuracy by 0.6%. This is because the weather conditions
+to 0.7% deviation. Furthermore, by comparing subplots (a) and (b), we
+observe a minor improvement in estimation accuracy when a model is
+fitted to the data to account for weather differences. The model fitting
+is most beneficial for the shorter dataset (randomized and stopping when
+all criteria are met, typically 24-36 weeks of data), where it improves
+estimation accuracy by 0.4%. This is because the weather conditions
 between the first and second year for each of the locations are
@@
 
-We also plotted the mean error in Figure \@ref(fig:meanerr) and we
-noticed that the uncertainty range associated with the conventional
-method is significantly larger compared to the randomized method that
-only takes 36 weeks according to subplot (c) in Figure
-\@ref(fig:abstract).
-
-```{r meanerr, fig.cap="Distribution of mean deviation from the ground-truth savings calculated by the two M&V methods", fig.height=6, fig.width=12}
+We also plotted the estimated deviation without absolute calculation in
+Figure \@ref(fig:meanerr) and we noticed that the uncertainty range
+associated with the conventional method is significantly larger compared
+to the randomized method that only takes 36 weeks according to subplot
+(c) in Figure \@ref(fig:abstract).
+
+```{r meanerr, fig.cap="Distribution of estimated deviation from the ground-truth savings calculated by the two M&V methods", fig.height=8, fig.width=12}
 # plot for the stable subset
@@
        x = NULL, 
-       y = "Mean error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with no model normalization on weather")) +
+       y = "Error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with measured weather")) +
   coord_cartesian(ylim = c(-20, 20)) +
@@
        x = NULL, 
-       y = "Mean error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with TOWT model normalized on TMY weather")) +
+       y = "Estimated error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
   coord_cartesian(ylim = c(-20, 20)) +
@@
           labels = c("a)", "b)")) +
-  plot_annotation(title = "Mean error in savings estimation by different M&V methods")
+  plot_annotation(title = "Estimated error in savings estimation by different M&V methods")
 
@@
 Additionally in the literature, we found another M&V study using over
-500+ commercial buildings, but their study scope was to compare the
-predictive performance of a variety of M&V baseline models from simple
-weekly mean to more complex machine learning methods using the
-normalized mean bias error (NMBE) [@granderson_accuracy_2016]. Since we
-also calculated the normalized savings estimation error, we compared our
-normalized accuracy for 600+ buildings with the TOWT model and TMY
-weather with their TOWT model prediction assessment in the following
-table, highlighting that the model and overall building datasets are
-performing similarly.
+500+ commercial buildings (on a different open-source dataset), but
+their study scope was to compare the predictive performance of a variety
+of M&V baseline models from simple weekly mean to more complex machine
+learning methods using the normalized mean bias error (NMBE)
+[@granderson_accuracy_2016]. Since we also calculated the normalized
+savings estimation error, we compared our normalized accuracy for 600+
+buildings with the TOWT model and TMY weather with their TOWT model
+prediction assessment in the following table, highlighting that the
+model and overall building datasets are performing similarly.
 
@@
 
-```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect when buildings are subject to static non-rountine events and gradual operational drift.", fig.height=6, fig.width=12}
+```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect when buildings are subject to static non-rountine events and gradual operational drift.", fig.height=5, fig.width=12}
 
@@
   geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
-  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
-            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
+  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
+            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
+            size = 4) +
   scale_y_continuous(expand = c(0, 0), 
@@
        x = NULL, 
-       y = "Mean effect in fractional savings",
-       title = "Mean effect of non-routine events when using different M&V methods", 
-       subtitle = str_glue("All {A_building} buildings with no model normalization on weather")) +
+       y = "Estimated effect in fractional savings",
+       title = "Estimated effect of non-routine events when using different M&V methods", 
+       subtitle = str_glue("All {A_building} buildings with measured weather")) +
   coord_cartesian(ylim = c(-18, 18)) +
@@
 deviation around -1.25%, our analysis on conventional method shows that
-the typical deviation is -1.3%, meaning that the typical building is
+the typical deviation is -1.5%, meaning that the typical building is
 decreasing energy use slightly in this dataset in the absence of any
@@
 intervention to this building and use conventional M&V to measure the
-effect, they would 'detect' 1.3% energy savings even when the
+effect, they would 'detect' 1.5% energy savings even when the
 intervention had no actual impact on energy performance.
@@
 To evaluate modeling accuracy, we used the Coefficient of Variation of
-Root-Mean Squared Error (CV(RMSE)) as the primary error metric. Since
-CV(RMSE) is a normalized measure, it enables direct comparison across
-different model-fitting results. According to ASHRAE Guideline 14,
-whole-building baseline models should achieve a CV(RMSE) below 25% for
-computing savings or 30% when less than 12 months of measurements are
-available [@ashrae_ashrae_2023]. In our study, Figure \@ref(fig:towtacc)
-compares the model fitting accuracy of the two M&V methods and it shows
-around 87.5% of the models fitted in the randomized M&V method are code
-compliant while 80% for the conventional M&V method. As the quantity of
-the training set is the same, this highlights the impact of data
-sampling on model fitting accuracy, and for TOWT, the distribution of
-outdoor weather conditions sampled in the training dataset is
-influential. The conventional method trains the model using the entire
-first-year dataset, if the location is heating or cooling dominant, the
-data-driven model is trained using colder or warmer outdoor conditions
-leading to biased model coefficients. Although the random sampling in
-our study does not specifically block the outdoor temperature factor,
-the CV(RMSE) distribution shown in the figure indicates a significant
-improvement in model fitting accuracy.
-
-```{r towtacc, fig.cap="TOWT model fitting accuracy distribution for all buildings included (with each data point representing one building)", fig.width=8, fig.height=4}
+Root-Mean Squared Error (CV(RMSE)) as the primary error metric. CV(RMSE)
+is a normalized measure, thus it enables direct comparison across
+different modeling results. Since we have the ground-true baseline
+measurements in the post-retrofit period, we can assess the error caused
+by model adjustment (i.e. how accurate is the 'counterfactual' baseline
+provided by the two M&V methods). Figure \@ref(fig:towtacc) shows the
+difference between the true post-retrofit baseline and adjusted baseline
+through TOWT modeling and since the quantity of the training set is the
+same, this highlights the impact of the data sampling technique. For
+example, if a non-routine event happens, such as a floor of tenant
+moving out, which influences electricity demand but are independent from
+outdoor weather conditions and time-of-week factors, the TOWT model can
+yield significant deviations. This is particularly the case for the
+conventional M&V method since it samples continuously throughout the
+pre-retrofit period. However, randomized sampling is less impacted by
+the risk since it samples only 50% of the pre-retrofit baseline
+throughout a blocking period. Consequently, we noticed as shown in the
+figure, there is a noticeable improvement in modeling accuracy when
+using the randomized method.
+
+```{r towtacc, fig.cap="TOWT modeling accuracy distribution for all buildings included (with each data point representing one building)", fig.width=8, fig.height=4}
 
@@
             aes(x = method, y = mean, label = paste0(round(mean, digits = 0), " %"))) +
-  geom_hline(yintercept = 25, color = "red", lty = "dashed") +
-  annotate(geom = "text", 
-           x = 0.5, 
-           y = 27, 
-           size = 4, 
-           color = "red", 
-           label = paste0(25, " %")) +
   scale_fill_manual(values = ls_colors) +
@@
 #### ABS-INT ####
+# sprt sequence plot for different sampling intervals
 df_seq_interval_fs_keep_stable %<>% 
@@
   geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8, linewidth = 0.15) +
   geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
-  geom_text(data = . %>% 
-              group_by(group, interval) %>% 
-              summarise(mean = mean(diff, inf.rm = T)) %>% 
-              ungroup(), 
-            aes(x = interval, 
-                y = mean, group = group,
-                label = paste0(round(mean, digits = 1), " %")), 
+  geom_text(data = . %>% group_by(group, interval) %>% summarise(median = median(diff, inf.rm = T)) %>% ungroup(), 
+            aes(x = interval, y = median, group = group,
+                label = paste0(round(median, digits = 1), " %")), 
+            position = position_dodge(width = 0.75), 
             color = "white", 
-            position = position_dodge(width = 0.75)) +
+            size = 4) +
   scale_y_continuous(expand = c(0, 0), 
@@
   scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
+  scale_color_manual(values = c("grey80", "grey80")) + 
   labs(fill = NULL, 
@@
        y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with no model normalization on weather")) +
+       subtitle = str_glue("All {A_building} buildings with measured weather")) +
   coord_cartesian(ylim = c(0, 15)) +
@@
   geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8, linewidth = 0.15) +
   geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
-  geom_text(data = . %>% 
-              group_by(group, interval) %>% 
-              summarise(mean = mean(diff, na.rm = T)) %>% 
-              ungroup(), 
-            aes(x = interval, 
-                y = mean, 
-                group = group,
-                label = paste0(round(mean, digits = 1), " %")),
+  geom_text(data = . %>% group_by(group, interval) %>% summarise(median = median(diff, na.rm = T)) %>% ungroup(), 
+            aes(x = interval, y = median, group = group,
+                label = paste0(round(median, digits = 1), " %")), 
+            position = position_dodge(width = 0.75), 
             color = "white", 
-            position = position_dodge(width = 0.75)) +
+            size = 4) +
   scale_y_continuous(expand = c(0, 0), 
@@
        y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with TOWT model normalized on TMY weather")) +
+       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
   coord_cartesian(ylim = c(0, 15)) +
@@
   plot_annotation(title = "Absolute error in savings estimation of randomized M&V after satisfying all stopping criteria")
-
 ```
@@
 when calculating the distribution shown in the darker-colored set. The
-results suggest a gradual increase in error from 2.4% to 3.7% when the
-sampling interval increases from daily to weekly intervals. This is due
-to the fact that increasing the sample interval means reducing
-randomization. To quantify the combined impact of data loss due to
-preventing the carryover effect, we generated the lighter-colored set,
-which drops all non-consecutive days. As a result, we noticed a
-significant increase in estimation error when sampling daily, which led
-to a similar accuracy compared to weekly intervals. Therefore,
-considering this trade-off, we recommend using a two-day sampling
-interval if the carryover effect is likely but expected to last less
-than one day, and then normalizing the estimation via energy modeling.
-We also included the mean deviation distribution and the accuracy
-distribution calculated after all stopping criteria were satisfied in
-the supplementary material and the results show the same indication.
-
-```{r intervaltime, fig.cap="Comparison of time required to satisfy all stopping criteria when considering dropping non-consecutive days (or not)", fig.height=6, fig.width=12}
-# timeline to finish sequential test
+results suggest without dropping non-consecutive days in subplot (a), a
+gradual increase in error from 1.6% to 2.3% (1.2% to 1.9% with TOWT
+model normalization in subplot (b)) when the sampling interval increases
+from daily to weekly intervals. This is due to the fact that increasing
+the sample interval means reducing randomization. To quantify the
+combined impact of data loss due to preventing the carryover effect, we
+generated the lighter-colored set, which drops all non-consecutive days.
+As a result, we noticed an increase in estimation error when sampling
+daily to 2.4% (to 1.9% when TOWT is fitted), which led to a similar
+accuracy compared to weekly intervals. Therefore, considering this
+trade-off, we recommend using a two-day sampling interval if the
+carryover effect is likely but expected to last less than one day, and
+then normalizing the estimation via energy modeling. We also included
+the deviation distribution without absolute calculation and the
+deviation distribution calculated throughout the two years in the
+supplementary material and the results show the same indication.
+
+```{r intervaltime, fig.cap="Comparison of time required to satisfy all stopping criteria when considering dropping non-consecutive days (or not)", fig.height=8, fig.width=12}
 df_drop <- bind_rows(df_seq_interval_tl_drop_stable, df_seq_interval_tl_drop_variable) %>%
@@
   geom_bar(stat = "identity", position = "dodge") +
-  facet_wrap(~interval, nrow = 1) +
-  geom_text(aes(x = n_weeks, y = perc + 2, label = paste0(round(perc, digits = 0), "%"), group = group), position = position_dodge(1)) +
+  facet_wrap(~interval, nrow = 2) +
+  geom_text(aes(x = n_weeks, y = perc + 2, label = paste0(round(perc, digits = 0), "%"), group = group), 
+            position = position_dodge(1), 
+            size = 4) +
   scale_fill_manual(values = ls_colors) +
@@
        fill = NULL,
-       title = "Percentage of buildings using randomized M&V when satisfying all stopping critiera", 
+       title = "Percentage of buildings satisfying all stopping critiera", 
        subtitle = "Counted at the end of each 12-week blocking period") +
@@
 particularly at shorter intervals. For instance, when sampling at a
-one-day interval, 90% of buildings meet all criteria within 24 weeks if
+one-day interval, 89% of buildings meet all criteria within 24 weeks if
 no data are dropped, but this number decreases significantly to 52% when
-non-consecutive days (account for 1/3 of the data) are excluded. As the
-sampling interval increases, this difference gradually diminishes: when
-sampling at a weekly interval, removing non-consecutive days has no
-impact on the timeline of meeting stopping criteria.
+non-consecutive days (account for 1/3 of the data) are excluded, and 3%
+of the buildings require another 12 weeks to satisfy the weather
+criterion. As the sampling interval increases, this difference gradually
+diminishes: when sampling at a weekly interval, removing non-consecutive
+days has no impact on the timeline of meeting stopping criteria.
 
@@
 
-Another advantage of using randomized M&V is the flexibility of changing
-the sampling ratio after the target savings are detected. For instance,
-the building owner can continue sampling at a 50%/50% ratio between the
-baseline and the intervention to further reduce the uncertainty
-associated with the savings. Alternatively, they could switch to 100%
-intervention to maximize cost savings, though this approach risks the
-baseline becoming outdated over time, which is another trade-off to
-consider. A middle-ground approach may be to continue sampling at a
-20%/80% ratio or 10%/90% between the baseline and the intervention. To
-compare those possible choices to a building owner, figure
-\@ref(fig:abscont) shows the results of changing the original 50%/50%
-sampling ratio after satisfying all stopping criteria and continuing a
-new ratio till the end of the two-year period. Those results were
-calculated without dropping non-consecutive days and include the initial
-50%/50% sampled data. On the far left, we also provided the case when
-the building owner continues with no change in sampling ratio. The
-results show that as the sampled baseline days and intervention days
-become unbalanced, the savings estimation error grows from 1% to 5%.
-This is reasonable since sampling at 10%/90% is similar to the
-conventional M&V method. And on the far right, we show again the
-accuracy of the conventional M&V method, which is still worse then
-sampling at 10%/90%. Also, We include the mean deviation version in the
-supplementary material.
+In our previous study [@raftery2024], we recommended that building
+owners adjust the sampling ratio after meeting all stopping criteria, as
+continuing baseline sampling helps maintain M&V accuracy but may reduce
+the potential for energy savings. To further examine its impact on M&V
+accuracy, Figure \@ref(fig:abscont) compares two sampling ratios
+(20%/80% and 10%/90%) and two assocaited scenarios. The "Continued"
+group follows the standard randomized M&V framework with an initial
+50%/50% sampling ratio, then switches to the indicated ratio after
+satisfying the stopping criteria, continuing this until the end of the
+two-year period. In contrast, the "Randomized" group uses the indicated
+ratio from the outset and continues throughout the two year. Subplot (a)
+shows that absolute error is significantly lower without the initial
+50%/50% sampling, due to baseline drift and non-routine events that most
+buildings experience over two years (as noted in Section
+\@ref(filter-and-clean-dataset)). Subplot (b) confirms this finding even
+after weather normalization.
 
@@
 #### ABS-CONT ####
-conv_S <- df_fs_stable %>% 
-  filter(method != "rand") %>% 
-  pivot_wider(names_from = method, values_from = savings) %>% 
-  mutate(diff = abs(true - conv), 
-         ratio = "conv") %>% 
-  select(name, ratio, diff)
-
-rand_cont_S <- df_cont_stable %>% 
-  select(-cont_tmy) %>% 
+df_MW_S <- df_cont_stable %>% 
   left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = abs(savings - cont_fs)) %>% 
-  select(name, diff, ratio)
-
-rand_final_S <- df_fs_stable %>% 
-  filter(method != "conv") %>% 
-  pivot_wider(names_from = method, values_from = savings) %>% 
-  mutate(diff = abs(true - rand), 
-         ratio = 0) %>% 
-  select(name, diff, ratio)
-
-df_MW_S <- rbind(rand_cont_S, rand_final_S, conv_S)
+  mutate(diff = abs(savings - cont_fs), 
+         diff_all = abs(savings - cont_fs_all)) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio)))))
 
 # plot for the variable subset
-conv_V <- df_fs_variable %>% 
-  filter(method != "rand") %>% 
-  pivot_wider(names_from = method, values_from = savings) %>% 
-  mutate(diff = abs(true - conv), 
-         ratio = "conv") %>% 
-  select(name, ratio, diff)
-
-rand_cont_V <- df_cont_variable %>% 
-  select(-cont_tmy) %>% 
+df_MW_V <- df_cont_variable %>% 
   left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = abs(savings - cont_fs)) %>% 
-  select(name, diff, ratio)
-
-rand_final_V <- df_fs_variable %>% 
-  filter(method != "conv") %>% 
-  pivot_wider(names_from = method, values_from = savings) %>% 
-  mutate(diff = abs(true - rand), 
-         ratio = 0) %>% 
-  select(name, diff, ratio)
-
-df_MW_V <- rbind(rand_cont_V, rand_final_V, conv_V)
-
+  mutate(diff = abs(savings - cont_fs), 
+         diff_all = abs(savings - cont_fs_all)) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio)))))
+
+# plot for combined dataset
 df_MW_A <- rbind(df_MW_S, df_MW_V)
 
-df_MW_A %>% 
+p_top <- df_MW_A %>% 
   mutate(ratio = as.factor(ratio), 
-         ratio = recode_factor(ratio, "0" = "Randomized\n(50/50)", "1" = "Randomized\n(20/80)", "2" = "Randomized\n(10/90)", "conv" = "Conventional")) %>% 
-  ggplot(aes(x = ratio, y = diff, fill = ratio)) +
+         ratio = recode_factor(ratio, "1_1" = "Continued\n(20/80)", "1_2" = "Randomized\n(20/80)", "2_1" = "Continued\n(10/90)", "2_2" = "Randomized\n(10/90)")) %>% 
+  ggplot(aes(x = ratio, y = values, fill = ratio)) +
   geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
@@
               group_by(ratio) %>% 
-              summarise(mean = mean(diff)) %>% 
+              summarise(median = median(values)) %>% 
               ungroup(), 
-            aes(x = ratio, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
+            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %"))) +
   scale_y_continuous(expand = c(0, 0), 
@@
        y = "Absolute error in fractional savings", 
-       title = "Absolute error in savings estimation fo randomized M&V using different sampling ratio",
-       subtitle = str_glue("All {A_building} buildings with with no model normalization on weather")) +
-  coord_cartesian(ylim = c(0, 20)) +
+       subtitle = str_glue("All {A_building} buildings with measured weather conditions")) +
+  coord_cartesian(ylim = c(0, 15)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "none",
+        axis.text.x = element_blank(), 
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+# TMY
+# plot for the stable subset
+df_TW_S <- df_cont_stable %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - cont_tmy), 
+         diff_all = abs(savings - cont_tmy_all)) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio)))))
+
+# plot for the variable subset
+df_TW_V <- df_cont_variable %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - cont_tmy), 
+         diff_all = abs(savings - cont_tmy_all)) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio)))))
+
+# plot for combined dataset
+df_TW_A <- rbind(df_TW_S, df_TW_V)
+
+p_bottom <- df_TW_A %>% 
+  mutate(ratio = as.factor(ratio), 
+         ratio = recode_factor(ratio, "1_1" = "Continued\n(20/80)", "1_2" = "Randomized\n(20/80)", "2_1" = "Continued\n(10/90)", "2_2" = "Randomized\n(10/90)")) %>% 
+  ggplot(aes(x = ratio, y = values, fill = ratio)) +
+  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
+  geom_lv(k = 4, outlier.shape = NA) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio)) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% 
+              group_by(ratio) %>% 
+              summarise(median = median(values)) %>% 
+              ungroup(), 
+            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %"))) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80")) + 
+  labs(fill = NULL, 
+       x = NULL, 
+       y = "Absolute error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
+  coord_cartesian(ylim = c(0, 15)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
         plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+ggarrange(p_top, p_bottom, 
+          ncol = 1, nrow = 2,
+          labels = c("a)", "b)")) +
+  plot_annotation(title = "Absolute error in savings estimation by different sampling ratio")
 ```
 
+To further illustrate this point, we plotted results in Figure
+\@ref(fig:stablecont) using only the \'stable\' building subset (66
+buildings), as defined in Section \@ref(filter-and-clean-dataset). The
+differences between sampling strategies are much reduced in this subset
+and can be largely corrected through weather normalization modeling.
+This suggests that, consistent with our earlier findings regarding the
+convention M&V method, initial baseline measurements provide limited
+value in maintaining M&V accuracy when the building\'s baseline
+performance changes over time.
+
+```{r stablecont, fig.cap="Comparison of error distribution under different sampling ratio only for the buildings in 'stable' subset", fig.height=6, fig.width=12}
+p_top <- df_MW_S %>% 
+  mutate(ratio = as.factor(ratio), 
+         ratio = recode_factor(ratio, "1_1" = "Continued\n(20/80)", "1_2" = "Randomized\n(20/80)", "2_1" = "Continued\n(10/90)", "2_2" = "Randomized\n(10/90)")) %>% 
+  ggplot(aes(x = ratio, y = values, fill = ratio)) +
+  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
+  geom_lv(k = 4, outlier.shape = NA) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio)) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% 
+              group_by(ratio) %>% 
+              summarise(median = median(values)) %>% 
+              ungroup(), 
+            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %"))) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80")) + 
+  labs(fill = NULL, 
+       x = NULL, 
+       y = "Absolute error in fractional savings", 
+       subtitle = str_glue("All {S_building} buildings with measured weather conditions")) +
+  coord_cartesian(ylim = c(0, 12)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "none",
+        axis.text.x = element_blank(), 
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+p_bottom <- df_TW_S %>% 
+  mutate(ratio = as.factor(ratio), 
+         ratio = recode_factor(ratio, "1_1" = "Continued\n(20/80)", "1_2" = "Randomized\n(20/80)", "2_1" = "Continued\n(10/90)", "2_2" = "Randomized\n(10/90)")) %>% 
+  ggplot(aes(x = ratio, y = values, fill = ratio)) +
+  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
+  geom_lv(k = 4, outlier.shape = NA) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio)) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% 
+              group_by(ratio) %>% 
+              summarise(median = median(values)) %>% 
+              ungroup(), 
+            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %"))) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80")) + 
+  labs(fill = NULL, 
+       x = NULL, 
+       y = "Absolute error in fractional savings", 
+       subtitle = str_glue("All {S_building} buildings with TOWT model and TMY weather")) +
+  coord_cartesian(ylim = c(0, 12)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "none",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+ggarrange(p_top, p_bottom, 
+          ncol = 1, nrow = 2,
+          labels = c("a)", "b)")) +
+  plot_annotation(title = "Absolute error for the subset buildings with 2-year stable usage")
+```
+
+However, as shown in Figure \@ref(fig:abstract), the highest accuracy is
+achieved when sampling at a 50%/50% ratio. Therefore, we continue to
+recommend that building owners follow the proposed framework and
+maintain a 50%/50% sampling ratio until all stopping criteria are met.
+The additional results from Figure \@ref(fig:abscont) and Figure
+\@ref(fig:stablecont) further adds to the guidance: if building baseline
+changes, retrofit savings estimation should simply discard the existing
+baseline measurements and revert back to 50%/50% sampling ratio as
+described in the proposed framework.
+
 ## Blocking period impact
 
-Selecting different blocking periods affects how frequently the a M&V
-analyst evaluats stopping criteria, thus influencing the duration of the
-measurement and verification (M&V) process. Here, we summarize the
-advantages and disadvantages of shorter versus longer blocking periods.
-As discussed in our previous study, the second step of this randomized
-M&V method involves choosing an appropriate blocking period to minimize
-potential biases caused by incomplete randomization. For most practical
-scenarios, we recommend a blocking period of approximately 12 to 16
-weeks, as it typically covers the duration of a season, ensuring a
-representative sampling of operational conditions. In general, longer
-blocking periods are advantageous when evaluating multiple intervention
-strategies or using sampling intervals greater than one day. In such
-cases, extended blocking ensures a balanced distribution of both
-day-of-week and outdoor weather conditions.
+Selecting different blocking periods affects how frequently the M&V
+analyst evaluates stopping criteria and may influence the decision when
+a target saving is detected. Here, we summarize the advantages and
+disadvantages of shorter versus longer blocking periods. As discussed in
+our previous study, the second step of this randomized M&V method
+involves choosing an appropriate blocking period to minimize potential
+biases caused by incomplete randomization. For most practical scenarios,
+we recommend a blocking period of approximately 12 to 16 weeks, as it
+typically covers the duration of a season, ensuring a representative
+sampling of operational conditions. In general, longer blocking periods
+are advantageous when evaluating multiple intervention strategies or
+using sampling intervals greater than one day. In such cases, extended
+blocking ensures a balanced distribution of both day-of-week and outdoor
+weather conditions.
 
 Opting for a shorter blocking period, such as 6 weeks, has the potential
-benefit of allowing the stopping criteria to be met more rapidly. We
-plotted the percentage of buildings finishing all stopping criteria when
-the blocking period is 6 weeks in Figure \@ref(fig:6weekblock). As a
+benefit of allowing the stopping criteria to be met earlier. We plotted
+the percentage of buildings finishing all stopping criteria when the
+blocking period is set as 6 weeks in Figure \@ref(fig:6weekblock). As a
 result, when compared to Figure \@ref(fig:intervaltime), the only
 difference we found is when sampling at daily interval and not dropping
-non-consecutive days, 22% of buildings can get an estimate earlier. This
-indicates most buildings require 9 months to meet outdoor weather
-conditions starting from January.
+non-consecutive days, 12% of buildings can get an estimate earlier. This
+means those buildings are likely to meet all stopping criteria between
+12 weeks and 18 weeks but when sampled over a 12-week block, the savings
+can only be determined at week 24. Although sampled baseline and
+intervention days can be different when the blocking period is changed,
+the results indicate most buildings still require 9 months to meet
+outdoor weather conditions when starting M&V from January.
 
@@
 
-In this study, we consistently set the baseline period to begin on
-January 1st, which offers the advantage of capturing colder outdoor
-weather conditions early in the M&V process. However, when the baseline
-measurements misses the initial winter period, meeting the required
-weather condition criteria can potentially take a longer time. To assess
-this impact, we repeated the randomized M&V analysis for all buildings
-using a new start date in March. As shown in Figure  \@ref(fig:startT),
-the new M&V process results in a noticeable delay in satisfying the
-stopping criteria. Compared to Figure  \@ref(fig:intervaltime), where a
-January start allowed most buildings to finish within 36 weeks, a March
-start extended the required duration to 48 weeks for the majority of
-buildings. Furthermore, when a longer sampling interval is used or
-non-consecutive days are excluded (to avoid carryover effects), some
-buildings may require up to 60 weeks to complete the randomized M&V. As
-expected, to collect sufficient data representing winter conditions, the
-sampling may need to continue until the next January. However, as the
-overall data collection period extends, the influence of dropping
+In this study, we set the baseline period to begin on January 1st, which
+offers the advantage of capturing colder outdoor weather conditions
+early in the M&V process. However, when the baseline measurements misses
+the initial winter period, meeting the required weather condition
+criteria can potentially take a longer time. To assess this impact, we
+repeated the randomized M&V analysis for all buildings using a new start
+date in March. As shown in Figure \@ref(fig:startT), the new M&V process
+results in a noticeable delay in satisfying the stopping criteria.
+Compared to Figure \@ref(fig:intervaltime), where a January start
+allowed most buildings to finish within 36 weeks, a March start extended
+the required duration to 48 weeks for the majority of buildings.
+Furthermore, when a longer sampling interval is used or non-consecutive
+days are excluded (to avoid carryover effects), some buildings may
+require up to 60 weeks to complete the randomized M&V. As expected, to
+collect sufficient data representing winter conditions, the sampling may
+need to continue until the next January (around week 48). However, as
+the overall data collection period extends, the influence of dropping
 non-consecutive days diminishes and becomes less of a concern.
@@
 42 (mid-December), which is one blocking period earlier than having a
-12-week blocking period. This result further highlights that the overall
-duration of the randomized M&V process is strongly driven by outdoor
-weather. Accordingly, selecting an optimal blocking period should
-consider historical or typical meteorological weather file to
-effectively minimize the time required.
+12-week blocking period. In both figures, we noticed a small proportion
+of buildings are less affected by changing the M&V start time, they are
+mostly located in climate zones with less outdoor temperature
+variations. This result further highlights that the overall duration of
+the randomized M&V process is strongly driven by outdoor weather.
+Accordingly, selecting an optimal blocking period should consider
+historical or typical meteorological weather file to effectively
+minimize the time required.
 
@@
 full range of baseline and intervention measurements under normal
-operating conditions, typically taking 6-9 months for each phase. Most
+operating conditions, typically taking 6 - 9 months for each phase. Most
 importantly, we verified that with a reduced M&V timeline, the
 randomized method can estimate savings more accurately by showing that
-the absolute error is only 1 - 2 % for a typical building. Whereas for
+the absolute error is only 1 - 2% for a typical building. Whereas for
 the much longer two-year conventional method, the estimation error was
-approximately 6% in a typical building. We also evaluated the impact of
+approximately 5% in a typical building. We also evaluated the impact of
 non-routine events especially more gradual and subtle operational drift
@@
 negligible impact on the savings estimated using the randomized method,
-demonstrating robustness to this issue. Lastly, we consider scenarios
-where there is a known carryover effect from switching between
+demonstrating robustness to this issue. Furthermore, we consider
+scenarios where there is a known carryover effect from switching between
 strategies in the building and assess the impact of dropping samples
@@
 Through that process, we showed that for the typical building, if a
-carryover is likely to be present, then using a 3-day sampling interval
-and dropping non-consecutive samples yields the optimal design.
+carryover is likely to be present, then using a 2-day sampling interval
+and dropping non-consecutive samples yields the optimal design. Lastly,
+we investigated the time required to complete the randomized M&V method
+under different scenarios. We found measuring a sufficient range of
+outdoor weather condition remains the most stringent criterion. But even
+for the most non-ideal case, the randomized method can still produce an
+estimate quicker than 2 years.
 
@@
 random sampling. We also provide a detailed assessment of using
-different sampling ratios, those results provide useful considerations
-to building analysts when generating randomized schedules for a variety
-of use cases.
+different sampling ratios, or a blocking period, those results provide
+useful considerations to building analysts when generating randomized
+schedules for a variety of use cases.
 

### 2025-06-08 code/1_analysis.R

@@ -55,9 +55,9 @@ source(paste0(function_path, "rand_seq.R"))
 
 # define parameters
 # section run control
-run_params <- list(type = "stable", 
+run_params <- list(type = "variable", 
                    sprt = T, 
-                   sprt_cont = F, 
+                   sprt_cont = T, 
                    interval = F, 
                    null = F)
 
@@ -423,32 +423,32 @@ for (n in 1:(nrow(all_names))){
                                 "site" = site, 
                                 "conv" = cv_rmse)
   
-  # base_proj %>%
-  #   ggplot() +
-  #   geom_point(aes(x = datetime, y = eload, color = "Measurement"), alpha = 0.2, size = 0.2) +
-  #   geom_point(aes(x = datetime, y = towt, color = "Prediction"), alpha = 0.2, size = 0.2) +
-  #   geom_smooth(aes(x = datetime, y = eload, color = "Measurement"), formula = y ~ x, method = "loess", linewidth = 0.7) +
-  #   geom_smooth(aes(x = datetime, y = towt, color = "Prediction"), formula = y ~ x, method = "loess", linewidth = 0.7) +
-  #   annotate(geom = "text",
-  #            x = median(base_proj$datetime),
-  #            y = median(base_proj$eload) + 2 * sd(base_proj$eload),
-  #            label = paste0("CV(RMSE): ", round(cv_rmse, digits = 2), "%")) +
-  #   scale_color_brewer(palette = "Set1") +
-  #   scale_x_datetime(date_breaks = "2 months",
-  #                    date_labels = "%b")  +
-  #   scale_y_continuous(expand = c(0, 0),
-  #                      breaks = breaks_pretty(n = 3),
-  #                      labels = number_format(suffix = " kW")) +
-  #   labs(x = NULL,
-  #        y = NULL,
-  #        color = NULL,
-  #        title = "TOWT model prediction results") +
-  #   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-  #         legend.direction = "horizontal",
-  #         legend.position = "bottom",
-  #         plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-  # 
-  # ggsave(filename = "towt_acc.png", path = sitefigs_path, units = "in", height = 6, width = 10, dpi = 300)
+  base_proj %>%
+    ggplot() +
+    geom_line(aes(x = datetime, y = eload, color = "Measurement"), alpha = 0.2, linewidth = 0.2) +
[…]

### 2025-06-08 code/graphical.R

@@ -312,8 +312,8 @@ p_top <- df_MW_A %>%
   geom_lv(k = 4, outlier.shape = NA) +
   geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
   geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
-  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
-            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
+  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
+            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %"))) +
   scale_y_continuous(expand = c(0, 0), 
                      breaks = breaks_pretty(n = 4), 
                      labels = number_format(suffix = " %")) +
@@ -400,8 +400,8 @@ p_middle <- df_TW_A %>%
   geom_lv(k = 4, outlier.shape = NA) +
   geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
   geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
-  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
-            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
+  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
+            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %"))) +
   scale_y_continuous(expand = c(0, 0), 
                      breaks = breaks_pretty(n = 4), 
                      labels = number_format(suffix = " %")) +
@@ -528,8 +528,8 @@ p1 <- df_MW_S %>%
   geom_lv(k = 4, outlier.shape = NA) +
   geom_boxplot(outlier.shape = NA, coef = 0, fill = "#00000000", aes(color = method)) +
   geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
-  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
[…]

### 2025-06-13 code/supplementary.Rmd
updated manuscript for review and created suppl rmd file

--- 2025-06-08 code/manuscript.Rmd
@@
   - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
-abstract: Conventional measurement and verification (M&V) methods rely on pre- and post- retrofit comparison to estimate energy savings.
-They are often time-consuming and unreliable, especially when non-routine events which are static changes or more gradual changes denoted as operational drift occur during the M&V process.
-Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies conventional M&V suggested by industry guidelines, the results can be largely confounded.
-In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method which randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily).
-We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention.
-Our results show that compared to the conventional method, the randomized method provides faster and more accurate savings estimations.
-Additionally we found that when non-routine events are present, the randomized M&V approach estimates savings that are much closer to the ground-truth savings than the conventional M&V method, demonstrating much improved reliability.
-We also assessed the impact of starting the M&V at different time of the year, continuing randomization with different sampling ratio after satisfying all criteria, dropping samples affected by carryover when switching between strategies, and identify the optimal sampling interval for that scenario using this large dataset.
+abstract: Conventional measurement and verification (M&V) methods rely on comparing pre- and post-retrofit performance to estimate energy savings.
+They are often time-consuming and unreliable, especially when non-routine events (step changes or more gradual changes in how a building operates) occur during the M&V process.
+Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies the conventional M&V method, the results can be confounded.
+In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method that randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily).
+We tested this novel randomized M&V method on a large public dataset (639 buildings) covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention.
+The results show that compared to the conventional method, the randomized method provides more accurate savings estimations with a 74% reduction in error and is even faster.
+Additionally, we found that when non-routine events are present, the randomized method estimates savings that are much closer to the ground-truth savings than the conventional method, demonstrating much improved reliability.
+We also assessed the impact of normalizing for different weather, starting the M&V at different time of the year, continuing randomization with a different sampling ratio after satisfying all criteria, dropping samples affected by carryover when switching between strategies, and identified the optimal sampling interval for that scenario using this large dataset.
 
@@
                "Randomized\n(10/90)" = "#6baed6", 
-               "Continued\n(20/80)" = "#6baed6", 
-               "Continued\n(10/90)" = "#6baed6",
+               "Continued\n(20/80)" = "#c6dbef", 
+               "Continued\n(10/90)" = "#c6dbef",
                "Daily\nsampling" = "#66c2a4",
@@
 study, we will focus on the energy savings quantified at the
-whole-building level where the measurements are obtained from utility
+whole-building level, where the measurements are obtained from utility
 bills or whole-building meters, however the overall conclusions would
@@
 challenging. Consequently, those limitations reduce the feasibility of
-M&V and complicates the quantification of estimation uncertainty, thus
+M&V and complicate the quantification of estimation uncertainty, thus
 impacting the accuracy and timeliness of savings assessments.
@@
 randomized design to the field of M&V in buildings was first used in
-[@raftery2018], which compared the energy performance on two different
+[@raftery2018], which compared the energy performance of two different
 supply air temperature reset strategies in a building. Subsequent
 improvements to the M&V method include adding a sequential evaluation
-framework and defined stopping criteria to end the M&V period early
+framework and defining stopping criteria to end the M&V period early
 after reaching a target level of uncertainty, among other improvements,
@@
 ensuring balanced sampling across days of the week and seasons. For
-example, given a 10-week M&V period for 1 intervention, the balanced
-randomized schedule would equally sample 5 Mondays with the baseline
-strategy in operation and 5 Mondays with the intervention strategy in
-operation. The limitation of this method is that it is only applicable
-to a subset of retrofit projects where interventions can be easily
-switched on and off. However, for all applicable use cases, it allows
-analysts to detect energy savings sequentially shortly after the test
-begins. In addition, once the desired estimate uncertainty target is
-achieved, analysts can terminate the M&V study and switch to 100%
-intervention. The key advantage of randomization is that if control
-strategies are sampled with equal probability, the influence of other
-confounding factors such as occupancy change (commonly known as one of
-the non-routine events) and other more subtle long term changes in
-building energy use are likely to be evenly distributed among
-measurements, leading to a more accurate and unbiased assessment of the
-intervention effect.
+example, given a 10-week M&V period for one intervention, the balanced
+randomized schedule would equally sample five Mondays with the baseline
+strategy in operation and five Mondays with the intervention strategy in
+operation. The major limitation of this method is that it is only
+applicable to a subset of retrofit projects where interventions can be
+easily switched on and off, which is a small subset of all the
+interventions performed in buildings. However, for all applicable use
+cases, it allows analysts to detect energy savings sequentially shortly
+after the test begins. In addition, once the desired estimate
+uncertainty target is achieved, analysts can terminate the M&V study and
+switch to 100% intervention. The key advantage of randomization is that
+if control strategies are sampled with equal probability, the influence
+of other confounding factors such as occupancy change (a common
+non-routine event) and other more subtle long term changes in building
+energy use are likely to be evenly distributed among measurements,
+leading to a more accurate and unbiased assessment of the intervention
+effect.
 
@@
 dataset designed to advance research and development in building energy
-efficiency and control strategies acting as a test-bed for modeling,
+efficiency and control strategies, acting as a test-bed for modeling,
 simulation, and algorithm development [@miller_building_2020]. BGD2
-contains over 1000 buildings' metadata and realistic operational
+contains over 1000 buildings\' metadata and realistic operational
 measurements between 2016 and 2017 from across North America and Europe,
@@
 mostly used the metered electricity and outdoor weather measurements for
-running the M&V methods.
+evaluating the M&V methods.
 
@@
 25% for 12-month data collection. Thus, most research related to M&V for
-whole-building approach focuses on the accuracy of baseline modeling,
-exploring model performance from simple regression models to more
-complex machine learning techniques. One study reviewed various models
-suitable for M&V applications as well as selected input features
+the whole-building approach focuses on the accuracy of baseline
+modeling, exploring model performance from simple regression models to
+more complex machine learning techniques. One study reviewed various
+models suitable for M&V applications as well as selected input features
 [@alrobaie_review_2022] and another study provided a definitive
@@
 directly associated with calculated savings, for instance, accounting
-for the potential bias that the baseline model might deteriorate (i.e.
-become 'stale') over the extended period of pre- and post-analysis.
+for the potential bias that the baseline model might deteriorate (i.e.,
+become \'stale\') over the extended period of pre- and post-analysis.
 
@@
 
-Changes in a building that influences its energy usage are commonly
-refered to as non-routine events in M&V and are typically unrelated to
-the intervention strategy. Current guidelines define non-routine events
-as 'static factors' that needed to be adjusted after projecting baseline
-in the post-retrofit period [@ashrae_ashrae_2023;
+Changes in a building that influence its energy usage and are unrelated
+to the intervention being studied by the M&V method are commonly
+referred to as non-routine events. Current guidelines define non-routine
+events as \'static factors\' that need to be adjusted after projecting
+the baseline in the post-retrofit period [@ashrae_ashrae_2023;
 @efficiency_valuation_organisation_international_2012]. A common
@@
 equipment and operations during the reporting period, which is
-labor-intensive and error-prone. ASHRAE guideline 14 recommends
+labor-intensive and error-prone. ASHRAE Guideline 14 recommends
 performing engineering calculations or computer software simulations to
 adjust the post-retrofit baseline. Additionally, it is relatively rare
-to have access to all measurements needed for adjustments such as
+to have access to all measurements needed for adjustments, such as
 occupancy, and thus the analyst normally assumes those factors remain
-unchanged throughout the study. For example, some studies uses linear
-interpolation to estimate counter-factual baseline for demand response
-program [@beil_round-trip_2015; @keskar_commercial_2020]. Consequently,
-one study points out that it is inaccurate to assume no change in the
-operating conditions during the response period
-[@huang_experimental_2023]. Other studies realized such limitation in
+unchanged throughout the study. For example, some studies use linear
+interpolation to estimate counter-factual baseline for the demand
+response program [@beil_round-trip_2015; @keskar_commercial_2020].
+Consequently, one study points out that it is inaccurate to assume no
+change in the operating conditions during the response period (Huang,
+Katipamula, and Lutes 2023). Other studies realized such limitation in
 field study, and emphasized the importance of requiring matched groups
@@
 buildings, which are hardly noticed. We define those gradual changes as
-'operational drift' in this study. One likely encountered operational
+\'operational drift\' in this study. One likely encountered operational
 drift situation is the filter clogging in air handling units due to
@@
 formally defined in the literature or standards, but in this study, we
-consider it as a special type of non-routine events as it also
-represents a change in facility operation but in a longer term.
+consider it as a special type of non-routine event, as it also
+represents a change in facility operation that occurs gradually over a
+longer period of time.
 
@@
 intervention is a control strategy that adjusts the chilled water
-plant's supply water temperature, which can be commonly found in the
+plant\'s supply water temperature, which can be commonly found in the
 literature [@duarte_field_2023; @lee_simulationoptimization_2012;
@@
 made it even simpler by adjusting the setpoint based on outdoor weather
-conditions with more description shown in Section
-\@ref(apply-virtue-intervention). By conducting such an analysis, we aim
-to:
+conditions (see description in Section
+\@ref(apply-virtue-intervention)). By conducting such an analysis, we
+aim to:
 
@@
     contains real-world building energy usage data collected over two
-    years, inherently contains varying degrees of usage changes. By
-    virtually introducing intervention effects into the existing
+    years, which inherently contains varying degrees of usage changes.
+    By virtually introducing intervention effects into the existing
     dataset, we can calculate the ground-truth energy savings, which
@@
     ensured the reproducibility of the method by making the analysis
-    code open source including randomized schedule generation,
+    code open source, including randomized schedule generation,
     sequential statistical analysis, energy modeling and normalized
@@
     provided by the Building Performance Database
-    [@lawrenceberkeleynationallab; @mathew_big-data_2015].
+    [@lawrence_berkeley_national_laboratory_buildings_2015;
+    @mathew_big-data_2015].
 
@@
 
-For all qualified buildings, we further added one more stringent
-criterion to explore all buildings with very stable electricity usage
-between the two years:
-
-1.  No statistically significant difference (P-value \> 0.05) between
-    the two-year electricity usage, which are sometimes assumed to be
-    the case for whole-building measurement and verification.
-
-As a result, only 66 buildings are labeled in this 'stable' subset (out
-of total of 600 buildings) implying the assumption that buildings are
-typically stable over the timeframe involved in conventional M&V is
-rarely true. As mentioned in Section \@ref(non-routine-events-impact),
-such variability is largely associated with non-routine events,
-especially operational drifts since the assessment, even under ideal
-situations, last for two years. And according to the statistics shown
-here, it is a typical case observed in real buildings applications.
+For all qualified buildings, we also create a feature to identify
+buildings with very stable electricity usage overall two year period by
+identifying those with no statistically significant difference (P-value
+\> 0.05) between the two-year electricity usage. Applying this filter to
+real buildings, only 66 buildings are labeled in this \'stable\' subset
+out of a total of 600 buildings. While we did not adjust for weather
+differences between the two years when generating this high-level
+feature, this result highlights that the assumption that buildings are
+typically stable over the timeframe involved in conventional M&V methods
+is rarely true. As mentioned in \@ref(non-routine-events-impact), such
+variability is largely associated with non-routine events, especially
+operational drifts since the assessment, even under ideal situations,
+lasts for two years. And according to the statistics shown here, it is a
+typical case observed in real buildings applications.
 
@@
 |                        |                                                                  |       |
-|-------------------|----------------------------------|-------------------|
+|--------------|--------------------------------------------|--------------|
 | Parameter              | Description                                                      | Value |
@@
 M&V methods in the previous study [@raftery2024]. The conventional
-method is a pre- and post- comparison on a 12-month baseline and
-12-month intervention timeline. The randomized method defines sampling
+method is a pre- and post-comparison on a 12-month baseline and 12-month
+intervention timeline. The randomized method defines sampling
 requirements as:
@@
 -   A minimum and maximum of 12 and 108 weeks respectively. The
-    randomized schedule covers the entire two-year period but stopping
-    criteria enables an early stop at the end of the satisfied blocking
-    period.
+    randomized schedule covers the entire two-year period but the
+    stopping criteria enable an early stop at the end of the satisfied
+    blocking period.
 
 -   At least 80% of the dry-bulb temperature range in the annual TMY
-    data sampled by both strategies.
+    data is sampled by both strategies.
 
 -   90% confidence that energy savings exceed or do not exceed 0% using
-    the SPRT test. Medium effect size (d = 0.5) quantified by cohen's d
-    and calculated SPRT statistics either falls below the lower
-    threshold or exceeds the upper threshold.
+    the SPRT test. Medium effect size (d = 0.5) quantified by Cohen\'s d
+    and calculated SPRT statistics either fall below the lower threshold
+    or exceed the upper threshold.
 
@@
   plot_annotation(title = "Overall accuracy and time required for savings estimation by different M&V methods")
+
+MW_summary <- df_MW_A %>% 
+  pivot_wider(names_from = "method", values_from = "diff") %>% 
+  group_by(name) %>% 
+  summarise(improv_eob = (conv_diff - rand_eob_diff) / conv_diff * 100, 
+            improv_final = (conv_diff - rand_final_diff) / conv_diff * 100) %>% 
+  ungroup() %>% 
+  summarise(across(c(improv_eob, improv_final), median))
+
+TW_summary <- df_TW_A %>% 
+  pivot_wider(names_from = "method", values_from = "diff") %>% 
+  group_by(name) %>% 
+  summarise(improv_eob = (conv_diff - rand_eob_diff) / conv_diff * 100, 
+            improv_final = (conv_diff - rand_final_diff) / conv_diff * 100, 
+            improv = ifelse(improv_eob >= 0, 1, 0)) %>% 
+  ungroup() %>% 
+  mutate(perc = sum(improv)) %>% 
+  summarise(across(c(improv_eob, improv_final, perc), median))
 ```
@@
 conditions, and subplot b) shows the savings normalized on the typical
-meteorological weather of the building location after using TOWT model.
-As a result, the conventional M&V method exhibits a median deviation of
-5% in savings estimation, whereas the proposed randomized method
-demonstrates significantly smaller deviations. If the analyst stops
-immediately after satisfying all stopping criteria, the deviation is
-reduced to approximately 1.6%. Extending the M&V period to match the
-length of the conventional method (24 months) further improves accuracy
+meteorological weather of the building location after using a
+time-of-week-temperature (TOWT) model [@mathieu_quantifying_2011]. As a
+result, the conventional M&V method exhibits a median deviation of 5% in
+savings estimation, whereas the randomized method demonstrates
+significantly smaller deviations. If the analyst stops immediately after
+satisfying all stopping criteria, the deviation is reduced to
+approximately 1.6%. Extending the M&V period to match the same time
+period of the conventional method (24 months) further improves accuracy
 to 0.7% deviation. Furthermore, by comparing subplots (a) and (b), we
@@
 of adjusting for temperature as an independent variable in the model is
-relatively small. Subplot (c) compares the overall timeline of the two
-M&V methods. The results show that almost all buildings achieve accurate
-M&V results and meet the stopping criteria within 24 weeks (for this
-intervention), with the remainder meeting all stopping criteria within
-36 weeks. The additional 12 weeks is usually driven by a need to span
-sufficient weather conditions representative of a full year of data.
-Meanwhile, the red arrow highlights the excess time required by the
-conventional M&V method (24 months). Thus, the randomized M&V method
-finds more accurate results in approximately a quarter of the time
-required for conventional M&V. Even if it was the case that an existing
-12-month baseline data was available at the start of the M&V process,
-randomized M&V would still obtain a more accurate result, in less time,
-than obtaining the 12-month retrofit dataset required by conventional
-M&V.
-
-We also plotted the estimated deviation without absolute calculation in
-Figure \@ref(fig:meanerr) and we noticed that the uncertainty range
-associated with the conventional method is significantly larger compared
-to the randomized method that only takes 36 weeks according to subplot
-(c) in Figure \@ref(fig:abstract).
+relatively small. We found 83% of all filtered buildings show less error
+when using the randomized M&V, and by comparing the error distribution
+in subplot (b), we noticed a median of 74% error reduction at the time
+of satisfying all stopping criteria. Subplot (c) compares the overall
+timeline of the two M&V methods. The results show that almost all
+buildings achieve accurate M&V results and meet the stopping criteria
+within 24 weeks (for this intervention), with the remainder meeting all
+stopping criteria within 36 weeks. The additional 12 weeks are usually
+driven by a need to span sufficient weather conditions representative of
+a full year of data. Meanwhile, the red arrow highlights the excess time
+required by the conventional M&V method (24 months). Thus, the
+randomized M&V method finds more accurate results in approximately a
+quarter of the time required for conventional M&V. Even if it was the
+case that an existing 12-month baseline data was available at the start
+of the M&V process, randomized M&V would still obtain a more accurate
+result, in less time, than obtaining the 12-month retrofit dataset
+required by conventional M&V.
+
+Figure \@ref(fig:meanerr) shows the biased (instead of absolute)
+estimated deviation, highlighting that the uncertainty range associated
+with the conventional method is significantly larger compared to even
+the randomized method that only takes 36 weeks, according to subplot (c)
+in Figure \@ref(fig:abstract).
 
@@
 |                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
-|-------------------|----------------------------------|-------------------|
+|-----------------|--------------------------------------|-----------------|
 | 25th percentile | -5.85                                               | `r percentiles$Q1`     |
@@
 reason why we observe baseline electricity usage changes throughout the
-two year period. To evaluate the robustness of the two M&V methods to
+two-year period. To evaluate the robustness of the two M&V methods to
 the effect of operational drift and non-routine events, we repeated the
 M&V run on the original dataset prior to the application of the reset
-control. In other words, there is no intervention applied here, so the
-most reliable M&V method will be the one that detects closer to 0 kW
-savings using the measurements. Therefore, the $Acc$ here are calculated
-as:
+control. In other words, there *is no intervention applied here*, so the
+most reliable M&V method will be the one that detects the closest to 0
+kW savings using the measurements. Therefore, the $Acc$ here are
+calculated as:
 
@@
 
-Figure \@ref(fig:nosaving) shows the randomized M&V method (both at the
-time of satisfying all stopping criteria and at the end of a two-year
-period) produced consistent results. We included the absolute deviation
-($|ACC|$) figure in the supplementary material. Similarly to Figure
-\@ref(fig:abstract), stopping when meeting all stopping criteria
+Figure \@ref(fig:nosaving) shows that the randomized M&V method (both at
+the time of satisfying all stopping criteria and at the end of a
+two-year period) produced consistent results. We included the absolute
+deviation ($|ACC|$) figure in the supplementary material. Similarly to
+Figure \@ref(fig:abstract), stopping when meeting all stopping criteria
 generally yields slightly larger uncertainty than continuing for 24
@@
 [@granderson_accuracy_2016] where they obtained a median normalized
-deviation around -1.25%, our analysis on conventional method shows that
-the typical deviation is -1.5%, meaning that the typical building is
-decreasing energy use slightly in this dataset in the absence of any
+deviation around -1.25%, our analysis on the conventional method shows
+that the typical deviation is -1.5%, meaning that the typical building
+is decreasing energy use slightly in this dataset in the absence of any
 known intervention. Thus, if a building analyst were to apply an
 intervention to this building and use conventional M&V to measure the
-effect, they would 'detect' 1.5% energy savings even when the
-intervention had no actual impact on energy performance.
+effect, they would typically \'detect\' 1.5% energy savings even when
+the intervention had no actual impact on energy performance.
 
@@
 measurements in the post-retrofit period, we can assess the error caused
-by model adjustment (i.e. how accurate is the 'counterfactual' baseline
-provided by the two M&V methods). Figure \@ref(fig:towtacc) shows the
-difference between the true post-retrofit baseline and adjusted baseline
-through TOWT modeling and since the quantity of the training set is the
-same, this highlights the impact of the data sampling technique. For
-example, if a non-routine event happens, such as a floor of tenant
-moving out, which influences electricity demand but are independent from
-outdoor weather conditions and time-of-week factors, the TOWT model can
-yield significant deviations. This is particularly the case for the
-conventional M&V method since it samples continuously throughout the
-pre-retrofit period. However, randomized sampling is less impacted by
-the risk since it samples only 50% of the pre-retrofit baseline
-throughout a blocking period. Consequently, we noticed as shown in the
-figure, there is a noticeable improvement in modeling accuracy when
-using the randomized method.
+by model adjustment (i.e., how accurate is the \'counterfactual\'
+baseline provided by the two M&V methods). Figure \@ref(fig:towtacc)
+shows the difference between the true post-retrofit baseline and
+adjusted baseline through TOWT modeling and since the quantity of the
+training set is the same, this highlights the impact of the data
+sampling technique. For example, if a non-routine event happens (such as
+a tenant occupying an entire floor of a building moving out), the TOWT
+model can yield significant deviations. This is particularly the case
+for the conventional M&V method since it samples continuously throughout
+the pre-retrofit period. However, randomized sampling is less impacted
+by these events since it samples only 50% of the pre-retrofit baseline
+throughout a blocking period. Consequently, the figure shows, there is a
+noticeable improvement in modeling accuracy when using the randomized
+method.
 
@@
 can be a substantial time-lagged effect associated with interventions.
-This means that the effect of one control strategy may persist after
-switching to another. To mitigate such carryover effects, building
-analysts may choose to drop non-consecutive days in the switchback
-experiment. For example, consider a heat pump hot water heater where the
-tank is charged at very different times of the day by the intervention
-strategy compared to the baseline strategy. When switching strategy, the
-tank may be much warmer or cooler than typical for that strategy, thus
-influencing subsequent measurements, particularly the performance
-measured on the first day (i.e., a non-consecutive day). In such cases,
-increasing the sampling interval (e.g., sample strategies every 2 or 3
-days instead of daily) and discarding measurements from those
-non-consecutive days can eliminate carryover. However, this approach
-reduces the total data collected: for instance, sampling every 3 days
-would drop 1/6 of measurements and sampling every 2 day means would
-remove 1/4 of measurements. Therefore, by increasing the sampling
-interval, there is an accuracy penalty due to a reduction in how often
-samples occur, and when decreasing the sampling interval, there is an
-accuracy penalty from dropping more non-consecutive days. To quantify
-such trade-off, \@ref(fig:absinterval) shows the accuracy of different
-sampling intervals and the consequence of dropping the non-consecutive
-days in the dataset at the time when randomized M&V produces a result
-(after satisfying all stopping criteria).
+This means that the effect of one control strategy may persist, or
+\'carryover\', after switching to another. To mitigate such carryover
+effects, building analysts may choose to drop non-consecutive days in
+the switchback experiment. For example, consider a heat pump hot water
+heater where the tank is charged at very different times of the day by
+the intervention strategy compared to the baseline strategy. When
+switching strategy, the tank may be much warmer or cooler than typical
+for that strategy, thus influencing subsequent measurements,
+particularly the performance measured on the first day (i.e., a
+non-consecutive day). In such cases, increasing the sampling interval
+(e.g., sample strategies every 2 or 3 days instead of daily) and
+discarding measurements from those non-consecutive days can eliminate
+carryover. However, this approach reduces the total data collected: for
+instance, sampling every 3 days and dropping days with non-consecutive
+control strategies would drop 1/6th  of measurements on average.
+Sampling every 2 days means dropping a quarter of the measurements.
+Therefore, by increasing the sampling interval, there is an accuracy
+penalty due to a reduction in randomization (how often samples occur),
+and when decreasing the sampling interval, there is an accuracy penalty
+from dropping more non-consecutive days. To quantify such trade-off,
+Figure \@ref(fig:absinterval) shows the accuracy of different sampling
+intervals and the consequence of dropping the non-consecutive days in
+the dataset at the time when randomized M&V produces a result (after
+satisfying all stopping criteria).
 
@@
 when calculating the distribution shown in the darker-colored set. The
-results suggest without dropping non-consecutive days in subplot (a), a
-gradual increase in error from 1.6% to 2.3% (1.2% to 1.9% with TOWT
-model normalization in subplot (b)) when the sampling interval increases
-from daily to weekly intervals. This is due to the fact that increasing
-the sample interval means reducing randomization. To quantify the
-combined impact of data loss due to preventing the carryover effect, we
-generated the lighter-colored set, which drops all non-consecutive days.
-As a result, we noticed an increase in estimation error when sampling
-daily to 2.4% (to 1.9% when TOWT is fitted), which led to a similar
-accuracy compared to weekly intervals. Therefore, considering this
-trade-off, we recommend using a two-day sampling interval if the
+results suggest that without dropping non-consecutive days in subplot
+(a), a gradual increase in error from 1.6% to 2.3% (1.2% to 1.9% with
+weather normalization in subplot (b) when the sampling interval
+increases from daily to weekly intervals. This is due to the fact that
+increasing the sample interval means reducing randomization. To quantify
+the combined impact of data loss due to preventing the carryover effect,
+we generated the lighter-colored set, which drops all non-consecutive
+days. As a result, we noticed an increase in estimation error when
+sampling daily to 2.4% (to 1.9% when TOWT is fitted), which led to a
+similar accuracy compared to weekly intervals. Therefore, considering
+this trade-off, we recommend using a two-day sampling interval if the
 carryover effect is likely but expected to last less than one day, and
-then normalizing the estimation via energy modeling. We also included
-the deviation distribution without absolute calculation and the
-deviation distribution calculated throughout the two years in the
-supplementary material and the results show the same indication.
+then normalizing the estimation via energy modeling. The supplementary
+material contains similar figures showing biased deviation distribution,
+as well as the same results when continuing the randomized M&V for 24
+months. Those results indicate the same conclusion; that sampling every
+two days and dropping non-consecutive samples yields the optimal result
+when carryover effects are present.
+
+Sampling at different intervals and excluding non-consecutive days also
+impact other stopping criteria, such as the time needed to satisfy
+statistical uncertainty and weather conditions thresholds. In Figure
+\@ref(fig:intervaltime), we compared the percentage of buildings
+satisfying all stopping criteria under different sampling intervals,
+with or without dropping non-consecutive days. The results indicate that
+randomized M&V concludes more quickly when all data are retained,
+particularly at shorter intervals, as expected. For instance, when
+sampling at a one-day interval, 89% of buildings meet all criteria
+within 24 weeks if no data are dropped, but this number decreases
+significantly to 52% when non-consecutive days (account for 1/3 of the
+data) are excluded, and 3% of the buildings require another 12 weeks to
+satisfy the weather criterion. As the sampling interval increases, this
+difference gradually diminishes: when sampling at a weekly interval,
+removing non-consecutive days has no impact on the timeline of meeting
+stopping criteria.
 
@@
 
-Sampling at different intervals and excluding non-consecutive days also
-impact other stopping criteria, such as the time needed to satisfy
-statistical uncertainty and weather conditions thresholds. In Figure
-\@ref(fig:intervaltime), we compared the percentage of buildings
-satisfying all stopping criteria under different sampling intervals,
-with or without dropping non-consecutive days. The results indicate that
-randomized M&V concludes more quickly when all data are retained,
-particularly at shorter intervals. For instance, when sampling at a
-one-day interval, 89% of buildings meet all criteria within 24 weeks if
-no data are dropped, but this number decreases significantly to 52% when
-non-consecutive days (account for 1/3 of the data) are excluded, and 3%
-of the buildings require another 12 weeks to satisfy the weather
-criterion. As the sampling interval increases, this difference gradually
-diminishes: when sampling at a weekly interval, removing non-consecutive
-days has no impact on the timeline of meeting stopping criteria.
-
 ## Sampling ratio impact
@@
 owners adjust the sampling ratio after meeting all stopping criteria, as
-continuing baseline sampling helps maintain M&V accuracy but may reduce
-the potential for energy savings. To further examine its impact on M&V
-accuracy, Figure \@ref(fig:abscont) compares two sampling ratios
-(20%/80% and 10%/90%) and two assocaited scenarios. The "Continued"
-group follows the standard randomized M&V framework with an initial
-50%/50% sampling ratio, then switches to the indicated ratio after
-satisfying the stopping criteria, continuing this until the end of the
-two-year period. In contrast, the "Randomized" group uses the indicated
-ratio from the outset and continues throughout the two year. Subplot (a)
-shows that absolute error is significantly lower without the initial
-50%/50% sampling, due to baseline drift and non-routine events that most
-buildings experience over two years (as noted in Section
-\@ref(filter-and-clean-dataset)). Subplot (b) confirms this finding even
-after weather normalization.
-
-```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months", fig.height=6, fig.width=12}
+continuing baseline sampling helps maintain M&V accuracy with limited
+impact on the energy savings actually achieved by the building. To
+further examine its impact on M&V accuracy, Figure \@ref(fig:abscont)
+compares two sampling ratios (20%/80% and 10%/90%) and two associated
+scenarios with the previously shown results when sampling at 50%/50%.
+The \"Continued\" group follows the standard randomized M&V framework
+with an initial 50%/50% sampling ratio, then switches to the indicated
+sampling ratio after satisfying the stopping criteria, continuing this
+until the end of the two-year period. In contrast, the \"Randomized\"
+group uses the indicated ratio from the outset and continues throughout
+the two year. Subplot (a) shows that absolute error is significantly
+lower without the initial 50%/50% sampling, due to baseline drift and
+non-routine events that most buildings experience over two years (as
+noted in Section \@ref(filter-and-clean-dataset)). Subplot (b) confirms
+this finding even after weather normalization.
+
+```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months (previous results of sampling at 50%/50% shown on the left side of the red dashed line)", fig.height=6, fig.width=12}
 
@@
                                ifelse(ratio == 2 & type == "diff", "2_1", 
-                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio)))))
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_stable %>% 
+              filter(method != "conv") %>% 
+              pivot_wider(names_from = method, values_from = savings) %>% 
+              mutate(values = abs(true - rand), 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
 
@@
                                ifelse(ratio == 2 & type == "diff", "2_1", 
-                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio)))))
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_variable %>% 
+              filter(method != "conv") %>% 
+              pivot_wider(names_from = method, values_from = savings) %>% 
+              mutate(values = abs(true - rand), 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
 
@@
   mutate(ratio = as.factor(ratio), 
-         ratio = recode_factor(ratio, "1_1" = "Continued\n(20/80)", "1_2" = "Randomized\n(20/80)", "2_1" = "Continued\n(10/90)", "2_2" = "Randomized\n(10/90)")) %>% 
+         ratio = recode_factor(ratio, 
+                               "0" = "Randomized\n(50/50)", 
+                               "1_1" = "Continued\n(20/80)", 
+                               "1_2" = "Randomized\n(20/80)", 
+                               "2_1" = "Continued\n(10/90)", 
+                               "2_2" = "Randomized\n(10/90)")) %>% 
   ggplot(aes(x = ratio, y = values, fill = ratio)) +
   geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
-  geom_lv(k = 4, outlier.shape = NA) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio)) +
-  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75) +
+  geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
   geom_text(data = . %>% 
@@
               ungroup(), 
-            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %"))) +
+            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %")), 
+            size = 4) +
   scale_y_continuous(expand = c(0, 0), 
@@
   scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80")) + 
+  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80", "grey80")) + 
   labs(fill = NULL, 
@@
                                ifelse(ratio == 2 & type == "diff", "2_1", 
-                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio)))))
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_tmy_stable %>% 
+              left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+              mutate(values = abs(savings - rand), 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
 
@@
                                ifelse(ratio == 2 & type == "diff", "2_1", 
-                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio)))))
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_tmy_variable %>% 
+              left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+              mutate(values = abs(savings - rand), 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
 
@@
 To further illustrate this point, we plotted results in Figure
-\@ref(fig:stablecont) using only the \'stable\' building subset (66
+\@ref(fig:stablecont) using only the 'stable' building subset (66
 buildings), as defined in Section \@ref(filter-and-clean-dataset). The
@@
 This suggests that, consistent with our earlier findings regarding the
-convention M&V method, initial baseline measurements provide limited
-value in maintaining M&V accuracy when the building\'s baseline
+conventional M&V method, initial baseline measurements provide limited
+value in maintaining M&V accuracy when the building's baseline
 performance changes over time.
 
-```{r stablecont, fig.cap="Comparison of error distribution under different sampling ratio only for the buildings in 'stable' subset", fig.height=6, fig.width=12}
+```{r stablecont, fig.cap="Comparison of error distribution under different sampling ratio only for the buildings in 'stable' subset" (results of sampling at 50%/50% shown on the left side of the red dashed line), fig.height=6, fig.width=12}
 p_top <- df_MW_S %>% 
   mutate(ratio = as.factor(ratio), 
-         ratio = recode_factor(ratio, "1_1" = "Continued\n(20/80)", "1_2" = "Randomized\n(20/80)", "2_1" = "Continued\n(10/90)", "2_2" = "Randomized\n(10/90)")) %>% 
+         ratio = recode_factor(ratio, 
+                               "0" = "Randomized\n(50/50)",
+                               "1_1" = "Continued\n(20/80)", 
+                               "1_2" = "Randomized\n(20/80)", 
+                               "2_1" = "Continued\n(10/90)", 
+                               "2_2" = "Randomized\n(10/90)")) %>% 
   ggplot(aes(x = ratio, y = values, fill = ratio)) +
   geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
-  geom_lv(k = 4, outlier.shape = NA) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio)) +
-  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75) +
+  geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
   geom_text(data = . %>% 
@@
   mutate(ratio = as.factor(ratio), 
-         ratio = recode_factor(ratio, "1_1" = "Continued\n(20/80)", "1_2" = "Randomized\n(20/80)", "2_1" = "Continued\n(10/90)", "2_2" = "Randomized\n(10/90)")) %>% 
+         ratio = recode_factor(ratio, 
+                               "0" = "Randomized\n(50/50)",
+                               "1_1" = "Continued\n(20/80)", 
+                               "1_2" = "Randomized\n(20/80)", 
+                               "2_1" = "Continued\n(10/90)", 
+                               "2_2" = "Randomized\n(10/90)")) %>% 
   ggplot(aes(x = ratio, y = values, fill = ratio)) +
   geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
-  geom_lv(k = 4, outlier.shape = NA) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio)) +
-  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75) +
+  geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
   geom_text(data = . %>% 
@@
 The additional results from Figure \@ref(fig:abscont) and Figure
-\@ref(fig:stablecont) further adds to the guidance: if building baseline
+\@ref(fig:stablecont) further add to the guidance: if building baseline
 changes, retrofit savings estimation should simply discard the existing
@@
 analyst evaluates stopping criteria and may influence the decision when
-a target saving is detected. Here, we summarize the advantages and
-disadvantages of shorter versus longer blocking periods. As discussed in
-our previous study, the second step of this randomized M&V method
-involves choosing an appropriate blocking period to minimize potential
-biases caused by incomplete randomization. For most practical scenarios,
-we recommend a blocking period of approximately 12 to 16 weeks, as it
-typically covers the duration of a season, ensuring a representative
-sampling of operational conditions. In general, longer blocking periods
-are advantageous when evaluating multiple intervention strategies or
-using sampling intervals greater than one day. In such cases, extended
-blocking ensures a balanced distribution of both day-of-week and outdoor
-weather conditions.
+a savings is detected at a sufficient level of certainty. Here, we
+summarize the advantages and disadvantages of shorter versus longer
+blocking periods. As discussed in our previous study, the second step of
+this randomized M&V method involves choosing an appropriate blocking
+period to minimize potential biases caused by incomplete randomization.
+For most practical scenarios, we recommend a blocking period of
+approximately 12 to 16 weeks, as it typically covers the duration of a
+season, ensuring a representative sampling of operational conditions. In
+general, longer blocking periods are advantageous when evaluating
+multiple intervention strategies or using sampling intervals greater
+than one day. In such cases, extended blocking ensures a balanced
+distribution of both day-of-week and outdoor weather conditions.
 
@@
 result, when compared to Figure \@ref(fig:intervaltime), the only
-difference we found is when sampling at daily interval and not dropping
-non-consecutive days, 12% of buildings can get an estimate earlier. This
-means those buildings are likely to meet all stopping criteria between
-12 weeks and 18 weeks but when sampled over a 12-week block, the savings
-can only be determined at week 24. Although sampled baseline and
-intervention days can be different when the blocking period is changed,
-the results indicate most buildings still require 9 months to meet
-outdoor weather conditions when starting M&V from January.
+difference we found is when sampling at a daily interval and not
+dropping non-consecutive days, 12% of buildings can get an estimate
+earlier. This means those buildings are likely to meet all stopping
+criteria between 12 weeks and 18 weeks but when sampled over a 12-week
+block, the savings can only be determined at week 24. Although sampled
+baseline and intervention days can be different when the blocking period
+is changed, the results indicate most buildings still require 9 months
+to meet outdoor weather conditions when starting M&V from January.
 
@@
 offers the advantage of capturing colder outdoor weather conditions
-early in the M&V process. However, when the baseline measurements misses
+early in the M&V process. However, when the baseline measurements miss
 the initial winter period, meeting the required weather condition
@@
 42 (mid-December), which is one blocking period earlier than having a
-12-week blocking period. In both figures, we noticed a small proportion
-of buildings are less affected by changing the M&V start time, they are
-mostly located in climate zones with less outdoor temperature
-variations. This result further highlights that the overall duration of
-the randomized M&V process is strongly driven by outdoor weather.
-Accordingly, selecting an optimal blocking period should consider
-historical or typical meteorological weather file to effectively
-minimize the time required.
+12-week blocking period. In both figures, we noticed that a small
+proportion of buildings are less affected by changing the M&V start
+time, they are mostly located in climate zones with less outdoor
+temperature variations. This result further highlights that the overall
+duration of the randomized M&V process is strongly driven by outdoor
+weather. Accordingly, selecting an optimal blocking period should
+consider a historical or typical meteorological weather file to
+effectively minimize the time required. In general, starting M&V in the
+middle of the winter or summer periods will yield results spanning the
+full range of weather conditions faster than starting at other times of
+year, such as during a swing season.
 
@@
 
-1.  The simulated intervention remains somewhat generic and simplistic
-    considering the diverse building types and climate zones in the BDG2
-    dataset. As the primary focus is to accurately detect the
-    intervention (or any intervention) effect so for simplicity, we
-    applied the parameters listed in Table 1 uniformly across all
-    buildings. Yet, in some cases, raising the water temperature by 1°C
-    might yield more or less than the specified 8% electricity savings
-    due to target buildings' demand flexibility. For example, one field
-    validation of HVAC optimization shows additional savings might be
-    limited if other best practice controls are already implemented
-    [@granderson2018].
-2.  For the main analysis, we assumed no carryover effect and a daily
-    sampling interval would suffice for most commercial buildings, but
-    certainly exceptions exist. However, in the discussion, we addressed
-    such concern by comparing different sampling strategies and the
-    impact of dropping measurements from non-consecutive days.
+1.  The simulated intervention (chilled water temperature reset) remains
+    somewhat generic and simplistic, considering the diverse building
+    types and climate zones in the BDG2 dataset. As the primary focus is
+    to accurately detect the intervention (or any intervention) effect
+    so for simplicity, we applied the parameters listed in Table 1
+    uniformly across all buildings.
+2.  For the main analysis, we assumed no carryover effect and that a
+    daily sampling interval would suffice for most commercial buildings,
+    but certainly exceptions exist. However, in the discussion, we
+    addressed such concern by comparing different sampling strategies
+    and the impact of dropping measurements from non-consecutive days.
 
@@
 ASHRAE Guideline 14, using a large, open-source commercial building
-dataset. The porposed M&V method leverages the randomized experimental
+dataset. The randomized M&V method leverages the randomized experimental
 design concept from other scientific fields, along with statistical
-sequential inference techniques, to determine when target savings are
-detected. We used a virtual control retrofit case---resetting the
-chilled water setpoint based on outdoor weather conditions---and applied
-it to over 600 filtered commercial buildings. By comparing the savings
-estimations of the conventional method with the novel randomized method,
-we found that the randomized approach provides faster and more robust
-savings estimations.
+sequential inference techniques, to determine when savings are detected
+at sufficient certainty. We used a virtual control retrofit
+case---resetting the chilled water setpoint based on outdoor weather
+conditions---and applied it to over 600 filtered commercial buildings.
+By comparing the savings estimations of the conventional method with the
+novel randomized method, we found that the randomized approach provides
+faster and more robust savings estimations.
 
 Specifically, we showed that throughout 11 different locations assessed
-in this study, the randomized M&V can provide a saving estimation by 36
-weeks (with the majority finished by 24 weeks) once all stopping
+in this study, the randomized M&V can provide a saving estimation after
+36 weeks (with the majority finished by 24 weeks) once all stopping
 criteria are satisfied. In contrast, the conventional method requires a
@@
 approximately 5% in a typical building. We also evaluated the impact of
-non-routine events especially more gradual and subtle operational drift
+non-routine events, especially more gradual and subtle operational drift
 on the proposed M&V method and the results show that baseline changes in
 the post-retrofit period can deviate savings estimated using the
-conventional method. As a comparison, we found those events have a very
+conventional method. In contrast, we found that those events have a very
 negligible impact on the savings estimated using the randomized method,
@@
 where the strategy is non-consecutive and varying the sampling interval.
-Through that process, we showed that for the typical building, if a
-carryover is likely to be present, then using a 2-day sampling interval
-and dropping non-consecutive samples yields the optimal design. Lastly,
-we investigated the time required to complete the randomized M&V method
-under different scenarios. We found measuring a sufficient range of
-outdoor weather condition remains the most stringent criterion. But even
-for the most non-ideal case, the randomized method can still produce an
-estimate quicker than 2 years.
-
-We also discussed the impact of model fitting accuracy when using the
-two methods and we found a significant improvement in our method due to
-random sampling. We also provide a detailed assessment of using
-different sampling ratios, or a blocking period, those results provide
-useful considerations to building analysts when generating randomized
-schedules for a variety of use cases.
+Through that process, we showed that for the typical building, if
+carryover is present, using a 2-day sampling interval and dropping days
+with non-consecutive control strategies operating yields the optimal
+design for the median building in this dataset. Lastly, we investigated
+the time required to complete the randomized M&V method under different
+scenarios. We found that measuring a sufficient range of outdoor weather
+conditions remains the most stringent criterion. But even for the most
+non-ideal case, the randomized method can still produce an estimate much
+faster than 2 years, and produces an estimate with far lower error.
 

### 2025-06-17 code/manuscript.Rmd
Added graphical abstract and updated the manuscript

@@
   - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
-abstract: Conventional measurement and verification (M&V) methods rely on pre- and post- retrofit comparison to estimate energy savings.
-They are often time-consuming and unreliable, especially when non-routine events which are static changes or more gradual changes denoted as operational drift occur during the M&V process.
-Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies conventional M&V suggested by industry guidelines, the results can be largely confounded.
-In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method which randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily).
-We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention.
-Our results show that compared to the conventional method, the randomized method provides faster and more accurate savings estimations.
-Additionally we found that when non-routine events are present, the randomized M&V approach estimates savings that are much closer to the ground-truth savings than the conventional M&V method, demonstrating much improved reliability.
-We also assessed the impact of starting the M&V at different time of the year, continuing randomization with different sampling ratio after satisfying all criteria, dropping samples affected by carryover when switching between strategies, and identify the optimal sampling interval for that scenario using this large dataset.
+abstract: Conventional measurement and verification (M&V) methods rely on comparing pre- and post-retrofit performance to estimate energy savings.
+They are often time-consuming and unreliable, especially when non-routine events (step changes or more gradual changes in how a building operates) occur during the M&V process.
+Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies the conventional M&V method, the results can be confounded.
+In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method that randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily).
+We tested this novel randomized M&V method on a large public dataset (639 buildings) covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention.
+The results show that compared to the conventional method, the randomized method provides more accurate savings estimations with a meidum of 74% accuracy improvement and is even faster.
+Additionally, we found that when non-routine events are present, the randomized method estimates savings that are much closer to the ground-truth savings than the conventional method, demonstrating much improved reliability.
+We also assessed the impact of normalizing for different weather, starting the M&V at different time of the year, continuing randomization with a different sampling ratio after satisfying all criteria, dropping samples affected by carryover when switching between strategies, and identified the optimal sampling interval for that scenario using this large dataset.
 
@@
 
+### Graphical abstract {.unnumbered}
+
+```{r graphicalabstract, out.width = "80%"}
+
+knitr::include_graphics("../figs/manuscript/graphical_abs.jpg", dpi = 900)
+```
+
+### Keywords {.unnumbered}
+
+Measurement and verification (M&V); Controls; Randomized switchback
+design; Uncertainty quantification; Building energy analysis; Heating
+ventilation and air conditioning
+
+### Highlights {.unnumbered}
+
+-   Verified the improvements of the randomized switchback method using
+    a large open-source dataset
+-   Addressed the influence of non-routine events and also more subtle
+    long-term 'operational drift' in buildings
+-   Investigated the trade-off of selecting longer sampling intervals to
+    account for carryover effect and weather dependency of the method
+-   Discussed the impact of selecting different sampling ratio on the
+    M&V accuracy
+
 # Introduction
@@
 dataset designed to advance research and development in building energy
-efficiency and control strategies acting as a test-bed for modeling,
+efficiency and control strategies, acting as a test-bed for modeling,
 simulation, and algorithm development [@miller_building_2020]. BGD2
@@
 directly associated with calculated savings, for instance, accounting
-for the potential bias that the baseline model might deteriorate (i.e.
+for the potential bias that the baseline model might deteriorate (i.e.,
 become 'stale') over the extended period of pre- and post-analysis.
@@
 
-Changes in a building that influences its energy usage are commonly
-refered to as non-routine events in M&V and are typically unrelated to
-the intervention strategy. Current guidelines define non-routine events
-as 'static factors' that needed to be adjusted after projecting baseline
-in the post-retrofit period [@ashrae_ashrae_2023;
+Changes in a building that influence its energy usage and are unrelated
+to the intervention being studied by the M&V method are commonly
+referred to as non-routine events. Current guidelines define non-routine
+events as 'static factors' that need to be adjusted after projecting the
+baseline in the post-retrofit period [@ashrae_ashrae_2023;
 @efficiency_valuation_organisation_international_2012]. A common
@@
 
-```{r flowchart, fig.cap = "Flow chart showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="75%"}
+```{r flowchart, fig.cap = "Flow chart showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="70%"}
 
@@
 
-For all qualified buildings, we further added one more stringent
-criterion to explore all buildings with very stable electricity usage
-between the two years:
-
-1.  No statistically significant difference (P-value \> 0.05) between
-    the two-year electricity usage, which are sometimes assumed to be
-    the case for whole-building measurement and verification.
-
-As a result, only 66 buildings are labeled in this 'stable' subset (out
-of total of 600 buildings) implying the assumption that buildings are
-typically stable over the timeframe involved in conventional M&V is
-rarely true. As mentioned in Section \@ref(non-routine-events-impact),
-such variability is largely associated with non-routine events,
-especially operational drifts since the assessment, even under ideal
-situations, last for two years. And according to the statistics shown
-here, it is a typical case observed in real buildings applications.
+For all qualified buildings, we also create a feature to identify
+buildings with very stable electricity usage overall by identifying
+those with no statistically significant difference (P-value \> 0.05)
+between the two-year electricity usage. Applying this filter to real
+buildings, only 66 buildings are labeled in this 'stable' subset out of
+a total of more than 600 buildings. While we did not adjust for weather
+differences between the two years when generating this high-level
+feature, this result highlights that the assumption that buildings are
+typically stable over the timeframe involved in conventional M&V methods
+is rarely true. As mentioned in \@ref(non-routine-events-impact), such
+variability is largely associated with non-routine events, especially
+operational drifts since the assessment, even under ideal situations,
+lasts for two years. And according to the statistics shown here, it is a
+typical case observed in real buildings applications.
 
@@
 
-|                        |                                                                  |       |
-|-------------------|----------------------------------|-------------------|
 | Parameter              | Description                                                      | Value |
+|------------------|------------------------------------|------------------|
 | $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
@@
 -   A minimum and maximum of 12 and 108 weeks respectively. The
-    randomized schedule covers the entire two-year period but stopping
-    criteria enables an early stop at the end of the satisfied blocking
-    period.
+    randomized schedule covers the entire two-year period but the
+    stopping criteria enable an early stop at the end of the satisfied
+    blocking period.
 
 -   At least 80% of the dry-bulb temperature range in the annual TMY
-    data sampled by both strategies.
+    data is sampled by both strategies.
 
 -   90% confidence that energy savings exceed or do not exceed 0% using
-    the SPRT test. Medium effect size (d = 0.5) quantified by cohen's d
-    and calculated SPRT statistics either falls below the lower
-    threshold or exceeds the upper threshold.
+    the sequential probability ratio test (SPRT). Medium effect size (d
+    = 0.5) quantified by Cohen's d and calculated SPRT statistics either
+    fall below the lower threshold or exceed the upper threshold.
 
@@
 
+Figure \@ref(fig:abstract) shows the overall results of M&V methods
+comparison and for clarity, we plotted conventional M&V results in grey,
+randomized M&V results after satisfying all stopping criteria in green,
+and the randomized M&V results sampled for the entire 24 months in blue.
+Subplots (a) and (b) calculate the savings estimation error as the
+absolute deviation from the true savings (i.e. $|Acc|$) and show the
+results distribution using the conventional M&V method in the first
+column, the randomized M&V method that stops after satisfying all
+stopping criteria in the second column, and the randomized M&V method
+continues 50%/50% sampling throughout 2 years in the third column. In
+addition, subplot (a) shows the savings estimated from measured weather
+conditions, and subplot b) shows the savings normalized on the typical
+meteorological weather of the building location after using a
+time-of-week-temperature (TOWT) model [@mathieu_quantifying_2011]. As a
+result, the conventional M&V method exhibits a median deviation of 5% in
+savings estimation, whereas the randomized method demonstrates
+significantly smaller deviations. If the analyst stops immediately after
+satisfying all stopping criteria, the deviation is reduced to
+approximately 1.5%. Extending the M&V period to match the same time
+period of the conventional method (24 months) further improves accuracy
+to 0.7% deviation. Furthermore, by comparing subplots (a) and (b), we
+observed a minor improvement in estimation accuracy when a model is
+fitted to the data to account for weather differences. The model fitting
+is most beneficial for the shorter dataset (randomized and stopping when
+all criteria are met, typically 24-36 weeks of data), where it reduces
+deviation from 1.5% to 1.2% (around 20% improvement). This is because
+the weather conditions between the first and second year for each of the
+locations are relatively similar on average over this time period, and
+thus the effect of adjusting for temperature as an independent variable
+in the model is relatively small. We found 83% of all filtered buildings
+show less error when using the randomized M&V, and by comparing the
+error distribution in subplot (b), we noticed the medium deviation drops
+from 4.6% to 1.2% (around 74% improvement) at the time of satisfying all
+stopping criteria. Subplot (c) compares the overall timeline of the two
+M&V methods. The results show that almost all buildings achieve accurate
+M&V results and meet the stopping criteria within 24 weeks (for this
+intervention), with the remainder meeting all stopping criteria within
+36 weeks. The additional 12 weeks are usually driven by a need to span
+sufficient weather conditions representative of a full year of data.
+Meanwhile, the red arrow highlights the excess time required by the
+conventional M&V method (24 months). Thus, the randomized M&V method
+finds more accurate results in approximately a quarter of the time
+required for conventional M&V. Even if it was the case that an existing
+12-month baseline data was available at the start of the M&V process,
+randomized M&V would still obtain a more accurate result, in less time,
+than obtaining the 12-month retrofit dataset required by conventional
+M&V.
+
 ```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period).", fig.height=11, fig.width=12}
@@
   geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
-  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
   geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
@@
   plot_annotation(title = "Overall accuracy and time required for savings estimation by different M&V methods")
+
+MW_summary <- df_MW_A %>% 
+  pivot_wider(names_from = "method", values_from = "diff") %>% 
+  group_by(name) %>% 
+  summarise(improv_eob = (conv_diff - rand_eob_diff), 
+            improv_final = (conv_diff - rand_final_diff)) %>% 
+  ungroup() %>% 
+  summarise(across(c(improv_eob, improv_final), median))
+
+TW_summary <- df_TW_A %>% 
+  pivot_wider(names_from = "method", values_from = "diff") %>% 
+  group_by(name) %>% 
+  summarise(improv_eob = (conv_diff - rand_eob_diff), 
+            improv_final = (conv_diff - rand_final_diff), 
+            improv = ifelse(improv_eob >= 0, 1, 0)) %>% 
+  ungroup() %>% 
+  mutate(perc = sum(improv)) %>% 
+  summarise(across(c(improv_eob, improv_final, perc), median))
 ```
 
-Figure \@ref(fig:abstract) shows the overall results of M&V methods
-comparison and for clarity, we plotted conventional M&V results in grey,
-randomized M&V results after satisfying all stopping criteria in green,
-and the randomized M&V results sampled for the entire 24 months in blue.
-Subplots (a) and (b) calculate the savings estimation error as the
-absolute deviation from the true savings (i.e. $|Acc|$) and show the
-results distribution using the conventional M&V method in the first
-column, the randomized M&V method that stops after satisfying all
-stopping criteria in the second column, and the randomized M&V method
-continues 50%/50% sampling throughout 2 years in the third column. In
-addition, subplot (a) shows the savings estimated from measured weather
-conditions, and subplot b) shows the savings normalized on the typical
-meteorological weather of the building location after using TOWT model.
-As a result, the conventional M&V method exhibits a median deviation of
-5% in savings estimation, whereas the proposed randomized method
-demonstrates significantly smaller deviations. If the analyst stops
-immediately after satisfying all stopping criteria, the deviation is
-reduced to approximately 1.6%. Extending the M&V period to match the
-length of the conventional method (24 months) further improves accuracy
-to 0.7% deviation. Furthermore, by comparing subplots (a) and (b), we
-observe a minor improvement in estimation accuracy when a model is
-fitted to the data to account for weather differences. The model fitting
-is most beneficial for the shorter dataset (randomized and stopping when
-all criteria are met, typically 24-36 weeks of data), where it improves
-estimation accuracy by 0.4%. This is because the weather conditions
-between the first and second year for each of the locations are
-relatively similar on average over this time period, and thus the effect
-of adjusting for temperature as an independent variable in the model is
-relatively small. Subplot (c) compares the overall timeline of the two
-M&V methods. The results show that almost all buildings achieve accurate
-M&V results and meet the stopping criteria within 24 weeks (for this
-intervention), with the remainder meeting all stopping criteria within
-36 weeks. The additional 12 weeks is usually driven by a need to span
-sufficient weather conditions representative of a full year of data.
-Meanwhile, the red arrow highlights the excess time required by the
-conventional M&V method (24 months). Thus, the randomized M&V method
-finds more accurate results in approximately a quarter of the time
-required for conventional M&V. Even if it was the case that an existing
-12-month baseline data was available at the start of the M&V process,
-randomized M&V would still obtain a more accurate result, in less time,
-than obtaining the 12-month retrofit dataset required by conventional
-M&V.
-
-We also plotted the estimated deviation without absolute calculation in
-Figure \@ref(fig:meanerr) and we noticed that the uncertainty range
-associated with the conventional method is significantly larger compared
-to the randomized method that only takes 36 weeks according to subplot
-(c) in Figure \@ref(fig:abstract).
-
-```{r meanerr, fig.cap="Distribution of estimated deviation from the ground-truth savings calculated by the two M&V methods", fig.height=8, fig.width=12}
+Figure \@ref(fig:meanerr) shows the biased (instead of absolute)
+estimated deviation, highlighting that the uncertainty range associated
+with the conventional method is significantly larger compared to even
+the randomized method that only takes 36 weeks, according to subplot (c)
+in Figure \@ref(fig:abstract).
+
+```{r meanerr, fig.cap="Distribution of estimated deviation from the ground-truth savings calculated by the two M&V methods (red dashed line highlighting 0% error)", fig.height=8, fig.width=12}
 # plot for the stable subset
@@
 |                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
-|-------------------|----------------------------------|-------------------|
+|-------------------|-----------------------------------|-------------------|
 | 25th percentile | -5.85                                               | `r percentiles$Q1`     |
@@
 
-: Table 1. Savings accuracy comparison with a similar literature
+: Table 2. Savings accuracy comparison with a similar literature
 
 ## Non-routine events impact
+
+As mentioned in Section \@ref(filter-and-clean-dataset), the majority of
+the selected buildings are affected by operational drift, which is the
+reason why we observe baseline electricity usage changes throughout the
+two-year period. To evaluate the robustness of the two M&V methods to
+the effect of operational drift and non-routine events, we repeated the
+M&V run on the original dataset prior to the application of the reset
+control. In other words, there *is no intervention applied here*, so the
+most reliable M&V method will be the one that detects the closest to 0
+kW savings using the measurements. Therefore, the $Acc$ here are
+calculated as:
+
+$$
+Acc = 0-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
+$$
+
+Figure \@ref(fig:nosaving) shows that the randomized M&V method (both at
+the time of satisfying all stopping criteria and at the end of a
+two-year period) produced consistent results. We included the absolute
+deviation ($|ACC|$) figure in the supplementary material. Similarly to
+Figure \@ref(fig:abstract), stopping when meeting all stopping criteria
+generally yields slightly larger uncertainty than continuing for 24
+months. However, it still yields a more accurate result than the
+conventional method. Interestingly, similar to
+[@granderson_accuracy_2016] where they obtained a median normalized
+deviation around -1.25%, our analysis on the conventional method shows
+that the typical deviation is -1.5%, meaning that the typical building
+is decreasing energy use slightly in this dataset in the absence of any
+known intervention. Thus, if a building analyst were to apply an
+intervention to this building and use conventional M&V to measure the
+effect, they would typically 'detect' 1.5% energy savings even when the
+intervention had no actual impact on energy performance.
 
@@
 
-As mentioned in Section \@ref(filter-and-clean-dataset), the majority of
-the selected buildings are affected by operational drift, which is the
-reason why we observe baseline electricity usage changes throughout the
-two year period. To evaluate the robustness of the two M&V methods to
-the effect of operational drift and non-routine events, we repeated the
-M&V run on the original dataset prior to the application of the reset
-control. In other words, there is no intervention applied here, so the
-most reliable M&V method will be the one that detects closer to 0 kW
-savings using the measurements. Therefore, the $Acc$ here are calculated
-as:
-
-$$
-Acc = 0-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
-$$
-
-Figure \@ref(fig:nosaving) shows the randomized M&V method (both at the
-time of satisfying all stopping criteria and at the end of a two-year
-period) produced consistent results. We included the absolute deviation
-($|ACC|$) figure in the supplementary material. Similarly to Figure
-\@ref(fig:abstract), stopping when meeting all stopping criteria
-generally yields slightly larger uncertainty than continuing for 24
-months. However, it still yields a more accurate result than the
-conventional method. Interestingly, similar to
-[@granderson_accuracy_2016] where they obtained a median normalized
-deviation around -1.25%, our analysis on conventional method shows that
-the typical deviation is -1.5%, meaning that the typical building is
-decreasing energy use slightly in this dataset in the absence of any
-known intervention. Thus, if a building analyst were to apply an
-intervention to this building and use conventional M&V to measure the
-effect, they would 'detect' 1.5% energy savings even when the
-intervention had no actual impact on energy performance.
-
 # Discussion
@@
 measurements in the post-retrofit period, we can assess the error caused
-by model adjustment (i.e. how accurate is the 'counterfactual' baseline
+by model adjustment (i.e., how accurate is the 'counterfactual' baseline
 provided by the two M&V methods). Figure \@ref(fig:towtacc) shows the
@@
 same, this highlights the impact of the data sampling technique. For
-example, if a non-routine event happens, such as a floor of tenant
-moving out, which influences electricity demand but are independent from
-outdoor weather conditions and time-of-week factors, the TOWT model can
-yield significant deviations. This is particularly the case for the
+example, if a non-routine event happens (such as a tenant occupying an
+entire floor of a building moving out), the TOWT model can yield
+significant deviations. This is particularly the case for the
 conventional M&V method since it samples continuously throughout the
 pre-retrofit period. However, randomized sampling is less impacted by
-the risk since it samples only 50% of the pre-retrofit baseline
-throughout a blocking period. Consequently, we noticed as shown in the
-figure, there is a noticeable improvement in modeling accuracy when
-using the randomized method.
+these events since it samples only 50% of the pre-retrofit baseline
+throughout a blocking period. Consequently, the figure shows, there is a
+noticeable improvement in modeling accuracy when using the randomized
+method.
 
@@
 can be a substantial time-lagged effect associated with interventions.
-This means that the effect of one control strategy may persist after
-switching to another. To mitigate such carryover effects, building
-analysts may choose to drop non-consecutive days in the switchback
-experiment. For example, consider a heat pump hot water heater where the
-tank is charged at very different times of the day by the intervention
-strategy compared to the baseline strategy. When switching strategy, the
-tank may be much warmer or cooler than typical for that strategy, thus
-influencing subsequent measurements, particularly the performance
-measured on the first day (i.e., a non-consecutive day). In such cases,
-increasing the sampling interval (e.g., sample strategies every 2 or 3
-days instead of daily) and discarding measurements from those
-non-consecutive days can eliminate carryover. However, this approach
-reduces the total data collected: for instance, sampling every 3 days
-would drop 1/6 of measurements and sampling every 2 day means would
-remove 1/4 of measurements. Therefore, by increasing the sampling
-interval, there is an accuracy penalty due to a reduction in how often
-samples occur, and when decreasing the sampling interval, there is an
-accuracy penalty from dropping more non-consecutive days. To quantify
-such trade-off, \@ref(fig:absinterval) shows the accuracy of different
-sampling intervals and the consequence of dropping the non-consecutive
-days in the dataset at the time when randomized M&V produces a result
-(after satisfying all stopping criteria).
+This means that the effect of one control strategy may persist, or
+'carryover', after switching to another. To mitigate such carryover
+effects, building analysts may choose to drop non-consecutive days in
+the switchback experiment. For example, consider a heat pump water
+heater where the tank is charged at very different times of the day by
+the intervention strategy compared to the baseline strategy. When
+switching strategy, the tank may be much warmer or cooler than typical
+for that strategy, thus influencing subsequent measurements,
+particularly the performance measured on the first day (i.e., a
+non-consecutive day). In such cases, increasing the sampling interval
+(e.g., sample strategies every 2 or 3 days instead of daily) and
+discarding measurements from those non-consecutive days can eliminate
+carryover. However, this approach reduces the total data collected: for
+instance, sampling every 3 days and dropping days with non-consecutive
+control strategies would drop 1/6 of measurements on average. Sampling
+every 2 days means dropping a quarter of the measurements. Therefore, by
+increasing the sampling interval, there is an accuracy penalty due to a
+reduction in randomization (how often samples occur), and when
+decreasing the sampling interval, there is an accuracy penalty from
+dropping more non-consecutive days. To quantify such trade-off, Figure
+\@ref(fig:absinterval) shows the accuracy of different sampling
+intervals and the consequence of dropping the non-consecutive days in
+the dataset at the time when randomized M&V produces a result (after
+satisfying all stopping criteria).
 
@@
   geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8, linewidth = 0.15) +
-  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
   geom_text(data = . %>% group_by(group, interval) %>% summarise(median = median(diff, inf.rm = T)) %>% ungroup(), 
@@
   geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8, linewidth = 0.15) +
-  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
   geom_text(data = . %>% group_by(group, interval) %>% summarise(median = median(diff, na.rm = T)) %>% ungroup(), 
@@
 when calculating the distribution shown in the darker-colored set. The
-results suggest without dropping non-consecutive days in subplot (a), a
-gradual increase in error from 1.6% to 2.3% (1.2% to 1.9% with TOWT
-model normalization in subplot (b)) when the sampling interval increases
-from daily to weekly intervals. This is due to the fact that increasing
-the sample interval means reducing randomization. To quantify the
-combined impact of data loss due to preventing the carryover effect, we
-generated the lighter-colored set, which drops all non-consecutive days.
-As a result, we noticed an increase in estimation error when sampling
-daily to 2.4% (to 1.9% when TOWT is fitted), which led to a similar
-accuracy compared to weekly intervals. Therefore, considering this
-trade-off, we recommend using a two-day sampling interval if the
-carryover effect is likely but expected to last less than one day, and
-then normalizing the estimation via energy modeling. We also included
-the deviation distribution without absolute calculation and the
-deviation distribution calculated throughout the two years in the
-supplementary material and the results show the same indication.
+results suggest that without dropping non-consecutive days in subplot
+(a), a gradual increase in error from 1.5% to 2.3% (1.2% to 1.9% with
+weather normalization in subplot (b) when the sampling interval
+increases from daily to weekly intervals. This is due to the fact that
+increasing the sample interval means reducing randomization. To quantify
+the combined impact of data loss due to preventing the carryover effect,
+we generated the lighter-colored set, which drops all non-consecutive
+days. As a result, we noticed an increase in estimation error when
+sampling daily to 2.4% (to 1.9% when TOWT is fitted), which led to a
+similar accuracy compared to weekly intervals. Therefore, considering
+this trade-off, we recommend using a two-day sampling interval (or
+three-day sampling interval if the building owner prefers less frequent
+switchover) if the carryover effect is likely but expected to last less
+than one day, and then normalizing the estimation via energy modeling.
+The supplementary material contains similar figures showing biased
+deviation distribution, as well as the same results when continuing the
+randomized M&V for 24 months. Those results indicate the same
+conclusion; that sampling every two days and dropping non-consecutive
+samples yields the optimal result when carryover effects are present.
+
+Sampling at different intervals and excluding non-consecutive days also
+impact other stopping criteria, such as the time needed to satisfy
+statistical uncertainty and weather conditions thresholds. In Figure
+\@ref(fig:intervaltime), we compared the percentage of buildings
+satisfying all stopping criteria under different sampling intervals,
+with or without dropping non-consecutive days. The results indicate that
+randomized M&V concludes more quickly when all data are retained,
+particularly at shorter intervals, as expected. For instance, when
+sampling at a one-day interval, 90% of buildings meet all criteria
+within 24 weeks if no data are dropped, but this number decreases
+significantly to 52% when non-consecutive days (account for 1/3 of the
+data) are excluded, and 3% of the buildings require another 12 weeks to
+satisfy the weather criterion. As the sampling interval increases, this
+difference gradually diminishes: when sampling at a weekly interval,
+removing non-consecutive days has no impact on the timeline of meeting
+stopping criteria.
 
@@
 owners adjust the sampling ratio after meeting all stopping criteria, as
-continuing baseline sampling helps maintain M&V accuracy but may reduce
-the potential for energy savings. To further examine its impact on M&V
-accuracy, Figure \@ref(fig:abscont) compares two sampling ratios
-(20%/80% and 10%/90%) and two assocaited scenarios. The "Continued"
-group follows the standard randomized M&V framework with an initial
-50%/50% sampling ratio, then switches to the indicated ratio after
-satisfying the stopping criteria, continuing this until the end of the
-two-year period. In contrast, the "Randomized" group uses the indicated
-ratio from the outset and continues throughout the two year. Subplot (a)
-shows that absolute error is significantly lower without the initial
-50%/50% sampling, due to baseline drift and non-routine events that most
+continuing baseline sampling helps maintain M&V accuracy with limited
+impact on the energy savings actually achieved by the building. To
+further examine its impact on M&V accuracy, Figure \@ref(fig:abscont)
+compares two sampling ratios (20%/80% and 10%/90%) and two associated
+scenarios with the previously shown results when sampling at 50%/50%.
+The "Continued" group (shown in a lighter color) follows the standard
+randomized M&V framework with an initial 50%/50% sampling ratio, then
+switches to the indicated sampling ratio after satisfying the stopping
+criteria, continuing this until the end of the two-year period. In
+contrast, the "Randomized" group uses the indicated ratio from the
+outset and continues throughout the two year. Subplot (a) shows that
+absolute error is significantly lower without the initial 50%/50%
+sampling, due to baseline drift and non-routine events that most
 buildings experience over two years (as noted in Section
@@
 
-```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months", fig.height=6, fig.width=12}
+```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months (previous results of sampling at 50%/50% shown on the left side of the red dashed line)", fig.height=6, fig.width=12}
 
@@
 
-To further illustrate this point, we plotted results in Figure
-\@ref(fig:stablecont) using only the \'stable\' building subset (66
-buildings), as defined in Section \@ref(filter-and-clean-dataset). The
-differences between sampling strategies are much reduced in this subset
-and can be largely corrected through weather normalization modeling.
-This suggests that, consistent with our earlier findings regarding the
-convention M&V method, initial baseline measurements provide limited
-value in maintaining M&V accuracy when the building\'s baseline
-performance changes over time.
-
-```{r stablecont, fig.cap="Comparison of error distribution under different sampling ratio only for the buildings in 'stable' subset", fig.height=6, fig.width=12}
-p_top <- df_MW_S %>% 
-  mutate(ratio = as.factor(ratio), 
-         ratio = recode_factor(ratio, "1_1" = "Continued\n(20/80)", "1_2" = "Randomized\n(20/80)", "2_1" = "Continued\n(10/90)", "2_2" = "Randomized\n(10/90)")) %>% 
-  ggplot(aes(x = ratio, y = values, fill = ratio)) +
-  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
-  geom_lv(k = 4, outlier.shape = NA) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio)) +
-  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
-  geom_text(data = . %>% 
-              group_by(ratio) %>% 
-              summarise(median = median(values)) %>% 
-              ungroup(), 
-            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %"))) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 4), 
-                     labels = number_format(suffix = " %")) +
-  scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80")) + 
-  labs(fill = NULL, 
-       x = NULL, 
-       y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {S_building} buildings with measured weather conditions")) +
-  coord_cartesian(ylim = c(0, 12)) +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.position = "none",
-        axis.text.x = element_blank(), 
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-p_bottom <- df_TW_S %>% 
-  mutate(ratio = as.factor(ratio), 
-         ratio = recode_factor(ratio, "1_1" = "Continued\n(20/80)", "1_2" = "Randomized\n(20/80)", "2_1" = "Continued\n(10/90)", "2_2" = "Randomized\n(10/90)")) %>% 
-  ggplot(aes(x = ratio, y = values, fill = ratio)) +
-  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
-  geom_lv(k = 4, outlier.shape = NA) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio)) +
-  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
-  geom_text(data = . %>% 
-              group_by(ratio) %>% 
-              summarise(median = median(values)) %>% 
-              ungroup(), 
-            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %"))) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 4), 
-                     labels = number_format(suffix = " %")) +
-  scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80")) + 
-  labs(fill = NULL, 
-       x = NULL, 
-       y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {S_building} buildings with TOWT model and TMY weather")) +
-  coord_cartesian(ylim = c(0, 12)) +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.position = "none",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-ggarrange(p_top, p_bottom, 
-          ncol = 1, nrow = 2,
-          labels = c("a)", "b)")) +
-  plot_annotation(title = "Absolute error for the subset buildings with 2-year stable usage")
-```
-
-However, as shown in Figure \@ref(fig:abstract), the highest accuracy is
-achieved when sampling at a 50%/50% ratio. Therefore, we continue to
-recommend that building owners follow the proposed framework and
-maintain a 50%/50% sampling ratio until all stopping criteria are met.
-The additional results from Figure \@ref(fig:abscont) and Figure
-\@ref(fig:stablecont) further adds to the guidance: if building baseline
-changes, retrofit savings estimation should simply discard the existing
-baseline measurements and revert back to 50%/50% sampling ratio as
-described in the proposed framework.
+We further illustrated this point in the supplementary material where we
+plotted results using only the 'stable' building subset (66 buildings),
+as defined in Section \@ref(filter-and-clean-dataset). Since they are
+less influenced by non-routine events, the differences between sampling
+strategies are much reduced in this subset and can be largely corrected
+through weather normalization modeling.
+
+As shown in the previous Figure \@ref(fig:abstract), the highest
+accuracy is achieved when sampling at a 50%/50% ratio. Therefore, we
+continue to recommend that building owners follow the proposed framework
+and maintain a 50%/50% sampling ratio until all stopping criteria are
+met. The additional results from Figure \@ref(fig:abscont) further add
+to the guidance: if building baseline changes, retrofit savings
+estimation should simply discard the existing baseline measurements and
+revert back to 50%/50% sampling ratio as described in our framework. On
+the other hand, if the building owner has decided to test for two years
+and would like to capitalize more savings at the same time, sampling
+between the baseline and intervention at 10%/90% from start is
+recommended.
 
@@
 Similarly, we evaluated the time required to satisfy all stopping
-criteria when using a shorter blocking period, as shown in Figure
-\@ref(fig:6weekstartT). Compared to Figure \@ref(fig:startT), a
-significant proportion of buildings completed the randomized M&V by week
-42 (mid-December), which is one blocking period earlier than having a
-12-week blocking period. In both figures, we noticed a small proportion
-of buildings are less affected by changing the M&V start time, they are
-mostly located in climate zones with less outdoor temperature
-variations. This result further highlights that the overall duration of
-the randomized M&V process is strongly driven by outdoor weather.
-Accordingly, selecting an optimal blocking period should consider
-historical or typical meteorological weather file to effectively
-minimize the time required.
-
-```{r 6weekstartT, fig.cap = "Time required to satisfy all stopping criteria starting from March and using a shorter blocking period of 6 weeks"}
-
-knitr::include_graphics(paste0(fig_path, "6week_start_time.png"))
-```
+criteria when using a shorter blocking period, and included the plot in
+the supplementary material. We found that compared to Figure
+\@ref(fig:startT), a significant proportion of buildings completed the
+randomized M&V by week 42 (mid-December), which is one blocking period
+earlier than having a 12-week blocking period. This result further
+highlights that, in general, starting M&V in the middle of the winter or
+summer periods will yield results spanning the full range of weather
+conditions faster than starting at other times of year, such as during a
+swing season.
 

### 2025-06-27 code/manuscript.Rmd
updated supplementary material and manuscript

@@
 We tested this novel randomized M&V method on a large public dataset (639 buildings) covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention.
-The results show that compared to the conventional method, the randomized method provides more accurate savings estimations with a meidum of 74% accuracy improvement and is even faster.
+The results show that compared to the conventional method, the randomized method provides more accurate savings estimations with a median of 74% accuracy improvement and is even faster (typically 36 weeks instead of 2 years).
 Additionally, we found that when non-routine events are present, the randomized method estimates savings that are much closer to the ground-truth savings than the conventional method, demonstrating much improved reliability.
-We also assessed the impact of normalizing for different weather, starting the M&V at different time of the year, continuing randomization with a different sampling ratio after satisfying all criteria, dropping samples affected by carryover when switching between strategies, and identified the optimal sampling interval for that scenario using this large dataset.
+We also assessed the impact of normalizing for different weather, starting the M&V at different dates of the year, continuing randomization with a different sampling ratio after satisfying all criteria, and dropping samples affected by carryover effects when switching between strategies.
+For each scenario, we identified the optimal sampling interval using the large dataset.
 
@@
 bills or whole-building meters, however the overall conclusions would
-also apply to isolation methods. Typically, the process begins with
-baseline measurements taken during the year before implementing the
-energy-efficiency retrofit, followed by the same measurement procedure
-during the year after the intervention. After collecting two years of
-data, an M&V analyst fits an energy prediction model, using variables
-such as outdoor temperature and time [@mathieu_quantifying_2011] to
-project baseline energy consumption in the post-retrofit period. The
-difference between the counter-factual baseline and the measured
-intervention period energy consumption represents the energy savings.
-One of the key drawbacks of this method is its reliance on a two-year
-timeline to quantify savings, during which baseline measurements can
-become outdated due to changes in building performance caused by
-non-routine events unrelated to the intervention. Specifically, an M&V
-analyst should account for two types of baseline changes: (1) static
-changes, which are known building operational changes: such as
-renovation, equipment addition or removal; and (2) gradual changes,
-which often occur subtly and unknown, such as the incremental adoption
-of LED lighting or more efficient plug loads. Identifying and
-quantifying these changes are particularly labor-intensive and
-challenging. Consequently, those limitations reduce the feasibility of
-M&V and complicate the quantification of estimation uncertainty, thus
-impacting the accuracy and timeliness of savings assessments.
+also apply to isolation methods. In conventional practice, the process
+begins with baseline measurements taken during the year before
+implementing the energy-efficiency retrofit, followed by the same
+measurement procedure during the year after the intervention. After
+collecting two years of data, an M&V analyst fits an energy prediction
+model, using variables such as outdoor temperature and time
+[@mathieu_quantifying_2011] to project baseline energy consumption in
+the post-retrofit period. The difference between the counter-factual
+baseline and the measured intervention period energy consumption
+represents the energy savings. One of the key drawbacks of this method
+is its reliance on a two-year timeline to quantify savings, during which
+baseline measurements can become outdated due to changes in building
+performance caused by non-routine events unrelated to the intervention.
+Specifically, an M&V analyst should account for two types of baseline
+changes: (1) static changes, which are known building operational
+changes: such as renovation, equipment addition or removal, and changes
+related to tenant turnover; and (2) gradual changes, which often occur
+subtly and unknown, such as the incremental adoption of LED lighting or
+more efficient plug loads. Identifying and quantifying these changes are
+particularly labor-intensive and challenging. Consequently, those
+limitations reduce the feasibility of M&V and complicate the
+quantification of estimation uncertainty, thus impacting the accuracy
+and timeliness of savings assessments.
 
@@
 framework and defining stopping criteria to end the M&V period early
-after reaching a target level of uncertainty, among other improvements,
-such as using blocked randomization by weekday. The full framework is
-detailed in a previously published study [@raftery2024] along with a
-case study application to a real building, and other example use cases.
-In summary, this method provides M&V analysts with a randomized schedule
-that alternates between baseline and intervention implementation while
-ensuring balanced sampling across days of the week and seasons. For
-example, given a 10-week M&V period for one intervention, the balanced
-randomized schedule would equally sample five Mondays with the baseline
-strategy in operation and five Mondays with the intervention strategy in
-operation. The major limitation of this method is that it is only
-applicable to a subset of retrofit projects where interventions can be
-easily switched on and off, which is a small subset of all the
-interventions performed in buildings. However, for all applicable use
-cases, it allows analysts to detect energy savings sequentially shortly
-after the test begins. In addition, once the desired estimate
-uncertainty target is achieved, analysts can terminate the M&V study and
-switch to 100% intervention. The key advantage of randomization is that
-if control strategies are sampled with equal probability, the influence
-of other confounding factors such as occupancy change (a common
-non-routine event) and other more subtle long term changes in building
-energy use are likely to be evenly distributed among measurements,
-leading to a more accurate and unbiased assessment of the intervention
-effect.
+after reaching a target level of uncertainty, using blocked
+randomization by weekday. The full framework is detailed in a previously
+published study [@raftery2024] along with a case study application to a
+real building, and other example use cases. In summary, this method
+provides M&V analysts with a randomized schedule that alternates between
+baseline and intervention implementation while ensuring balanced
+sampling across days of the week and seasons. For example, given a
+10-week M&V period for one intervention, the balanced randomized
+schedule would equally sample five Mondays with the baseline strategy in
+operation and five Mondays with the intervention strategy in operation.
+The major limitation of this method is that it is only applicable to a
+subset of retrofit projects where interventions can be easily switched
+on and off, which is a small subset of all the interventions performed
+in buildings. However, for all applicable use cases, it allows analysts
+to detect energy savings sequentially shortly after the test begins. In
+addition, once the desired estimate uncertainty target is achieved,
+analysts can terminate the M&V study and switch to 100% intervention.
+The key advantage of randomization is that if control strategies are
+sampled with equal probability, the influence of other confounding
+factors such as occupancy change (a common non-routine event) and other
+more subtle long term changes in building energy use are likely to be
+evenly distributed among measurements, leading to a more accurate and
+unbiased assessment of the intervention effect.
 
@@
 energy to maintain the required duct static pressure [@feng_newly_2019;
-@zhai_full-scale_2017]. If M&V analysts are unaware of such changes and
-lack an appropriate adjustment method (e.g., replacing filters before
-the intervention begins), the savings could be overestimated as the
+@zhai_full-scale_2017]. We found a similar concept briefly defined in
+Guideline 14 as the 'load creep' [@ashrae_ashrae_2023] during multi-year
+M&V project. However, currently there is no detailed instruction
+provided. If M&V analysts are unaware of such changes and lack an
+appropriate adjustment method (e.g., replacing filters before the
+intervention begins), the savings could be overestimated as the
 decreased energy use is incorrectly attributed to the intervention
@@
 those with no statistically significant difference (P-value \> 0.05)
-between the two-year electricity usage. Applying this filter to real
+between their two-year electricity usage. Applying this filter to real
 buildings, only 66 buildings are labeled in this 'stable' subset out of
@@
 | Parameter              | Description                                                      | Value |
-|------------------|------------------------------------|------------------|
+|-------------------|-----------------------------------|-------------------|
 | $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
@@
 
--   A minimum and maximum of 12 and 108 weeks respectively. The
+-   A minimum and maximum of 12 and 108 weeks, respectively. The
     randomized schedule covers the entire two-year period but the
@@
 Acc = \frac{E_{baseline} - E_{intervention}}{E_{baseline}}-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
-$$where $E$ indicates true electricity usage and $E'$ indicates the
-estimated electricity usage either through the conventional method or
-the new randomized method. We present the distribution of savings
-estimation accuracy using boxen plots, also known as letter plots, which
-is an advanced variation of the box plot designed to extend beyond the
-interquartile range (IQR) by progressively dividing the data into
-smaller percentiles, revealing more detail in the tails of the
-distribution. We set the division parameter to k = 4, meaning the entire
-box area represents the data distribution from 6.25% to 93.75%. As the
-steps move closer to the center line (50% median), the distribution
-range progressively narrows with the next step representing 12.5% to
-87.5%, followed by 25% to 75%. In addition, for each boxen plot, we also
-showed the mean value of the distribution on top for reference.
+$$where $E$ indicates true electricity usage ($E_{intervention}$ is the
+'true' electricity usage calculated from Figure \@ref(fig:chwst) and
+above equations) and $E'$ indicates the estimated electricity usage
+either through the conventional method or the new randomized method. We
+present the distribution of savings estimation accuracy using boxen
+plots, also known as letter plots, which is an advanced variation of the
+box plot designed to extend beyond the interquartile range (IQR) by
+progressively dividing the data into smaller percentiles, revealing more
+detail in the tails of the distribution. We set the division parameter
+to k = 4, meaning the entire box area represents the data distribution
+from 6.25% to 93.75%. As the steps move closer to the center line (50%
+median), the distribution range progressively narrows with the next step
+representing 12.5% to 87.5%, followed by 25% to 75%. In addition, for
+each boxen plot, we also showed the mean value of the distribution on
+top for reference.
 
@@
 savings estimation error, we compared our normalized accuracy for 600+
-buildings with the TOWT model and TMY weather with their TOWT model
+buildings with the TOWT model and TMY weather againts their TOWT model
 prediction assessment in the following table, highlighting that the
@@
 decreasing the sampling interval, there is an accuracy penalty from
-dropping more non-consecutive days. To quantify such trade-off, Figure
-\@ref(fig:absinterval) shows the accuracy of different sampling
-intervals and the consequence of dropping the non-consecutive days in
-the dataset at the time when randomized M&V produces a result (after
-satisfying all stopping criteria).
+dropping more non-consecutive days. The dataset contains no specific
+information about the buildings' thermal mass condition, but to quantify
+such trade-off, we compared the accuracy of different sampling intervals
+and the consequence of dropping the non-consecutive days at the time
+when randomized M&V produces a result (after satisfying all stopping
+criteria) in Figure \@ref(fig:absinterval).
 
@@
 
-```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months (previous results of sampling at 50%/50% shown on the left side of the red dashed line)", fig.height=6, fig.width=12}
+```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months (previous results of sampling at 50%/50% shown on the left side of the red dashed line)", fig.height=8, fig.width=12}
 
@@
 the initial winter period, meeting the required weather condition
-criteria can potentially take a longer time. To assess this impact, we
-repeated the randomized M&V analysis for all buildings using a new start
-date in March. As shown in Figure \@ref(fig:startT), the new M&V process
-results in a noticeable delay in satisfying the stopping criteria.
-Compared to Figure \@ref(fig:intervaltime), where a January start
-allowed most buildings to finish within 36 weeks, a March start extended
-the required duration to 48 weeks for the majority of buildings.
-Furthermore, when a longer sampling interval is used or non-consecutive
-days are excluded (to avoid carryover effects), some buildings may
-require up to 60 weeks to complete the randomized M&V. As expected, to
-collect sufficient data representing winter conditions, the sampling may
-need to continue until the next January (around week 48). However, as
-the overall data collection period extends, the influence of dropping
-non-consecutive days diminishes and becomes less of a concern.
+criteria can potentially take a longer time. A previous study also
+showed that M&V accuracy, especially model fitting, depends on both
+monitoring duration and starting month, with optimal modeling achieved
+when sufficient weather variation is captured [@singh_predicting_2014].
+To assess this impact in our switchback randomization, we repeated the
+randomized M&V analysis for all buildings using a new start date in
+March. As shown in Figure \@ref(fig:startT), the new M&V process results
+in a noticeable delay in satisfying the stopping criteria. Compared to
+Figure \@ref(fig:intervaltime), where a January start allowed most
+buildings to finish within 36 weeks, a March start extended the required
+duration to 48 weeks for the majority of buildings. Furthermore, when a
+longer sampling interval is used or non-consecutive days are excluded
+(to avoid carryover effects), some buildings may require up to 60 weeks
+to complete the randomized M&V. As expected, to collect sufficient data
+representing winter conditions, the sampling may need to continue until
+the next January (around week 48). However, as the overall data
+collection period extends, the influence of dropping non-consecutive
+days diminishes and becomes less of a concern.
 
@@
 Specifically, we showed that throughout 11 different locations assessed
-in this study, the randomized M&V can provide a saving estimation after
+in this study, the randomized M&V can provide a savings estimation after
 36 weeks (with the majority finished by 24 weeks) once all stopping
@@
 strategies in the building and assess the impact of dropping samples
-where the strategy is non-consecutive and varying the sampling interval.
+when the strategy is non-consecutive and varying the sampling interval.
 Through that process, we showed that for the typical building, if
@@
 scenarios. We found that measuring a sufficient range of outdoor weather
-conditions remains the most stringent criterion. But even for the most
-non-ideal case, the randomized method can still produce an estimate much
-faster than 2 years, and produces an estimate with far lower error.
+conditions remains the most stringent criterion. However, even in the
+most non-ideal case, the randomized method can still produce an estimate
+much faster than two years, and produces an estimate with far lower
+error.
 

### 2025-06-27 code/supplementary.Rmd
updated supplementary material and manuscript

--- 2025-05-14 code/manuscript.Rmd
@@
 ---
-title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
-author:
-  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
-abstract: Conventional measurement and verification (M&V) methods rely on pre- and post- retrofit comparison to estimate energy savings.
-They are often time-consuming and unreliable, especially when non-routine events including static changes and gradual changes denoted as operational drift occurs during the M&V process.
-Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies conventional M&V suggested by industry guidelines, the results can be largely confounded.
-In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method which randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily).
-We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention.
-Our results show that compared to the conventional method, the randomized method provides faster and more accurate savings estimations.
-Additionally we found that when non-routine events are present, the randomized M&V approach estimates savings that are much closer to the ground-truth savings than the conventional M&V method, demonstrating much improved reliability.
-We also assessed the impact of dropping samples affected by carryover when switching between strategies, and identify the optimal sampling interval for that scenario using this large dataset.
+title: "Supplementary material"
 
@@
     wrap: 72
-bibliography: references.bib
-biblio-style: apalike
 
@@
                       dev = "jpeg",
-                      # cache = T,
                       dpi = 300,
@@
                       fig.pos = "b", 
-                      fig.path = "../figs/manuscript/figs/")
+                      fig.path = "../figs/supplementary/figs/")
 
@@
                "Randomized\n(10/90)" = "#6baed6", 
+               "Continued\n(20/80)" = "#c6dbef", 
+               "Continued\n(10/90)" = "#c6dbef",
                "Daily\nsampling" = "#66c2a4",
@@
 
-# Introduction
-
-## Conventional M&V
-
-Measurement and Verification (M&V) is the process of quantifying energy
-savings from energy efficiency projects by comparing actual energy
-consumption against a baseline, adjusting for factors like weather and
-occupancy. This process ensures that improvements in energy performance
-are accurately evaluated. In the United States, practitioners often
-refer to ASHRAE Guideline 14, the International Performance Measurement
-and Verification Protocol (IPMVP), and the Federal Energy Management
-Program (FEMP) for standard guidelines [@doe_mv_2008;
-@efficiency_valuation_organisation_international_2012;
-@ashrae_ashrae_2023]. These guidelines outline standardized methods for
-quantifying energy savings, whether through calibrated simulations or
-monitored measurements for specific equipment or systems (isolation
-methods) or for entire buildings (whole-building methods). In this
-study, we will focus on the energy savings quantified at the
-whole-building level where the measurements are obtained from utility
-bills or whole-building meters, however the overall conclusions would
-also apply to isolation methods. Typically, the process begins with
-baseline measurements taken during the year before implementing the
-energy-efficiency retrofit, followed by the same measurement procedure
-during the year after the intervention. After collecting two years of
-data, an M&V analyst fits an energy prediction model, using variables
-such as outdoor temperature and time [@mathieu_quantifying_2011] to
-project baseline energy consumption in the post-retrofit period. The
-difference between the counter-factual baseline and the measured
-intervention period energy consumption represents the energy savings.
-One of the key drawbacks of this method is its reliance on a two-year
-timeline to quantify savings, during which baseline measurements can
-become outdated due to changes in building performance caused by
-non-routine events unrelated to the intervention. Specifically, an M&V
-analyst should account for two types of baseline changes: (1) static
-changes, which are known building operational changes: such as
-renovation, equipment addition or removal; and (2) gradual changes,
-which often occur subtly and unknown, such as the incremental adoption
-of LED lighting or more efficient plug loads. Identifying and
-quantifying these changes are particularly labor-intensive and
-challenging. Consequently, those limitations reduce the feasibility of
-M&V and complicates the quantification of estimation uncertainty, thus
-impacting the accuracy and timeliness of savings assessments.
-
-## Randomized M&V
-
-To address the limitations of conventional M&V methods and the
-challenges posed by non-routine events, we proposed a M&V method that
-adopts the randomized crossover design [@raftery2024], a gold standard
-from medical, agricultural and online controlled experimental (i.e. A/B
-testing) studies [@blackston2019; @duan2013; @gupta2019; @mie2022;
-@munro; @raseduzzaman2017]. An early application of this randomized
-design to the field of M&V in buildings was first used in
-[@raftery2018], which compared the energy performance on two different
-supply air temperature reset strategies in a building. Subsequent
-improvements to the M&V method include adding a sequential evaluation
-framework and defined stopping criteria to end the M&V period early
-after reaching a target level of uncertainty, among other improvements,
-such as using blocked randomization by weekday. The full framework is
-detailed in a previously published study [@raftery2024] along with a
-case study application to a real building, and other example use cases.
-In summary, this method provides M&V analysts with a randomized schedule
-that alternates between baseline and intervention implementation while
-ensuring balanced sampling across days of the week and seasons. For
-example, given a 10-week M&V period for 1 intervention, the balanced
-randomized schedule would equally sample 5 Mondays with the baseline
-strategy in operation and 5 Mondays with the intervention strategy in
-operation. The limitation of this method is that it is only applicable
-to a subset of retrofit projects where interventions can be easily
-switched on and off. However, for all applicable use cases, it allows
-analysts to detect energy savings sequentially shortly after the test
-begins. In addition, once the desired estimate uncertainty target is
-achieved, analysts can terminate the M&V study and switch to 100%
-intervention. The key advantage of randomization is that if control
-strategies are sampled with equal probability, the influence of other
-confounding factors such as occupancy change (commonly known as one of
-the non-routine events) and other more subtle long term changes in
-building energy use are likely to be evenly distributed among
-measurements, leading to a more accurate and unbiased assessment of the
-intervention effect.
-
-## BDG2 dataset
-
-To demonstrate and differences between the two M&V methods, the Building
-Genome Dataset 2 (BGD2) is used, which is an extensive open-access
-dataset designed to advance research and development in building energy
-efficiency and control strategies acting as a test-bed for modeling,
-simulation, and algorithm development [@miller_building_2020]. BGD2
-contains over 1000 buildings' metadata and realistic operational
-measurements between 2016 and 2017 from across North America and Europe,
-making it one of the most comprehensive collections of building-related
-data available for scientific use. The dataset includes various
-commercial building types such as offices, education facilities, public,
-and retail buildings, and provides detailed information on their
-physical characteristics (e.g. energy ratings, heating types and floor
-area) and hourly measurements of chilled and hot water, electricity, gas
-usage as well as site outdoor weather conditions. In this study, we
-mostly use the metered electricity and outdoor weather measurements for
-running the M&V methods.
-
-## Literature review
-
-### Whole building approach
-
-ASHRAE Guideline 14 and the IPMVP provide options for whole-building M&V
-use cases. For code compliance using the prescriptive option, 12 months
-of baseline and 12 months of post-retrofit measurements are required. In
-addition, the expected savings should be larger than 10% and the
-baseline model fitting accuracy should have CV(RMSE) lower than 20%.
-Thus, most research related to M&V for whole-building approach focuses
-on the accuracy of baseline modeling, exploring model performance from
-simple regression models to more complex machine learning techniques.
-One study reviewed various models suitable for M&V applications as well
-as selected input features [@alrobaie_review_2022] and another study
-provided a definitive methodology to apply machine learning models for
-M&V use cases [@gallagher_development_2018]. In addition, a few studies
-investigated the critical performance metrics to evaluate the developed
-baseline models [@granderson_automated_2015;
-@granderson_development_2014] and compared a variety of models using
-those metrics [@granderson_accuracy_2016]. These studies made
-significant contributions by emphasizing the uncertainty associated with
-the model-fitting process, a key factor in accurately determining energy
-savings. Furthermore, other researchers addressed this issue by
-leveraging statistical formulation and inference to improve baseline
-energy models [@burkhart_measurement_2014; @heo_calibration_2012;
-@walter_uncertainty_2014]. However, a gap still remains in the
-literature regarding the rigorous quantification of uncertainties
-directly associated with calculated savings, for instance, accounting
-for the potential bias that the baseline model might deteriorate (i.e.
-become 'stale') over the extended period of pre- and post-analysis.
-
-### Non-routine events impact
-
-Non-routine events in M&V commonly refer to unexpected changes in a
-building that influence a building's energy usage and are typically
-unrelated to the intervention strategy. Current guidelines define
-non-routine events as 'static factors' that needed to be adjusted after
-projecting baseline in the post-retrofit period [@ashrae_ashrae_2023;
-@efficiency_valuation_organisation_international_2012]. A common
-non-routine event in energy-saving M&V projects is a change in occupancy
-or a significant shift in occupant behavior, equipment run time, and
-operating conditions (e.g. set points, lighting, and ventilation
-levels). However, current standards or guidelines only provide a general
-approach for consideration. For example, IPMVP requires the facility
-owner and the M&V analyst to periodically perform inspections of all
-equipment and operations during the reporting period, which is
-labor-intensive and error-prone. ASHRAE guideline 14 recommends
-performing engineering calculations or computer software simulations to
-adjust the post-retrofit baseline. Additionally, it is relatively rare
-to have access to all measurements needed for adjustments such as
-occupancy, and thus the analyst normally assumes those factors remain
-unchanged throughout the study. For example, some studies uses linear
-interpolation to estimate counter-factual baseline for demand response
-program [@beil_round-trip_2015; @keskar_commercial_2020]. Consequently,
-one study points out that it is inaccurate to assume no change in the
-operating conditions during the response period
-[@huang_experimental_2023]. Other studies realized such limitation in
-field study, and emphasized the importance of requiring matched groups
-to control for exogenous factors beyond weather differences when
-comparing between baseline and intervention
-[@demand_side_analytics_population_2022; @huang_experimental_2023].
-
-Overall, those non-routine events mostly refer to clearly observable
-changes, but there are also more gradual and subtle changes in
-buildings, which are hardly noticed. We define those gradual changes as
-'operational drift' in this study. One likely encountered operational
-drift situation is the filter clogging in air handling units due to
-particle accumulation. This causes supply fans to gradually consume more
-energy to maintain the required duct static pressure [@feng_newly_2019;
-@zhai_full-scale_2017]. If M&V analysts are unaware of such changes and
-lack an appropriate adjustment method (e.g., replacing filters before
-the intervention begins), the savings could be overestimated as the
-decreased energy use is incorrectly attributed to the intervention
-rather than the filter replacement. An operational drift has not been
-formally defined in the literature or standards, but in this study, we
-consider it as a special type of non-routine events as it also
-represents a change in facility operation but in a longer term.
-
-## Objectives
-
-As mentioned, the goal of an M&V project is to determine the effect,
-typically energy savings, of some intervention in the building. In this
-study, we focus on switchable interventions and an example of such an
-intervention is a control strategy that adjusts the chilled water
-plant's supply water temperature, which can be commonly found in the
-literature [@duarte_field_2023; @lee_simulationoptimization_2012;
-@qiu_chilled_2022; @jin2007; @taylor_optimizing_2012]. In our case, we
-make it even simpler by adjusting the setpoint based on outdoor weather
-conditions with more description shown in Section
-\@ref(apply-virtue-intervention). By conducting such an analysis, we aim
-to:
-
-1.  Determine, using a large sample of buildings, how much more
-    accurately the randomized M&V method would estimate the savings of
-    the proposed intervention compared to the conventional M&V, and how
-    much more quickly it would reach a result.
-
-2.  Determine how much more robust the randomized method is compared to
-    the conventional method when static non-routine events and
-    operational drift are present in the dataset. The BDG2 dataset
-    contains real-world building energy usage data collected over two
-    years, inherently contains varying degrees of usage changes. By
-    virtually introducing intervention effects into the existing
-    dataset, we can calculate the ground-truth energy savings, which
-    serve as a reference point for method comparison.
-
-3.  Open-source implementation of the proposed randomized M&V method. We
-    ensured the reproducibility of the method by making the analysis
-    code open source including randomized schedule generation,
-    sequential statistical analysis, energy modeling and normalized
-    saving calculation. In addition, we also included code for extended
-    use cases of the randomized method, such as changing sampling ratio
-    and sampling intervals. Using the available open resources, building
-    analysts should be able to seamlessly integrate and apply them in
-    their own M&V projects.
-
-# Method
-
-We outlined the methodology of the study in Figure \@ref(fig:flowchart)
-and extended several key components in this section.
-
-```{r flowchart, fig.cap = "Flow chart showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="75%"}
-
-knitr::include_graphics(paste0(fig_path, "flowchart.png"))
-```
-
-## Filter and clean dataset
-
-We extracted the electricity measurements from the BDG2 dataset and
-filter out all qualified buildings based on the following criteria:
-
-1.  Missing values \< 1000: given the hourly resolution of all
-    measurements, this is equivalent to 1.5 months of missing days.
-
-2.  Mean electricity usage \> 0 kWh: target buildings should have active
-    electricity usage.
-
-3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
-    decrease in building electricity usage in the second year should be
-    less than 25% of that in the first year.
-
-4.  Electric EUI \< 750 kWh/$m^2$: excludes buildings at the top 5% of
-    electrical energy usage intensity according to the statistics
-    provided by the Building Performance Database
-    [@lawrenceberkeleynationallab; @mathew_big-data_2015].
-
-5.  Warehouse and parking types are excluded: target buildings have less
-    demand flexibility to implement a chilled water set point reset
-    control.
-
-```{r stablesite, fig.cap = "Site summary of the filtered buildings from BDG2 dataset (counts < 10 are omitted for visualization; left: aggregated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}
-
-set3 <- colorRampPalette(brewer.pal('Set3',n=12))
-type_colors <- setNames(set3(13), all_types_variable$type)
-
-stable_set <- df_energy_stable %>%
-  group_by(type) %>%
-  distinct(name) %>%
-  summarise(stable = n()) %>%
-  ungroup()
-
-variable_set <- df_energy_variable %>%
-  group_by(type) %>%
-  distinct(name) %>%
-  summarise(variable = n()) %>%
-  ungroup()
-
-p1 <- variable_set %>% 
-  left_join(stable_set, by = "type") %>% 
-  mutate(stable = replace_na(stable, 0)) %>% 
-  group_by(type) %>% 
-  summarise(n = stable + variable) %>% 
-  ungroup() %>% 
-  mutate(type = fct_reorder(type, n, .desc = F)) %>%
-  ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
-  geom_col() +
-  geom_text(aes(label = ifelse(n > 10, n, "")), color = "black", position = position_stack(vjust = 0.5)) +
-  scale_y_continuous(expand = c(0, 0),
-                     breaks = breaks_pretty(n = 4)) +
-  scale_x_discrete(expand = c(0, 0.1)) +
-  scale_fill_manual(values = type_colors) +
-  labs(x = NULL,
-       y = NULL,
-       subtitle = "Combining all locations",
-       fill = NULL) +
-  theme(axis.text.x = element_blank(),
-        panel.grid.major.y = element_line(color = "grey80"),
-        legend.direction = "horizontal",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-stable_set <- df_energy_stable %>% 
-  left_join(df_loc, by = "site") %>% 
-  group_by(type, location) %>%
-  distinct(name) %>%
-  summarise(stable = n()) %>%
-  ungroup()
-
-variable_set <- df_energy_variable %>% 
-   left_join(df_loc, by = "site") %>% 
-  group_by(type, location) %>%
-  distinct(name) %>%
-  summarise(variable = n()) %>%
-  ungroup()
-
-p2 <- variable_set %>% 
-  left_join(stable_set, by = c("type", "location")) %>% 
-  mutate(stable = replace_na(stable, 0)) %>% 
-  group_by(location, type) %>% 
-  summarise(n = stable + variable) %>% 
-  ungroup() %>% 
-  group_by(location) %>%
-  mutate(proportion = n / sum(n),
-         total = sum(n),
-         ymax = cumsum(proportion),
-         ymin = c(0, head(ymax, n = -1))) %>%
-  mutate(label_pos = (ymax + ymin) / 2) %>%
-  ungroup() %>%
-  group_by(type) %>%
-  mutate(order = sum(n)) %>%
-  ungroup() %>%
-  mutate(type = fct_reorder(type, order, .desc = F)) %>%
-  ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3.2, fill = type)) +
-  geom_rect() +
-  coord_polar(theta = "y") +
-  xlim(c(2, 4)) +
-  facet_wrap(~ location, nrow = 3) +
-  labs(x = NULL,
-       y = NULL,
-       subtitle = "For each location",
-       fill = NULL) +
-  scale_fill_manual(values = type_colors) +
-  geom_text(aes(x = 3.7, y = label_pos, label = ifelse(n > 10, n, ""))) +
-  geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black", check_overlap = T) +
-  theme(legend.direction = "horizontal",
-        axis.text = element_blank(),
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-ggarrange(p1, p2,
-          ncol = 2, nrow = 1,
-          widths = c(0.2, 1),
-          common.legend = TRUE,
-          legend = "bottom") +
-  plot_annotation(title = "Summary of filtered buildings and types")
-```
-
-For all qualified buildings, we further added one more stringent
-criterion to explore all buildings with very stable electricity usage
-between the two years:
-
-1.  No statistically significant difference (P-value \> 0.05) between
-    the two-year electricity usage, which are sometimes assumed to be
-    the case for whole-building measurement and verification.
-
-As a result, there are only 66 buildings were labeled in this 'stable'
-subset (out of total of 600 buildings) implying that the assumption that
-buildings are typically stable over the timeframe involved in
-conventional M&V is rarely true. As mentioned in Section
-\@ref(non-routine-events-impact), such variability is largely associated
-with non-routine events, especially operational drifts since the
-assessment last for two years. And according to the statistics shown
-here, it is a typical case observed in real buildings applications.
-
-## Apply virtue intervention
-
-Figure \@ref(fig:chwst) shows the algorithm for the proposed control
-intervention that reset the chiller supply temperature based on the
-outdoor weather conditions, which can be commonly found in the
-literature [@lee_chilled_2022; @congradac_recognition_2012]. For both
-strategies, we assume that the chiller operates when the outdoor
-temperature exceeds 10°C. The baseline strategy, representing the
-existing measurements from the dataset, operates with a constant water
-supply temperature. The intervention strategy, as illustrated in the
-figure, adjusts the water supply temperature dynamically, resetting it
-from 7°C to 12°C.
-
-```{r chwst, fig.cap = "Proposed intervention strategy: chilled water supply temperature reset based on outdoor temperature.", fig.width=8, fig.height=4}
-
-data.frame(
-  OAT = c(10, 15, 25, 30),  
-  Baseline = c(6, 6, 6, 6),      
-  Intervention = c(12, 12, 7, 7)
-  ) %>% 
-  pivot_longer(cols = c("Baseline", "Intervention"), names_to = "strategy", values_to = "value") %>% 
-  ggplot(aes(x = OAT, y = value, color = strategy)) +
-  geom_line(linewidth = 1.2) +
-  scale_color_manual(values = ls_colors) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = c(4, 6, 8, 10, 12),
-                     limits = c(5, 12.5),
-                     labels = number_format(suffix = " °C")) +
-  scale_x_continuous(expand = c(0, 0), 
-                     breaks = c(5, 10, 15, 20, 25, 30), 
-                     labels = number_format(suffix = " °C")) +
-  geom_vline(xintercept = 10, lty = "dashed", color = "red") +
-  annotate(geom = "text", x = 7.5, y = 8, label = "No chiller operation\nbelow 10 °C") +
-  coord_cartesian(xlim = c(5, 30)) +
-  labs(x = "Outdoor temperature", 
-       y = "Chilled water supply temperature", 
-       color = NULL, 
-       title = "Chilled water supply temperature reset intervention") +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.direction = "horizontal",
-        legend.position = "bottom",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-```
-
-We used a simple relationship to map the chilled water supply
-temperature reset to the electrical energy savings: $$
-\text{savings} = mean(E_{CHW}) \times 25\% \times \text{perc_savings}
-$$
-
-We assume on average, HVAC systems account for approximately 50% of a
-building's total electricity consumption, and the chilled water plant
-further consumes around 50% of the HVAC electricity. While this
-assumption largely simplifies the diverse energy usage across various
-building types, for the scope of this paper, we assume that 25% of the
-total building electricity is used by the chilled water plant, $E_{CHW}$
-[@us2012commercial]. Typically, the savings from an intervention are not
-proportional to the building's hourly electricity usage, which is
-generally the challenge for M&V. To address this, we mapped the
-resulting electricity savings as a percentage of the plant's normal
-operation, calculated as its mean electricity usage over the two-year
-period. This percentage is influenced by factors such as outdoor
-temperature ($OAT$), intervention supply water temperature
-($T_{interv}$), baseline supply water temperature ($T_{base}$) and hour
-of the day ($\delta_{occ}$, binary indicator whether during peak hours
-from 9 AM to 4 PM).
-
-$$
-\text{perc_savings = }
-\begin{cases}
-0, & \text{ if } OAT < 10°C \\
-\left[ \beta_{\text{temp}} \times (\text{T}_{\text{interv}} - \text{T}_{\text{base}}) \right] \times 
-\left[ \beta_{\text{occ}} \times \delta_{occ} + \beta_{\text{unocc}} \times (1 - \delta_{occ})\right], & \text{ otherwise}
-\end{cases}
-$$
-
-Parameters and their pre-defined values are summarized in the table
-below. For simplicity, those parameters were not rigorously calibrated
-for each building and were applied uniformly in the filtered dataset.
-Interested readers can also change the parameters in our open-source
-code to simulate different scenarios.
-
-|                        |                                                                  |       |
-|-------------------|------------------------------------|-------------------|
-| Parameter              | Description                                                      | Value |
-| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
-| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
-| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |
-
-: Table 1. Parameters for calculating the intervention savings.
-
-## Run M&V methods
-
-We described in detail the workflow of both conventional and randomized
-M&V methods in the previous study [@raftery2024]. The conventional
-method is a pre- and post- comparison on a 12-month baseline and
-12-month intervention timeline. The randomized method defines sampling
-requirements as:
-
--   Use a daily sampling interval with the sampling time at midnight
-    each day.
-
--   Block by day of the week with a block period of 12 weeks.
-
-and stopping criteria as:
-
--   A minimum and maximum of 12 and 108 weeks respectively. The
-    randomized schedule covers the entire two-year period but stopping
-    criteria enables an early stop at the end of the satisfied blocking
-    period.
-
--   At least 80% of the dry-bulb temperature range in the annual TMY
-    data sampled by both strategies.
-
--   No carryover effect using a t-test with a p-value not exceeding a
-    defined significance threshold of 0.05.
-
--   90% confidence that energy savings exceed or do not exceed 0% using
-    the SPRT test. Medium effect size (d = 0.5) quantified by cohen's d
-    and calculated SPRT statistics either falls below the lower
-    threshold or exceeds the upper threshold.
-
--   Sample strategies at an equal sampling ratio (50% baseline, 50%
-    intervention).
-
-# Results
-
-The accuracy ($Acc$) is calculated as: $$
-Acc = \frac{E_{baseline} - E_{intervention}}{E_{baseline}}-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
-$$where $E$ indicates true electricity usage and $E'$ indicates the
-estimated electricity usage either through the conventional method or
-the new randomized method. We present the distribution of savings
-estimation accuracy using boxen plots, also known as letter plots, which
-is an advanced variation of the box plot designed to extend beyond the
-interquartile range (IQR) by progressively dividing the data into
-smaller percentiles, revealing more detail in the tails of the
-distribution. We set the division parameter to k = 4, meaning the entire
-box area represents the data distribution from 6.25% to 93.75%. As the
-steps move closer to the center line (50% median), the distribution
-range progressively narrows with the next step representing 12.5% to
-87.5%, followed by 25% to 75%. In addition, for each boxen plot, we also
-showed the mean value of the distribution on top for reference.
-
-```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period).", fig.height=12, fig.width=12}
-
-#### GA ####
-rand_eob_S <- df_seq_fs_stable %>% 
+```{r absnull, fig.cap="Comparison between the two M&V methods in detecting no intervention effect (expressed without absolute calculation) when buildings are subject to baseline change.", fig.height=6, fig.width=12}
+#### ABS-NULL ####
+rand_eob_S <- df_seq_fs_null_stable %>% 
   filter(seq == "eob") %>% 
-  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = abs(savings - fs), 
+  mutate(diff = abs(0 - fs), 
          method = "rand_eob") %>% 
@@
 
-conv_S <- df_fs_stable %>% 
+conv_S <- df_fs_null_stable %>% 
   filter(method != "rand") %>% 
@@
 
-rand_final_S <- df_fs_stable %>% 
+rand_final_S <- df_fs_null_stable %>% 
   filter(method != "conv") %>% 
@@
 # plot for the variable subset
-rand_eob_V <- df_seq_fs_variable %>% 
+rand_eob_V <- df_seq_fs_null_variable %>% 
   filter(seq == "eob") %>% 
-  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site"))%>% 
-  mutate(diff = abs(savings - fs), 
+  mutate(diff = abs(0 - fs), 
          method = "rand_eob") %>% 
@@
 
-conv_V <- df_fs_variable %>% 
+conv_V <- df_fs_null_variable %>% 
   filter(method != "rand") %>% 
@@
 
-rand_final_V <- df_fs_variable %>% 
+rand_final_V <- df_fs_null_variable %>% 
   filter(method != "conv") %>% 
@@
 
-p_top <- df_MW_A %>% 
+df_MW_A %>% 
   mutate(method = as.factor(method), 
@@
   geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
-  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
-  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff, na.rm = T)) %>% ungroup(), 
-            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 4), 
-                     labels = number_format(suffix = " %")) +
-  scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
-  labs(fill = NULL, 
-       color = NULL, 
-       x = NULL, 
-       y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with no model normalization on weather")) +
-  coord_cartesian(ylim = c(0, 23)) +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.position = "none",
-        axis.text.x = element_blank(),
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-# TMY
-# plot for the stable subset
-rand_eob_S <- df_sprt_all_stable %>% 
-  filter(seq == "eob") %>% 
-  select(-c(seq, n_weeks)) %>% 
-  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = abs(savings - annual), 
-         method = "rand_eob") %>% 
-  select(name, method, diff)
-
-conv_S <- df_fs_tmy_stable %>% 
-  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = abs(savings - conv), 
-         method = "conv") %>% 
-  select(name, method, diff)
-
-rand_final_S <- df_fs_tmy_stable %>% 
-  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = abs(savings - rand), 
-         method = "rand_final") %>% 
-  select(name, method, diff)
-
-df_TW_S <- rbind(rand_eob_S, conv_S, rand_final_S)
-
-# plot for the variable subset
-rand_eob_V <- df_sprt_all_variable %>% 
-  filter(seq == "eob") %>% 
-  select(-c(seq, n_weeks)) %>% 
-  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = abs(savings - annual), 
-         method = "rand_eob") %>% 
-  select(name, method, diff)
-
-conv_V <- df_fs_tmy_variable %>% 
-  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = abs(savings - conv), 
-         method = "conv") %>% 
-  select(name, method, diff)
-
-rand_final_V <- df_fs_tmy_variable %>% 
-  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = abs(savings - rand), 
-         method = "rand_final") %>% 
-  select(name, method, diff)
-
-df_TW_V <- rbind(rand_eob_V, conv_V, rand_final_V)
-
-# plot for combined dataset
-rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
-  select(name, rand_eob_diff = diff)
-
-conv_A <- bind_rows(conv_V, conv_S) %>% 
-  select(name, conv_diff = diff)
-
-rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
-  select(name, rand_final_diff = diff)
-
-df_TW_A <- rand_eob_A %>% 
-  left_join(conv_A, by = "name") %>% 
-  left_join(rand_final_A, by = "name") %>% 
-  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")
-
-p_middle <- df_TW_A %>% 
-  mutate(method = as.factor(method), 
-         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
-  ggplot(aes(x = method, y = diff, fill = method)) +
-  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
-  geom_lv(k = 4, outlier.shape = NA) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
-  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
-  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
-            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
+  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
+            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
+            size = 4) +
   scale_y_continuous(expand = c(0, 0), 
@@
        x = NULL, 
-       y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with TOWT model normalized on TMY weather")) +
+       y = "Absolute effect in fractional savings", 
+       title = "Absolute effect of non-routine events when using different M&V methods", 
+       subtitle = str_glue("All {A_building} buildings wither measured weather")) +
   coord_cartesian(ylim = c(0, 23)) +
@@
         plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-p_accuracy <- ggarrange(p_top, p_middle, 
-                        ncol = 1, nrow = 2,
-                        labels = c("a)", "b)"),
-                        common.legend = T, 
-                        legend = "bottom")
-
-# Timeline plot
-df_time <- df_sprt_all_variable %>% 
-  filter(seq != "final") %>% 
-  bind_rows(df_sprt_all_stable %>% filter(seq != "final")) %>% 
-  select(name, seq, n_weeks)
-
-count <- list()
-n <- 1
-for (i in seq(0, 36, by = 3)){
-
-  df_sprt <- df_time %>% 
-    filter(seq == "sprt",
-           n_weeks <= i)
-
-  df_eob <- df_time %>% 
-    filter(seq == "eob",
-           n_weeks <= i)
-
-  df_temp <- df_time %>% 
-    filter(seq == "temp", 
-           n_weeks <= i)
-
-  count[[n]] <- tibble("n_weeks" = i, 
-                       "eob" = nrow(df_eob), 
-                       "temp" = nrow(df_temp), 
-                       "sprt" = nrow(df_sprt))
-
-  n <- n + 1
-}
-
-count <- bind_rows(count)
-
-p_timeline <- count %>% 
-  ggplot() +
-  geom_bar(data = . %>% 
-             filter(n_weeks == 24 | n_weeks == 36),
-           aes(x = n_weeks, y = eob, fill = "Buildings finishing randomized M&V"), 
-           stat = "identity", 
-           position = "stack",
-           alpha = 0.8, 
-           width = 3) +
-  geom_line(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"),
-            alpha = 0.2) +
-  geom_point(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"), 
-             size = 1.5) +
-  geom_line(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"),
-            alpha = 0.2) +
-  geom_point(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"), 
-             size = 1.5, 
-             shape = 17) +
-  geom_segment(aes(x = 36.5, y = max(eob), xend = 95.5, yend = max(eob)),
-               arrow = arrow(length = unit(0.25, "in")),   
-               linewidth = 1.1,  
-               color = "#fb8072") + 
-  annotate(geom = "text", 
-           x = 66, 
-           y = max(count$eob) - 30, 
-           size = 5,
-           label = "Excess time by conventional M&V") +
-  geom_vline(xintercept = c(12, 24, 36, 48, 96), lty = "dashed", color = "grey80") +
-  annotate(geom = "text", 
-           x = seq(6, 48, by = 12), 
-           y = 200, 
-           label = paste0("12-week\nblock"), 
-           alpha = 0.5) +
-  scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = ls_colors) +
-  scale_x_continuous(expand = c(0, 0), 
-                     limits = c(0, 100),
-                     breaks = c(12, 24, 36, 48, 96), 
-                     labels = c("12 weeks", "24 weeks", "36 weeks", "1 year", "2 years")) +
-  coord_cartesian(ylim = c(0, 650)) +
-  labs(x = NULL, 
-       y = "Number of buildings", 
-       fill = NULL, 
-       color = NULL, 
-       subtitle = "timeline comparison of satisfying all stopping criteria") +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.direction = "horizontal",
-        legend.position = "bottom",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-ggarrange(p_accuracy, p_timeline, 
-          ncol = 1, nrow = 2,
-          heights = c(2.2, 1),
-          labels = c("", "c)"),
-          common.legend = F, 
-          legend = "bottom") +
-  plot_annotation(title = "Overall accuracy and time required for savings estimation by different M&V methods")
 ```
 
-Figure \@ref(fig:abstract) shows the overall results of M&V methods
-comparison and for clarity, we plotted conventional M&V results in grey,
-randomized M&V results after satisfying all stopping criteria in green,
-and the randomized M&V results sampled for the entire 24 months in blue.
-Subplots (a) and (b) calculate the savings estimation error as the
-absolute deviation from the true savings (i.e. $|Acc|$) and show the
-results distribution using the conventional M&V method in the first
-column, the randomized M&V method that stops after satisfying all
-stopping criteria in the second column, and the randomized M&V method
-continues 50%/50% sampling throughout 2 years in the third column. In
-addition, subplot (a) shows the savings estimated from measured weather
-conditions, and subplot b) shows the savings normalized on the typical
-meteorological weather of the building location after using TOWT model.
-As a result, the conventional M&V method exhibits an average deviation
-of 6% in savings estimation, whereas the proposed randomized method
-demonstrates significantly smaller deviations. If the analyst stops
-immediately after satisfying all stopping criteria, the deviation is
-reduced to approximately 2%. Extending the M&V period to match the
-length of the conventional method (24 months) further improves accuracy
-to 1%. Furthermore, by comparing subplots (a) and (b), we observe a
-minor improvement in estimation accuracy when a model is fitted to the
-data to account for weather differences. The model fitting is most
-beneficial for the shorter dataset (randomized and stopping when all
-criteria are met, typically 24-36 weeks of data), where it improves
-estimation accuracy by 0.6%. This is because the weather conditions
-between the first and second year for each of the locations are
-relatively similar on average over this time period, and thus the effect
-of adjusting for temperature as an independent variable in the model is
-relatively small. Subplot (c) compares the overall timeline of the two
-M&V methods. The results show that almost all buildings achieve accurate
-M&V results and meet the stopping criteria within 24 weeks (for this
-intervention), with the remainder meeting all stopping criteria within
-36 weeks. The additional 12 weeks is usually driven by a need to span
-sufficient weather conditions representative of a full year of data.
-Meanwhile, the red arrow highlights the excess time required by the
-conventional M&V method (24 months). Thus, the randomized M&V method
-finds more accurate results in approximately a quarter of the time
-required for conventional M&V. Even if it was the case that an existing
-12-month baseline data was available at the start of the M&V process,
-randomized M&V would still obtain a more accurate result, in less time,
-than obtaining the 12-month retrofit dataset required by conventional
-M&V.
-
-We also plotted the mean error in Figure \@ref(fig:meanerr) and we
-noticed that the uncertainty range associated with the conventional
-method is significantly larger compared to the randomized method that
-only takes 36 weeks according to subplot (c) in Figure
-\@ref(fig:abstract).
-
-```{r meanerr, fig.cap="Distribution of mean deviation from the ground-truth savings calculated by the two M&V methods", fig.height=6, fig.width=12}
-# plot for the stable subset
-rand_eob_S <- df_seq_fs_stable %>% 
-  filter(seq == "eob") %>% 
+```{r meancont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy (expressed without absolute calculation for errors)", fig.height=8, fig.width=12}
+
+#### MEAN-CONT ####
+df_MW_S <- df_cont_stable %>% 
   left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = savings - fs, 
-         method = "rand_eob") %>% 
-  select(name, method, diff)
-
-conv_S <- df_fs_stable %>% 
-  filter(method != "rand") %>% 
-  pivot_wider(names_from = method, values_from = savings) %>% 
-  mutate(diff = true - conv, 
-         method = "conv") %>% 
-  select(name, method, diff)
-
-rand_final_S <- df_fs_stable %>% 
+  mutate(diff = savings - cont_fs, 
+         diff_all = savings - cont_fs_all) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_stable %>% 
               filter(method != "conv") %>% 
               pivot_wider(names_from = method, values_from = savings) %>% 
-  mutate(diff = true - rand, 
-         method = "rand_final") %>% 
-  select(name, method, diff)
-
-df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)
+              mutate(values = true - rand, 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
 
 # plot for the variable subset
-rand_eob_V <- df_seq_fs_variable %>% 
-  filter(seq == "eob") %>% 
-  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site"))%>% 
-  mutate(diff = savings - fs, 
-         method = "rand_eob") %>% 
-  select(name, method, diff)
-
-conv_V <- df_fs_variable %>% 
-  filter(method != "rand") %>% 
-  pivot_wider(names_from = method, values_from = savings) %>% 
-  mutate(diff = true - conv, 
-         method = "conv") %>% 
-  select(name, method, diff)
-
-rand_final_V <- df_fs_variable %>% 
+df_MW_V <- df_cont_variable %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = savings - cont_fs, 
+         diff_all = savings - cont_fs_all) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_variable %>% 
               filter(method != "conv") %>% 
               pivot_wider(names_from = method, values_from = savings) %>% 
-  mutate(diff = true - rand, 
-         method = "rand_final") %>% 
-  select(name, method, diff)
-
-df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)
+              mutate(values = true - rand, 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
 
 # plot for combined dataset
-rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
-  select(name, rand_eob_diff = diff)
-
-conv_A <- bind_rows(conv_V, conv_S) %>% 
-  select(name, conv_diff = diff)
-
-rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
-  select(name, rand_final_diff = diff)
-
-df_MW_A <- rand_eob_A %>% 
-  left_join(conv_A, by = "name") %>% 
-  left_join(rand_final_A, by = "name") %>% 
-  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")
+df_MW_A <- rbind(df_MW_S, df_MW_V)
 
 p_top <- df_MW_A %>% 
-  mutate(method = as.factor(method), 
-         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
-  ggplot(aes(x = method, y = diff, fill = method)) +
+  mutate(ratio = as.factor(ratio), 
+         ratio = recode_factor(ratio, 
+                               "0" = "Randomized\n(50/50)", 
+                               "1_1" = "Continued\n(20/80)", 
+                               "1_2" = "Randomized\n(20/80)", 
+                               "2_1" = "Continued\n(10/90)", 
+                               "2_2" = "Randomized\n(10/90)")) %>% 
+  ggplot(aes(x = ratio, y = values, fill = ratio)) +
   geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
-  geom_lv(k = 4, outlier.shape = NA) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
-  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
-  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
-            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
+  geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75, linewidth = 0.15) +
+  geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% 
+              group_by(ratio) %>% 
+              summarise(median = median(values)) %>% 
+              ungroup(), 
+            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %")), 
+            size = 4) +
   scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 4), 
+                     breaks = breaks_pretty(n = 5), 
                      labels = number_format(suffix = " %")) +
   scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
+  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80", "grey80")) + 
   labs(fill = NULL, 
-       color = NULL, 
        x = NULL, 
-       y = "Mean error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with no model normalization on weather")) +
-  coord_cartesian(ylim = c(-20, 20)) +
+       y = "Absolute error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with measured weather conditions")) +
+  coord_cartesian(ylim = c(-12, 15)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
 # plot for the stable subset
-rand_eob_S <- df_sprt_all_stable %>% 
-  filter(seq == "eob") %>% 
-  select(-c(seq, n_weeks)) %>% 
+df_TW_S <- df_cont_stable %>% 
   left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = savings - annual, 
-         method = "rand_eob") %>% 
-  select(name, method, diff)
-
-conv_S <- df_fs_tmy_stable %>% 
+  mutate(diff = savings - cont_tmy, 
+         diff_all = savings - cont_tmy_all) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_tmy_stable %>% 
               left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = savings - conv, 
-         method = "conv") %>% 
-  select(name, method, diff)
-
-rand_final_S <- df_fs_tmy_stable %>% 
-  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = savings - rand, 
-         method = "rand_final") %>% 
-  select(name, method, diff)
-
-df_TW_S <- rbind(rand_eob_S, conv_S, rand_final_S)
+              mutate(values = savings - rand, 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
 
 # plot for the variable subset
-rand_eob_V <- df_sprt_all_variable %>% 
-  filter(seq == "eob") %>% 
-  select(-c(seq, n_weeks)) %>% 
+df_TW_V <- df_cont_variable %>% 
   left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = savings - annual, 
-         method = "rand_eob") %>% 
-  select(name, method, diff)
-
-conv_V <- df_fs_tmy_variable %>% 
+  mutate(diff = savings - cont_tmy, 
+         diff_all = savings - cont_tmy_all) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_tmy_variable %>% 
               left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = savings - conv, 
-         method = "conv") %>% 
-  select(name, method, diff)
-
-rand_final_V <- df_fs_tmy_variable %>% 
-  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = savings - rand, 
-         method = "rand_final") %>% 
-  select(name, method, diff)
-
-df_TW_V <- rbind(rand_eob_V, conv_V, rand_final_V)
+              mutate(values = savings - rand, 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
 
 # plot for combined dataset
-rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
-  select(name, rand_eob_diff = diff)
-
-conv_A <- bind_rows(conv_V, conv_S) %>% 
-  select(name, conv_diff = diff)
-
-rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
-  select(name, rand_final_diff = diff)
-
-df_TW_A <- rand_eob_A %>% 
-  left_join(conv_A, by = "name") %>% 
-  left_join(rand_final_A, by = "name") %>% 
-  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")
-
-p_middle <- df_TW_A %>% 
-  mutate(method = as.factor(method), 
-         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
-  ggplot(aes(x = method, y = diff, fill = method)) +
+df_TW_A <- rbind(df_TW_S, df_TW_V)
+
+p_bottom <- df_TW_A %>% 
+  mutate(ratio = as.factor(ratio), 
+         ratio = recode_factor(ratio, 
+                               "0" = "Randomized\n(50/50)", 
+                               "1_1" = "Continued\n(20/80)", 
+                               "1_2" = "Randomized\n(20/80)", 
+                               "2_1" = "Continued\n(10/90)", 
+                               "2_2" = "Randomized\n(10/90)")) %>% 
+  ggplot(aes(x = ratio, y = values, fill = ratio)) +
   geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
-  geom_lv(k = 4, outlier.shape = NA) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
-  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
-  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
-            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
+  geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75, linewidth = 0.15) +
+  geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% 
+              group_by(ratio) %>% 
+              summarise(median = median(values)) %>% 
+              ungroup(), 
+            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %")), 
+            size = 4) +
   scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 4), 
+                     breaks = breaks_pretty(n = 5), 
                      labels = number_format(suffix = " %")) +
   scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
+  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80", "grey80")) + 
   labs(fill = NULL, 
        x = NULL, 
-       y = "Mean error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with TOWT model normalized on TMY weather")) +
-  coord_cartesian(ylim = c(-20, 20)) +
+       y = "Absolute error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
+  coord_cartesian(ylim = c(-12, 15)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
 
-ggarrange(p_top, p_middle, 
+ggarrange(p_top, p_bottom, 
           ncol = 1, nrow = 2,
           labels = c("a)", "b)")) +
-  plot_annotation(title = "Mean error in savings estimation by different M&V methods")
-
-percentiles <- df_TW_A %>%
-  filter(method == "conv_diff") %>% 
-  summarise(
-    Q1     = round(quantile(diff, 0.25, na.rm = TRUE), digits = 2),
-    Median = round(quantile(diff, 0.50, na.rm = TRUE), digits = 2),
-    Q3     = round(quantile(diff, 0.75, na.rm = TRUE), digits = 2)
-  )
+  plot_annotation(title = "Absolute error in savings estimation by different sampling ratio")
 ```
 
-Additionally in the literature, we found another M&V study using over
-500+ commercial buildings, but their study scope was to compare the
-predictive performance of a variety of M&V baseline models from simple
-weekly mean to more complex machine learning methods using the
-normalized mean bias error (NMBE) [@granderson_accuracy_2016]. Since we
-also calculated the normalized savings estimation error, we compared our
-normalized accuracy for 600+ buildings with the TOWT model and TMY
-weather with their TOWT model prediction assessment in the following
-table, highlighting that the model and overall building datasets are
-performing similarly.
-
-|                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
-|-------------------|----------------------------------|-------------------|
-| 25th percentile | -5.85                                               | `r percentiles$Q1`     |
-| 50th percentile | -1.25                                               | `r percentiles$Median` |
-| 75th percentile | 3.86                                                | `r percentiles$Q3`     |
-
-: Table 1. Savings accuracy comparison with a similar literature
-
-## Non-routine events impact
-
-```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect when buildings are subject to static non-rountine events and gradual operational drift.", fig.height=6, fig.width=12}
-
-#### MEAN-NULL ####
-rand_eob_S <- df_seq_fs_null_stable %>% 
-  filter(seq == "eob") %>% 
-  mutate(diff = 0 - fs, 
-         method = "rand_eob") %>% 
-  select(name, method, diff)
-
-conv_S <- df_fs_null_stable %>% 
-  filter(method != "rand") %>% 
-  pivot_wider(names_from = method, values_from = savings) %>% 
-  mutate(diff = true - conv, 
-         method = "conv") %>% 
-  select(name, method, diff)
-
-rand_final_S <- df_fs_null_stable %>% 
-  filter(method != "conv") %>% 
-  pivot_wider(names_from = method, values_from = savings) %>% 
-  mutate(diff = true - rand, 
-         method = "rand_final") %>% 
-  select(name, method, diff)
-
-df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)
-
-# plot for the variable subset
-rand_eob_V <- df_seq_fs_null_variable %>% 
-  filter(seq == "eob") %>% 
-  mutate(diff = 0 - fs, 
-         method = "rand_eob") %>% 
-  select(name, method, diff)
-
-conv_V <- df_fs_null_variable %>% 
-  filter(method != "rand") %>% 
-  pivot_wider(names_from = method, values_from = savings) %>% 
-  mutate(diff = true - conv, 
-         method = "conv") %>% 
-  select(name, method, diff)
-
-rand_final_V <- df_fs_null_variable %>% 
-  filter(method != "conv") %>% 
-  pivot_wider(names_from = method, values_from = savings) %>% 
-  mutate(diff = true - rand, 
-         method = "rand_final") %>% 
-  select(name, method, diff)
-
-df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)
-
-# plot for combined dataset
-rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
-  select(name, rand_eob_diff = diff)
-
-conv_A <- bind_rows(conv_V, conv_S) %>% 
-  select(name, conv_diff = diff)
-
-rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
-  select(name, rand_final_diff = diff)
-
-df_MW_A <- rand_eob_A %>% 
-  left_join(conv_A, by = "name") %>% 
-  left_join(rand_final_A, by = "name") %>% 
-  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")
-
-df_MW_A %>% 
-  mutate(method = as.factor(method), 
-         method = recode_factor(method, 
-                                "conv_diff" = "Conventional", 
-                                "rand_eob_diff" = "Randomized", 
-                                "rand_final_diff" = "Randomized\n(24 months)")) %>% 
-  ggplot(aes(x = method, y = diff, fill = method)) +
-  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
-  geom_lv(k = 4, outlier.shape = NA) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
-  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
-  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
-            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 6), 
-                     labels = number_format(suffix = " %")) +
-  scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
-  labs(fill = NULL, 
-       x = NULL, 
-       y = "Mean effect in fractional savings",
-       title = "Mean effect of non-routine events when using different M&V methods", 
-       subtitle = str_glue("All {A_building} buildings with no model normalization on weather")) +
-  coord_cartesian(ylim = c(-18, 18)) +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.position = "none",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-```
-
-As mentioned in Section \@ref(filter-and-clean-dataset), the majority of
-the selected buildings are affected by operational drift, which is the
-reason why we observe baseline electricity usage changes throughout the
-two year period. To evaluate the robustness of the two M&V methods to
-the effect of operational drift and non-routine events, we repeated the
-M&V run on the original dataset prior to the application of the reset
-control. In other words, there is no intervention applied here, so the
-most reliable M&V method will be the one that detects closer to 0 kW
-savings using the measurements. Therefore, the $Acc$ here are calculated
-as:
-
-$$
-Acc = 0-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
-$$
-
-Figure \@ref(fig:nosaving) shows the randomized M&V method (both at the
-time of satisfying all stopping criteria and at the end of a two-year
-period) produced consistent results. We included the absolute deviation
-($|ACC|$) figure in the supplementary material. Similarly to Figure
-\@ref(fig:abstract), stopping when meeting all stopping criteria
-generally yields slightly larger uncertainty than continuing for 24
-months. However, it still yields a more accurate result than the
-conventional method. Interestingly, similar to
-[@granderson_accuracy_2016] where they obtained a median normalized
-deviation around -1.25%, our analysis on conventional method shows that
-the typical deviation is -1.3%, meaning that the typical building is
-decreasing energy use slightly in this dataset in the absence of any
-known intervention. Thus, if a building analyst were to apply an
-intervention to this building and use conventional M&V to measure the
-effect, they would 'detect' 1.3% energy savings even when the
-intervention had no actual impact on energy performance.
-
-# Discussion
-
-## TOWT modeling accuracy
-
-To evaluate modeling accuracy, we used the Coefficient of Variation of
-Root-Mean Squared Error (CV(RMSE)) as the primary error metric. Since
-CV(RMSE) is a normalized measure, it enables direct comparison across
-different model-fitting results. According to ASHRAE Guideline 14,
-whole-building baseline models should achieve a CV(RMSE) below 25% for
-computing savings or 30% when less than 12 months of measurements are
-available [@ashrae_ashrae_2023]. In our study, Figure \@ref(fig:towtacc)
-compares the model fitting accuracy of the two M&V methods and it shows
-around 87.5% of the models fitted in the randomized M&V method are code
-compliant while 80% for the conventional M&V method. As the quantity of
-the training set is the same, this highlights the impact of data
-sampling on model fitting accuracy, and for TOWT, the distribution of
-outdoor weather conditions sampled in the training dataset is
-influential. The conventional method trains the model using the entire
-first-year dataset, if the location is heating or cooling dominant, the
-data-driven model is trained using colder or warmer outdoor conditions
-leading to biased model coefficients. Although the random sampling in
-our study does not specifically block the outdoor temperature factor,
-the CV(RMSE) distribution shown in the figure indicates a significant
-improvement in model fitting accuracy.
-
-```{r towtacc, fig.cap="TOWT model fitting accuracy distribution for all buildings included (with each data point representing one building)", fig.width=8, fig.height=4}
-
-cvrmse <- df_model_acc_stable %>% 
-  bind_rows(df_model_acc_variable) %>% 
-  pivot_longer(c(conv, rand), 
-               names_to = "method", 
-               values_to = "cvrmse") %>% 
-  mutate(method = as.factor(method), 
-         method = recode_factor(method, 
-                              "rand" = "Randomized\n(24 months)", 
-                              "conv" = "Conventional"))
-
-result <- t.test(filter(cvrmse, method == "Randomized\n(24 months)")$cvrmse, 
-                 filter(cvrmse, method == "Conventional")$cvrmse, 
-                 exact = TRUE,
-                 alternative = "less",
-                 mu = 0,
-                 conf.int = TRUE, 
-                 conf.level = 0.90,
-                 var.equal = TRUE)$p.value
-
-cvrmse %>% 
-  ggplot(aes(x = method, y = cvrmse, fill = method)) +
-  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
-  geom_lv(k = 4, outlier.shape = NA) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
-  annotate(geom = "text", 
-           x = 1.5, 
-           y = 40, 
-           size = 4, 
-           label = paste0("P value = ", round(result, digits = 3))) +
-  geom_errorbarh(aes(xmin = 1.2, xmax = 1.8, y = 38), height = 2, linewidth = 0.25) +
-  geom_text(data = . %>% 
-              group_by(method) %>% 
-              summarise(mean = mean(cvrmse, na.rm = T)) %>% 
-              ungroup(), 
-            aes(x = method, y = mean, label = paste0(round(mean, digits = 0), " %"))) +
-  geom_hline(yintercept = 25, color = "red", lty = "dashed") +
-  annotate(geom = "text", 
-           x = 0.5, 
-           y = 27, 
-           size = 4, 
-           color = "red", 
-           label = paste0(25, " %")) +
-  scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = c("grey80", "grey80")) + 
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 3), 
-                     labels = number_format(suffix = " %")) +
-  coord_cartesian(ylim = c(0, 50)) +
-  labs(x = NULL, 
-       y = NULL, 
-       color = NULL, 
-       title = "TOWT model fitting accuracy by different M&V methods", 
-       subtitle = "CV(RMSE)") +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.direction = "horizontal",
-        legend.position = "none",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-```
-
-## Sampling interval impact
-
-In some cases, particularly those involving large thermal mass and
-active thermal energy storage systems or operating continuously, there
-can be a substantial time-lagged effect associated with interventions.
-This means that the effect of one control strategy may persist after
-switching to another. To mitigate such carryover effects, building
-analysts may choose to drop non-consecutive days in the switchback
-experiment. For example, consider a heat pump hot water heater where the
-tank is charged at very different times of the day by the intervention
-strategy compared to the baseline strategy. When switching strategy, the
-tank may be much warmer or cooler than typical for that strategy, thus
-influencing subsequent measurements, particularly the performance
-measured on the first day (i.e., a non-consecutive day). In such cases,
-increasing the sampling interval (e.g., sample strategies every 2 or 3
-days instead of daily) and discarding measurements from those
-non-consecutive days can eliminate carryover. However, this approach
-reduces the total data collected: for instance, sampling every 3 days
-would drop 1/6 of measurements and sampling every 2 day means would
-remove 1/4 of measurements. Therefore, by increasing the sampling
-interval, there is an accuracy penalty due to a reduction in how often
-samples occur, and when decreasing the sampling interval, there is an
-accuracy penalty from dropping more non-consecutive days. To quantify
-such trade-off, \@ref(fig:absinterval) shows the accuracy of different
-sampling intervals and the consequence of dropping the non-consecutive
-days in the dataset at the time when randomized M&V produces a result
-(after satisfying all stopping criteria).
-
-```{r absinterval, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy (dropped: all non-consecutive days were dropped; kept: all measurements were kept)", fig.height=9, fig.width=12}
-
-#### ABS-INT ####
-df_seq_interval_fs_keep_stable %<>% 
-  select(-c(sprt, temp, final)) %>% 
-  bind_rows(df_seq_fs_stable %>% 
-              filter(seq == "eob") %>% 
-              select(-seq) %>% 
-              mutate(interval = 1) %>% 
-              rename(eob = "fs"))
-
-df_seq_interval_fs_keep_variable %<>% 
-  select(-c(sprt, temp, final)) %>% 
-  bind_rows(df_seq_fs_variable %>% 
-              filter(seq == "eob") %>% 
-              select(-seq) %>% 
-              mutate(interval = 1) %>% 
-              rename(eob = "fs"))
-
-df_drop <- df_seq_interval_fs_drop_stable %>% 
-  bind_rows(df_seq_interval_fs_drop_variable) %>% 
-  select(name, eob, interval) %>% 
-  mutate(group = "drop")
-
-df_keep <- df_seq_interval_fs_keep_stable %>% 
-  bind_rows(df_seq_interval_fs_keep_variable) %>% 
-  select(name, eob, interval) %>% 
-  mutate(group = "keep")
+```{r absint2y, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy over the course of two years (dropped: all non-consecutive days were dropped; kept: all measurements were kept)", fig.height=9, fig.width=12}
+# dropped version
+df_interval_keep_stable %<>% 
+  left_join(df_fs_stable %>% filter(method == "rand"), by = c("name", "site")) %>% 
+  select(-c("scenario", "method")) %>% 
+  rename("interval_fs_1" = "savings") %>% 
+  left_join(df_fs_tmy_stable, by = c("name", "site")) %>% 
+  select(-conv) %>% 
+  rename("interval_tmy_1" = "rand")
+
+df_interval_keep_variable %<>% 
+  left_join(df_fs_variable %>% filter(method == "rand"), by = c("name", "site")) %>% 
+  select(-c("scenario", "method")) %>% 
+  rename("interval_fs_1" = "savings") %>% 
+  left_join(df_fs_tmy_variable, by = c("name", "site")) %>% 
+  select(-conv) %>% 
+  rename("interval_tmy_1" = "rand")
+
+df_drop <- df_interval_drop_stable %>% 
+  bind_rows(df_interval_drop_variable) %>% 
+  select(name, contains("fs")) %>% 
+  rename("drop_1" = "interval_fs_1", 
+         "drop_2" = "interval_fs_2", 
+         "drop_3" = "interval_fs_3", 
+         "drop_7" = "interval_fs_7")
+
+df_keep <- df_interval_keep_stable %>% 
+  bind_rows(df_interval_keep_variable) %>% 
+  select(name, contains("fs")) %>% 
+  rename("keep_1" = "interval_fs_1",
+         "keep_2" = "interval_fs_2", 
+         "keep_3" = "interval_fs_3", 
+         "keep_7" = "interval_fs_7")
 
 df_MW_A <- df_drop %>% 
-  bind_rows(df_keep) %>% 
+  left_join(df_keep, by = c("name")) %>%
   left_join(rbind(df_fs_stable %>% 
@@
                     filter(scenario == "ref" & method == "true")), by = "name") %>% 
-  mutate(diff = abs(savings - eob)) %>% 
-  filter(is.finite(diff))
+  mutate(drop_1 = abs(savings - drop_1),
+         drop_2 = abs(savings - drop_2), 
+         drop_3 = abs(savings - drop_3), 
+         drop_7 = abs(savings - drop_7), 
+         keep_1 = abs(savings - keep_1),
+         keep_2 = abs(savings - keep_2), 
+         keep_3 = abs(savings - keep_3), 
+         keep_7 = abs(savings - keep_7)) %>% 
+  pivot_longer(cols = contains("drop") | contains("keep"),  
+               names_to = c("group", "interval"),
+               names_pattern = "(drop|keep)_(\\d+)",
+               values_to = "diff")
 
@@
   geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8) +
-  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
-  geom_text(data = . %>% 
-              group_by(group, interval) %>% 
-              summarise(mean = mean(diff, inf.rm = T)) %>% 
-              ungroup(), 
-            aes(x = interval, 
-                y = mean, group = group,
-                label = paste0(round(mean, digits = 1), " %")), 
-            color = "white",
-            position = position_dodge(width = 0.75)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8, linewidth = 0.15) +
+  geom_text(data = . %>% group_by(group, interval) %>% summarise(median = median(diff, na.rm = T)) %>% ungroup(), 
+            aes(x = interval, y = median, group = group,
+                label = paste0(round(median, digits = 1), " %")), 
+            position = position_dodge(width = 0.75), 
+            size = 4, 
+            color = "white") +
   scale_y_continuous(expand = c(0, 0), 
@@
        y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with no model normalization on weather")) +
-  coord_cartesian(ylim = c(0, 15)) +
+       subtitle = str_glue("All {A_building} buildings with measured weather")) +
+  coord_cartesian(ylim = c(0, 8)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
 
-# with tmy version
-df_seq_interval_nm_keep_stable %<>% 
-  select(-c(sprt, temp, final)) %>% 
-  bind_rows(df_sprt_all_stable %>% 
-              filter(seq == "eob") %>% 
-              select(-c(seq, n_weeks)) %>% 
-              mutate(interval = 1) %>% 
-              rename(eob = "annual"))
-
-df_seq_interval_nm_keep_variable %<>% 
-  select(-c(sprt, temp, final)) %>% 
-  bind_rows(df_sprt_all_variable %>% 
-              filter(seq == "eob") %>% 
-              select(-c(seq, n_weeks)) %>% 
-              mutate(interval = 1) %>% 
-              rename(eob = "annual"))
-
-df_drop <- df_seq_interval_nm_drop_stable %>% 
-  bind_rows(df_seq_interval_nm_drop_variable) %>% 
-  select(name, eob, interval) %>% 
-  mutate(group = "drop")
-
-df_keep <- df_seq_interval_nm_keep_stable %>% 
-  bind_rows(df_seq_interval_nm_keep_variable) %>% 
-  select(name, eob, interval) %>% 
-  mutate(group = "keep")
+# TMY version
+df_drop <- df_interval_drop_stable %>% 
+  bind_rows(df_interval_drop_variable) %>% 
+  select(name, contains("tmy")) %>% 
+  rename("drop_1" = "interval_tmy_1",
+         "drop_2" = "interval_tmy_2", 
+         "drop_3" = "interval_tmy_3", 
+         "drop_7" = "interval_tmy_7")
+
+df_keep <- df_interval_keep_stable %>% 
+  bind_rows(df_interval_keep_variable) %>% 
+  select(name, contains("tmy")) %>% 
+  rename("keep_1" = "interval_tmy_1",
+         "keep_2" = "interval_tmy_2", 
+         "keep_3" = "interval_tmy_3", 
+         "keep_7" = "interval_tmy_7")
 
 df_TW_A <- df_drop %>% 
-  bind_rows(df_keep) %>% 
+  left_join(df_keep, by = c("name")) %>%
   left_join(rbind(df_fs_stable %>% 
@@
   geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8) +
-  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
-  geom_text(data = . %>% 
-              group_by(group, interval) %>% 
-              summarise(mean = mean(diff, na.rm = T)) %>% 
-              ungroup(), 
-            aes(x = interval, 
-                y = mean, 
-                group = group,
-                label = paste0(round(mean, digits = 1), " %")),
-            color = "white",
-            position = position_dodge(width = 0.75)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8, linewidth = 0.15) +
+  geom_text(data = . %>% group_by(group, interval) %>% summarise(median = median(diff, na.rm = T)) %>% ungroup(), 
+            aes(x = interval, y = median, group = group,
+                label = paste0(round(median, digits = 1), " %")), 
+            position = position_dodge(width = 0.75), 
+            size = 4,
+            color = "white") +
   scale_y_continuous(expand = c(0, 0), 
@@
        y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with TOWT model normalized on TMY weather")) +
-  coord_cartesian(ylim = c(0, 15)) +
+       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
+  coord_cartesian(ylim = c(0, 8)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
           legend = "bottom") +
-  plot_annotation(title = "Absolute error in savings estimation of randomized M&V after satisfying all stopping criteria")
-
+  plot_annotation(title = "Absolute error in savings estimation of ranodmized M&V by different sampling intervals")
 ```
 
-To quantify the impact of reduced randomness, all measurements are kept
-when calculating the distribution shown in the darker-colored set. The
-results suggest a gradual increase in error from 2.4% to 3.7% when the
-sampling interval increases from daily to weekly intervals. This is due
-to the fact that increasing the sample interval means reducing
-randomization. To quantify the combined impact of data loss due to
-preventing the carryover effect, we generated the lighter-colored set,
-which drops all non-consecutive days. As a result, we noticed a
-significant increase in estimation error when sampling daily, which led
-to a similar accuracy compared to weekly intervals. Therefore,
-considering this trade-off, we recommend using a two-day sampling
-interval if the carryover effect is likely but expected to last less
-than one day, and then normalizing the estimation via energy modeling.
-We also included the mean deviation distribution and the accuracy
-distribution calculated after all stopping criteria were satisfied in
-the supplementary material and the results show the same indication.
-
-```{r intervaltime, fig.cap="Comparison of time required to satisfy all stopping criteria when considering dropping non-consecutive days (or not)", fig.height=6, fig.width=12}
-# timeline to finish sequential test
-df_drop <- bind_rows(df_seq_interval_tl_drop_stable, df_seq_interval_tl_drop_variable) %>%
-  mutate(group = "drop", 
-         temp = ifelse(base_temp >= interv_temp, base_temp, interv_temp)) %>% 
-  pivot_longer(c(sprt, temp, eob, final), names_to = "seq", values_to = "n_weeks") %>% 
-  select(name, seq, n_weeks, interval, group)
-
-interval_1_keep <- bind_rows(df_sprt_all_stable, df_sprt_all_variable) %>%
-  select(c(name, seq, n_weeks)) %>%
-  filter(seq != "final") %>%
-  mutate(seq = as.factor(seq)) %>%
-  mutate(interval = 1, 
-         group = "keep")
-
-df_keep <- bind_rows(df_seq_interval_tl_keep_stable, df_seq_interval_tl_keep_variable) %>%
-  mutate(group = "keep", 
-         temp = ifelse(base_temp >= interv_temp, base_temp, interv_temp)) %>% 
-  pivot_longer(c(sprt, temp, eob, final), names_to = "seq", values_to = "n_weeks") %>% 
-  select(name, seq, n_weeks, interval, group) %>% 
-  bind_rows(interval_1_keep)
-
-df_MW_A <- bind_rows(df_keep, df_drop) %>% 
-  filter(seq != "final") %>% 
+```{r meanintspt, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy after satisfying all stopping criteria (without absolute calculation)", fig.height=9, fig.width=12}
+# sprt sequence plot for different sampling intervals
+df_seq_interval_fs_keep_stable %<>% 
+  select(-c(sprt, temp, final)) %>% 
+  bind_rows(df_seq_fs_stable %>% 
               filter(seq == "eob") %>% 
-  mutate(seq = as.factor(seq), 
-         seq = recode_factor(seq, 
-                             "eob" = "Buildings finishing all criteria"), 
-         interval = as.factor(interval), 
+              select(-seq) %>% 
+              mutate(interval = 1) %>% 
+              rename(eob = "fs"))
+
+df_seq_interval_fs_keep_variable %<>% 
+  select(-c(sprt, temp, final)) %>% 
+  bind_rows(df_seq_fs_variable %>% 
+              filter(seq == "eob") %>% 
+              select(-seq) %>% 
+              mutate(interval = 1) %>% 
+              rename(eob = "fs"))
+
+df_drop <- df_seq_interval_fs_drop_stable %>% 
+  bind_rows(df_seq_interval_fs_drop_variable) %>% 
+  select(name, eob, interval) %>% 
+  mutate(group = "drop")
+
+df_keep <- df_seq_interval_fs_keep_stable %>% 
+  bind_rows(df_seq_interval_fs_keep_variable) %>% 
+  select(name, eob, interval) %>% 
+  mutate(group = "keep")
+
+df_MW_A <- df_drop %>% 
+  bind_rows(df_keep) %>% 
+  left_join(rbind(df_fs_stable %>% 
+                    filter(scenario == "ref" & method == "true"), 
+                  df_fs_variable %>% 
+                    filter(scenario == "ref" & method == "true")), by = "name") %>% 
+  mutate(diff = (savings - eob)) %>% 
+  filter(is.finite(diff))
+
+p_top <- df_MW_A %>% 
+  mutate(interval = as.factor(interval), 
          interval = recode_factor(interval, 
@@
          group = as.factor(group), 
-         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept"), 
-         n_weeks = as.factor(n_weeks), 
-         n_weeks = recode_factor(n_weeks, 
-                                 "12" = "12\nweeks", 
-                                 "24" = "24\nweeks", 
-                                 "36" = "36\nweeks", 
-                                 "48" = "48\nweeks")) %>% 
-  group_by(n_weeks, group, interval) %>% 
-  summarise(n = n()) %>%
-  ungroup()
-
-levA <- levels(df_MW_A$n_weeks)
-levB <- levels(df_MW_A$group)
-levC <- levels(df_MW_A$interval)
-
-# 2. Expand to all combinations of factorA and factorB
-all_combinations <- expand_grid(
-  n_weeks = levA,
-  group = levB, 
-  interval = levC, 
-)
-
-df_MW_A <- all_combinations %>%
-  left_join(df_MW_A, by = c("n_weeks", "group", "interval")) %>% 
-  mutate(n = ifelse(is.na(n), 0, n)) %>% 
-  group_by(group, interval) %>% 
-  mutate(n = ifelse(n > sum(n) * 0.01, n, 0), 
-         perc = n / sum(n) * 100) %>% 
-  ungroup()
-
-df_MW_A %>% 
-  ggplot(aes(x = n_weeks, y = perc, fill = group)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  facet_wrap(~interval, nrow = 1) +
-  geom_text(aes(x = n_weeks, y = perc + 2, label = paste0(round(perc, digits = 0), "%"), group = group), position = position_dodge(1)) +
+         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
+  ggplot(aes(x = interval, y = diff, fill = group)) +
+  geom_jitter(alpha = 0.8, 
+              size = 0.5, 
+              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
+  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8, linewidth = 0.15) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% group_by(group, interval) %>% summarise(median = median(diff, inf.rm = T)) %>% ungroup(), 
+            aes(x = interval, y = median, group = group,
+                label = paste0(round(median, digits = 1), " %")), 
+            position = position_dodge(width = 0.75), 
+            size = 4, 
+            color = "white") +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 5), 
+                     labels = number_format(suffix = " %")) +
   scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = ls_colors) +
-  scale_y_continuous(breaks = breaks_pretty(n = 4), 
+  scale_color_manual(values = c("grey80", "grey80")) + 
+  labs(fill = NULL, 
+       color = NULL, 
+       x = NULL, 
+       y = "Estimated error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with measured weather")) +
+  coord_cartesian(ylim = c(-12, 12)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "bottom",
+        axis.text.x = element_blank(),
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+# with tmy version
+df_seq_interval_nm_keep_stable %<>% 
+  select(-c(sprt, temp, final)) %>% 
+  bind_rows(df_sprt_all_stable %>% 
+              filter(seq == "eob") %>% 
+              select(-c(seq, n_weeks)) %>% 
+              mutate(interval = 1) %>% 
+              rename(eob = "annual"))
+
+df_seq_interval_nm_keep_variable %<>% 
+  select(-c(sprt, temp, final)) %>% 
+  bind_rows(df_sprt_all_variable %>% 
+              filter(seq == "eob") %>% 
+              select(-c(seq, n_weeks)) %>% 
+              mutate(interval = 1) %>% 
+              rename(eob = "annual"))
+
+df_drop <- df_seq_interval_nm_drop_stable %>% 
+  bind_rows(df_seq_interval_nm_drop_variable) %>% 
+  select(name, eob, interval) %>% 
+  mutate(group = "drop")
+
+df_keep <- df_seq_interval_nm_keep_stable %>% 
+  bind_rows(df_seq_interval_nm_keep_variable) %>% 
+  select(name, eob, interval) %>% 
+  mutate(group = "keep")
+
+df_TW_A <- df_drop %>% 
+  bind_rows(df_keep) %>% 
+  left_join(rbind(df_fs_stable %>% 
+                    filter(scenario == "ref" & method == "true"), 
+                  df_fs_variable %>% 
+                    filter(scenario == "ref" & method == "true")), by = "name") %>% 
+  mutate(diff = (savings - eob)) %>% 
+  filter(is.finite(diff))
+
+p_bottom <- df_TW_A %>% 
+  mutate(interval = as.factor(interval), 
+         interval = recode_factor(interval, 
+                                  "1" = "1-day\nsampling",
+                                  "2" = "2-day\nsampling", 
+                                  "3" = "3-day\nsampling", 
+                                  "7" = "7-day\nsampling"), 
+         group = as.factor(group), 
+         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
+  ggplot(aes(x = interval, y = diff, fill = group)) +
+  geom_jitter(alpha = 0.8, 
+              size = 0.5, 
+              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
+  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8, linewidth = 0.15) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% group_by(group, interval) %>% summarise(median = median(diff, na.rm = T)) %>% ungroup(), 
+            aes(x = interval, y = median, group = group,
+                label = paste0(round(median, digits = 1), " %")), 
+            position = position_dodge(width = 0.75), 
+            size = 4, 
+            color = "white") +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 5), 
                      labels = number_format(suffix = " %")) +
-  labs(x = NULL,
-       y = NULL,
-       fill = NULL,
-       title = "Percentage of buildings using randomized M&V when satisfying all stopping critiera", 
-       subtitle = "Counted at the end of each 12-week blocking period") +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80")) + 
+  labs(fill = NULL, 
+       color = NULL, 
+       x = NULL, 
+       y = "Esimated error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
+  coord_cartesian(ylim = c(-12, 12)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
         plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+ggarrange(p_top, p_bottom, 
+          ncol = 1, nrow = 2, 
+          align = "hv", 
+          labels = c("a)", "b)"),
+          common.legend = T, 
+          legend = "bottom") +
+  plot_annotation(title = "Estimated error in savings estimation of randomized M&V after satisfying all stopping criteria")
+
 ```
 
-Sampling at different intervals and excluding non-consecutive days also
-impact other stopping criteria, such as the time needed to satisfy
-statistical uncertainty and weather conditions thresholds. In Figure
-\@ref(fig:intervaltime), we compared the percentage of buildings
-satisfying all stopping criteria under different sampling intervals,
-with or without dropping non-consecutive days. The results indicate that
-randomized M&V concludes more quickly when all data are retained,
-particularly at shorter intervals. For instance, when sampling at a
-one-day interval, 90% of buildings meet all criteria within 24 weeks if
-no data are dropped, but this number decreases significantly to 52% when
-non-consecutive days (account for 1/3 of the data) are excluded. As the
-sampling interval increases, this difference gradually diminishes: when
-sampling at a weekly interval, removing non-consecutive days has no
-impact on the timeline of meeting stopping criteria.
-
-## Sampling ratio impact
-
-Another advantage of using randomized M&V is the flexibility of changing
-the sampling ratio after the target savings are detected. For instance,
-the building owner can continue sampling at a 50%/50% ratio between the
-baseline and the intervention to further reduce the uncertainty
-associated with the savings. Alternatively, they could switch to 100%
-intervention to maximize cost savings, though this approach risks the
-baseline becoming outdated over time, which is another trade-off to
-consider. A middle-ground approach may be to continue sampling at a
-20%/80% ratio or 10%/90% between the baseline and the intervention. To
-compare those possible choices to a building owner, figure
-\@ref(fig:abscont) shows the results of changing the original 50%/50%
-sampling ratio after satisfying all stopping criteria and continuing a
-new ratio till the end of the two-year period. Those results were
-calculated without dropping non-consecutive days and include the initial
-50%/50% sampled data. On the far left, we also provided the case when
-the building owner continues with no change in sampling ratio. The
-results show that as the sampled baseline days and intervention days
-become unbalanced, the savings estimation error grows from 1% to 5%.
-This is reasonable since sampling at 10%/90% is similar to the
-conventional M&V method. And on the far right, we show again the
-accuracy of the conventional M&V method, which is still worse then
-sampling at 10%/90%. Also, We include the mean deviation version in the
-supplementary material.
-
-```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months", fig.height=6, fig.width=12}
-
-#### ABS-CONT ####
-conv_S <- df_fs_stable %>% 
-  filter(method != "rand") %>% 
-  pivot_wider(names_from = method, values_from = savings) %>% 
-  mutate(diff = abs(true - conv), 
-         ratio = "conv") %>% 
-  select(name, ratio, diff)
-
-rand_cont_S <- df_cont_stable %>% 
-  select(-cont_tmy) %>% 
-  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = abs(savings - cont_fs)) %>% 
-  select(name, diff, ratio)
-
-rand_final_S <- df_fs_stable %>% 
-  filter(method != "conv") %>% 
-  pivot_wider(names_from = method, values_from = savings) %>% 
-  mutate(diff = abs(true - rand), 
-         ratio = 0) %>% 
-  select(name, diff, ratio)
-
-df_MW_S <- rbind(rand_cont_S, rand_final_S, conv_S)
-
-# plot for the variable subset
-conv_V <- df_fs_variable %>% 
-  filter(method != "rand") %>% 
-  pivot_wider(names_from = method, values_from = savings) %>% 
-  mutate(diff = abs(true - conv), 
-         ratio = "conv") %>% 
-  select(name, ratio, diff)
-
-rand_cont_V <- df_cont_variable %>% 
-  select(-cont_tmy) %>% 
-  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = abs(savings - cont_fs)) %>% 
-  select(name, diff, ratio)
-
-rand_final_V <- df_fs_variable %>% 
-  filter(method != "conv") %>% 
-  pivot_wider(names_from = method, values_from = savings) %>% 
-  mutate(diff = abs(true - rand), 
-         ratio = 0) %>% 
-  select(name, diff, ratio)
-
-df_MW_V <- rbind(rand_cont_V, rand_final_V, conv_V)
-
-df_MW_A <- rbind(df_MW_S, df_MW_V)
-
-df_MW_A %>% 
-  mutate(ratio = as.factor(ratio), 
-         ratio = recode_factor(ratio, "0" = "Randomized\n(50/50)", "1" = "Randomized\n(20/80)", "2" = "Randomized\n(10/90)", "conv" = "Conventional")) %>% 
-  ggplot(aes(x = ratio, y = diff, fill = ratio)) +
-  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
-  geom_lv(k = 4, outlier.shape = NA) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio)) +
+```{r meanint2y, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy over the course of the two years (without absolute calculation)", fig.width=12, fig.height=9}
+# dropped version
+df_drop <- df_interval_drop_stable %>% 
+  bind_rows(df_interval_drop_variable) %>% 
+  select(name, contains("fs")) %>% 
+  rename("drop_1" = "interval_fs_1", 
+         "drop_2" = "interval_fs_2", 
+         "drop_3" = "interval_fs_3", 
+         "drop_7" = "interval_fs_7")
+
+df_keep <- df_interval_keep_stable %>% 
+  bind_rows(df_interval_keep_variable) %>% 
+  select(name, contains("fs")) %>% 
+  rename("keep_1" = "interval_fs_1",
+         "keep_2" = "interval_fs_2", 
+         "keep_3" = "interval_fs_3", 
+         "keep_7" = "interval_fs_7")
+
+df_MW_A <- df_drop %>% 
+  left_join(df_keep, by = c("name")) %>%
+  left_join(rbind(df_fs_stable %>% 
+                    filter(scenario == "ref" & method == "true"), 
+                  df_fs_variable %>% 
+                    filter(scenario == "ref" & method == "true")), by = "name") %>% 
+  mutate(drop_1 = savings - drop_1,
+         drop_2 = savings - drop_2, 
+         drop_3 = savings - drop_3, 
+         drop_7 = savings - drop_7, 
+         keep_1 = savings - keep_1,
+         keep_2 = savings - keep_2, 
+         keep_3 = savings - keep_3, 
+         keep_7 = savings - keep_7) %>% 
+  pivot_longer(cols = contains("drop") | contains("keep"),  
+               names_to = c("group", "interval"),
+               names_pattern = "(drop|keep)_(\\d+)",
+               values_to = "diff")
+
+p_top <- df_MW_A %>% 
+  mutate(interval = as.factor(interval), 
+         interval = recode_factor(interval, 
+                                  "1" = "1-day\nsampling",
+                                  "2" = "2-day\nsampling", 
+                                  "3" = "3-day\nsampling", 
+                                  "7" = "7-day\nsampling"), 
+         group = as.factor(group), 
+         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
+  ggplot(aes(x = interval, y = diff, fill = group)) +
+  geom_jitter(alpha = 0.8, 
+              size = 0.5, 
+              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
+  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8, linewidth = 0.15) +
   geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
-  geom_text(data = . %>% 
-              group_by(ratio) %>% 
-              summarise(mean = mean(diff)) %>% 
-              ungroup(), 
-            aes(x = ratio, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
+  geom_text(data = . %>% group_by(group, interval) %>% summarise(median = median(diff, na.rm = T)) %>% ungroup(), 
+            aes(x = interval, y = median, group = group,
+                label = paste0(round(median, digits = 1), " %")), 
+            position = position_dodge(width = 0.75), 
+            size = 4, 
+            color = "white") +
   scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 4), 
+                     breaks = breaks_pretty(n = 5), 
                      labels = number_format(suffix = " %")) +
   scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80")) + 
+  scale_color_manual(values = c("grey80", "grey80")) + 
   labs(fill = NULL, 
+       color = NULL, 
        x = NULL, 
-       y = "Absolute error in fractional savings", 
-       title = "Absolute error in savings estimation fo randomized M&V using different sampling ratio",
-       subtitle = str_glue("All {A_building} buildings with with no model normalization on weather")) +
-  coord_cartesian(ylim = c(0, 20)) +
+       y = "Estimated error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with measured weather")) +
+  coord_cartesian(ylim = c(-10, 10)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.position = "none",
+        legend.position = "bottom",
+        axis.text.x = element_blank(),
         plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+# TMY version
+df_drop <- df_interval_drop_stable %>% 
+  bind_rows(df_interval_drop_variable) %>% 
+  select(name, contains("tmy")) %>% 
+  rename("drop_1" = "interval_tmy_1",
+         "drop_2" = "interval_tmy_2", 
+         "drop_3" = "interval_tmy_3", 
+         "drop_7" = "interval_tmy_7")
+
+df_keep <- df_interval_keep_stable %>% 
+  bind_rows(df_interval_keep_variable) %>% 
+  select(name, contains("tmy")) %>% 
+  rename("keep_1" = "interval_tmy_1",
+         "keep_2" = "interval_tmy_2", 
+         "keep_3" = "interval_tmy_3", 
+         "keep_7" = "interval_tmy_7")
+
+df_TW_A <- df_drop %>% 
+  left_join(df_keep, by = c("name")) %>%
+  left_join(rbind(df_fs_stable %>% 
+                    filter(scenario == "ref" & method == "true"), 
+                  df_fs_variable %>% 
+                    filter(scenario == "ref" & method == "true")), by = "name") %>% 
+  mutate(drop_1 = savings - drop_1,
+         drop_2 = savings - drop_2, 
+         drop_3 = savings - drop_3, 
+         drop_7 = savings - drop_7, 
+         keep_1 = savings - keep_1,
+         keep_2 = savings - keep_2, 
+         keep_3 = savings - keep_3, 
+         keep_7 = savings - keep_7) %>% 
+  pivot_longer(cols = contains("drop") | contains("keep"),  
+               names_to = c("group", "interval"),
+               names_pattern = "(drop|keep)_(\\d+)",
+               values_to = "diff")
+
+p_bottom <- df_TW_A %>% 
+  mutate(interval = as.factor(interval), 
+         interval = recode_factor(interval, 
+                                  "1" = "1-day\nsampling",
+                                  "2" = "2-day\nsampling", 
+                                  "3" = "3-day\nsampling", 
+                                  "7" = "7-day\nsampling"), 
+         group = as.factor(group), 
+         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
+  ggplot(aes(x = interval, y = diff, fill = group)) +
+  geom_jitter(alpha = 0.8, 
+              size = 0.5, 
+              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
+  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8, linewidth = 0.15) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% group_by(group, interval) %>% summarise(median = median(diff, na.rm = T)) %>% ungroup(), 
+            aes(x = interval, y = median, group = group,
+                label = paste0(round(median, digits = 1), " %")), 
+            position = position_dodge(width = 0.75), 
+            size = 4, 
+            color = "white") +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 5), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80")) + 
+  labs(fill = NULL, 
+       color = NULL, 
+       x = NULL, 
+       y = "Estimated error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
+  coord_cartesian(ylim = c(-10, 10)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "bottom",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+ggarrange(p_top, p_bottom, 
+          ncol = 1, nrow = 2, 
+          align = "hv", 
+          labels = c("a)", "b)"),
+          common.legend = T, 
+          legend = "bottom") +
+  plot_annotation(title = "Estimated error in savings estimation of randomized M&V by different sampling intervals")
 ```
-
-## Blocking period impact
-
-Selecting different blocking periods affects how frequently the a M&V
-analyst evaluats stopping criteria, thus influencing the duration of the
-measurement and verification (M&V) process. Here, we summarize the
-advantages and disadvantages of shorter versus longer blocking periods.
-As discussed in our previous study, the second step of this randomized
-M&V method involves choosing an appropriate blocking period to minimize
-potential biases caused by incomplete randomization. For most practical
-scenarios, we recommend a blocking period of approximately 12 to 16
-weeks, as it typically covers the duration of a season, ensuring a
-representative sampling of operational conditions. In general, longer
-blocking periods are advantageous when evaluating multiple intervention
-strategies or using sampling intervals greater than one day. In such
-cases, extended blocking ensures a balanced distribution of both
-day-of-week and outdoor weather conditions.
-
-Opting for a shorter blocking period, such as 6 weeks, has the potential
-benefit of allowing the stopping criteria to be met more rapidly. We
-plotted the percentage of buildings finishing all stopping criteria when
-the blocking period is 6 weeks in Figure \@ref(fig:6weekblock). As a
-result, when compared to Figure \@ref(fig:intervaltime), the only
-difference we found is when sampling at daily interval and not dropping
-non-consecutive days, 22% of buildings can get an estimate earlier. This
-indicates most buildings require 9 months to meet outdoor weather
-conditions starting from January.
-
-```{r 6weekblock, fig.cap = "Time required to satisfy all stopping criteria when using a 6-week blocking period"}
-
-knitr::include_graphics(paste0(fig_path, "6week_time.png"))
-```
-
-Sampling at a shorter blocking period also means that the analyst must
-perform the hypothesis tests more frequently, which can necessitate
-adjusting the significance level using methods such as the Bonferroni
-correction to control for increased type-I error risk. On the plus side,
-a shorter blocking period can reduce the adverse impact of missing or
-problematic data. For example, if an intervention requires fine-tuning
-after initial implementation, removing data from the first blocking
-period would be less detrimental compared to longer blocking periods,
-preserving the overall integrity of the experiment.
-
-## M&V starting time impact
-
-In this study, we consistently set the baseline period to begin on
-January 1st, which offers the advantage of capturing colder outdoor
-weather conditions early in the M&V process. However, when the baseline
-measurements misses the initial winter period, meeting the required
-weather condition criteria can potentially take a longer time. To assess
-this impact, we repeated the randomized M&V analysis for all buildings
-using a new start date in March. As shown in Figure  \@ref(fig:startT),
-the new M&V process results in a noticeable delay in satisfying the
-stopping criteria. Compared to Figure  \@ref(fig:intervaltime), where a
-January start allowed most buildings to finish within 36 weeks, a March
-start extended the required duration to 48 weeks for the majority of
-buildings. Furthermore, when a longer sampling interval is used or
-non-consecutive days are excluded (to avoid carryover effects), some
-buildings may require up to 60 weeks to complete the randomized M&V. As
-expected, to collect sufficient data representing winter conditions, the
-sampling may need to continue until the next January. However, as the
-overall data collection period extends, the influence of dropping
-non-consecutive days diminishes and becomes less of a concern.
-
-```{r startT, fig.cap = "Time required to satisfy all stopping criteria starting from March"}
-
-knitr::include_graphics(paste0(fig_path, "startT.png"))
-```
-
-Similarly, we evaluated the time required to satisfy all stopping
-criteria when using a shorter blocking period, as shown in Figure
- \@ref(fig:6weekstartT). Compared to Figure  \@ref(fig:startT), a
-significant proportion of buildings completed the randomized M&V by week
-42 (mid-December), which is one blocking period earlier than having a
-12-week blocking period. This result further highlights that the overall
-duration of the randomized M&V process is strongly driven by outdoor
-weather. Accordingly, selecting an optimal blocking period should
-consider historical or typical meteorological weather file to
-effectively minimize the time required.
 
@@
 ```
-
-## Limitations
-
-Aside from the general limitation that this method only applies to the
-subset of interventions in buildings that are easily switchable, we
-identify two key limitations in this study:
-
-1.  The simulated intervention remains somewhat generic and simplistic
-    considering the diverse building types and climate zones in the BDG2
-    dataset. As the primary focus is to accurately detect the
-    intervention (or any intervention) effect so for simplicity, we
-    applied the parameters listed in Table 1 uniformly across all
-    buildings. Yet, in some cases, raising the water temperature by 1°C
-    might yield more or less than the specified 8% electricity savings
-    due to target buildings' demand flexibility. For example, one field
-    validation of HVAC optimization shows additional savings might be
-    limited if other best practice controls are already implemented
-    [@granderson2018].
-2.  For the main analysis, we assumed no carryover effect and a daily
-    sampling interval would suffice for most commercial buildings, but
-    certainly exceptions exist. However, in the discussion, we addressed
-    such concern by comparing different sampling strategies and the
-    impact of dropping measurements from non-consecutive days.
-
-# Conclusion
-
-This research demonstrated the application of a randomized
-whole-building measurement and verification (M&V) method, comparing its
-performance to the conventional approach described in the IPMVP and in
-ASHRAE Guideline 14, using a large, open-source commercial building
-dataset. The porposed M&V method leverages the randomized experimental
-design concept from other scientific fields, along with statistical
-sequential inference techniques, to determine when target savings are
-detected. We used a virtual control retrofit case---resetting the
-chilled water setpoint based on outdoor weather conditions---and applied
-it to over 600 filtered commercial buildings. By comparing the savings
-estimations of the conventional method with the novel randomized method,
-we found that the randomized approach provides faster and more robust
-savings estimations.
-
-Specifically, we showed that throughout 11 different locations assessed
-in this study, the randomized M&V can provide a saving estimation by 36
-weeks (with the majority finished by 24 weeks) once all stopping
-criteria are satisfied. In contrast, the conventional method requires a
-full range of baseline and intervention measurements under normal
-operating conditions, typically taking 6-9 months for each phase. Most
-importantly, we verified that with a reduced M&V timeline, the
-randomized method can estimate savings more accurately by showing that
-the absolute error is only 1 - 2 % for a typical building. Whereas for
-the much longer two-year conventional method, the estimation error was
-approximately 6% in a typical building. We also evaluated the impact of
-non-routine events especially more gradual and subtle operational drift
-on the proposed M&V method and the results show that baseline changes in
-the post-retrofit period can deviate savings estimated using the
-conventional method. As a comparison, we found those events have a very
-negligible impact on the savings estimated using the randomized method,
-demonstrating robustness to this issue. Lastly, we consider scenarios
-where there is a known carryover effect from switching between
-strategies in the building and assess the impact of dropping samples
-where the strategy is non-consecutive and varying the sampling interval.
-Through that process, we showed that for the typical building, if a
-carryover is likely to be present, then using a 3-day sampling interval
-and dropping non-consecutive samples yields the optimal design.
-
-We also discussed the impact of model fitting accuracy when using the
-two methods and we found a significant improvement in our method due to
-random sampling. We also provide a detailed assessment of using
-different sampling ratios, those results provide useful considerations
-to building analysts when generating randomized schedules for a variety
-of use cases.
-
-# CRediT authorship contribution statement
-
-**Aoyu Zou**: Conceptualization, Data curation, Formal analysis,
-Methodology, Investigation, Software, Writing - original draft,
-Writing - review & editing. **Paul Raftery**: Conceptualization, Funding
-acquisition, Formal analysis, Methodology, Investigation, Project
-administration, Supervision, Writing - review & editing. **Stefano
-Schiavon**: Conceptualization, Methodology, Investigation, Supervision,
-Writing - review & editing. **Carlos Duarte**: Methodology,
-Investigation, Supervision, Writing - review & editing. **Gail Brager**:
-Supervision, Writing - review & editing.
-
-# Reproducibility
-
-A reproducible example with analysis code is available (MIT license) at
-<https://github.com/CenterForTheBuiltEnvironment/genome_mnv>.
-
-# Declaration of competing interest
-
-The authors declare that they have no known competing financial
-interests or personal relationships that could have appeared to
-influence the work reported in this paper.
-
-# Acknowledgements
-
-# References

### 2025-07-06 code/manuscript.Rmd
temp commit for figure update

@@
 author:
-  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
+  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^ ^[Department of Architecture, College of Environmental Design, University of California, Berkeley, USA] ^[Department of Civil and Environmental Engineering, College of Engineering, University of California, Berkeley, USA], Carlos Duarte ^1^, Gail Brager ^1^
 abstract: Conventional measurement and verification (M&V) methods rely on comparing pre- and post-retrofit performance to estimate energy savings.
-They are often time-consuming and unreliable, especially when non-routine events (step changes or more gradual changes in how a building operates) occur during the M&V process.
+They are often time-consuming and unreliable, especially when non-routine events, as step changes or more gradual changes in how a building operates, occur during the M&V process.
 Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies the conventional M&V method, the results can be confounded.
-In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method that randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily).
+In this study, we demonstrated that switchable interventions, such as most HVAC control retrofits, can benefit from a new M&V method that randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily).
 We tested this novel randomized M&V method on a large public dataset (639 buildings) covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention.
-The results show that compared to the conventional method, the randomized method provides more accurate savings estimations with a median of 74% accuracy improvement and is even faster (typically 36 weeks instead of 2 years).
+The results show that compared to the conventional method, the randomized method provides more accurate savings estimations with a median of 74% accuracy improvement and is even faster (typically 36 weeks instead of 104 weeks, ~65% reduction).
 Additionally, we found that when non-routine events are present, the randomized method estimates savings that are much closer to the ground-truth savings than the conventional method, demonstrating much improved reliability.
@@
 # define base ggplot theme
-theme_update(plot.title = element_text(size = 14, colour = "grey20", face = "bold", hjust = 0.5),
-             plot.subtitle = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
+theme_update(plot.title = element_text(size = 16, colour = "grey20", face = "bold", hjust = 0.5),
+             plot.subtitle = element_text(size = 14, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
              plot.caption = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5),
@@
              panel.grid.major = element_blank(),
-             axis.text = element_text(size = 10),
-             strip.text = element_text(size = 10, color = "grey20", face = "bold"),
+             axis.text = element_text(size = 12),
+             legend.text = element_text(size = 12),
+             strip.text = element_text(size = 12, color = "grey20", face = "bold"),
              strip.background = element_blank())
@@
 
-```{r graphicalabstract, out.width = "80%"}
-
-knitr::include_graphics("../figs/manuscript/graphical_abs.jpg", dpi = 900)
+```{r graphicalabstract, out.width = "100%"}
+
+knitr::include_graphics(paste0(fig_path, "graphical_abs.jpg"), dpi = 900)
 ```
@@
 
--   Verified the improvements of the randomized switchback method using
-    a large open-source dataset
+-   Randomized switchback method leads to time reduction and increased
+    accuracy
 -   Addressed the influence of non-routine events and also more subtle
@@
 standard from medical, agricultural and online controlled experimental
-(i.e. A/B testing) studies [@blackston2019; @duan2013; @gupta2019;
-@mie2022; @munro; @raseduzzaman2017]. An early application of this
-randomized design to the field of M&V in buildings was first used in
-[@raftery2018], which compared the energy performance of two different
-supply air temperature reset strategies in a building. Subsequent
-improvements to the M&V method include adding a sequential evaluation
-framework and defining stopping criteria to end the M&V period early
-after reaching a target level of uncertainty, using blocked
+(i.e. A/B testing) studies [@blackston_comparison_2019; @duan2013;
+@gupta2019; @mie2022; @munro; @raseduzzaman2017]. An early application
+of this randomized design to the field of M&V in buildings was first
+used in [@raftery2018], which compared the energy performance of two
+different supply air temperature reset strategies in a building.
+Subsequent improvements to the M&V method include adding a sequential
+evaluation framework and defining stopping criteria to end the M&V
+period early after reaching a target level of uncertainty, using blocked
 randomization by weekday. The full framework is detailed in a previously
@@
 
-```{r stablesite, fig.cap = "Site summary of the filtered buildings from BDG2 dataset (counts < 10 are omitted for visualization; left: aggregated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}
-
-set3 <- colorRampPalette(brewer.pal('Set3',n=12))
-type_colors <- setNames(set3(13), all_types_variable$type)
-
-stable_set <- df_energy_stable %>%
-  group_by(type) %>%
-  distinct(name) %>%
-  summarise(stable = n()) %>%
-  ungroup()
-
-variable_set <- df_energy_variable %>%
-  group_by(type) %>%
-  distinct(name) %>%
-  summarise(variable = n()) %>%
-  ungroup()
-
-p1 <- variable_set %>% 
-  left_join(stable_set, by = "type") %>% 
-  mutate(stable = replace_na(stable, 0)) %>% 
-  group_by(type) %>% 
-  summarise(n = stable + variable) %>% 
-  ungroup() %>% 
-  mutate(type = fct_reorder(type, n, .desc = F)) %>%
-  ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
-  geom_col() +
-  geom_text(aes(label = ifelse(n > 10, n, "")), color = "black", position = position_stack(vjust = 0.5)) +
-  scale_y_continuous(expand = c(0, 0),
-                     breaks = breaks_pretty(n = 4)) +
-  scale_x_discrete(expand = c(0, 0.1)) +
-  scale_fill_manual(values = type_colors) +
-  labs(x = NULL,
-       y = NULL,
-       subtitle = "Combining all locations",
-       fill = NULL) +
-  theme(axis.text.x = element_blank(),
-        panel.grid.major.y = element_line(color = "grey80"),
-        legend.direction = "horizontal",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-stable_set <- df_energy_stable %>% 
-  left_join(df_loc, by = "site") %>% 
-  group_by(type, location) %>%
-  distinct(name) %>%
-  summarise(stable = n()) %>%
-  ungroup()
-
-variable_set <- df_energy_variable %>% 
-   left_join(df_loc, by = "site") %>% 
-  group_by(type, location) %>%
-  distinct(name) %>%
-  summarise(variable = n()) %>%
-  ungroup()
-
-p2 <- variable_set %>% 
-  left_join(stable_set, by = c("type", "location")) %>% 
-  mutate(stable = replace_na(stable, 0)) %>% 
-  group_by(location, type) %>% 
-  summarise(n = stable + variable) %>% 
-  ungroup() %>% 
-  group_by(location) %>%
-  mutate(proportion = n / sum(n),
-         total = sum(n),
-         ymax = cumsum(proportion),
-         ymin = c(0, head(ymax, n = -1))) %>%
-  mutate(label_pos = (ymax + ymin) / 2) %>%
-  ungroup() %>%
-  group_by(type) %>%
-  mutate(order = sum(n)) %>%
-  ungroup() %>%
-  mutate(type = fct_reorder(type, order, .desc = F)) %>%
-  ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3.2, fill = type)) +
-  geom_rect() +
-  coord_polar(theta = "y") +
-  xlim(c(2, 4)) +
-  facet_wrap(~ location, nrow = 3) +
-  labs(x = NULL,
-       y = NULL,
-       subtitle = "For each location",
-       fill = NULL) +
-  scale_fill_manual(values = type_colors) +
-  geom_text(aes(x = 3.7, y = label_pos, label = ifelse(n > 10, n, ""))) +
-  geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black", check_overlap = T) +
-  theme(legend.direction = "horizontal",
-        axis.text = element_blank(),
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-ggarrange(p1, p2,
-          ncol = 2, nrow = 1,
-          widths = c(0.2, 1),
-          common.legend = TRUE,
-          legend = "bottom") +
-  plot_annotation(title = "Summary of filtered buildings and types")
-```
-
 For all qualified buildings, we also create a feature to identify
@@
   geom_vline(xintercept = 10, lty = "dashed", color = "red") +
-  annotate(geom = "text", x = 7.5, y = 8, label = "No chiller operation\nbelow 10 °C") +
+  annotate(geom = "text", x = 7.5, y = 8, label = "No chiller\noperation\nbelow 10 °C", size = 5) +
   coord_cartesian(xlim = c(5, 30)) +
@@
 | Parameter              | Description                                                      | Value |
-|-------------------|-----------------------------------|-------------------|
+|----------------|-----------------------------------------|----------------|
 | $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
@@
 
-```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period).", fig.height=11, fig.width=12}
+```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period).", fig.height=10, fig.width=9}
 
@@
   ggplot(aes(x = method, y = diff, fill = method)) +
-  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
   geom_lv(k = 4, outlier.shape = NA) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method), linewidth = 0.25) +
   geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
             aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
-            size = 4) +
+            size = 5) +
   scale_y_continuous(expand = c(0, 0), 
@@
            x = 66, 
-           y = max(count$eob) - 30, 
+           y = max(count$eob) - 35, 
            size = 5,
@@
            y = 200, 
+           size = 4, 
            label = paste0("12-week\nblock"), 
@@
           legend = "bottom") +
-  plot_annotation(title = "Overall accuracy and time required for savings estimation by different M&V methods")
+  plot_annotation(title = "Overall savings estimation comparison by different M&V methods")
 
@@
 
-```{r meanerr, fig.cap="Distribution of estimated deviation from the ground-truth savings calculated by the two M&V methods (red dashed line highlighting 0% error)", fig.height=8, fig.width=12}
+```{r meanerr, fig.cap="Distribution of estimated deviation from the ground-truth savings calculated by the two M&V methods (red dashed line highlighting 0% error)", fig.height=7, fig.width=9}
 # plot for the stable subset
@@
   ggplot(aes(x = method, y = diff, fill = method)) +
-  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
   geom_lv(k = 4, outlier.shape = NA) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method), linewidth = 0.25) +
   geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
@@
             aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
-            size = 4) +
+            size = 5) +
   scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 4), 
+                     breaks = breaks_pretty(n = 5), 
                      labels = number_format(suffix = " %")) +
@@
        x = NULL, 
-       y = "Error in fractional savings", 
+       y = "Estimated error in fractional savings", 
        subtitle = str_glue("All {A_building} buildings with measured weather")) +
-  coord_cartesian(ylim = c(-20, 20)) +
+  coord_cartesian(ylim = c(-15, 15)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
        subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
-  coord_cartesian(ylim = c(-20, 20)) +
+  coord_cartesian(ylim = c(-15, 15)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
 |                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
-|-------------------|----------------------------------|-------------------|
+|------------------|-------------------------------------|------------------|
 | 25th percentile | -5.85                                               | `r percentiles$Q1`     |
@@
 
-```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect when buildings are subject to static non-rountine events and gradual operational drift.", fig.height=5, fig.width=12}
+```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect when buildings are subject to static non-rountine events and gradual operational drift.", fig.height=4, fig.width=9}
 
@@
   ggplot(aes(x = method, y = diff, fill = method)) +
-  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
   geom_lv(k = 4, outlier.shape = NA) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method), linewidth = 0.25) +
   geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
@@
             aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
-            size = 4) +
+            size = 5) +
   scale_y_continuous(expand = c(0, 0), 
@@
        y = "Estimated effect in fractional savings",
-       title = "Estimated effect of non-routine events when using different M&V methods", 
+       title = "Estimated effect of non-routine events by different M&V methods", 
        subtitle = str_glue("All {A_building} buildings with measured weather")) +
-  coord_cartesian(ylim = c(-18, 18)) +
+  coord_cartesian(ylim = c(-16, 16)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
 
-```{r towtacc, fig.cap="TOWT modeling accuracy distribution for all buildings included (with each data point representing one building)", fig.width=8, fig.height=4}
+```{r towtacc, fig.cap="TOWT modeling accuracy distribution for all buildings included (with each data point representing one building)", fig.width=9, fig.height=4}
 
@@
   ggplot(aes(x = method, y = cvrmse, fill = method)) +
-  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
   geom_lv(k = 4, outlier.shape = NA) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method), linewidth = 0.25) +
   annotate(geom = "text", 
@@
            size = 4, 
-           label = paste0("P value = ", round(result, digits = 3))) +
+           label = ifelse(result < 0.001, "P value < 0.001", paste0("P value = ", result))) +
   geom_errorbarh(aes(xmin = 1.2, xmax = 1.8, y = 38), height = 2, linewidth = 0.25) +
@@
               group_by(method) %>% 
-              summarise(mean = mean(cvrmse, na.rm = T)) %>% 
+              summarise(median = median(cvrmse, na.rm = T)) %>% 
               ungroup(), 
-            aes(x = method, y = mean, label = paste0(round(mean, digits = 0), " %"))) +
+            aes(x = method, y = median, label = paste0(round(median, digits = 0), " %")), 
+            size = 5) +
   scale_fill_manual(values = ls_colors) +
@@
                      labels = number_format(suffix = " %")) +
-  coord_cartesian(ylim = c(0, 50)) +
+  coord_cartesian(ylim = c(0, 42)) +
   labs(x = NULL, 
-       y = NULL, 
+       y = "CV(RMSE)", 
        color = NULL, 
-       title = "TOWT model fitting accuracy by different M&V methods", 
-       subtitle = "CV(RMSE)") +
+       title = "TOWT model fitting accuracy by different M&V methods") +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
 
-```{r absinterval, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy (dropped: all non-consecutive days were dropped; kept: all measurements were kept)", fig.height=9, fig.width=12}
+```{r absinterval, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy (dropped: all non-consecutive days were dropped; kept: all measurements were kept)", fig.height=8, fig.width=9}
 
@@
   ggplot(aes(x = interval, y = diff, fill = group)) +
-  geom_jitter(alpha = 0.8, 
-              size = 0.5, 
-              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
   geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
@@
             color = "white", 
-            size = 4) +
+            size = 5) +
   scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 4), 
+                     breaks = breaks_pretty(n = 5), 
                      labels = number_format(suffix = " %")) +
@@
        subtitle = str_glue("All {A_building} buildings with measured weather")) +
-  coord_cartesian(ylim = c(0, 15)) +
+  coord_cartesian(ylim = c(0, 12)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
        subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
-  coord_cartesian(ylim = c(0, 15)) +
+  coord_cartesian(ylim = c(0, 12)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
           legend = "bottom") +
-  plot_annotation(title = "Absolute error in savings estimation of randomized M&V after satisfying all stopping criteria")
+  plot_annotation(title = "Absolute error in savings estimation of randomized M&V\nafter satisfying all stopping criteria")
 ```
@@
 
-```{r intervaltime, fig.cap="Comparison of time required to satisfy all stopping criteria when considering dropping non-consecutive days (or not)", fig.height=8, fig.width=12}
+```{r intervaltime, fig.cap="Comparison of time required to satisfy all stopping criteria when considering dropping non-consecutive days (or not)", fig.height=7, fig.width=9}
 df_drop <- bind_rows(df_seq_interval_tl_drop_stable, df_seq_interval_tl_drop_variable) %>%
@@
   facet_wrap(~interval, nrow = 2) +
-  geom_text(aes(x = n_weeks, y = perc + 2, label = paste0(round(perc, digits = 0), "%"), group = group), 
+  geom_text(aes(x = n_weeks, y = perc + 5, label = paste0(round(perc, digits = 0), "%"), group = group), 
             position = position_dodge(1), 
-            size = 4) +
+            size = 5) +
   scale_fill_manual(values = ls_colors) +
@@
 
-```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months (previous results of sampling at 50%/50% shown on the left side of the red dashed line)", fig.height=8, fig.width=12}
+```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months (previous results of sampling at 50%/50% shown on the left side of the red dashed line)", fig.height=4, fig.width=9}
 
@@
 
-p_top <- df_MW_A %>% 
+df_MW_A %>% 
   mutate(ratio = as.factor(ratio), 
@@
   ggplot(aes(x = ratio, y = values, fill = ratio)) +
-  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
   geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75, linewidth = 0.15) +
   geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
@@
             aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %")), 
-            size = 4) +
+            size = 5) +
   scale_y_continuous(expand = c(0, 0), 
@@
        y = "Absolute error in fractional savings", 
+       title = "Absolute error in savings estimation by different sampling ratio",
        subtitle = str_glue("All {A_building} buildings with measured weather conditions")) +
@@
         legend.position = "none",
-        axis.text.x = element_blank(), 
         plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-# TMY
-# plot for the stable subset
-df_TW_S <- df_cont_stable %>% 
-  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = abs(savings - cont_tmy), 
-         diff_all = abs(savings - cont_tmy_all)) %>% 
-  select(name, diff, diff_all, ratio) %>% 
-  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
-  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
-                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
-                               ifelse(ratio == 2 & type == "diff", "2_1", 
-                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
-  bind_rows(df_fs_tmy_stable %>% 
-              left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-              mutate(values = abs(savings - rand), 
-                     ratio = "0", 
-                     type = "diff_ref") %>% 
-              select(name, values, ratio, type))
-
-# plot for the variable subset
-df_TW_V <- df_cont_variable %>% 
-  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = abs(savings - cont_tmy), 
-         diff_all = abs(savings - cont_tmy_all)) %>% 
-  select(name, diff, diff_all, ratio) %>% 
-  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
-  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
-                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
-                               ifelse(ratio == 2 & type == "diff", "2_1", 
-                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
-  bind_rows(df_fs_tmy_variable %>% 
-              left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-              mutate(values = abs(savings - rand), 
-                     ratio = "0", 
-                     type = "diff_ref") %>% 
-              select(name, values, ratio, type))
-
-# plot for combined dataset
-df_TW_A <- rbind(df_TW_S, df_TW_V)
-
-p_bottom <- df_TW_A %>% 
-  mutate(ratio = as.factor(ratio), 
-         ratio = recode_factor(ratio, 
-                               "0" = "Randomized\n(50/50)", 
-                               "1_1" = "Continued\n(20/80)", 
-                               "1_2" = "Randomized\n(20/80)", 
-                               "2_1" = "Continued\n(10/90)", 
-                               "2_2" = "Randomized\n(10/90)")) %>% 
-  ggplot(aes(x = ratio, y = values, fill = ratio)) +
-  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
-  geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75) +
-  geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
-  geom_text(data = . %>% 
-              group_by(ratio) %>% 
-              summarise(median = median(values)) %>% 
-              ungroup(), 
-            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %")), 
-            size = 4) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 4), 
-                     labels = number_format(suffix = " %")) +
-  scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80", "grey80")) + 
-  labs(fill = NULL, 
-       x = NULL, 
-       y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
-  coord_cartesian(ylim = c(0, 15)) +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.position = "none",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-ggarrange(p_top, p_bottom, 
-          ncol = 1, nrow = 2,
-          labels = c("a)", "b)")) +
-  plot_annotation(title = "Absolute error in savings estimation by different sampling ratio")
 ```

### 2025-07-06 code/supplementary.Rmd
temp commit for figure update

@@
 readfile_stable <- str_glue("../readfiles/stable/")
-fig_path = "../figs/manuscript/"
+fig_path = "../figs/supplementary/"
 
@@
 
-```{r absnull, fig.cap="Comparison between the two M&V methods in detecting no intervention effect (expressed without absolute calculation) when buildings are subject to baseline change.", fig.height=6, fig.width=12}
+```{r stablesite, fig.cap = "Site summary of the filtered buildings from BDG2 dataset (counts < 10 are omitted for visualization; left: aggregated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}
+
+set3 <- colorRampPalette(brewer.pal('Set3',n=12))
+type_colors <- setNames(set3(13), all_types_variable$type)
+
+stable_set <- df_energy_stable %>%
+  group_by(type) %>%
+  distinct(name) %>%
+  summarise(stable = n()) %>%
+  ungroup()
+
+variable_set <- df_energy_variable %>%
+  group_by(type) %>%
+  distinct(name) %>%
+  summarise(variable = n()) %>%
+  ungroup()
+
+p1 <- variable_set %>% 
+  left_join(stable_set, by = "type") %>% 
+  mutate(stable = replace_na(stable, 0)) %>% 
+  group_by(type) %>% 
+  summarise(n = stable + variable) %>% 
+  ungroup() %>% 
+  mutate(type = fct_reorder(type, n, .desc = F)) %>%
+  ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
+  geom_col() +
+  geom_text(aes(label = ifelse(n > 10, n, "")), color = "black", position = position_stack(vjust = 0.5)) +
+  scale_y_continuous(expand = c(0, 0),
+                     breaks = breaks_pretty(n = 4)) +
+  scale_x_discrete(expand = c(0, 0.1)) +
+  scale_fill_manual(values = type_colors) +
+  labs(x = NULL,
+       y = NULL,
+       subtitle = "Combining all",
+       fill = NULL) +
+  theme(axis.text.x = element_blank(),
+        panel.grid.major.y = element_line(color = "grey80"),
+        legend.direction = "horizontal",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+stable_set <- df_energy_stable %>% 
+  left_join(df_loc, by = "site") %>% 
+  group_by(type, location) %>%
+  distinct(name) %>%
+  summarise(stable = n()) %>%
+  ungroup()
+
+variable_set <- df_energy_variable %>% 
+   left_join(df_loc, by = "site") %>% 
+  group_by(type, location) %>%
+  distinct(name) %>%
+  summarise(variable = n()) %>%
+  ungroup()
+
+p2 <- variable_set %>% 
+  left_join(stable_set, by = c("type", "location")) %>% 
+  mutate(stable = replace_na(stable, 0)) %>% 
+  group_by(location, type) %>% 
+  summarise(n = stable + variable) %>% 
+  ungroup() %>% 
+  group_by(location) %>%
+  mutate(proportion = n / sum(n),
+         total = sum(n),
+         ymax = cumsum(proportion),
+         ymin = c(0, head(ymax, n = -1))) %>%
+  mutate(label_pos = (ymax + ymin) / 2) %>%
+  ungroup() %>%
+  group_by(type) %>%
+  mutate(order = sum(n)) %>%
+  ungroup() %>%
+  mutate(type = fct_reorder(type, order, .desc = F)) %>%
+  ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3.2, fill = type)) +
+  geom_rect() +
+  coord_polar(theta = "y") +
+  xlim(c(2, 4)) +
+  facet_wrap(~ location, nrow = 3) +
+  labs(x = NULL,
+       y = NULL,
+       subtitle = "For each location",
+       fill = NULL) +
+  scale_fill_manual(values = type_colors) +
+  geom_text(aes(x = 3.7, y = label_pos, label = ifelse(n > 10, n, ""))) +
+  geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black", check_overlap = T) +
+  theme(legend.direction = "horizontal",
+        axis.text = element_blank(),
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+ggarrange(p1, p2,
+          ncol = 2, nrow = 1,
+          widths = c(0.2, 1),
+          common.legend = TRUE,
+          legend = "bottom") +
+  plot_annotation(title = "Summary of filtered buildings and types")
+```
+
+```{r chwst, fig.cap = "Proposed intervention strategy: chilled water supply temperature reset based on outdoor temperature.", fig.width=8, fig.height=4}
+
+data.frame(
+  OAT = c(10, 15, 25, 30),  
+  Baseline = c(6, 6, 6, 6),      
+  Intervention = c(12, 12, 7, 7)
+  ) %>% 
+  pivot_longer(cols = c("Baseline", "Intervention"), names_to = "strategy", values_to = "value") %>% 
+  ggplot(aes(x = OAT, y = value, color = strategy)) +
+  geom_line(linewidth = 1.2) +
+  scale_color_manual(values = ls_colors) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = c(4, 6, 8, 10, 12),
+                     limits = c(5, 12.5),
+                     labels = number_format(suffix = " °C")) +
+  scale_x_continuous(expand = c(0, 0), 
+                     breaks = c(5, 10, 15, 20, 25, 30), 
+                     labels = number_format(suffix = " °C")) +
+  geom_vline(xintercept = 10, lty = "dashed", color = "red") +
+  annotate(geom = "text", x = 7.5, y = 8, label = "No chiller\noperation\nbelow 10 °C", size = 5) +
+  coord_cartesian(xlim = c(5, 30)) +
+  labs(x = "Outdoor temperature", 
+       y = "Chilled water supply temperature", 
+       color = NULL, 
+       title = "Chilled water supply temperature reset intervention") +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.direction = "horizontal",
+        legend.position = "bottom",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+```
+
+```{r meanerr, fig.cap="Distribution of estimated deviation from the ground-truth savings calculated by the two M&V methods (red dashed line highlighting 0% error)", fig.height=7, fig.width=9}
+# plot for the stable subset
+rand_eob_S <- df_seq_fs_stable %>% 
+  filter(seq == "eob") %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = savings - fs, 
+         method = "rand_eob") %>% 
+  select(name, method, diff)
+
+conv_S <- df_fs_stable %>% 
+  filter(method != "rand") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = true - conv, 
+         method = "conv") %>% 
+  select(name, method, diff)
+
+rand_final_S <- df_fs_stable %>% 
+  filter(method != "conv") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = true - rand, 
+         method = "rand_final") %>% 
+  select(name, method, diff)
+
+df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)
+
+# plot for the variable subset
+rand_eob_V <- df_seq_fs_variable %>% 
+  filter(seq == "eob") %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site"))%>% 
+  mutate(diff = savings - fs, 
+         method = "rand_eob") %>% 
+  select(name, method, diff)
+
+conv_V <- df_fs_variable %>% 
+  filter(method != "rand") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = true - conv, 
+         method = "conv") %>% 
+  select(name, method, diff)
+
+rand_final_V <- df_fs_variable %>% 
+  filter(method != "conv") %>% 
+  pivot_wider(names_from = method, values_from = savings) %>% 
+  mutate(diff = true - rand, 
+         method = "rand_final") %>% 
+  select(name, method, diff)
+
+df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)
+
+# plot for combined dataset
+rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
+  select(name, rand_eob_diff = diff)
+
+conv_A <- bind_rows(conv_V, conv_S) %>% 
+  select(name, conv_diff = diff)
+
+rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
+  select(name, rand_final_diff = diff)
+
+df_MW_A <- rand_eob_A %>% 
+  left_join(conv_A, by = "name") %>% 
+  left_join(rand_final_A, by = "name") %>% 
+  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")
+
+p_top <- df_MW_A %>% 
+  mutate(method = as.factor(method), 
+         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
+  ggplot(aes(x = method, y = diff, fill = method)) +
+  geom_lv(k = 4, outlier.shape = NA) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method), linewidth = 0.25) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
+  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
+            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
+            size = 5) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 5), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
+  labs(fill = NULL, 
+       color = NULL, 
+       x = NULL, 
+       y = "Estimated error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with measured weather")) +
+  coord_cartesian(ylim = c(-15, 15)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "none",
+        axis.text.x = element_blank(),
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+# TMY
+# plot for the stable subset
+rand_eob_S <- df_sprt_all_stable %>% 
+  filter(seq == "eob") %>% 
+  select(-c(seq, n_weeks)) %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = savings - annual, 
+         method = "rand_eob") %>% 
+  select(name, method, diff)
+
+conv_S <- df_fs_tmy_stable %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = savings - conv, 
+         method = "conv") %>% 
+  select(name, method, diff)
+
+rand_final_S <- df_fs_tmy_stable %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = savings - rand, 
+         method = "rand_final") %>% 
+  select(name, method, diff)
+
+df_TW_S <- rbind(rand_eob_S, conv_S, rand_final_S)
+
+# plot for the variable subset
+rand_eob_V <- df_sprt_all_variable %>% 
+  filter(seq == "eob") %>% 
+  select(-c(seq, n_weeks)) %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = savings - annual, 
+         method = "rand_eob") %>% 
+  select(name, method, diff)
+
+conv_V <- df_fs_tmy_variable %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = savings - conv, 
+         method = "conv") %>% 
+  select(name, method, diff)
+
+rand_final_V <- df_fs_tmy_variable %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = savings - rand, 
+         method = "rand_final") %>% 
+  select(name, method, diff)
+
+df_TW_V <- rbind(rand_eob_V, conv_V, rand_final_V)
+
+# plot for combined dataset
+rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
+  select(name, rand_eob_diff = diff)
+
+conv_A <- bind_rows(conv_V, conv_S) %>% 
+  select(name, conv_diff = diff)
+
+rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
+  select(name, rand_final_diff = diff)
+
+df_TW_A <- rand_eob_A %>% 
+  left_join(conv_A, by = "name") %>% 
+  left_join(rand_final_A, by = "name") %>% 
+  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")
+
+p_middle <- df_TW_A %>% 
+  mutate(method = as.factor(method), 
+         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
+  ggplot(aes(x = method, y = diff, fill = method)) +
+  geom_lv(k = 4, outlier.shape = NA) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method), linewidth = 0.25) +
+  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
+  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
+            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
+            size = 5) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 5), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
+  labs(fill = NULL, 
+       x = NULL, 
+       y = "Estimated error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
+  coord_cartesian(ylim = c(-15, 15)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "none",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+ggarrange(p_top, p_middle, 
+          ncol = 1, nrow = 2,
+          labels = c("a)", "b)")) +
+  plot_annotation(title = "Estimated error in savings estimation by different M&V methods")
+
+percentiles <- df_TW_A %>%
+  filter(method == "conv_diff") %>% 
+  summarise(
+    Q1     = round(quantile(diff, 0.25, na.rm = TRUE), digits = 2),
+    Median = round(quantile(diff, 0.50, na.rm = TRUE), digits = 2),
+    Q3     = round(quantile(diff, 0.75, na.rm = TRUE), digits = 2)
+  )
+```
+
+```{r absnull, fig.cap="Comparison between the two M&V methods in detecting no intervention effect (expressed without absolute calculation) when buildings are subject to baseline change.", fig.height=4, fig.width=9}
 #### ABS-NULL ####
@@
   ggplot(aes(x = method, y = diff, fill = method)) +
-  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
   geom_lv(k = 4, outlier.shape = NA) +
@@
        y = "Absolute effect in fractional savings", 
-       title = "Absolute effect of non-routine events when using different M&V methods", 
+       title = "Estimated absolute effect of non-routine events by different M&V methods", 
        subtitle = str_glue("All {A_building} buildings wither measured weather")) +
@@
 
-```{r meancont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy (expressed without absolute calculation for errors)", fig.height=8, fig.width=12}
+```{r meancont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy (expressed without absolute calculation for errors)", fig.height=7, fig.width=9}
 
@@
   ggplot(aes(x = ratio, y = values, fill = ratio)) +
-  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
   geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
@@
 
-```{r absint2y, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy over the course of two years (dropped: all non-consecutive days were dropped; kept: all measurements were kept)", fig.height=9, fig.width=12}
+```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months (previous results of sampling at 50%/50% shown on the left side of the red dashed line)", fig.height=7, fig.width=9}
+
+#### ABS-CONT ####
+df_MW_S <- df_cont_stable %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - cont_fs), 
+         diff_all = abs(savings - cont_fs_all)) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_stable %>% 
+              filter(method != "conv") %>% 
+              pivot_wider(names_from = method, values_from = savings) %>% 
+              mutate(values = abs(true - rand), 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
+
+# plot for the variable subset
+df_MW_V <- df_cont_variable %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - cont_fs), 
+         diff_all = abs(savings - cont_fs_all)) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_variable %>% 
+              filter(method != "conv") %>% 
+              pivot_wider(names_from = method, values_from = savings) %>% 
+              mutate(values = abs(true - rand), 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
+
+# plot for combined dataset
+df_MW_A <- rbind(df_MW_S, df_MW_V)
+
+p_top <- df_MW_A %>% 
+  mutate(ratio = as.factor(ratio), 
+         ratio = recode_factor(ratio, 
+                               "0" = "Randomized\n(50/50)", 
+                               "1_1" = "Continued\n(20/80)", 
+                               "1_2" = "Randomized\n(20/80)", 
+                               "2_1" = "Continued\n(10/90)", 
+                               "2_2" = "Randomized\n(10/90)")) %>% 
+  ggplot(aes(x = ratio, y = values, fill = ratio)) +
+  geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75) +
+  geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% 
+              group_by(ratio) %>% 
+              summarise(median = median(values)) %>% 
+              ungroup(), 
+            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %")), 
+            size = 5) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80", "grey80")) + 
+  labs(fill = NULL, 
+       x = NULL, 
+       y = "Absolute error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with measured weather conditions")) +
+  coord_cartesian(ylim = c(0, 15)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "none",
+        axis.text.x = element_blank(), 
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+# TMY
+# plot for the stable subset
+df_TW_S <- df_cont_stable %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - cont_tmy), 
+         diff_all = abs(savings - cont_tmy_all)) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_tmy_stable %>% 
+              left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+              mutate(values = abs(savings - rand), 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
+
+# plot for the variable subset
+df_TW_V <- df_cont_variable %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - cont_tmy), 
+         diff_all = abs(savings - cont_tmy_all)) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_tmy_variable %>% 
+              left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+              mutate(values = abs(savings - rand), 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
+
+# plot for combined dataset
+df_TW_A <- rbind(df_TW_S, df_TW_V)
+
+p_bottom <- df_TW_A %>% 
+  mutate(ratio = as.factor(ratio), 
+         ratio = recode_factor(ratio, 
+                               "0" = "Randomized\n(50/50)", 
+                               "1_1" = "Continued\n(20/80)", 
+                               "1_2" = "Randomized\n(20/80)", 
+                               "2_1" = "Continued\n(10/90)", 
+                               "2_2" = "Randomized\n(10/90)")) %>% 
+  ggplot(aes(x = ratio, y = values, fill = ratio)) +
+  geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75) +
+  geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% 
+              group_by(ratio) %>% 
+              summarise(median = median(values)) %>% 
+              ungroup(), 
+            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %")), 
+            size = 5) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80", "grey80")) + 
+  labs(fill = NULL, 
+       x = NULL, 
+       y = "Absolute error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
+  coord_cartesian(ylim = c(0, 15)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "none",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+ggarrange(p_top, p_bottom, 
+          ncol = 1, nrow = 2,
+          labels = c("a)", "b)")) +
+  plot_annotation(title = "Absolute error in savings estimation by different sampling ratio")
+```
+
+```{r absint2y, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy over the course of two years (dropped: all non-consecutive days were dropped; kept: all measurements were kept)", fig.height=8, fig.width=9}
 # dropped version
@@
             position = position_dodge(width = 0.75), 
-            size = 4, 
+            size = 5, 
             color = "white") +
@@
        subtitle = str_glue("All {A_building} buildings with measured weather")) +
-  coord_cartesian(ylim = c(0, 8)) +
+  coord_cartesian(ylim = c(0, 6)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
             position = position_dodge(width = 0.75), 
-            size = 4,
+            size = 5,
             color = "white") +
@@
        subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
-  coord_cartesian(ylim = c(0, 8)) +
+  coord_cartesian(ylim = c(0, 6)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
           legend = "bottom") +
-  plot_annotation(title = "Absolute error in savings estimation of ranodmized M&V by different sampling intervals")
+  plot_annotation(title = "Absolute error in savings estimation of ranodmized M&V\nafter 2 years")
 ```
 
-```{r meanintspt, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy after satisfying all stopping criteria (without absolute calculation)", fig.height=9, fig.width=12}
+```{r meanintspt, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy after satisfying all stopping criteria (without absolute calculation)", fig.height=8, fig.width=9}
 # sprt sequence plot for different sampling intervals
@@
        subtitle = str_glue("All {A_building} buildings with measured weather")) +
-  coord_cartesian(ylim = c(-12, 12)) +
+  coord_cartesian(ylim = c(-10, 10)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
        subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
-  coord_cartesian(ylim = c(-12, 12)) +
+  coord_cartesian(ylim = c(-10, 10)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
           legend = "bottom") +
-  plot_annotation(title = "Estimated error in savings estimation of randomized M&V after satisfying all stopping criteria")
+  plot_annotation(title = "Estimated error in savings estimation of randomized M&V\nafter satisfying all stopping criteria")
 
@@
 
-```{r meanint2y, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy over the course of the two years (without absolute calculation)", fig.width=12, fig.height=9}
+```{r meanint2y, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy over the course of the two years (without absolute calculation)", fig.width=9, fig.height=8}
 # dropped version
@@
        subtitle = str_glue("All {A_building} buildings with measured weather")) +
-  coord_cartesian(ylim = c(-10, 10)) +
+  coord_cartesian(ylim = c(-5, 5)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
        subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
-  coord_cartesian(ylim = c(-10, 10)) +
+  coord_cartesian(ylim = c(-5, 5)) +
   theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
@@
           legend = "bottom") +
-  plot_annotation(title = "Estimated error in savings estimation of randomized M&V by different sampling intervals")
+  plot_annotation(title = "Estimated error in savings estimation of randomized M&V\nafter 2 years")
 ```

### 2025-07-06 code/0_schedule.R
updated timeline figure for manuscript

@@ -1,134 +1,0 @@
-# Block design for running different sampling intervals
-# Code written by Aoyu Zou
-
-
-
-
-#### Setup ####
-# use pacman
-require(pacman)
-
-# load libraries
-pacman::p_load(tidyverse, lubridate, here, scales, slider, patchwork, # general
-               broom, ggpmisc, ggpubr, # linear models
-               RMV2.0, # TOWT model
-               sprtt, effsize) # sequential testing
-
-# turn off scientific notation
-options(scipen = 999)
-
-# set default theme for ggplot
-theme_set(theme_minimal())
-
-# define base ggplot theme
-theme_update(plot.title = element_text(size = 14, colour = "grey20", face = "bold", hjust = 0.5),
-             plot.subtitle = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
-             plot.caption = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5),
-             plot.background = element_rect(fill = "white", colour = NA),
-             panel.grid.minor = element_blank(),
-             panel.grid.major = element_blank(),
-             axis.text = element_text(size = 10),
-             strip.text = element_text(size = 10, color = "grey20", face = "bold"),
-             strip.background = element_blank())
-
-
-
-
-
-#### READ ####
-function_path <- "../functions/"
-output_path <- "../readfiles/"
-source(paste0(function_path, "rand_seq.R"))
-
-
-
-
-
-#### GENERATION ####
-# adding random sampling schedules
-block_params <- list(start_date = "2016-01-01",
-                     n_weeks = 108,
-                     n_seasons = 18, 
-                     block_unit = 6, 
-                     pool = 20)
-
-# schedule generation
-exclude <- data.frame(date = NA)
-
-# consecutive sampling
-schedule_2_design <- data.frame()
-schedule_3_design <- data.frame()
-schedule_7_design <- data.frame()
-i <- 1
-
-while (i <= block_params$pool){
-  
-  schedule <- data.frame()
-  for (sample_start in as.list(seq(as.Date(block_params$start_date), as.Date(block_params$start_date) + weeks(block_params$n_weeks), by = block_params$block_unit * 7))){
-
-    sample_end = sample_start + weeks(block_params$block_unit) - days(1)
[…]

### 2025-07-06 code/2_stablevisual.R

@@ -1,1192 +1,0 @@
-# Genome 2 M&V analysis
-# written by Aoyu
-
-
-
-#### SETUP ####
-
-# use pacman
-require(pacman)
-
-# load libraries
-pacman::p_load(tidyverse, lubridate, scales, slider, cowplot, patchwork, RColorBrewer, # general
-               broom, ggpmisc, ggpubr) # plotting
-
-# turn off scientific notation
-options(scipen = 999)
-
-# set default theme for ggplot
-theme_set(theme_minimal())
-
-# define base ggplot theme
-theme_update(plot.title = element_text(size = 14, color = "grey20", face = "bold", hjust = 0.5),
-             plot.subtitle = element_text(size = 10, color = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
-             plot.caption = element_text(size = 8, color = "grey20", face = "italic", hjust = 0.5),
-             plot.background = element_rect(fill = "white", color = NA),
-             panel.grid.minor = element_blank(),
-             panel.grid.major = element_blank(),
-             strip.text = element_text(size = 10, color = "grey20", face = "bold"),
-             strip.background = element_blank())
-
-# define strategy colors
-ls_colors <- c("Baseline" = "#99d8c9",
-               "Measured baseline" = "#99d8c9",
-               "Adjusted baseline" = "#99d8c9",
-               "Projected baseline\n(no change)" = "#1b9e77",
-               "Intervention" = "#fdbb84",
-               "Measured interv" = "#fdbb84",
-               "True savings" = "black",
-               "True normalized savings" = "black")
-
-# define parameters
-run_params <- list(type = "stable", 
-                   site = F)
-
-sprt_param <- list(baseline = "Baseline",
-                   strategy = "Intervention",
-                   parameter = "power_ave",
-                   label = "power",
-                   n_weeks = 48)
-
-block_params <- list(start_date = "2016-01-01",
-                     n_weeks = 108,
-                     n_seasons = 9, 
-                     block_unit = 12)
-
-
-
-
-#### FUNCTIONS ####
-# Function defined to read downloaded tmy files
-get_tmy <- function(all_sites){
[…]

### 2025-07-06 code/graphical.R

@@ -14,14 +14,15 @@ here::i_am("manuscript.rmd")
 theme_set(theme_minimal())
 
 # define base ggplot theme
-theme_update(plot.title = element_text(size = 14, colour = "grey20", face = "bold", hjust = 0.5),
-             plot.subtitle = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
+theme_update(plot.title = element_text(size = 16, colour = "grey20", face = "bold", hjust = 0.5),
+             plot.subtitle = element_text(size = 14, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
              plot.caption = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5),
              plot.background = element_rect(fill = "white", colour = NA),
              panel.grid.minor = element_blank(),
              panel.grid.major = element_blank(),
-             axis.text = element_text(size = 10),
-             strip.text = element_text(size = 10, color = "grey20", face = "bold"),
+             axis.text = element_text(size = 12),
+             legend.text = element_text(size = 12),
+             strip.text = element_text(size = 12, color = "grey20", face = "bold"),
              strip.background = element_blank())
 
 # colors
@@ -1920,7 +1921,7 @@ df_MW_A %>%
   ggplot(aes(x = n_weeks, y = perc, fill = group)) +
   geom_bar(stat = "identity", position = "dodge") +
   facet_wrap(~interval, nrow = 2) +
-  geom_text(aes(x = n_weeks, y = perc + 2, label = paste0(round(perc, digits = 0), "%"), group = group), position = position_dodge(1)) +
+  geom_text(aes(x = n_weeks, y = perc + 5, label = paste0(round(perc, digits = 0), "%"), group = group), position = position_dodge(1), size = 5) +
   scale_fill_manual(values = ls_colors) +
   scale_color_manual(values = ls_colors) +
   scale_y_continuous(breaks = breaks_pretty(n = 4), 

### 2025-07-06 code/0_schedule.R
updated figure for manuscript

@@ -1,134 +1,0 @@
-# Block design for running different sampling intervals
-# Code written by Aoyu Zou
-
-
-
-
-#### Setup ####
-# use pacman
-require(pacman)
-
-# load libraries
-pacman::p_load(tidyverse, lubridate, here, scales, slider, patchwork, # general
-               broom, ggpmisc, ggpubr, # linear models
-               RMV2.0, # TOWT model
-               sprtt, effsize) # sequential testing
-
-# turn off scientific notation
-options(scipen = 999)
-
-# set default theme for ggplot
-theme_set(theme_minimal())
-
-# define base ggplot theme
-theme_update(plot.title = element_text(size = 14, colour = "grey20", face = "bold", hjust = 0.5),
-             plot.subtitle = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
-             plot.caption = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5),
-             plot.background = element_rect(fill = "white", colour = NA),
-             panel.grid.minor = element_blank(),
-             panel.grid.major = element_blank(),
-             axis.text = element_text(size = 10),
-             strip.text = element_text(size = 10, color = "grey20", face = "bold"),
-             strip.background = element_blank())
-
-
-
-
-
-#### READ ####
-function_path <- "../functions/"
-output_path <- "../readfiles/"
-source(paste0(function_path, "rand_seq.R"))
-
-
-
-
-
-#### GENERATION ####
-# adding random sampling schedules
-block_params <- list(start_date = "2016-03-01",
-                     n_weeks = 96,
-                     n_seasons = 8, 
-                     block_unit = 12, 
-                     pool = 20)
-
-# schedule generation
-exclude <- data.frame(date = NA)
-
-# consecutive sampling
-schedule_2_design <- data.frame()
-schedule_3_design <- data.frame()
-schedule_7_design <- data.frame()
-i <- 1
-
-while (i <= block_params$pool){
-  
-  schedule <- data.frame()
-  for (sample_start in as.list(seq(as.Date(block_params$start_date), as.Date(block_params$start_date) + weeks(block_params$n_weeks), by = block_params$block_unit * 7))){
-
-    sample_end = sample_start + weeks(block_params$block_unit) - days(1)
[…]

### 2025-07-06 code/0_stabledata.R

@@ -1,124 +1,0 @@
-# Genome 2 M&V analysis
-# written by Aoyu
-
-
-
-#### SETUP ####
-
-# use pacman
-require(pacman)
-
-# load libraries
-pacman::p_load(tidyverse, lubridate, scales, slider, cowplot, patchwork, RColorBrewer, # general
-               broom, ggpmisc, ggpubr, # linear models
-               RMV2.0, # TOWT model
-               sprtt, effsize) # sequential testing
-
-# turn off scientific notation
-options(scipen = 999)
-
-# set default theme for ggplot
-theme_set(theme_minimal())
-
-# define base ggplot theme
-theme_update(plot.title = element_text(size = 14, colour = "grey20", face = "bold", hjust = 0.5),
-             plot.subtitle = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
-             plot.caption = element_text(size = 8, colour = "grey20", face = "italic", hjust = 0.5),
-             plot.background = element_rect(fill = "white", colour = NA),
-             panel.grid.minor = element_blank(),
-             panel.grid.major = element_blank(),
-             strip.text = element_text(size = 10, color = "grey20", face = "bold"),
-             strip.background = element_blank())
-
-# T test
-perform_t_test <- function(df_pre, df_post, building) {
-  t.test(df_pre[[building]], df_post[[building]], alternative = "two.sided")
-}
-
-
-
-
-#### READ DATA ####
-data_path <- "../../buds-lab-building-data-genome-project-2/data/meters/cleaned/"
-meta_path <- "../../buds-lab-building-data-genome-project-2/data/metadata/"
-weather_path <- "../../buds-lab-building-data-genome-project-2/data/weather/"
-output_path <- "../readfiles/stable/"
-
-df_elec <- read_csv(paste0(data_path, "electricity_cleaned.csv"))
-df_meta <- read_csv(paste0(meta_path, "metadata.csv"))
-df_weather <- read_csv(paste0(weather_path, "weather.csv"))
-
-
-
-
-#### FILTER ####
-# Focus
-remove_type <- c("parking", "warehouse", "utility")
-remove_site <- c("Eagle", "Bobcat", "Swan", "Hog")
-
-# NAs
[…]

### 2025-07-06 code/0_variabledata.R

@@ -1,146 +1,0 @@
-# Genome 2 M&V analysis
-# written by Aoyu
-
-
-
-#### SETUP ####
-
-# use pacman
-require(pacman)
-
-# load libraries
-pacman::p_load(tidyverse, lubridate, scales, slider, cowplot, patchwork, RColorBrewer, # general
-               broom, ggpmisc, ggpubr, # linear models
-               RMV2.0, # TOWT model
-               sprtt, effsize) # sequential testing
-
-# turn off scientific notation
-options(scipen = 999)
-
-# set default theme for ggplot
-theme_set(theme_minimal())
-
-# define base ggplot theme
-theme_update(plot.title = element_text(size = 14, colour = "grey20", face = "bold", hjust = 0.5),
-             plot.subtitle = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
-             plot.caption = element_text(size = 8, colour = "grey20", face = "italic", hjust = 0.5),
-             plot.background = element_rect(fill = "white", colour = NA),
-             panel.grid.minor = element_blank(),
-             panel.grid.major = element_blank(),
-             strip.text = element_text(size = 10, color = "grey20", face = "bold"),
-             strip.background = element_blank())
-
-# T test
-perform_t_test <- function(df_pre, df_post, building) {
-  t.test(df_pre[[building]], df_post[[building]], alternative = "two.sided")
-}
-
-perform_mean_diff <- function(df_pre, df_post, building) {
-  abs(mean(df_pre[[building]], na.rm = T) - mean(df_post[[building]], na.rm = T)) / mean(df_pre[[building]], na.rm = T) * 100
-}
-
-perform_mean_est <- function(df_pre, df_post, building) {
-  0.5 * (sum(df_pre[[building]], na.rm = T) + sum(df_post[[building]], na.rm = T))
-}
-
-#### READ DATA ####
-data_path <- "../../buds-lab-building-data-genome-project-2/data/meters/cleaned/"
-meta_path <- "../../buds-lab-building-data-genome-project-2/data/metadata/"
-weather_path <- "../../buds-lab-building-data-genome-project-2/data/weather/"
-output_path <- "../readfiles/variable/"
-
[…]

### 2025-07-06 code/graphical.R

@@ -8,20 +8,21 @@ pacman::p_load(tidyverse, lubridate, here, stats, zoo, scales, lvplot, ggpubr, g
 options(scipen = 999, digits = 15)
 
 # set directory
-here::i_am("manuscript.rmd")
+here::i_am("code.Rproj")
 
 # set default theme for ggplot
 theme_set(theme_minimal())
 
 # define base ggplot theme
-theme_update(plot.title = element_text(size = 14, colour = "grey20", face = "bold", hjust = 0.5),
-             plot.subtitle = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
+theme_update(plot.title = element_text(size = 16, colour = "grey20", face = "bold", hjust = 0.5),
+             plot.subtitle = element_text(size = 14, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
              plot.caption = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5),
              plot.background = element_rect(fill = "white", colour = NA),
              panel.grid.minor = element_blank(),
              panel.grid.major = element_blank(),
-             axis.text = element_text(size = 10),
-             strip.text = element_text(size = 10, color = "grey20", face = "bold"),
+             axis.text = element_text(size = 12),
+             legend.text = element_text(size = 12),
+             strip.text = element_text(size = 12, color = "grey20", face = "bold"),
              strip.background = element_blank())
 
 # colors
@@ -1553,7 +1554,7 @@ df_MW_A %>%
   ggplot(aes(x = n_weeks, y = perc, fill = group)) +
   geom_bar(stat = "identity", position = "dodge") +
   facet_wrap(~interval, nrow = 2) +
-  geom_text(aes(x = n_weeks, y = perc + 2, label = paste0(round(perc, digits = 0), "%"), group = group), position = position_dodge(1)) +
+  geom_text(aes(x = n_weeks, y = perc + 5, label = paste0(round(perc, digits = 0), "%"), group = group), position = position_dodge(1), size = 5) +
   scale_fill_manual(values = ls_colors) +
   scale_color_manual(values = ls_colors) +
   scale_y_continuous(breaks = breaks_pretty(n = 4), 
@@ -1567,4 +1568,4 @@ df_MW_A %>%
         legend.position = "bottom",
[…]

### 2025-07-06 code/0_schedule.R
updated figure for manuscript

@@ -1,134 +1,0 @@
-# Block design for running different sampling intervals
-# Code written by Aoyu Zou
-
-
-
-
-#### Setup ####
-# use pacman
-require(pacman)
-
-# load libraries
-pacman::p_load(tidyverse, lubridate, here, scales, slider, patchwork, # general
-               broom, ggpmisc, ggpubr, # linear models
-               RMV2.0, # TOWT model
-               sprtt, effsize) # sequential testing
-
-# turn off scientific notation
-options(scipen = 999)
-
-# set default theme for ggplot
-theme_set(theme_minimal())
-
-# define base ggplot theme
-theme_update(plot.title = element_text(size = 14, colour = "grey20", face = "bold", hjust = 0.5),
-             plot.subtitle = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
-             plot.caption = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5),
-             plot.background = element_rect(fill = "white", colour = NA),
-             panel.grid.minor = element_blank(),
-             panel.grid.major = element_blank(),
-             axis.text = element_text(size = 10),
-             strip.text = element_text(size = 10, color = "grey20", face = "bold"),
-             strip.background = element_blank())
-
-
-
-
-
-#### READ ####
-function_path <- "../functions/"
-output_path <- "../readfiles/"
-source(paste0(function_path, "rand_seq.R"))
-
-
-
-
-
-#### GENERATION ####
-# adding random sampling schedules
-block_params <- list(start_date = "2016-03-01",
-                     n_weeks = 96,
-                     n_seasons = 16, 
-                     block_unit = 6, 
-                     pool = 20)
-
-# schedule generation
-exclude <- data.frame(date = NA)
-
-# consecutive sampling
-schedule_2_design <- data.frame()
-schedule_3_design <- data.frame()
-schedule_7_design <- data.frame()
-i <- 1
-
-while (i <= block_params$pool){
-  
-  schedule <- data.frame()
-  for (sample_start in as.list(seq(as.Date(block_params$start_date), as.Date(block_params$start_date) + weeks(block_params$n_weeks), by = block_params$block_unit * 7))){
-
-    sample_end = sample_start + weeks(block_params$block_unit) - days(1)
[…]

### 2025-07-06 code/2_variablevisual.R

@@ -1,1251 +1,0 @@
-# Genome 2 M&V analysis
-# written by Aoyu
-
-
-
-#### SETUP ####
-
-# use pacman
-require(pacman)
-
-# load libraries
-pacman::p_load(tidyverse, lubridate, scales, slider, cowplot, patchwork, RColorBrewer, # general
-               broom, ggpmisc, ggpubr) # plotting
-
-# turn off scientific notation
-options(scipen = 999)
-
-# set default theme for ggplot
-theme_set(theme_minimal())
-
-# define base ggplot theme
-theme_update(plot.title = element_text(size = 14, color = "grey20", face = "bold", hjust = 0.5),
-             plot.subtitle = element_text(size = 10, color = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
-             plot.caption = element_text(size = 8, color = "grey20", face = "italic", hjust = 0.5),
-             plot.background = element_rect(fill = "white", color = NA),
-             panel.grid.minor = element_blank(),
-             panel.grid.major = element_blank(),
-             strip.text = element_text(size = 10, color = "grey20", face = "bold"),
-             strip.background = element_blank())
-
-# define strategy colors
-ls_colors <- c("Baseline" = "#99d8c9",
-               "Measured baseline" = "#99d8c9",
-               "Adjusted baseline" = "#99d8c9",
-               "Projected baseline\n(no change)" = "#1b9e77",
-               "Intervention" = "#fdbb84",
-               "Measured interv" = "#fdbb84",
-               "True savings" = "black",
-               "True normalized savings" = "black")
-
-# define parameters
-run_params <- list(type = "variable", 
-                   site = T)
-
-sprt_param <- list(baseline = "Baseline",
-                   strategy = "Intervention",
-                   parameter = "power_ave",
-                   label = "power",
-                   n_weeks = 48)
-
-block_params <- list(start_date = "2016-01-01",
-                     n_weeks = 108,
-                     n_seasons = 9, 
-                     block_unit = 12)
-
-
-
-#### FUNCTIONS ####
-# Function defined to read downloaded tmy files
[…]

### 2025-07-06 code/graphical.R

@@ -8,20 +8,21 @@ pacman::p_load(tidyverse, lubridate, here, stats, zoo, scales, lvplot, ggpubr, g
 options(scipen = 999, digits = 15)
 
 # set directory
-here::i_am("manuscript.rmd")
+here::i_am("code.Rproj")
 
 # set default theme for ggplot
 theme_set(theme_minimal())
 
 # define base ggplot theme
-theme_update(plot.title = element_text(size = 14, colour = "grey20", face = "bold", hjust = 0.5),
-             plot.subtitle = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
+theme_update(plot.title = element_text(size = 16, colour = "grey20", face = "bold", hjust = 0.5),
+             plot.subtitle = element_text(size = 14, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
              plot.caption = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5),
              plot.background = element_rect(fill = "white", colour = NA),
              panel.grid.minor = element_blank(),
              panel.grid.major = element_blank(),
-             axis.text = element_text(size = 10),
-             strip.text = element_text(size = 10, color = "grey20", face = "bold"),
+             axis.text = element_text(size = 12),
+             legend.text = element_text(size = 12),
+             strip.text = element_text(size = 12, color = "grey20", face = "bold"),
              strip.background = element_blank())
 
 # colors
@@ -1582,7 +1583,7 @@ df_MW_A %>%
   ggplot(aes(x = n_weeks, y = perc, fill = group)) +
   geom_bar(stat = "identity", position = "dodge") +
   facet_wrap(~interval, nrow = 2) +
-  geom_text(aes(x = n_weeks, y = perc + 2, label = paste0(round(perc, digits = 0), "%"), group = group), position = position_dodge(1)) +
+  geom_text(aes(x = n_weeks, y = perc + 5, label = paste0(round(perc, digits = 0), "%"), group = group), position = position_dodge(1), size = 5) +
   scale_fill_manual(values = ls_colors) +
   scale_color_manual(values = ls_colors) +
   scale_y_continuous(breaks = breaks_pretty(n = 4), 
@@ -1596,5 +1597,5 @@ df_MW_A %>%
         legend.position = "bottom",
[…]

### 2025-07-15 code/manuscript.Rmd
updaded review for manuscript and updated readme file

@@
 author:
-  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^ ^[Department of Architecture, College of Environmental Design, University of California, Berkeley, USA] ^[Department of Civil and Environmental Engineering, College of Engineering, University of California, Berkeley, USA], Carlos Duarte ^1^, Gail Brager ^1^
-abstract: Conventional measurement and verification (M&V) methods rely on comparing pre- and post-retrofit performance to estimate energy savings.
-They are often time-consuming and unreliable, especially when non-routine events, as step changes or more gradual changes in how a building operates, occur during the M&V process.
-Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies the conventional M&V method, the results can be confounded.
+  - Aoyu Zou:
+      institute: cbe
+      correspondence: true
+  - name: Paul Raftery
+    institute: cbe
+  - name: Stefano Schiavon
+    institute: [cbe, arch, cee]
+  - name: Carlos Duarte
+    institute: cbe
+  - name: Gail Brager
+    institute: [cbe, arch]
+institute:
+  - cbe: Center for the Built Environment, University of California Berkeley, USA
+  - arch: Department of Architecture, College of Environmental Design, University of California, Berkeley, USA
+  - cee: Department of Civil and Environmental Engineering, College of Engineering, University of California, Berkeley, USA
+abstract: Conventional measurement and verification (M&V) methods for estimating energy savings rely on comparing pre- and post-retrofit performance.
+They are often time-consuming and unreliable, especially when non-routine events, such as step changes or more gradual changes in building operation, occur during the M&V process.
+When those events are unrelated to the retrofit intervention but may significantly affect building energy consumption, then when the analyst applies the conventional M&V method the results can be confounded.
 In this study, we demonstrated that switchable interventions, such as most HVAC control retrofits, can benefit from a new M&V method that randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily).
 We tested this novel randomized M&V method on a large public dataset (639 buildings) covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention.
-The results show that compared to the conventional method, the randomized method provides more accurate savings estimations with a median of 74% accuracy improvement and is even faster (typically 36 weeks instead of 104 weeks, ~65% reduction).
-Additionally, we found that when non-routine events are present, the randomized method estimates savings that are much closer to the ground-truth savings than the conventional method, demonstrating much improved reliability.
+The results show that, compared to the conventional method, the randomized method provides more accurate savings estimations with a median of 74% accuracy improvement and is even faster (typically 36 weeks instead of 104 weeks, ~65% reduction).
+Additionally, we found that when non-routine events are present (e.g., occupancy pattern change), the randomized method estimates savings that are much closer to the ground-truth savings than the conventional method, demonstrating much improved reliability.
 We also assessed the impact of normalizing for different weather, starting the M&V at different dates of the year, continuing randomization with a different sampling ratio after satisfying all criteria, and dropping samples affected by carryover effects when switching between strategies.
@@
     reference_docx: "../paper/template.docx"
+    number_sections: false
+    pandoc_args:
+      - '--lua-filter=scholarly-metadata.lua'
+      - '--lua-filter=author-info-blocks.lua'
+
+scholarly-metadata:
+  affiliations-location: footnote
+
 editor_options: 
@@
 
-Measurement and Verification (M&V) is the process of quantifying energy
-savings from energy efficiency projects by comparing actual energy
-consumption against a baseline, adjusting for factors like weather and
-occupancy. This process ensures that improvements in energy performance
-are accurately evaluated. In the United States, practitioners often
-refer to ASHRAE Guideline 14, the International Performance Measurement
-and Verification Protocol (IPMVP), and the Federal Energy Management
-Program (FEMP) for standard guidelines [@doe_mv_2008;
-@efficiency_valuation_organisation_international_2012;
+Measurement and Verification (M&V) is the process of quantifying
+savings, most commonly energy savings, but also potentially water and
+demand savings, from efficiency projects by comparing actual consumption
+against an established baseline, while adjusting for varying factors
+such as weather and occupancy. This process ensures that improvements in
+energy performance are accurately evaluated. In the United States,
+practitioners often refer to ASHRAE Guideline 14, the International
+Performance Measurement and Verification Protocol (IPMVP), and the
+Federal Energy Management Program (FEMP) for standard guidelines
+[@doe_mv_2008; @efficiency_valuation_organisation_international_2012;
 @ashrae_ashrae_2023]. These guidelines outline standardized methods for
@@
 [@mathieu_quantifying_2011] to project baseline energy consumption in
-the post-retrofit period. The difference between the counter-factual
-baseline and the measured intervention period energy consumption
-represents the energy savings. One of the key drawbacks of this method
-is its reliance on a two-year timeline to quantify savings, during which
-baseline measurements can become outdated due to changes in building
-performance caused by non-routine events unrelated to the intervention.
-Specifically, an M&V analyst should account for two types of baseline
-changes: (1) static changes, which are known building operational
-changes: such as renovation, equipment addition or removal, and changes
-related to tenant turnover; and (2) gradual changes, which often occur
-subtly and unknown, such as the incremental adoption of LED lighting or
-more efficient plug loads. Identifying and quantifying these changes are
-particularly labor-intensive and challenging. Consequently, those
-limitations reduce the feasibility of M&V and complicate the
+the post-retrofit time period (but without the retrofit itself). The
+difference between this counter-factual baseline and the measured energy
+consumption during the actual post-retrofit (intervention) period is the
+energy savings. However, other changes unrelated to the implemented
+strategies, such as occupancy patterns, can also contribute to the
+energy consumption difference. Although building analysts can apply
+routine adjustments to account for expected variations in independent
+variables commonly included in energy models, such as weather conditions
+and time, it is hard to accurately adjust for building performance
+change caused by non-routine events. Those events refer to less
+predictable and infrequent events that impact energy consumption but are
+unrelated to the specific intervention under study. Specifically, when
+using the conventional method, an M&V analyst should account for two
+additional types of baseline changes beyond routine adjustments: (1)
+known static changes to building operation, which are considered as
+typical non-routine events: including renovation, equipment addition or
+removal, and changes related to tenant turnover; and (2) gradual
+changes, which often occur subtly and unknown due to lack of accurate
+measurements to quantify, such as the incremental adoption of LED
+lighting or more efficient plug loads. Identifying and quantifying these
+changes are particularly labor-intensive and challenging. Consequently,
+those limitations reduce the feasibility of M&V and complicate the
 quantification of estimation uncertainty, thus impacting the accuracy
@@
 of this randomized design to the field of M&V in buildings was first
-used in [@raftery2018], which compared the energy performance of two
-different supply air temperature reset strategies in a building.
-Subsequent improvements to the M&V method include adding a sequential
-evaluation framework and defining stopping criteria to end the M&V
-period early after reaching a target level of uncertainty, using blocked
-randomization by weekday. The full framework is detailed in a previously
-published study [@raftery2024] along with a case study application to a
-real building, and other example use cases. In summary, this method
-provides M&V analysts with a randomized schedule that alternates between
-baseline and intervention implementation while ensuring balanced
-sampling across days of the week and seasons. For example, given a
-10-week M&V period for one intervention, the balanced randomized
-schedule would equally sample five Mondays with the baseline strategy in
-operation and five Mondays with the intervention strategy in operation.
+used in Raftery et al. [@raftery2018], which compared the energy
+performance of two different supply air temperature reset strategies in
+a building. Through random sampling, the study reached a conclusion of
+29% energy cost savings after just 6 months. Subsequent improvements to
+the M&V method included adding a sequential evaluation framework and
+defining stopping criteria to end the M&V period early after reaching a
+target level of uncertainty, using blocked randomization by weekday. The
+full framework is described in a previously published study
+[@raftery2024] along with a case study application to a real building,
+and other example use cases. In summary, this method provides M&V
+analysts with a randomized schedule that alternates between baseline and
+intervention implementation while ensuring balanced sampling across days
+of the week and seasons. For example, given a 10-week M&V period for one
+intervention, the balanced randomized schedule would equally sample five
+Mondays with the baseline strategy in operation and five Mondays with
+the intervention strategy in operation.
+
 The major limitation of this method is that it is only applicable to a
 subset of retrofit projects where interventions can be easily switched
-on and off, which is a small subset of all the interventions performed
-in buildings. However, for all applicable use cases, it allows analysts
-to detect energy savings sequentially shortly after the test begins. In
-addition, once the desired estimate uncertainty target is achieved,
-analysts can terminate the M&V study and switch to 100% intervention.
-The key advantage of randomization is that if control strategies are
-sampled with equal probability, the influence of other confounding
-factors such as occupancy change (a common non-routine event) and other
-more subtle long term changes in building energy use are likely to be
-evenly distributed among measurements, leading to a more accurate and
-unbiased assessment of the intervention effect.
+on and off (e.g., HVAC control retrofits), which is a small subset of
+all the interventions performed in buildings. However, for all
+applicable use cases, it allows analysts to detect energy savings
+sequentially shortly after the test begins. In addition, once the
+desired estimate uncertainty target is achieved, analysts can terminate
+the M&V study and switch to 100% intervention. The key advantage of
+randomization is that when control strategies are sampled with equal
+probability, the influence of confounding factors (e.g., occupancy
+change---a common non-routine event---and other more subtle long term
+changes in building energy use) are likely to be evenly distributed
+among measurements, leading to a more accurate and unbiased assessment
+of the intervention effect.
 
@@
 
-To demonstrate and differences between the two M&V methods, the Building
-Genome Dataset 2 (BGD2) is used, which is an extensive open-access
+To demonstrate the differences between the two M&V methods, we used the
+Building Genome Dataset 2 (BGD2), which is an extensive open-access
 dataset designed to advance research and development in building energy
@@
 simulation, and algorithm development [@miller_building_2020]. BGD2
-contains over 1000 buildings' metadata and realistic operational
-measurements between 2016 and 2017 from across North America and Europe,
-making it one of the most comprehensive collections of building-related
-data available for scientific use. The dataset includes various
-commercial building types such as offices, education facilities, public,
-and retail buildings, and provides detailed information on their
-physical characteristics (e.g. energy ratings, heating types and floor
-area) and hourly measurements of chilled and hot water, electricity, gas
-usage as well as site outdoor weather conditions. In this study, we
-mostly used the metered electricity and outdoor weather measurements for
-evaluating the M&V methods.
+contains over 1000 buildings' metadata and real operational measurements
+between 2016 and 2017 from across North America and Europe, making it
+one of the most comprehensive collections of building-related data
+available for scientific use. The dataset includes various commercial
+building types such as offices, education facilities, public, and retail
+buildings, and provides detailed information on their physical
+characteristics (e.g. energy ratings, heating types and floor area) and
+hourly measurements of chilled and hot water, electricity, gas usage as
+well as site outdoor weather conditions. In this study, we mostly used
+the metered electricity and outdoor weather measurements for evaluating
+the M&V methods.
 
@@
 use cases. For code compliance using the prescriptive option, 12 months
-of baseline and 12 months of post-retrofit measurements are required. In
-addition, the expected savings should be larger than 10% and the
-baseline energy model fitting accuracy should have CV(RMSE) lower than
-25% for 12-month data collection. Thus, most research related to M&V for
-the whole-building approach focuses on the accuracy of baseline
-modeling, exploring model performance from simple regression models to
-more complex machine learning techniques. One study reviewed various
-models suitable for M&V applications as well as selected input features
-[@alrobaie_review_2022] and another study provided a definitive
-methodology to apply machine learning models for M&V use cases
-[@gallagher_development_2018]. In addition, a few studies investigated
-the critical performance metrics to evaluate the developed baseline
-models [@granderson_automated_2015; @granderson_development_2014] and
-compared a variety of models using those metrics
-[@granderson_accuracy_2016]. These studies made significant
-contributions by emphasizing the uncertainty associated with the
-model-fitting process, a key factor in accurately determining energy
-savings. Furthermore, other researchers addressed this issue by
-leveraging statistical formulation and inference to improve baseline
-energy models [@burkhart_measurement_2014; @heo_calibration_2012;
-@walter_uncertainty_2014]. However, a gap still remains in the
-literature regarding the rigorous quantification of uncertainties
-directly associated with calculated savings, for instance, accounting
-for the potential bias that the baseline model might deteriorate (i.e.,
-become 'stale') over the extended period of pre- and post-analysis.
+of baseline and 12 months of post-retrofit measurements are required.
+Moreover, the expected savings should be larger than 10% and the
+baseline energy model fitting accuracy should have a Coefficient of
+Variation of the Root Mean Square Error, CV(RMSE), lower than 25% for
+12-month data collection. Thus, most research related to M&V for the
+whole-building approach focuses on the accuracy of baseline modeling,
+exploring model performance from simple regression models to more
+complex machine learning techniques. Alrobaie and Krarti
+[@alrobaie_review_2022] reviewed various models suitable for M&V
+applications as well as selected input features. Gallagher et al.
+[@gallagher_development_2018] provided a well-thought-out methodology to
+apply machine learning models for M&V use cases. Researchers at LBNL
+investigated the critical performance metrics to evaluate the developed
+baseline models [@granderson_automated_2015;
+@granderson_development_2014] and compared a variety of models using
+those metrics [@granderson_accuracy_2016]. These studies emphasized the
+uncertainty associated with the model-fitting process, a key factor in
+accurately determining energy savings. Furthermore, other researchers
+addressed this issue by leveraging statistical formulation and inference
+to improve baseline energy models [@burkhart_measurement_2014;
+@heo_calibration_2012; @walter_uncertainty_2014]. However, a gap still
+remains in the literature regarding how to rigorously quantify the
+uncertainties directly associated with calculated savings. For instance,
+this includes accounting for the potential bias that the baseline model
+might deteriorate over time, as well as the model no longer accurately
+representing the building energy behavior due to changes in operation,
+occupancy, equipment performance, etc.
 
@@
 
-Changes in a building that influence its energy usage and are unrelated
-to the intervention being studied by the M&V method are commonly
-referred to as non-routine events. Current guidelines define non-routine
-events as 'static factors' that need to be adjusted after projecting the
-baseline in the post-retrofit period [@ashrae_ashrae_2023;
+As mentioned earlier, non-routine events are changes in a building that
+influence its energy usage but are unrelated to the intervention being
+studied by the M&V. Current guidelines define non-routine events as
+'static factors' that need to be adjusted after projecting the baseline
+into the post-retrofit period [@ashrae_ashrae_2023;
 @efficiency_valuation_organisation_international_2012]. A common
@@
 operating conditions (e.g. set points, lighting, and ventilation
-levels). However, current standards or guidelines only provide a general
-approach for consideration. For example, IPMVP requires the facility
-owner and the M&V analyst to periodically perform inspections of all
-equipment and operations during the reporting period, which is
-labor-intensive and error-prone. ASHRAE Guideline 14 recommends
-performing engineering calculations or computer software simulations to
-adjust the post-retrofit baseline. Additionally, it is relatively rare
-to have access to all measurements needed for adjustments, such as
-occupancy, and thus the analyst normally assumes those factors remain
-unchanged throughout the study. For example, some studies use linear
-interpolation to estimate counter-factual baseline for the demand
-response program [@beil_round-trip_2015; @keskar_commercial_2020].
-Consequently, one study points out that it is inaccurate to assume no
-change in the operating conditions during the response period (Huang,
-Katipamula, and Lutes 2023). Other studies realized such limitation in
-field study, and emphasized the importance of requiring matched groups
-to control for exogenous factors beyond weather differences when
-comparing between baseline and intervention
+levels). However, current standards or guidelines only provide a
+reminder that non-routine events should be reported and properly
+adjusted. For example, IPMVP requires the facility owner and the M&V
+analyst to periodically perform inspections of all equipment and
+operations during the reporting period, which is labor-intensive and
+error-prone. ASHRAE Guideline 14 recommends performing engineering
+calculations or computer software simulations to adjust the
+post-retrofit baseline. Additionally, it is relatively rare to have
+access to all measurements needed for adjustments, such as occupancy,
+and thus the analyst normally assumes those factors remain unchanged
+throughout the study. For example, some studies use linear interpolation
+to estimate counter-factual baseline for the demand response program
+[@beil_round-trip_2015; @keskar_commercial_2020]. Consequently, it is
+inaccurate to assume no change in the operating conditions during the
+response period [@huang_experimental_2023]. This is a limitation in a
+field study, and it is important to require matched groups to control
+for exogenous factors beyond weather (which is usually taken into
+account) when comparing between baseline and intervention
 [@demand_side_analytics_population_2022; @huang_experimental_2023].
 
-Overall, those non-routine events mostly refer to clearly observable
+Overall, those non-routine events mainly refer to clearly observable
 changes, but there are also more gradual and subtle changes in
@@
 @zhai_full-scale_2017]. We found a similar concept briefly defined in
-Guideline 14 as the 'load creep' [@ashrae_ashrae_2023] during multi-year
-M&V project. However, currently there is no detailed instruction
-provided. If M&V analysts are unaware of such changes and lack an
-appropriate adjustment method (e.g., replacing filters before the
-intervention begins), the savings could be overestimated as the
-decreased energy use is incorrectly attributed to the intervention
-rather than the filter replacement. An operational drift has not been
-formally defined in the literature or standards, but in this study, we
-consider it as a special type of non-routine event, as it also
-represents a change in facility operation that occurs gradually over a
-longer period of time.
+Guideline 14 as the 'load creep' [@ashrae_ashrae_2023] during a
+multi-year M&V project. However, currently there are no detailed
+instructions provided on how to account for them. If M&V analysts are
+unaware of such changes and lack an appropriate adjustment method (e.g.,
+replacing filters before the intervention begins), the savings could be
+overestimated as the decreased energy use is incorrectly attributed to
+the intervention rather than the filter replacement. An operational
+drift has not been formally defined in the literature or standards; in
+this study, we consider it as a special type of non-routine event, as it
+also represents a change in facility operation that occurs gradually
+over a longer period of time.
 
@@
 
-As mentioned, the goal of an M&V project is to determine the effect,
-typically energy savings, of some intervention in the building. In this
-study, we focus on switchable interventions and an example of such an
-intervention is a control strategy that adjusts the chilled water
-plant's supply water temperature, which can be commonly found in the
-literature [@duarte_field_2023; @lee_simulationoptimization_2012;
-@qiu_chilled_2022; @jin2007; @taylor_optimizing_2012]. In our case, we
-made it even simpler by adjusting the setpoint based on outdoor weather
-conditions (see description in Section
-\@ref(apply-virtue-intervention)). By conducting such an analysis, we
+One goal of an M&V project is to determine the effect, typically energy
+savings, of an intervention in the building. In this study, we focus on
+switchable interventions. An example of such an intervention is a
+control strategy that strategically adjusts the chilled water plant's
+supply water temperature with proven ability to save energy
+[@duarte_field_2023; @lee_simulationoptimization_2012;
+@qiu_chilled_2022; @jin2007; @taylor_optimizing_2012]. We used a simple
+version of the control strategy in which the setpoint is adjusted based
+on outdoor temperature (see description in Section
+\@ref(apply-virtual-intervention)). By conducting such an analysis, we
 aim to:
 
-1.  Determine, using a large sample of buildings, how much more
-    accurately the randomized M&V method would estimate the savings of
-    the proposed intervention compared to the conventional M&V, and how
-    much more quickly it would reach a result.
+1.  Determine how much more accurately the randomized M&V method would
+    estimate the savings of the intervention compared to the
+    conventional M&V, and how much more quickly it would reach a result.
 
@@
     the conventional method when static non-routine events and
-    operational drift are present in the dataset. The BDG2 dataset
-    contains real-world building energy usage data collected over two
-    years, which inherently contains varying degrees of usage changes.
-    By virtually introducing intervention effects into the existing
-    dataset, we can calculate the ground-truth energy savings, which
-    serve as a reference point for method comparison.
-
-3.  Open-source implementation of the proposed randomized M&V method. We
-    ensured the reproducibility of the method by making the analysis
-    code open source, including randomized schedule generation,
-    sequential statistical analysis, energy modeling and normalized
-    saving calculation. In addition, we also included code for extended
-    use cases of the randomized method, such as changing sampling ratio
-    and sampling intervals. Using the available open resources, building
-    analysts should be able to seamlessly integrate and apply them in
-    their own M&V projects.
+    operational drift are present in the dataset.
+
+3.  Make the implementation of the proposed randomized M&V method
+    open-source.
+
+The significance of this study is that the BDG2 dataset contains
+real-world building energy usage data collected over two years, which
+inherently contains varying degrees of usage changes. In addition, by
+virtually introducing intervention effects into the existing dataset, we
+can calculate the ground-truth energy savings, which serve as a
+reference point for method comparison. Furthermore, we ensured the
+reproducibility of the method by making the analysis code open source,
+including randomized schedule generation, sequential statistical
+analysis, energy modeling and normalized saving calculation. This also
+covers the extended discussion below on sensitivity analysis and other
+use cases of the randomized method, such as changing sampling ratio and
+sampling intervals. Using the available open resources, building
+analysts should be able to seamlessly integrate and apply them in their
+own M&V projects.
 
@@
 We outlined the methodology of the study in Figure \@ref(fig:flowchart)
-and extended several key components in this section.
-
-```{r flowchart, fig.cap = "Flow chart showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="70%"}
+and extended several key components in this section. As a summary, we
+utilized the historical measurements from a dataset that contains a
+large number of real commercial buildings as the pre-retrofit baseline
+and simulated an intervention strategy. Then, we applied both
+conventional and randomized M&V methods to the modified dataset to
+estimate intervention savings, and compare which one produces results
+closer to the simulated intervention effect.
+
+```{r flowchart, fig.cap = "Flow chart showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="60%"}
 
@@
 We extracted the electricity measurements from the BDG2 dataset and
-filter out all qualified buildings based on the following criteria:
+filtered out all qualified buildings based on the following criteria:
 
@@
 3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
-    decrease in building electricity usage in the second year should be
-    less than 25% of that in the first year.
+    decrease in building electricity usage ($E$) in the second year
+    ($E_{2017}$) should be less than 25% of that in the first year
+    ($E_{2016}$).
 
@@
 
-For all qualified buildings, we also create a feature to identify
-buildings with very stable electricity usage overall by identifying
-those with no statistically significant difference (P-value \> 0.05)
-between their two-year electricity usage. Applying this filter to real
+As a result, we kept a total of 639 buildings from 10 different
+locations for the subsequent analysis, and we included a more detailed
+summary of the building features in the supplementary material. For all
+qualified buildings, we also create a feature to identify buildings with
+very stable electricity usage overall by identifying those with no
+statistically significant difference (P-value \> 0.05) between their
+pre- and post-retrofit electricity usage. Applying this filter to real
 buildings, only 66 buildings are labeled in this 'stable' subset out of
-a total of more than 600 buildings. While we did not adjust for weather
-differences between the two years when generating this high-level
-feature, this result highlights that the assumption that buildings are
-typically stable over the timeframe involved in conventional M&V methods
-is rarely true. As mentioned in \@ref(non-routine-events-impact), such
-variability is largely associated with non-routine events, especially
-operational drifts since the assessment, even under ideal situations,
-lasts for two years. And according to the statistics shown here, it is a
-typical case observed in real buildings applications.
-
-## Apply virtue intervention
-
-Figure \@ref(fig:chwst) shows the algorithm for the proposed control
-intervention that resets the chiller supply temperature based on the
-outdoor weather conditions, which can be commonly found in the
+a total of more than 600 buildings, corresponding to 11%. While we did
+not adjust for weather differences between the two years when generating
+this high-level feature, this result highlights that the assumption that
+buildings are typically stable over the timeframe involved in
+conventional M&V methods is rarely true; in 89% of the cases this
+assumption is not valid. As mentioned in Section
+\@ref(non-routine-events-impact), such variability is largely associated
+with non-routine events---especially operational drifts---since the
+assessment, even under ideal situations, lasts for two years. According
+to the statistics shown here, it is a typical case observed in real
+buildings applications.
+
+## Apply **virtual** intervention
+
+The proposed control intervention resets the chiller supply temperature
+based on the outdoor temperature, which can be commonly found in the
 literature [@lee_chilled_2022; @congradac_recognition_2012]. For both
@@
 existing measurements from the dataset, operates with a constant water
-supply temperature. The intervention strategy, as illustrated in the
-figure, adjusts the water supply temperature dynamically, resetting it
-from 7°C to 12°C.
-
-```{r chwst, fig.cap = "Proposed intervention strategy: chilled water supply temperature reset based on outdoor temperature.", fig.width=8, fig.height=4}
-
-data.frame(
-  OAT = c(10, 15, 25, 30),  
-  Baseline = c(6, 6, 6, 6),      
-  Intervention = c(12, 12, 7, 7)
-  ) %>% 
-  pivot_longer(cols = c("Baseline", "Intervention"), names_to = "strategy", values_to = "value") %>% 
-  ggplot(aes(x = OAT, y = value, color = strategy)) +
-  geom_line(linewidth = 1.2) +
-  scale_color_manual(values = ls_colors) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = c(4, 6, 8, 10, 12),
-                     limits = c(5, 12.5),
-                     labels = number_format(suffix = " °C")) +
-  scale_x_continuous(expand = c(0, 0), 
-                     breaks = c(5, 10, 15, 20, 25, 30), 
-                     labels = number_format(suffix = " °C")) +
-  geom_vline(xintercept = 10, lty = "dashed", color = "red") +
-  annotate(geom = "text", x = 7.5, y = 8, label = "No chiller\noperation\nbelow 10 °C", size = 5) +
-  coord_cartesian(xlim = c(5, 30)) +
-  labs(x = "Outdoor temperature", 
-       y = "Chilled water supply temperature", 
-       color = NULL, 
-       title = "Chilled water supply temperature reset intervention") +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.direction = "horizontal",
-        legend.position = "bottom",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-```
+supply temperature. The intervention strategy adjusts the water supply
+temperature dynamically, resetting it from 7°C to 12°C. We included a
+visual summary of the algorithm in the supplementary material.
 
 We used a simple relationship to map the chilled water supply
-temperature reset to the electrical energy savings: $$
-\text{savings} = mean(E_{CHW}) \times 25\% \times \text{perc_savings}
+temperature reset to the electrical energy savings:$$
+\text{savings} = mean(E_{CHW}) \times 25\% \times \text{perc}
 $$
@@
 building's total electricity consumption, and the chilled water plant
-further consumes around 50% of the HVAC electricity. While this
-assumption largely simplifies the diverse energy usage across various
-building types, for the scope of this paper, we assume that 25% of the
-total building electricity is used by the chilled water plant, $E_{CHW}$
-[@us2012commercial]. Typically, the savings from an intervention are not
-proportional to the building's hourly electricity usage, which is
-generally the challenge for M&V. To address this, we mapped the
-resulting electricity savings as a percentage of the plant's normal
-operation, calculated as its mean electricity usage over the two-year
-period. This percentage is influenced by factors such as outdoor
-temperature ($OAT$), intervention supply water temperature
-($T_{interv}$), baseline supply water temperature ($T_{base}$) and hour
-of the day ($\delta_{occ}$, binary indicator whether during peak hours
-from 9 AM to 4 PM).
+$E_{CHW}$ further consumes around 50% of the HVAC electricity, or 25% of
+the total building electricity [@us2012commercial]. We recognize that
+this assumption, used for the scope of this analysis, largely simplifies
+the diverse energy usage across various building types. Typically, the
+savings from an intervention are not proportional to the building's
+hourly electricity usage, which is generally a challenge for M&V. To
+address this, we mapped the resulting electricity savings as a
+percentage of the plant's normal operation, calculated as its mean
+electricity usage over the two-year period. This percentage is
+influenced by factors such as outdoor temperature ($OAT$), intervention
+supply water temperature ($T_{interv}$), baseline supply water
+temperature ($T_{base}$) and hour of the day ($\delta_{occ}$, binary
+indicator whether during peak hours from 9 AM to 4 PM).
 
 $$
-\text{perc_savings = }
+\text{perc = }
 \begin{cases}
@@
 Parameters and their pre-defined values are summarized in the table
-below. For simplicity, those parameters were not rigorously calibrated
-for each building and were applied uniformly in the filtered dataset.
-Interested readers can also change the parameters in our open-source
-code to simulate different scenarios.
+below, they were obtained from a trial supply water setpoint reset
+experiment on a campus office building [@zou_quantifying_2025]. For
+simplicity, those parameters were not rigorously calibrated for each
+building and were applied uniformly in the filtered dataset. Interested
+readers can also change the parameters in our open-source code to
+simulate different scenarios.
 
 | Parameter              | Description                                                      | Value |
-|----------------|-----------------------------------------|----------------|
+|------------------|------------------------------------|------------------|
 | $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
@@
 
-## Run M&V methods
+## M&V methods and results analysis
 
@@
 intervention timeline. The randomized method defines sampling
-requirements as:
-
--   Use a daily sampling interval with the sampling time at midnight
-    each day.
-
--   Block by day of the week with a block period of 12 weeks.
-
-and stopping criteria as:
-
--   A minimum and maximum of 12 and 108 weeks, respectively. The
-    randomized schedule covers the entire two-year period but the
-    stopping criteria enable an early stop at the end of the satisfied
-    blocking period.
-
--   At least 80% of the dry-bulb temperature range in the annual TMY
-    data is sampled by both strategies.
-
--   90% confidence that energy savings exceed or do not exceed 0% using
-    the sequential probability ratio test (SPRT). Medium effect size (d
-    = 0.5) quantified by Cohen's d and calculated SPRT statistics either
-    fall below the lower threshold or exceed the upper threshold.
-
--   Sample strategies at an equal sampling ratio (50% baseline, 50%
-    intervention).
-
-# Results
-
-The accuracy ($Acc$) is calculated as: $$
-Acc = \frac{E_{baseline} - E_{intervention}}{E_{baseline}}-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
-$$where $E$ indicates true electricity usage ($E_{intervention}$ is the
-'true' electricity usage calculated from Figure \@ref(fig:chwst) and
-above equations) and $E'$ indicates the estimated electricity usage
-either through the conventional method or the new randomized method. We
-present the distribution of savings estimation accuracy using boxen
-plots, also known as letter plots, which is an advanced variation of the
-box plot designed to extend beyond the interquartile range (IQR) by
+requirements as (a) use a daily sampling interval with the sampling time
+at midnight each day and (b) block by day of the week with a block
+period of 12 weeks.
+
+We used the following four stopping criteria: (a) a minimum and maximum
+of 12 and 108 weeks, respectively. The randomized schedule covers the
+entire two-year period but the stopping criteria enable an early stop at
+the end of the satisfied blocking period. (b) At least 80% of the
+dry-bulb temperature range in the annual TMY data is sampled by both
+strategies. (c) 90% confidence that energy savings exceed or do not
+exceed 0% using the sequential probability ratio test (SPRT). Medium
+effect size (d = 0.5) quantified by Cohen's d and calculated SPRT
+statistics either fall below the lower threshold or exceed the upper
+threshold. (d) sample strategies at an equal sampling ratio (50%
+baseline, 50% intervention).
+
+To summarize the M&V results and associated error to the methods, we
+present the distribution of savings estimation error using boxen plots,
+also known as letter plots, which is an advanced variation of the box
+plot designed to extend beyond the interquartile range (IQR) by
 progressively dividing the data into smaller percentiles, revealing more
@@
 representing 12.5% to 87.5%, followed by 25% to 75%. In addition, for
-each boxen plot, we also showed the mean value of the distribution on
+each boxen plot, we also showed the median value of the distribution on
 top for reference.
 
-Figure \@ref(fig:abstract) shows the overall results of M&V methods
-comparison and for clarity, we plotted conventional M&V results in grey,
-randomized M&V results after satisfying all stopping criteria in green,
-and the randomized M&V results sampled for the entire 24 months in blue.
+# Results
+
+The savings estimation error ($Err$) is calculated as: $$
+Err = \frac{E_{baseline} - E_{intervention}}{E_{baseline}}-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
+$$where $E$ indicates true electricity usage ($E_{intervention}$ is the
+'true' electricity usage calculated from Eq. (1) and (2)), and $E'$
+indicates the estimated electricity usage either through the
+conventional method or the new randomized method. Figure
+\@ref(fig:abstract) shows the overall results of M&V methods comparison
+and for clarity, we plotted conventional M&V results in grey, randomized
+M&V results after satisfying all stopping criteria in green, and the
+randomized M&V results sampled for the entire 24 months in blue.
 Subplots (a) and (b) calculate the savings estimation error as the
-absolute deviation from the true savings (i.e. $|Acc|$) and show the
+absolute deviation from the true savings (i.e. $|Err|$) and show the
 results distribution using the conventional M&V method in the first
@@
 meteorological weather of the building location after using a
-time-of-week-temperature (TOWT) model [@mathieu_quantifying_2011]. As a
-result, the conventional M&V method exhibits a median deviation of 5% in
-savings estimation, whereas the randomized method demonstrates
-significantly smaller deviations. If the analyst stops immediately after
-satisfying all stopping criteria, the deviation is reduced to
-approximately 1.5%. Extending the M&V period to match the same time
-period of the conventional method (24 months) further improves accuracy
-to 0.7% deviation. Furthermore, by comparing subplots (a) and (b), we
-observed a minor improvement in estimation accuracy when a model is
-fitted to the data to account for weather differences. The model fitting
-is most beneficial for the shorter dataset (randomized and stopping when
-all criteria are met, typically 24-36 weeks of data), where it reduces
-deviation from 1.5% to 1.2% (around 20% improvement). This is because
-the weather conditions between the first and second year for each of the
-locations are relatively similar on average over this time period, and
-thus the effect of adjusting for temperature as an independent variable
-in the model is relatively small. We found 83% of all filtered buildings
-show less error when using the randomized M&V, and by comparing the
-error distribution in subplot (b), we noticed the medium deviation drops
-from 4.6% to 1.2% (around 74% improvement) at the time of satisfying all
-stopping criteria. Subplot (c) compares the overall timeline of the two
-M&V methods. The results show that almost all buildings achieve accurate
-M&V results and meet the stopping criteria within 24 weeks (for this
-intervention), with the remainder meeting all stopping criteria within
-36 weeks. The additional 12 weeks are usually driven by a need to span
-sufficient weather conditions representative of a full year of data.
-Meanwhile, the red arrow highlights the excess time required by the
-conventional M&V method (24 months). Thus, the randomized M&V method
-finds more accurate results in approximately a quarter of the time
-required for conventional M&V. Even if it was the case that an existing
-12-month baseline data was available at the start of the M&V process,
-randomized M&V would still obtain a more accurate result, in less time,
-than obtaining the 12-month retrofit dataset required by conventional
-M&V.
-
-```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period).", fig.height=10, fig.width=9}
+time-of-week-temperature (TOWT) model [@mathieu_quantifyin
+
+The conventional M&V method exhibits a median deviation of 4.8% in
+savings estimation and 4.6% after model normalization, whereas the
+randomized method demonstrates significantly smaller deviations. If the
+analyst stops immediately after satisfying all stopping criteria, the
+deviation is reduced to approximately 1.5% based on measured weather and
+1.2% after normalization. Extending the M&V period to match the same
+time period of the conventional method (24 months) further improves
+accuracy to 0.7% deviation and 0.6% after normalization. Overall, we
+found 83% of all filtered buildings show less error when using the
+randomized M&V, and subplot (b) indicates around 74% improvement in
+accuracy (median error dropped from 4.6% to 1.2%) at the time of
+satisfying all stopping criteria.
+
+Furthermore, by comparing subplots (a) and (b), we observed a minor
+improvement in estimation accuracy when a model is fitted to the data to
+account for weather differences. The model fitting is most beneficial
+for the shorter dataset (randomized and stopping when all criteria are
+met, typically 24-36 weeks of data), where it reduces deviation from
+1.5% to 1.2% (around 20% improvement). This is because the weather
+conditions between the first and second year for each of the locations
+are relatively similar on average over this time period, and thus the
+effect of adjusting for temperature as an independent variable in the
+model is relatively small.
+
+Subplot (c) compares the overall timeline of the two M&V methods. The
+results show that almost all buildings achieve accurate M&V results and
+meet the stopping criteria within 24 weeks (for this intervention), with
+the remainder meeting all stopping criteria within 36 weeks. The
+additional 12 weeks are usually driven by a need to span sufficient
+weather conditions representative of a full year of data. The red arrow
+highlights the excess time required by the conventional M&V method (24
+months). Thus, the randomized M&V method finds more accurate results in
+approximately a quarter of the time required for conventional M&V. Even
+if it was the case that an existing 12-month baseline data was available
+at the start of the M&V process, randomized M&V would still obtain a
+more accurate result, in less time, than obtaining the 12-month retrofit
+dataset required by conventional M&V.
+
+```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period).
+Subplot (a) shows the savings estimated from measured weather conditions, and subplot b) shows the savings normalized on the typical meteorological weather of the building location after using a time-of-week-temperature (TOWT) model.
+Subplot (c) shows at the end of each blocking period the number of buildings satisfying all stopping criteria.
+The figure shows the randomized M&V method can provide much more accurate savings estimation than the convention method even at 36 weeks.", fig.height=10, fig.width=9}
 
@@
 
-Figure \@ref(fig:meanerr) shows the biased (instead of absolute)
-estimated deviation, highlighting that the uncertainty range associated
-with the conventional method is significantly larger compared to even
-the randomized method that only takes 36 weeks, according to subplot (c)
-in Figure \@ref(fig:abstract).
-
-```{r meanerr, fig.cap="Distribution of estimated deviation from the ground-truth savings calculated by the two M&V methods (red dashed line highlighting 0% error)", fig.height=7, fig.width=9}
+In the supplementary material, we also showed the signed/biased (instead
+of absolute) estimated deviation, highlighting that the uncertainty
+range associated with the conventional method is significantly larger
+compared to even the randomized method that only takes 36 weeks,
+according to subplot (c) in Figure \@ref(fig:abstract).
+
+```{r meanerr, fig.cap="Distribution of estimated deviation from the ground-truth savings calculated by the two M&V methods (red dashed line highlighting 0% error)", fig.height=7, fig.width=9, include = FALSE}
 # plot for the stable subset
@@
        x = NULL, 
-       y = "Estimated error in fractional savings", 
+       y = "Signed error in fractional savings", 
        subtitle = str_glue("All {A_building} buildings with measured weather")) +
@@
        x = NULL, 
-       y = "Estimated error in fractional savings", 
+       y = "Signed error in fractional savings", 
        subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
@@
           labels = c("a)", "b)")) +
-  plot_annotation(title = "Estimated error in savings estimation by different M&V methods")
+  plot_annotation(title = "Signed error in savings estimation by different M&V methods")
 
@@
 
-Additionally in the literature, we found another M&V study using over
-500+ commercial buildings (on a different open-source dataset), but
-their study scope was to compare the predictive performance of a variety
-of M&V baseline models from simple weekly mean to more complex machine
-learning methods using the normalized mean bias error (NMBE)
+Granderson et al. (2016), using over 500+ commercial buildings (on a
+different open-source dataset), compared the predictive performance of a
+variety of M&V baseline models from simple weekly mean to more complex
+machine learning methods using the normalized mean bias error (NMBE)
 [@granderson_accuracy_2016]. Since we also calculated the normalized
 savings estimation error, we compared our normalized accuracy for 600+
-buildings with the TOWT model and TMY weather againts their TOWT model
-prediction assessment in the following table, highlighting that the
-model and overall building datasets are performing similarly.
+buildings with the Time-Of-Week and Temperature (TOWT) model and Typical
+Meteorological Year (TMY) weather against their TOWT model prediction
+assessment in Table 2. The model and overall building datasets are
+performing similarly.
 
 |                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
-|------------------|-------------------------------------|------------------|
+|-------------------|-----------------------------------|-------------------|
 | 25th percentile | -5.85                                               | `r percentiles$Q1`     |
@@
 
-As mentioned in Section \@ref(filter-and-clean-dataset), the majority of
+As mentioned in Section \@ref(filter-and-clean-dataset), he majority of
 the selected buildings are affected by operational drift, which is the
@@
 M&V run on the original dataset prior to the application of the reset
-control. In other words, there *is no intervention applied here*, so the
+control. In other words, there is no intervention applied here, so the
 most reliable M&V method will be the one that detects the closest to 0
-kW savings using the measurements. Therefore, the $Acc$ here are
+kW savings using the measurements. Therefore, the $Err$ here are
 calculated as:
@@
 $$
-Acc = 0-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
+Err = 0-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
 $$
@@
 two-year period) produced consistent results. We included the absolute
-deviation ($|ACC|$) figure in the supplementary material. Similarly to
+error ($|Err|$) figure in the supplementary material. Similarly to
 Figure \@ref(fig:abstract), stopping when meeting all stopping criteria
@@
 deviation around -1.25%, our analysis on the conventional method shows
-that the typical deviation is -1.5%, meaning that the typical building
-is decreasing energy use slightly in this dataset in the absence of any
-known intervention. Thus, if a building analyst were to apply an
-intervention to this building and use conventional M&V to measure the
-effect, they would typically 'detect' 1.5% energy savings even when the
-intervention had no actual impact on energy performance.
+that the typical deviation is -1.5%, meaning that this method would
+predict that, in the absence of any known intervention, the typical
+building is decreasing energy use slightly in this dataset. Meanwhile,
+through random sampling, the influence of non-routine events is much
+closer to 0 with narrower uncertainty range.
 
@@
 
-# Discussion
+# Sensitivity analysis
 
@@
 is a normalized measure, thus it enables direct comparison across
-different modeling results. Since we have the ground-true baseline
+different modeling results. Since we have the ground-truth baseline
 measurements in the post-retrofit period, we can assess the error caused
@@
 difference between the true post-retrofit baseline and adjusted baseline
-through TOWT modeling and since the quantity of the training set is the
+through TOWT modeling; since the quantity of the training set is the
 same, this highlights the impact of the data sampling technique. For
@@
 these events since it samples only 50% of the pre-retrofit baseline
-throughout a blocking period. Consequently, the figure shows, there is a
+throughout a blocking period. Consequently, the figure shows there is a
 noticeable improvement in modeling accuracy when using the randomized
@@
 In some cases, particularly those involving large thermal mass and
-active thermal energy storage systems or operating continuously, there
-can be a substantial time-lagged effect associated with interventions.
-This means that the effect of one control strategy may persist, or
-'carryover', after switching to another. To mitigate such carryover
-effects, building analysts may choose to drop non-consecutive days in
-the switchback experiment. For example, consider a heat pump water
-heater where the tank is charged at very different times of the day by
-the intervention strategy compared to the baseline strategy. When
-switching strategy, the tank may be much warmer or cooler than typical
-for that strategy, thus influencing subsequent measurements,
-particularly the performance measured on the first day (i.e., a
-non-consecutive day). In such cases, increasing the sampling interval
-(e.g., sample strategies every 2 or 3 days instead of daily) and
-discarding measurements from those non-consecutive days can eliminate
-carryover. However, this approach reduces the total data collected: for
-instance, sampling every 3 days and dropping days with non-consecutive
-control strategies would drop 1/6 of measurements on average. Sampling
-every 2 days means dropping a quarter of the measurements. Therefore, by
-increasing the sampling interval, there is an accuracy penalty due to a
-reduction in randomization (how often samples occur), and when
-decreasing the sampling interval, there is an accuracy penalty from
-dropping more non-consecutive days. The dataset contains no specific
-information about the buildings' thermal mass condition, but to quantify
-such trade-off, we compared the accuracy of different sampling intervals
-and the consequence of dropping the non-consecutive days at the time
-when randomized M&V produces a result (after satisfying all stopping
-criteria) in Figure \@ref(fig:absinterval).
+active thermal energy storage systems or operating continuously without
+shutting down for switchover, there can be a substantial time-lagged
+effect associated with interventions. This means that the effect of one
+control strategy may persist, or 'carryover', after switching to
+another. To mitigate such carryover effects, building analysts may
+choose to drop a predetermined number of non-consecutive days in the
+switchback experiment. For example, consider a heat pump water heater
+where the tank is charged at very different times of the day by the
+intervention strategy compared to the baseline strategy. When switching
+the strategy, the tank may be much warmer or cooler than typical for
+that strategy, thus influencing subsequent measurements, particularly
+the performance measured on the first day (i.e., a non-consecutive day).
+In such cases, increasing the sampling interval (e.g., sample strategies
+every 2 or 3 days instead of daily) and discarding measurements from
+those non-consecutive days can eliminate carryover. However, this
+approach reduces the total data collected: for instance, sampling every
+3 days and dropping days with non-consecutive control strategies would
+drop 1/6 of measurements on average. Sampling every 2 days means
+dropping a quarter of the measurements. Therefore, by increasing the
+sampling interval, there is an accuracy penalty due to a reduction in
+randomization (how often samples occur); when decreasing the sampling
+interval, there is an accuracy penalty from dropping more
+non-consecutive days. The dataset contains no specific information about
+the buildings' thermal mass condition. So, to quantify such trade-off,
+we compared the accuracy of different sampling intervals and the
+consequence of dropping the non-consecutive days at the time when
+randomized M&V produces a result (after satisfying all stopping
+criteria). These results are shown in Figure \@ref(fig:absinterval).
 
@@
 
-To quantify the impact of reduced randomness, all measurements are kept
+To quantify the impact of reduced randomness, all measurements were kept
 when calculating the distribution shown in the darker-colored set. The
 results suggest that without dropping non-consecutive days in subplot
-(a), a gradual increase in error from 1.5% to 2.3% (1.2% to 1.9% with
-weather normalization in subplot (b) when the sampling interval
-increases from daily to weekly intervals. This is due to the fact that
-increasing the sample interval means reducing randomization. To quantify
-the combined impact of data loss due to preventing the carryover effect,
-we generated the lighter-colored set, which drops all non-consecutive
-days. As a result, we noticed an increase in estimation error when
-sampling daily to 2.4% (to 1.9% when TOWT is fitted), which led to a
-similar accuracy compared to weekly intervals. Therefore, considering
-this trade-off, we recommend using a two-day sampling interval (or
-three-day sampling interval if the building owner prefers less frequent
-switchover) if the carryover effect is likely but expected to last less
-than one day, and then normalizing the estimation via energy modeling.
+(a), there was a gradual increase in error from 1.5% to 2.3% (1.2% to
+1.9% with weather normalization in subplot (b)) when the sampling
+interval increases from daily to weekly intervals. This is due to the
+fact that increasing the sample interval means reducing randomization.
+
+To quantify the combined impact of data loss due to preventing the
+carryover effect, we generated the lighter-colored set, which drops all
+non-consecutive days. As a result, we noticed an increase in estimation
+error when sampling daily to 2.4% (to 1.9% when TOWT is fitted), which
+led to a similar accuracy compared to weekly intervals. Therefore,
+considering this trade-off, we recommend using a two-day sampling
+interval (or three-day sampling interval if the building owner prefers
+less frequent switchover) if the carryover effect is likely but expected
+to last less than one day, and then normalizing the estimation via
+energy modeling.
+
 The supplementary material contains similar figures showing biased
@@
 
-In our previous study [@raftery2024], we recommended that building
-owners adjust the sampling ratio after meeting all stopping criteria, as
-continuing baseline sampling helps maintain M&V accuracy with limited
-impact on the energy savings actually achieved by the building. To
-further examine its impact on M&V accuracy, Figure \@ref(fig:abscont)
-compares two sampling ratios (20%/80% and 10%/90%) and two associated
+The sampling ratio determines the likelihood that the M&V analyst will
+apply the baseline or intervention, and we recommend using an equal
+sampling ratio at the outset of the M&V process. In our previous study
+[@raftery2024], we also recommended that building owners adjust the
+sampling ratio by reducing the amount of baseline sampling after meeting
+all stopping criteria. This is because, although sampling the baseline
+allows for tracking changes within buildings, it can also limit
+opportunities for building owners to implement the intervention, thereby
+reducing potential savings. To further examine its impact on M&V
+accuracy, Figure \@ref(fig:abscont) compares two sampling ratios
+(baseline/intervention: 20%/80% and 10%/90%) and two associated
 scenarios with the previously shown results when sampling at 50%/50%.
@@
 contrast, the "Randomized" group uses the indicated ratio from the
-outset and continues throughout the two year. Subplot (a) shows that
-absolute error is significantly lower without the initial 50%/50%
+outset and continues throughout the two years. The figure below shows
+that absolute error is significantly lower without the initial 50%/50%
 sampling, due to baseline drift and non-routine events that most
 buildings experience over two years (as noted in Section
-\@ref(filter-and-clean-dataset)). Subplot (b) confirms this finding even
-after weather normalization.
+\@ref(filter-and-clean-dataset)). This is also the case even with
+weather normalization, the results of which are included in the
+supplementary material.
 
@@
 
-As shown in the previous Figure \@ref(fig:abstract), the highest
-accuracy is achieved when sampling at a 50%/50% ratio. Therefore, we
-continue to recommend that building owners follow the proposed framework
-and maintain a 50%/50% sampling ratio until all stopping criteria are
-met. The additional results from Figure \@ref(fig:abscont) further add
-to the guidance: if building baseline changes, retrofit savings
-estimation should simply discard the existing baseline measurements and
-revert back to 50%/50% sampling ratio as described in our framework. On
-the other hand, if the building owner has decided to test for two years
-and would like to capitalize more savings at the same time, sampling
-between the baseline and intervention at 10%/90% from start is
-recommended.
+As shown in the previous overall M&V method comparison Figure
+\@ref(fig:abstract) in the Results section (plotted again on the left
+side of the red dashed line), the highest accuracy is achieved when
+sampling at a 50%/50% ratio. Therefore, we continue to recommend that
+building owners follow the proposed framework and flow chart from our
+previous study [@raftery2024], and maintain a 50%/50% sampling ratio
+until all stopping criteria are met. If building owners choose to adjust
+the sampling ratio afterwards and a change in the building baseline is
+detected, the updated savings estimation should exclude the baseline
+measured prior to the sampling ratio change. As shown in the figure,
+additional baseline measurements before the ratio change do not
+contribute to reducing errors or uncertainty. The results also indicate
+that, if the building owner plans to conduct a two-year test and aims to
+maximize savings, it is recommended to implement a 10%/90% sampling
+ratio between the baseline and intervention from the outset.
 
@@
 
-Selecting different blocking periods affects how frequently the M&V
-analyst evaluates stopping criteria and may influence the decision when
-a savings is detected at a sufficient level of certainty. Here, we
-summarize the advantages and disadvantages of shorter versus longer
-blocking periods. As discussed in our previous study, the second step of
-this randomized M&V method involves choosing an appropriate blocking
-period to minimize potential biases caused by incomplete randomization.
-For most practical scenarios, we recommend a blocking period of
-approximately 12 to 16 weeks, as it typically covers the duration of a
-season, ensuring a representative sampling of operational conditions. In
-general, longer blocking periods are advantageous when evaluating
-multiple intervention strategies or using sampling intervals greater
-than one day. In such cases, extended blocking ensures a balanced
-distribution of both day-of-week and outdoor weather conditions.
+A blocking period refers to the given period during which an equal
+number of baseline and intervention days are sampled, and the associated
+day-of-week factors are balanced between the two. Selecting different
+blocking periods affects how frequently the M&V analyst evaluates
+stopping criteria and may influence the decision when the saving is
+detected at a sufficient level of certainty. Here, we summarize the
+advantages and disadvantages of shorter versus longer blocking periods.
+The second step of the randomized M&V method involves choosing an
+appropriate blocking period to minimize potential biases caused by
+incomplete randomization. For most practical scenarios, we recommend a
+blocking period of approximately 12 to 16 weeks, as it typically covers
+the duration of a season, ensuring a representative sampling of
+operational conditions. In general, longer blocking periods are
+advantageous when evaluating multiple intervention strategies or using
+sampling intervals greater than one day. In such cases, extended
+blocking ensures a balanced distribution of both day-of-week and outdoor
+weather conditions.
 
 Opting for a shorter blocking period, such as 6 weeks, has the potential
-benefit of allowing the stopping criteria to be met earlier. We plotted
-the percentage of buildings finishing all stopping criteria when the
-blocking period is set as 6 weeks in Figure \@ref(fig:6weekblock). As a
-result, when compared to Figure \@ref(fig:intervaltime), the only
-difference we found is when sampling at a daily interval and not
-dropping non-consecutive days, 12% of buildings can get an estimate
-earlier. This means those buildings are likely to meet all stopping
-criteria between 12 weeks and 18 weeks but when sampled over a 12-week
-block, the savings can only be determined at week 24. Although sampled
-baseline and intervention days can be different when the blocking period
-is changed, the results indicate most buildings still require 9 months
-to meet outdoor weather conditions when starting M&V from January.
+benefit of allowing the stopping criteria to be met earlier. We assessed
+the use of a 6-week blocking period to see whether buildings finished
+earlier in the original 12-week blocking period can reduce the time
+required with a shorter blocking period. We plotted the percentage of
+buildings finishing all stopping criteria when the blocking period is
+set as 6 weeks in Figure \@ref(fig:6weekblock). When compared to Figure
+\@ref(fig:intervaltime), the only difference we found is when sampling
+at a daily interval and not dropping non-consecutive days, 12% of
+buildings can get an estimate earlier. This means those buildings are
+likely to meet all stopping criteria between 12 weeks and 18 weeks but
+when sampled over a 12-week block, the savings can only be determined at
+week 24. Although sampled baseline and intervention days can be
+different when the blocking period is changed, the results indicate most
+buildings still require 9 months to meet outdoor weather conditions when
+starting M&V from January.
 
@@
 period would be less detrimental compared to longer blocking periods,
-preserving the overall integrity of the experiment.
+preserving the overall integrity of the experiment. Based on these
+analyses we recommend to use a blocking period of 6 weeks when sampled
+at a daily interval, but a longer blocking period such as 12 weeks when
+a longer sampling interval is required.
 
@@
 To assess this impact in our switchback randomization, we repeated the
-randomized M&V analysis for all buildings using a new start date in
-March. As shown in Figure \@ref(fig:startT), the new M&V process results
-in a noticeable delay in satisfying the stopping criteria. Compared to
-Figure \@ref(fig:intervaltime), where a January start allowed most
-buildings to finish within 36 weeks, a March start extended the required
-duration to 48 weeks for the majority of buildings. Furthermore, when a
-longer sampling interval is used or non-consecutive days are excluded
-(to avoid carryover effects), some buildings may require up to 60 weeks
-to complete the randomized M&V. As expected, to collect sufficient data
+randomized M&V analysis for all buildings using a new start date in the
+swing season in March. As shown in Figure \@ref(fig:startT), the new M&V
+process results in a noticeable delay in satisfying the stopping
+criteria. Compared to Figure \@ref(fig:intervaltime), where a January
+start allowed most buildings to finish within 36 weeks (\~8 months), a
+March start extended the required duration to 48 weeks (\~11 months) for
+the majority of buildings. Furthermore, when a longer sampling interval
+is used or non-consecutive days are excluded (to avoid carryover
+effects), some buildings may require up to 60 weeks (\~14 months) to
+complete the randomized M&V. As expected, to collect sufficient data
 representing winter conditions, the sampling may need to continue until
@@
 
-Aside from the general limitation that this method only applies to the
-subset of interventions in buildings that are easily switchable, we
-identify two key limitations in this study:
-
-1.  The simulated intervention (chilled water temperature reset) remains
-    somewhat generic and simplistic, considering the diverse building
-    types and climate zones in the BDG2 dataset. As the primary focus is
-    to accurately detect the intervention (or any intervention) effect
-    so for simplicity, we applied the parameters listed in Table 1
-    uniformly across all buildings.
-2.  For the main analysis, we assumed no carryover effect and that a
-    daily sampling interval would suffice for most commercial buildings,
-    but certainly exceptions exist. However, in the discussion, we
-    addressed such concern by comparing different sampling strategies
-    and the impact of dropping measurements from non-consecutive days.
+We identify two key limitations in this study. First, the simulated
+intervention (chilled water temperature reset) remains somewhat generic
+and simplistic, considering the diverse building types and climate zones
+in the BDG2 dataset. As the primary focus is to accurately detect the
+intervention (or any intervention) effect so for simplicity, we applied
+the parameters listed in Table 1 uniformly across all buildings.
+Secondly, for the main analysis, we assumed no carryover effect and that
+a daily sampling interval would suffice for most commercial buildings,
+but certainly exceptions exist. However, in the sensitivity analysis, we
+addressed such concern by comparing different sampling strategies and
+the impact of dropping measurements from non-consecutive days.
 
@@
 
-This research demonstrated the application of a randomized
-whole-building measurement and verification (M&V) method, comparing its
-performance to the conventional approach described in the IPMVP and in
-ASHRAE Guideline 14, using a large, open-source commercial building
-dataset. The randomized M&V method leverages the randomized experimental
-design concept from other scientific fields, along with statistical
-sequential inference techniques, to determine when savings are detected
-at sufficient certainty. We used a virtual control retrofit
-case---resetting the chilled water setpoint based on outdoor weather
-conditions---and applied it to over 600 filtered commercial buildings.
-By comparing the savings estimations of the conventional method with the
-novel randomized method, we found that the randomized approach provides
-faster and more robust savings estimations.
+We assessed the performance of a randomized whole-building measurement
+and verification (M&V) method by comparing it to the conventional M&V
+approach described in the IPMVP and in ASHRAE Guideline 14, using a
+large, open-source commercial building dataset. The randomized M&V
+method leverages the randomized experimental design concept from other
+scientific fields, along with statistical sequential inference
+techniques, to determine when savings are detected at sufficient
+certainty. We used a virtual control retrofit case, resetting the
+chilled water setpoint based on outdoor temperature, and applied it to
+over 600 commercial buildings. By comparing the savings estimations of
+the conventional method with the novel randomized method, we found that
+the randomized approach provides faster and more robust savings
+estimations.
 
@@
 full range of baseline and intervention measurements under normal
-operating conditions, typically taking 6 - 9 months for each phase. Most
-importantly, we verified that with a reduced M&V timeline, the
-randomized method can estimate savings more accurately by showing that
-the absolute error is only 1 - 2% for a typical building. Whereas for
-the much longer two-year conventional method, the estimation error was
-approximately 5% in a typical building. We also evaluated the impact of
-non-routine events, especially more gradual and subtle operational drift
-on the proposed M&V method and the results show that baseline changes in
-the post-retrofit period can deviate savings estimated using the
-conventional method. In contrast, we found that those events have a very
-negligible impact on the savings estimated using the randomized method,
-demonstrating robustness to this issue. Furthermore, we consider
-scenarios where there is a known carryover effect from switching between
-strategies in the building and assess the impact of dropping samples
-when the strategy is non-consecutive and varying the sampling interval.
-Through that process, we showed that for the typical building, if
-carryover is present, using a 2-day sampling interval and dropping days
-with non-consecutive control strategies operating yields the optimal
-design for the median building in this dataset. Lastly, we investigated
-the time required to complete the randomized M&V method under different
-scenarios. We found that measuring a sufficient range of outdoor weather
-conditions remains the most stringent criterion. However, even in the
-most non-ideal case, the randomized method can still produce an estimate
-much faster than two years, and produces an estimate with far lower
-error.
+operating conditions, typically requiring each operational strategy to
+be measured for a full year. Moreover, we verified that, even if the
+randomized M&V requires less time, it is more accurate showing an
+absolute error of 1 - 2%. Whereas for the much longer two-year
+conventional method, the estimation error was approximately 5%.
+
+We also evaluated the impact of non-routine events, particularly more
+gradual and subtle operational drift, on the proposed M&V method; the
+results show that baseline changes in the post-retrofit period are very
+common for realistic buildings and can deviate savings estimated using
+the conventional method. In contrast, we found that those events have a
+very negligible impact on the savings estimated using the randomized
+method, demonstrating robustness to this issue.
+
+Furthermore, we considered scenarios where there is a known carryover
+effect from switching between strategies in the building and assessed
+the impact of dropping samples when the strategy is non-consecutive and
+varying the sampling interval. Through that process, we showed that for
+the typical building, when carryover is present, using a 2-day sampling
+interval and excluding days with non-consecutive control strategies
+operating yields the optimal design for the median building in this
+dataset.
+
+Lastly, we investigated the sensitivity of time required to complete the
+randomized M&V method under different scenarios. We found that measuring
+a sufficient range of outdoor weather conditions remains the most
+stringent criterion. However, even in the most non-ideal case, the
+randomized method can still produce an estimate much faster than two
+years and  and with significantly lower error.
 
@@
 administration, Supervision, Writing - review & editing. **Stefano
-Schiavon**: Conceptualization, Methodology, Investigation, Supervision,
-Writing - review & editing. **Carlos Duarte**: Methodology,
-Investigation, Supervision, Writing - review & editing. **Gail Brager**:
-Supervision, Writing - review & editing.
+Schiavon**: Conceptualization, Methodology, Visualizations,
+Investigation, Supervision, Writing - review & editing. **Carlos
+Duarte**: Methodology, Investigation, Supervision, Writing - review &
+editing. **Gail Brager**: Supervision, Writing - review & editing.
 
@@
 
-The authors declare that they have no known competing financial
-interests or personal relationships that could have appeared to
-influence the work reported in this paper.
+The Center for the Built Environment (CBE) at the University of
+California, Berkeley, with which the authors are affiliated, is advised,
+and funded in part, by many partners that represent a diversity of
+organizations from the building industry, including manufacturers,
+building owners, facility managers, contractors, architects, engineers,
+government agencies, and utilities.
 

### 2025-07-15 code/supplementary.Rmd
updaded review for manuscript and updated readme file

@@
 
-```{r meanerr, fig.cap="Distribution of estimated deviation from the ground-truth savings calculated by the two M&V methods (red dashed line highlighting 0% error)", fig.height=7, fig.width=9}
+```{r meanerr, fig.cap="Distribution of estimated deviation (without absolute calculation) from the ground-truth savings calculated by the two M&V methods (red dashed line highlighting 0% error)", fig.height=7, fig.width=9}
 # plot for the stable subset
@@
 
-```{r absnull, fig.cap="Comparison between the two M&V methods in detecting no intervention effect (expressed without absolute calculation) when buildings are subject to baseline change.", fig.height=4, fig.width=9}
+```{r absnull, fig.cap="Comparison between the two M&V methods in detecting no intervention effect when buildings are subject to baseline change.", fig.height=4, fig.width=9}
 #### ABS-NULL ####
@@
 
-```{r meancont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy (expressed without absolute calculation for errors)", fig.height=7, fig.width=9}
+```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months (previous results of sampling at 50%/50% shown on the left side of the red dashed line)", fig.height=7, fig.width=9}
+
+df_MW_S <- df_cont_stable %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - cont_fs), 
+         diff_all = abs(savings - cont_fs_all)) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_stable %>% 
+              filter(method != "conv") %>% 
+              pivot_wider(names_from = method, values_from = savings) %>% 
+              mutate(values = abs(true - rand), 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
+
+# plot for the variable subset
+df_MW_V <- df_cont_variable %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - cont_fs), 
+         diff_all = abs(savings - cont_fs_all)) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_variable %>% 
+              filter(method != "conv") %>% 
+              pivot_wider(names_from = method, values_from = savings) %>% 
+              mutate(values = abs(true - rand), 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
+
+# plot for combined dataset
+df_MW_A <- rbind(df_MW_S, df_MW_V)
+
+p_top <- df_MW_A %>% 
+  mutate(ratio = as.factor(ratio), 
+         ratio = recode_factor(ratio, 
+                               "0" = "Randomized\n(50/50)", 
+                               "1_1" = "Continued\n(20/80)", 
+                               "1_2" = "Randomized\n(20/80)", 
+                               "2_1" = "Continued\n(10/90)", 
+                               "2_2" = "Randomized\n(10/90)")) %>% 
+  ggplot(aes(x = ratio, y = values, fill = ratio)) +
+  geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75) +
+  geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% 
+              group_by(ratio) %>% 
+              summarise(median = median(values)) %>% 
+              ungroup(), 
+            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %")), 
+            size = 5) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80", "grey80")) + 
+  labs(fill = NULL, 
+       x = NULL, 
+       y = "Absolute error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with measured weather conditions")) +
+  coord_cartesian(ylim = c(0, 15)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "none",
+        axis.text.x = element_blank(), 
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+# TMY
+# plot for the stable subset
+df_TW_S <- df_cont_stable %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - cont_tmy), 
+         diff_all = abs(savings - cont_tmy_all)) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_tmy_stable %>% 
+              left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+              mutate(values = abs(savings - rand), 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
+
+# plot for the variable subset
+df_TW_V <- df_cont_variable %>% 
+  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - cont_tmy), 
+         diff_all = abs(savings - cont_tmy_all)) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_tmy_variable %>% 
+              left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+              mutate(values = abs(savings - rand), 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
+
+# plot for combined dataset
+df_TW_A <- rbind(df_TW_S, df_TW_V)
+
+p_bottom <- df_TW_A %>% 
+  mutate(ratio = as.factor(ratio), 
+         ratio = recode_factor(ratio, 
+                               "0" = "Randomized\n(50/50)", 
+                               "1_1" = "Continued\n(20/80)", 
+                               "1_2" = "Randomized\n(20/80)", 
+                               "2_1" = "Continued\n(10/90)", 
+                               "2_2" = "Randomized\n(10/90)")) %>% 
+  ggplot(aes(x = ratio, y = values, fill = ratio)) +
+  geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75) +
+  geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% 
+              group_by(ratio) %>% 
+              summarise(median = median(values)) %>% 
+              ungroup(), 
+            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %")), 
+            size = 5) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80", "grey80")) + 
+  labs(fill = NULL, 
+       x = NULL, 
+       y = "Absolute error in fractional savings", 
+       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
+  coord_cartesian(ylim = c(0, 15)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "none",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+ggarrange(p_top, p_bottom, 
+          ncol = 1, nrow = 2,
+          labels = c("a)", "b)")) +
+  plot_annotation(title = "Absolute error in savings estimation by different sampling ratio\n(after 2 years)")
+```
+
+```{r stablecont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy using buildings with stable electricity usage over the entire 24 months (results of sampling at 50%/50% shown on the left side of the red dashed line)", fig.height=7, fig.width=9}
+
+#### ABS-CONT ####
+df_MW_S <- df_cont_stable %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - cont_fs), 
+         diff_all = abs(savings - cont_fs_all)) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_stable %>% 
+              filter(method != "conv") %>% 
+              pivot_wider(names_from = method, values_from = savings) %>% 
+              mutate(values = abs(true - rand), 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
+
+p_top <- df_MW_S %>% 
+  mutate(ratio = as.factor(ratio), 
+         ratio = recode_factor(ratio, 
+                               "0" = "Randomized\n(50/50)", 
+                               "1_1" = "Continued\n(20/80)", 
+                               "1_2" = "Randomized\n(20/80)", 
+                               "2_1" = "Continued\n(10/90)", 
+                               "2_2" = "Randomized\n(10/90)")) %>% 
+  ggplot(aes(x = ratio, y = values, fill = ratio)) +
+  geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75) +
+  geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% 
+              group_by(ratio) %>% 
+              summarise(median = median(values)) %>% 
+              ungroup(), 
+            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %")), 
+            size = 5) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80", "grey80")) + 
+  labs(fill = NULL, 
+       x = NULL, 
+       y = "Absolute error in fractional savings", 
+       subtitle = str_glue("{S_building} buildings in the stable subset with measured weather conditions")) +
+  coord_cartesian(ylim = c(0, 8)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "none",
+        axis.text.x = element_blank(), 
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+# TMY
+# plot for the stable subset
+df_TW_S <- df_cont_stable %>% 
+  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+  mutate(diff = abs(savings - cont_tmy), 
+         diff_all = abs(savings - cont_tmy_all)) %>% 
+  select(name, diff, diff_all, ratio) %>% 
+  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
+  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
+                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
+                               ifelse(ratio == 2 & type == "diff", "2_1", 
+                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
+  bind_rows(df_fs_tmy_stable %>% 
+              left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
+              mutate(values = abs(savings - rand), 
+                     ratio = "0", 
+                     type = "diff_ref") %>% 
+              select(name, values, ratio, type))
+
+p_bottom <- df_TW_S %>% 
+  mutate(ratio = as.factor(ratio), 
+         ratio = recode_factor(ratio, 
+                               "0" = "Randomized\n(50/50)", 
+                               "1_1" = "Continued\n(20/80)", 
+                               "1_2" = "Randomized\n(20/80)", 
+                               "2_1" = "Continued\n(10/90)", 
+                               "2_2" = "Randomized\n(10/90)")) %>% 
+  ggplot(aes(x = ratio, y = values, fill = ratio)) +
+  geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
+  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75) +
+  geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
+  geom_text(data = . %>% 
+              group_by(ratio) %>% 
+              summarise(median = median(values)) %>% 
+              ungroup(), 
+            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %")), 
+            size = 5) +
+  scale_y_continuous(expand = c(0, 0), 
+                     breaks = breaks_pretty(n = 4), 
+                     labels = number_format(suffix = " %")) +
+  scale_fill_manual(values = ls_colors) +
+  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80", "grey80")) + 
+  labs(fill = NULL, 
+       x = NULL, 
+       y = "Absolute error in fractional savings", 
+       subtitle = str_glue("{S_building} buildings in the stable subset with TOWT model and TMY weather")) +
+  coord_cartesian(ylim = c(0, 8)) +
+  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
+        legend.position = "none",
+        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
+
+ggarrange(p_top, p_bottom, 
+          ncol = 1, nrow = 2,
+          labels = c("a)", "b)")) +
+  plot_annotation(title = "Absolute error in savings estimation by different sampling ration(after 2 years)")
+```
+
+```{r meancont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy at the time of satisfying all stopping criteria (expressed without absolute calculation for errors)", fig.height=7, fig.width=9}
 
@@
        x = NULL, 
-       y = "Absolute error in fractional savings", 
+       y = "Signed error in fractional savings", 
        subtitle = str_glue("All {A_building} buildings with measured weather conditions")) +
@@
        x = NULL, 
-       y = "Absolute error in fractional savings", 
+       y = "Signed error in fractional savings", 
        subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
@@
           labels = c("a)", "b)")) +
-  plot_annotation(title = "Absolute error in savings estimation by different sampling ratio")
-```
-
-```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months (previous results of sampling at 50%/50% shown on the left side of the red dashed line)", fig.height=7, fig.width=9}
-
-#### ABS-CONT ####
-df_MW_S <- df_cont_stable %>% 
-  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = abs(savings - cont_fs), 
-         diff_all = abs(savings - cont_fs_all)) %>% 
-  select(name, diff, diff_all, ratio) %>% 
-  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
-  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
-                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
-                               ifelse(ratio == 2 & type == "diff", "2_1", 
-                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
-  bind_rows(df_fs_stable %>% 
-              filter(method != "conv") %>% 
-              pivot_wider(names_from = method, values_from = savings) %>% 
-              mutate(values = abs(true - rand), 
-                     ratio = "0", 
-                     type = "diff_ref") %>% 
-              select(name, values, ratio, type))
-
-# plot for the variable subset
-df_MW_V <- df_cont_variable %>% 
-  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = abs(savings - cont_fs), 
-         diff_all = abs(savings - cont_fs_all)) %>% 
-  select(name, diff, diff_all, ratio) %>% 
-  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
-  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
-                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
-                               ifelse(ratio == 2 & type == "diff", "2_1", 
-                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
-  bind_rows(df_fs_variable %>% 
-              filter(method != "conv") %>% 
-              pivot_wider(names_from = method, values_from = savings) %>% 
-              mutate(values = abs(true - rand), 
-                     ratio = "0", 
-                     type = "diff_ref") %>% 
-              select(name, values, ratio, type))
-
-# plot for combined dataset
-df_MW_A <- rbind(df_MW_S, df_MW_V)
-
-p_top <- df_MW_A %>% 
-  mutate(ratio = as.factor(ratio), 
-         ratio = recode_factor(ratio, 
-                               "0" = "Randomized\n(50/50)", 
-                               "1_1" = "Continued\n(20/80)", 
-                               "1_2" = "Randomized\n(20/80)", 
-                               "2_1" = "Continued\n(10/90)", 
-                               "2_2" = "Randomized\n(10/90)")) %>% 
-  ggplot(aes(x = ratio, y = values, fill = ratio)) +
-  geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75) +
-  geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
-  geom_text(data = . %>% 
-              group_by(ratio) %>% 
-              summarise(median = median(values)) %>% 
-              ungroup(), 
-            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %")), 
-            size = 5) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 4), 
-                     labels = number_format(suffix = " %")) +
-  scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80", "grey80")) + 
-  labs(fill = NULL, 
-       x = NULL, 
-       y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with measured weather conditions")) +
-  coord_cartesian(ylim = c(0, 15)) +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.position = "none",
-        axis.text.x = element_blank(), 
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-# TMY
-# plot for the stable subset
-df_TW_S <- df_cont_stable %>% 
-  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = abs(savings - cont_tmy), 
-         diff_all = abs(savings - cont_tmy_all)) %>% 
-  select(name, diff, diff_all, ratio) %>% 
-  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
-  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
-                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
-                               ifelse(ratio == 2 & type == "diff", "2_1", 
-                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
-  bind_rows(df_fs_tmy_stable %>% 
-              left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-              mutate(values = abs(savings - rand), 
-                     ratio = "0", 
-                     type = "diff_ref") %>% 
-              select(name, values, ratio, type))
-
-# plot for the variable subset
-df_TW_V <- df_cont_variable %>% 
-  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-  mutate(diff = abs(savings - cont_tmy), 
-         diff_all = abs(savings - cont_tmy_all)) %>% 
-  select(name, diff, diff_all, ratio) %>% 
-  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
-  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
-                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
-                               ifelse(ratio == 2 & type == "diff", "2_1", 
-                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
-  bind_rows(df_fs_tmy_variable %>% 
-              left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
-              mutate(values = abs(savings - rand), 
-                     ratio = "0", 
-                     type = "diff_ref") %>% 
-              select(name, values, ratio, type))
-
-# plot for combined dataset
-df_TW_A <- rbind(df_TW_S, df_TW_V)
-
-p_bottom <- df_TW_A %>% 
-  mutate(ratio = as.factor(ratio), 
-         ratio = recode_factor(ratio, 
-                               "0" = "Randomized\n(50/50)", 
-                               "1_1" = "Continued\n(20/80)", 
-                               "1_2" = "Randomized\n(20/80)", 
-                               "2_1" = "Continued\n(10/90)", 
-                               "2_2" = "Randomized\n(10/90)")) %>% 
-  ggplot(aes(x = ratio, y = values, fill = ratio)) +
-  geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
-  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75) +
-  geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
-  geom_text(data = . %>% 
-              group_by(ratio) %>% 
-              summarise(median = median(values)) %>% 
-              ungroup(), 
-            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %")), 
-            size = 5) +
-  scale_y_continuous(expand = c(0, 0), 
-                     breaks = breaks_pretty(n = 4), 
-                     labels = number_format(suffix = " %")) +
-  scale_fill_manual(values = ls_colors) +
-  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80", "grey80")) + 
-  labs(fill = NULL, 
-       x = NULL, 
-       y = "Absolute error in fractional savings", 
-       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
-  coord_cartesian(ylim = c(0, 15)) +
-  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
-        legend.position = "none",
-        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
-
-ggarrange(p_top, p_bottom, 
-          ncol = 1, nrow = 2,
-          labels = c("a)", "b)")) +
-  plot_annotation(title = "Absolute error in savings estimation by different sampling ratio")
+  plot_annotation(title = "Absolute error in savings estimation by different sampling ration\n(after satisfying all stopping criteria)")
 ```
@@
           legend = "bottom") +
-  plot_annotation(title = "Absolute error in savings estimation of ranodmized M&V\nafter 2 years")
+  plot_annotation(title = "Absolute error in savings estimation of ranodmized M&V\n(after 2 years)")
 ```
@@
        x = NULL, 
-       y = "Esimated error in fractional savings", 
+       y = "Signed error in fractional savings", 
        subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
@@
           legend = "bottom") +
-  plot_annotation(title = "Estimated error in savings estimation of randomized M&V\nafter satisfying all stopping criteria")
+  plot_annotation(title = "Signed error in savings estimation of randomized M&V\n(after satisfying all stopping criteria)")
 
@@
           legend = "bottom") +
-  plot_annotation(title = "Estimated error in savings estimation of randomized M&V\nafter 2 years")
+  plot_annotation(title = "Signed error in savings estimation of randomized M&V\n(after 2 years)")
 ```
